May 2025 - Question 1

This exercise is a comprehensive demonstration of **simple linear regression**, a statistical method used to model the relationship between two variables. The specific goal is to analyze how "humidity" influences "runs scored" by a sports team.

The process unfolds in a logical sequence:

1.  **Data Preparation and Visualization:** The exercise begins by reading data and creating a scatter plot to visually explore the relationship between the two variables.

2.  **Model Building:** A linear regression model is built to quantify this relationship, and the core statistical results (including the regression equation and R-squared value) are presented.

3.  **Prediction and Diagnostics:** The newly built model is then used to predict an outcome for a new data point, and a residual plot is created to visually verify that the model's underlying assumptions are valid.

4.  **Interpretation and Inference:** The final steps involve interpreting the key statistical outputs, such as the R-squared value and confidence intervals, and performing a formal hypothesis test (ANOVA) to confirm that the observed relationship is statistically significant and not due to random chance.

In short, the exercise is a complete workflow for applying and interpreting a linear regression model, moving from initial data exploration to formal statistical conclusions.

---

Q. 1)

This question involves analyzing the relationship between runs and humidity.

**i) Data Import**
The data is imported from a CSV file named "runs.csv".

```{r}
runs <- read.csv("C:\\ runs.csv")
```

**ii) Plotting the Data**
A scatter plot is created to visualize the relationship between runs and humidity.

```{r}
plot(runs, main = "Plot of runs vs humidity")
```

**iii) Correlation Analysis**
The Pearson correlation coefficient, $r$, is calculated to quantify the linear relationship.

```{r}
r <- cor(runs$Runs, runs$Humidity, method = "pearson")
```

$r = -0.7942107$
There appears to be a strong negative correlation between the humidity level and the runs scored by the team.

**iv) Linear Regression Model**
A linear regression model is fitted to predict runs based on humidity.

```{r}
model_runs <- lm(Runs ~ Humidity, data = runs)
summary(model_runs)
```

The summary output is:

```
Call:
lm(formula = Runs ~ Humidity, data = runs)
Residuals:
    Min      1Q  Median      3Q     Max 
-37.05  -14.49   -1.82   14.95   44.73 
Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept) 377.4457    32.2818  11.692 2.74e-12 ***
Humidity     -5.6969     0.8237  -6.916 1.62e-07 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Residual standard error: 20.38 on 28 degrees of freedom
Multiple R-squared: 0.6308,	Adjusted R-squared: 0.6176 
F-statistic: 47.83 on 1 and 28 DF,  p-value: 1.616e-07
```

The regression equation is:
$$\text{Runs} = 377.45 - 5.6969 \times \text{Humidity}$$

**v) Adding the Regression Line to the Plot**
The regression line is added to the scatter plot to visualize the model's fit.

```{r}
plot(runs, main = "Plot of runs vs humidity")
abline(model_runs)
```

**vi) Prediction**
The model is used to predict the number of runs when the humidity is 48.

```{r}
predict(model_runs, newdata = data.frame(Humidity = 48))
```

The predicted value is approximately 103.9961 runs.

**vii) Residual Analysis**
A residual plot is created to check the assumptions of the linear model.

```{r}
plot(model_runs$fitted.values, model_runs$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residual Plot")
abline(h = 0)
```

The residuals are randomly scattered around 0, which suggests that a linear model is appropriate.

**viii) $R^2$ Value**
The $R^2$ value is extracted from the model summary.

```{r}
summary(model_runs)$r.squared
```

$R^2 = 0.6307707$
This means that 63% of the variability in the number of runs can be explained by the humidity level, according to the model. The model does a moderately good job of explaining the relationship.

**ix) Confidence Intervals**
The confidence intervals for the model coefficients are calculated.

```{r}
confint(model_runs)
```

```
                 2.5 %     97.5 %
(Intercept) 311.319524 443.571913
Humidity    -7.384142  -4.009593
```

The 95% confidence interval for the intercept is $[311.31, 443.57]$.
The 95% confidence interval for the slope is $[-7.38, -4.01]$.

**x) Hypothesis Testing (ANOVA)**
An ANOVA test is performed to determine if there is a statistically significant linear relationship between humidity and runs.
The null and alternative hypotheses are:
$H\_0$: There is no linear relationship ($\\beta = 0$)
$H\_1$: There is a linear relationship ($\\beta \\neq 0$)
The ANOVA table is generated:

```{r}
anova_model <- anova(model_runs)
print(anova_model)
```

```
Analysis of Variance Table
Response: Runs
          Df  Sum Sq Mean Sq F value    Pr(>F)    
Humidity   1 19862.1 19862.1  47.834 1.616e-07 ***
Residuals 28 11626.0   415.2                      
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

The p-value is $1.616494 \times 10^{-7}$.
Since the p-value ($\< 0.05$) is very small, we reject the null hypothesis ($H\_0$). We conclude that humidity has a statistically significant linear relationship with the runs scored.