May 2025 - Question 2

This exercise consists of two distinct parts, each addressing a different type of statistical analysis.

**Part 1: Chi-Squared Test of Independence (Steps i & ii)**
This section focuses on analyzing categorical data. The goal is to determine if there is a statistically significant relationship between two variables: "exercise habit" (e.g., Sedentary, Very Active) and "workout intensity" (e.g., Light, Moderate, High). A contingency table is first created to organize the data, and then a chi-squared test is performed. The small p-value from this test leads to the conclusion that the two variables are not independent—in other words, a person's exercise habit is related to their workout intensity.

**Part 2: Paired Sample Analysis (Steps iii through viii)**
This section shifts to analyzing paired numerical data, specifically a person's cholesterol levels measured "before" and "after" a workout routine.
* **Initial Analysis:** The exercise starts by calculating and comparing the mean cholesterol levels to observe an initial reduction.
* **Correlation:** It then explores the relationship between the "before" and "after" measurements using different correlation coefficients (Pearson, Spearman, and Kendall), noting the distinctions between them.
* **Hypothesis Testing:** The core of this part is a series of paired t-tests.
    * A **two-sided t-test** (step vii) is used to see if *any* significant change occurred.
    * A **one-sided t-test** (step viii) is then used to specifically test the claim that the cholesterol levels were *reduced*. The very small p-value from this test provides strong statistical evidence to support the claim that the workout routine is effective.

---

In summary, the exercise provides a practical, two-part guide: the first half demonstrates how to test for a relationship between categorical variables, while the second half shows how to use a paired t-test to assess the impact of an intervention (like a workout routine) on a continuous variable.

---

**i) Data Representation**
A contingency table is created to display the relationship between exercise habits and workout intensity.

```{r}
exercise_data <- matrix(c(38, 120, 42,
                          232, 357, 73,
                          48, 29, 4,
                          15, 34, 8),
                        nrow = 4, byrow = TRUE)
rownames(exercise_data) <- c("Sedentary", "Moderately Active", "Very Active", "Extremely Active")
colnames(exercise_data) <- c("Light", "Moderate", "High")
print(exercise_data)
```

The resulting table is:

```
                  Light Moderate High
Sedentary            38      120   42
Moderately Active   232      357   73
Very Active          48       29    4
Extremely Active     15       34    8
```

**ii) Chi-Squared Test of Independence**
A chi-squared test is performed to determine if there is a relationship between exercise habits and workout intensity.
$H\_0$: Exercise habit and workout intensity are independent.
$H\_1$: Exercise habit and workout intensity are not independent.

```{r}
chisq.test(exercise_data)
```

The test results are:

```
Pearson's Chi-squared test
data: exercise_data
X-squared = 52.982, df = 6, p-value = 1.183e-09
```

The p-value is $1.183 \\times 10^{-9}$. Since the p-value ($\< 0.05$) is very small, we reject the null hypothesis ($H\_0$). We conclude that exercise habits and workout intensity are not independent.

**iii) Mean Cholesterol Levels**
The mean cholesterol levels are calculated for a group of people both before and after a new workout routine.

```{r}
before <- c(9.1, 8.0, 7.7, 10.0, 9.6, 7.9, 9.0, 7.1, 8.3, 9.6)
after <- c(8.2, 6.4, 6.6, 8.5, 8.0, 5.8, 7.8, 7.2, 6.7, 9.8)
mean(before)
mean(after)
```

The mean "before" cholesterol level is 8.63.
The mean "after" cholesterol level is 7.5.

**iv) Correlation Coefficients**
Three different correlation coefficients are calculated to measure the relationship between the "before" and "after" cholesterol levels.

```{r}
# Pearson
cor(before, after, method = "pearson")
# Spearman
cor(before, after, method = "spearman")
# Kendall
cor(before, after, method = "kendall")
```

Pearson: 0.7764615
Spearman: 0.8024353
Kendall: 0.6292532

**v) Interpretation of Correlation Coefficients**

  * **Pearson's correlation coefficient** measures the strength and direction of a *linear* relationship using the raw numerical data.
  * **Spearman's rank correlation** and **Kendall's Tau** measure the strength of a *monotonic* relationship. They are non-parametric methods that use the ranks of the data rather than the raw values. These are often more appropriate when the data is not perfectly linear or when the assumption of normality is not met, as they are less sensitive to outliers.

**vi) Observation from Mean Values**
The mean cholesterol level appears to have decreased after the workout routine.

**vii) Paired Two-Sided t-test**
A paired t-test is performed to determine if there is a statistically significant change in cholesterol levels.
$H\_0$: $\\mu\_{\\text{before}} = \\mu\_{\\text{after}}$ (no change)
$H\_1$: $\\mu\_{\\text{before}} \\neq \\mu\_{\\text{after}}$ (a change has occurred)

```{r}
t.test(before, after, paired = TRUE, var.equal = TRUE)
```

The test results are:

```
Paired t-test
data: before and after
t = 4.7558, df = 9, p-value = 0.001036
alternative hypothesis: true mean difference is not equal to 0
95 percent confidence interval:
0.5925029 1.6674971
sample estimates:
mean difference 
        1.13
```

The p-value is 0.001036. Since the p-value ($\< 0.05$) is very small, we reject the null hypothesis ($H\_0$). We conclude that the cholesterol levels have changed after the workout routine.

**viii) Paired One-Sided t-test**
To specifically test the claim that the workout routine *reduced* cholesterol levels, a one-sided paired t-test is performed.
$H\_0$: $\\mu\_{\\text{before}} = \\mu\_{\\text{after}}$ (no reduction)
$H\_1$: $\\mu\_{\\text{before}} \> \\mu\_{\\text{after}}$ (a reduction has occurred)

```{r}
t.test(before, after, paired = TRUE, alternative = "greater", conf.level = 0.99)
```

The test results are:

```
Paired t-test
data: before and after
t = 4.7558, df = 9, p-value = 0.0005179
alternative hypothesis: true mean difference is greater than 0
99 percent confidence interval:
0.4596157       Inf
sample estimates:
mean difference 
           1.13 
```

The p-value is 0.0005179. Since the p-value ($\< 0.01$) is very small, we reject the null hypothesis ($H\_0$). We conclude that the cholesterol level has been statistically reduced after the workout routine.