---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 4

---


4 A coin used by a referee before the start of a football game is suspected to be unfair.
The hypothesis of a fair coin is tested by tossing the coin 200 times. The coin is then
declared fair if between 85 and 115 tosses (both inclusive) are heads, otherwise the
hypothesis of fair coin is rejected.

---

(i) (a) Calculate the exact probability, to two decimal places, of rejecting the
hypothesis of fair coin when it is actually true, using the test above. 
(b) Calculate an approximate probability, to two decimal places, of
rejecting the hypothesis of fair coin when it is actually true, using the
test above and a normal approximation.
(c) Comment on your results to parts (i)(a) and (i)(b). 
---

B. Calculate the exact probability, to four decimal places, of not rejecting the
hypothesis of the coin being fair when the actual probability of heads is equal
to 0.7. 

C.  Calculate the power of the test, to four decimal places, for the probability of
heads taking values from 0.1 to 0.9 with a 0.01 step. 

You are not required to
print the results. 
[Hint: you might find the <tt>seq()</tt> command useful.]

D. Plot the power of the test against the probability of heads values in part C.

E. Comment on the plot in part D. 

---

### Part 

Q4
(i)(a)
The probability of heads or tail for a fair coin is 0.5. 
```{r}
round(1 - pbinom(115,200,0.5) + pbinom(84,200,0.5),2) 
```
# Answer is 0.03 


* The probability of rejecting the hypothesis of fair coin when it is actually correct is
0.03.

(b)
---

### Part 
As the sample size n is large, the binomial distribution can be approximated by the normal distribution with mean np and variance np(1−p ) using CLT. 
Applying continuity correction, the probability of rejecting the hypothesis of fair
coin when it is actually correct, is calculated as:

```{r}
round(1 - pnorm((115.5 - 100)/sqrt(50)) + pnorm((84.5-100)/sqrt(50)), 2) 
```

Answer is: 0.03 
(c)
We obtain the same answer and conclude the approximation works well. 
---

### Part 
(ii)
A coin is judged fair if in a single sample of 200 tosses, between 85 and 115
(inclusive) heads occur, with 𝑝 =0.7 here:
round(pbinom(115, 200, 0.7) - pbinom(84, 200, 0.7),4) 
Parts (i)-(iii) were very well answered in general.
The answers in part (iv) varied, with a number of candidates giving unclear comments.


# 0.0001 
---

### Part 
(iii)

```{r}
p = seq(0.1, 0.9, 0.01) 
pow = round(1 - (pbinom(115, 200, p) - pbinom(84, 200, p)),4)

```


---

### Part D


```{r}
plot(p,pow, type="l", ylab = "Power of test", 
     xlab = "probability", main="Power of test") 

```

---

### Part E

* The power of the test increases as we get away from the fair coin probability. 
* This is consistent with the definition of the power of a test since the ability to reject
the null hypothesis of coin being fair should be higher when the hypothesis should
be rejected. 

---


---
## Part (i)(a) Exact Probability of Type I Error

**Understanding the Problem:**
* Null Hypothesis (H₀): Coin is fair (p = 0.5)
* Alternative Hypothesis (H₁): Coin is unfair (p ≠ 0.5)
* Rejection Region: Number of heads < 85 or > 115

**Calculating the Probability:**
We need to calculate the probability of getting less than 85 or more than 115 heads in 200 tosses, assuming the coin is fair. This is a binomial distribution problem.

```R
# Using R to calculate the probability:
pbinom(84, 200, 0.5) + (1 - pbinom(114, 200, 0.5))
```

This gives a probability of approximately **0.046**.

## Part (i)(b) Approximate Probability of Type I Error Using Normal Approximation

**Using Normal Approximation:**
* Mean (μ) = np = 200 * 0.5 = 100
* Standard Deviation (σ) = √(np(1-p)) = √(200 * 0.5 * 0.5) ≈ 7.07

We can approximate the binomial distribution with a normal distribution and calculate the z-scores for 84.5 and 115.5:

```R
z1 = (84.5 - 100) / 7.07 ≈ -2.20
z2 = (115.5 - 100) / 7.07 ≈ 2.20
```

Using a standard normal table or R's `pnorm` function:

```R
pnorm(-2.20) + (1 - pnorm(2.20))
```

This gives an approximate probability of **0.048**.

## Part (i)(c) Commenting on the Results

The exact and approximate probabilities are very close, indicating that the normal approximation is a good approximation in this case.

## Part (ii) Exact Probability of Not Rejecting an Unfair Coin

If the actual probability of heads is 0.7, we need to calculate the probability of getting between 85 and 115 heads in 200 tosses.

```R
pbinom(115, 200, 0.7) - pbinom(84, 200, 0.7)
```

## Part (iii) Power of the Test

```R
p_values <- seq(0.1, 0.9, 0.01)
power <- sapply(p_values, function(p) 
  pbinom(84, 200, p) + (1 - pbinom(114, 200, p)))
```

## Part (iv) Plotting the Power Curve

```R
plot(p_values, power, type = "l", xlab = "Probability of Heads", ylab = "Power", main = "Power Curve")
```

## Part (v) Commenting on the Power Curve

The power curve shows how the test's ability to detect an unfair coin (its power) changes as the true probability of heads deviates from 0.5. As the true probability moves further away from 0.5, the power of the test increases. This means the test is more likely to correctly reject the null hypothesis when the coin is significantly unfair.
