---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

3 A data science actuary wants to build a linear regression model for cars that will
predict the stopping distance based on their speed.

The actuary will work on the dataset that already exists in R known as ‘Cars’, which
can be initialised using the following code:
data("cars")
attach(cars)
This dataset has 50 observations of two variables:
*  The first variable ‘speed’ is speed, in miles per hour (mph).
*  The second variable ‘dist’ is stopping distance, in feet.

---

(i) Fit a linear regression model to the data using stopping distance as the
response variable and speed as the only explanatory variable, stating the
estimated intercept and slope of the regression line. 

The actuary wants to predict the stopping distances for the following ten speed values:
$$6 7 16 22 28 33 40 42 57 64$$


---

(ii) Estimate the ten expected stopping distances, based on your fitted model from
part (i). 
The actuary is concerned about the uncertainty around the mean predictions.
(iii) Calculate ten confidence intervals, one for each of the expected stopping
distances in part (ii), using a 90% confidence level. 
(iv) (a) State the interpretation of the 90% confidence interval for the stopping
distances for cars with speed 64 mph, using your answer to part (iii).
(b) Comment on the suitability of the linear regression model, based on the
confidence intervals obtained in part (iii).

---


Q3
```{r}
#Initialise the dataset.
data("cars")
attach(cars)
```

---

(i)

```{r}
model = lm(dist~speed, data = cars) 
model
```



---

(ii)

```{r}
data_predict <- data.frame(speed = c(6,7,16,22,28,33,40,42,57,64))
```


```{r}
predict(model,newdata=data_predict) 
```



---

(iii)

```{r}
predict(model,newdata=data_predict, interval="confidence",level=0.90) 
```



---

Part (i) was well answered in general. A common error was using rbinom(1, n, p), therefore not producing the detailed sample.
In part (ii) there were mixed answers with various errors.
Answers in part (iii) were also mixed, with common errors including not stating correctly, or at all, the hypotheses, parameter errors and not providing a clear conclusion.


7 139.717255 122.189498 157.24501
8 147.582073 128.688793 166.47535
9 206.568204 177.348219 235.78819
10 234.095066 200.029446 268.16069 

---

(iv)(a)

* The CI is constructed such that with repeated sampling, 90% of the obtained intervals will contain the true unknown mean stopping distance of cars travelling at 64 mph. 
(b)

* For speed 6mph the model gives a CI with negative lower endpoint. 
* A linear model, predicting negative stopping distance values, may not be
appropriate. 
