---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---


1 (i) Simulate 10,000 samples from an Exp(3) distribution, each of size n = 10,
using the inverse transform method. You should save the generated values in
R for later use.

* You must use the command <tt>set.seed(3202)</tt> to initialise the random
number generator, before you start the simulation. 

(ii) Plot a histogram of the means of the samples generated in part (i), using an
appropriate option in R for plotting the histogram on the probability density
scale. 

---

### Exercises

(iii) Plot the Probability Density Function (PDF) of the sampling distribution of the
sample mean, under the Central Limit Theorem (CLT), corresponding to the
samples generated in part (i). 

The graph of the PDF should be superimposed
on the histogram produced in part (ii). 

(iv) Comment on the sampling distribution of the sample mean and the application
of the CLT in this case based on your answers to parts (ii) and (iii). 

---
Q1
(i)
R code:

```{r}
set.seed(3202)
n = 10; theta = 3
m1 = matrix(0,nrow=10^4,ncol=n) # Initialise matrix, each sample in a row
for (i in 1:10^4){ 
u = runif(n,0,1) 
m= - log(1-u)/theta # Generate samples in rows 
m1[i,] = m 
}

```

---

(ii)
R Code:

```{r}
xbar.a = apply(m1,1,mean) 
hist(xbar.a,freq=F,main="Sampling distribution of Xbar")
```

---
(iii)
Using the CLT, the sampling distribution should be N(1/3, 1/90). 
R code:

```{r}
hist(xbar.a,freq=F,main="Sampling distribution of Xbar")
x.grid = seq(0,1,length=1000) 
lines(x.grid,dnorm(x.grid,1/3,sqrt(1/90))) 

```

---

The curve shows the exact pdf of the sampling distribution. 
(iv)


---

* The sampling distribution is relatively asymmetrical compared to the normal. 
* The CLT does not work very well here 
as the sample size is small (n = 10).

---

---

---