---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 5

---
5 A medical study examines the dependence of Intensive Care Unit (ICU) admission
rates on a numbers of risk factors, for patients with a certain medical condition. The
data are available in the file ‘medical_data.RData’, which contains the following
variables for 16 groups of patients:
*  Rate: the rate of admission to ICU in a particular group of patients.
*  History: the medical history of the patients in a particular group (two levels;
0: short; 1: long).
*  Operation: indicator showing if patients in a particular group have had a recent
surgical operation (two levels; 0: no; 1: yes).
*  Comorbidity: an index of existing comorbidities in a particular group (four levels;
1–4 for comorbidities A–D).
*  Age: the median age of patients in a particular group.

---

(i) Plot the admission rates against Age, while also distinguishing between
different History levels on your graph. 
[Hint: you can use the points(…) command in R. Also, make sure that all
16 rates are shown in your graph, by adjusting the axes limits if necessary,
e.g. use xlim, ylim in R.]
(ii) Plot the admission rates against Age, while also distinguishing between
different Operation levels on your graph. 
[Hint: See hint in part (i).]

---

(iii) Comment on the impact of the variables History, Operation and Age on the
admission rates, based on your graphs from parts (i) and (ii).
(iv) Fit a normal Generalised Linear Model (GLM) to the data to investigate the
dependence of admission rates on the three following variables: History,
Operation and Age (when these variables are only included as main effects in
the model).
You should use the logarithmic link function in the model. 
(v) Comment on the impact of the variables History, Operation and Age on the
admission rates, based on the output of the model fitted in part (iv).

--- 
 
(vi) Fit a normal GLM to the data, including the variable Comorbidity as a main
effect, in addition to the other three variables included in part (iv). You should
use the logarithmic link function in the model. 
(vii) Comment on the impact of all four variables on the admission rates, based on
the output of the model fitted in part (vi). [4]
(viii) Determine which of the two models fitted in parts (iv) and (vi), should be
preferred for investigating the dependence of admission rates on the available
variables. 

---

(ix) Calculate the predicted rate of admission to ICU for a group of patients with:
*  short medical history.
*  no recent surgical operation.
*  comorbidity B.
*  median age 62.
The prediction should be based on the preferred model from part (viii). 
(x) Comment on a disadvantage of the GLMs used in parts (iv) and (vi), in terms
of predicting admission rates.

---

Q5
(i)
R code:
Parts (i)-(iv) were answered reasonably well, with common errors including incorrect parameterisations for the binomial distributions involved.
Part (v) was not well answered with a number of candidates not attempting it.

```{r}
plot(Age[History==0],Rate[History==0],pch=19,xlim=c(41,75),ylim=c(0.03,0.10),xlab="Age",ylab="Rate", main="Admission rates against age") 
points(Age[History==1],Rate[History==1],pch=19, col =
"red") 
```
Black bullets: History = 0; red bullets: History = 1. 
(ii)
R code:

```{r}
plot(Age[Operation==0],Rate[Operation==0],pch=19,xlim=c(41,
75),ylim=c(0.03, 0.10),xlab="Age",ylab="Rate") 
points(Age[Operation==1],Rate[Operation==1],pch=19, col = "red") 
```

Black bullets: Operation = 0; red bullets: operation = 1. 
(iii)
Age seems to have a clear increasing effect on rates. 
The effects of History and Operation are not clear. 
(iv)
R code:
m1 = glm(Rate ~ History + Operation + Age, family = gaussian(link="log")) 
summary(m1)
#Call:
#glm(formula = Rate ~ History + Operation + Age, family = gaussian(link = "log"))
#
#Deviance Residuals:
# Min 1Q Median 3Q Max
#-0.0096278 -0.0027215 -0.0004823 0.0030534 0.0165070
#
#Coefficients:
# Estimate Std. Error t value Pr(>|t|)
#(Intercept) -4.232037 0.184121 -22.985 2.73e-11 ***
#History1 0.008975 0.051115 0.176 0.864
#Operation1 0.010137 0.061309 0.165 0.871
#Age 0.025690 0.002616 9.822 4.35e-07 ***
#---
#Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#
#(Dispersion parameter for gaussian family taken to be 4.6878e-05)
#
# Null deviance: 0.00738638 on 15 degrees of freedom
#Residual deviance: 0.00056253 on 12 degrees of freedom
#AIC: -108.68 

---

(v)
History does not have a significant impact since p-value is large (0.864). 
Operation does not have a significant impact since p-value is large (0.871). 
Age has a highly significant 
positive effect since p-value is very small, 4.35e-07. 

---

(vi)
R code:
m2 = glm(Rate ~ History + Operation + Age + Comorbidity, family = gaussian(link="log")) 
summary(m2)


#Call:
#glm(formula = Rate ~ History + Operation + Age + Comorbidity,
# family = gaussian(link = "log"))
##
#Deviance Residuals:
# Min 1Q Median 3Q Max
#-0.005923 -0.002551 -0.001152 0.002753 0.006639
#
#Coefficients:
# Estimate Std. Error t value Pr(>|t|)
#(Intercept) -4.203456 0.151308 -27.781 4.93e-10 ***
#History1 0.005103 0.041287 0.124 0.9044
#Operation1 0.006746 0.050478 0.134 0.8966
#Age 0.027170 0.002302 11.803 8.86e-07 ***
#Comorbidity2 -0.182506 0.058996 -3.094 0.0129 *
#Comorbidity3 -0.149206 0.059872 -2.492 0.0343 *
#Comorbidity4 -0.125286 0.059462 -2.107 0.0644 .
#---
#Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#
#(Dispersion parameter for gaussian family taken to be 2.927809e-05)
#
# Null deviance: 0.0073864 on 15 degrees of freedom
#Residual deviance: 0.0002635 on 9 degrees of freedom
#AIC: -114.82 

---
(vii)
As before, History and Operation do not have a significant impact, while age still
has a highly significant positive effect. [1½]
Comorbidity has significant impact, 
with levels 2-3 (comorbidities B-C) having a significant reducing effect as
compared to level 1 (baseline, comorbidity A) 
and level 4 (comorbidity D) having a marginally (non-)significant negative impact. 

---

(viii)
The model with the 4 covariates in part (vi) should be preferred, 
as it has smaller AIC. 
(ix)
Predicted link function value is:
- 4.203456 - 0.182506 + 0.027170 * 62
= -2.701422 
and predicted value is:
exp(-2.701422) = 0.06711001 

---
(x)
The used GLM has a logarithmic link,
which implies that predicted rates can be above 1. 
