---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


An insurance company wants to study the association between the number of years their clients spent in education and their claim amounts.

Data from 25 randomly selected claims are contained in the file <tt>AmountYears.RData</tt> in the following two variables:

* <tt>ClaimAmount</tt> – this is the claim amount (in £).

* <tt>EducationYears</tt> – this is the number of years the client spent in education.

---

### Exercises

**A.** Plot the claim amounts against the years of education. 

**B.** A statistician suggests fitting a simple linear model to the data.
Fit a linear model to the data.  <br>Plot the regression line by adding it to the graph in ***Exercise A***. 

**C.** Another statistician looks at the plot in ***Exercise A*** and suggests a non-linear relationship.
Fit a model with a quadratic term added to the model fitted in ***Exercise B***.
<br><br>State the equation of the fitted model. 

**D.** Comment on the suitability of the quadratic model in ***Exercise C***, compared to
the model in ***Exercise B***, based on the output from ***Exercise C***. 

---

## Question 1 Solutions:


```{r}
load("StoredData/AmountYears.RData")
```


---

### Inspecting the Data

**Claim Amoount**

```{r}

length(ClaimAmount)

```


```{r}

summary(ClaimAmount)

```

---

### Inspecting the Data

**Claim Amoount**

```{r}
length(EducationYears)
```

```{r}
summary(EducationYears)
```


---

### Exercise A. 

**Plot the claim amounts against the years of education**

This code will produce a scatterplot showing the relationship between the number of years in education and the claim amounts.


```{r eval=FALSE}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in £)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")
```


Remark: always provide informative annotation on the graphs

---

```{r echo=FALSE,fig.width=10}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in £)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")
```

---

### Exercise B.

**Fit a linear model and plot the regression line**

Fit a linear model

```{r}
m1 <- lm(ClaimAmount ~ EducationYears)  
 
```

---

### Multiple $R^2$ and Adjusted $R^2$. 

When fitting linear models in R, you'll often come across two different types of $R^2$: Multiple $R^2$ and Adjusted $R^2$. 

Both provide insights into the goodness-of-fit of your model, but there is a key difference.


- **Multiple $R^2$:** Measures the proportion of variance explained by the model without considering the number of predictors.

- **Adjusted $R^2$:** Adjusts the $R^2$ value to account for the number of predictors, providing a more accurate measure of model fit, especially when comparing models with different numbers of predictors.

---

### Exercise B.

**Summary of linear model (m1):**


```{r}
coef(m1)
```

```{r eval=FALSE}
summary(m1)
```


---


```{r echo=FALSE}
summary(m1)
```

---

**Plot the regression line:**

This code adds the regression line to the scatterplot.

```{r eval=FALSE}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in £)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")

abline(m1, col = "red")
```


---

**Plot the regression line:**

```{r echo=FALSE,fig.width=10}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in £)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")

abline(m1, col = "red")
```

---

### The `I()` function

* The `I()` function in R is used to inhibit the interpretation or conversion of its argument. 

* It's often applied when you want to indicate that certain values should be treated "as is" without being altered or evaluated by R. 

* In model formulas, `I()` is used to indicate that an arithmetic operation should be performed directly rather than treated as part of the model structure. 

---

### Exercise C.

**Fit a quadratic model**

Fit a model with a quadratic term.

```{r}
m2 <- lm(ClaimAmount ~ EducationYears + I(EducationYears^2))  

```

---

### Exercise C.

**Summary of quadratic model (m2):**

```{r}
coef(m2)
```

```{r eval=FALSE}
summary(m2)
```

---

```{r echo=FALSE}
summary(m2)
```

---

### Exercise C.

**coefficients of quadratic model (m2):**

```{r}
coef(m2)
```

---

### Exercise C.

**State the equation of the fitted model:**

The fitted model is:

$$\text{ClaimAmount} = 609272 - 95999 \times \text{Educ_Yrs} + 4371 \times \text{Educ_Yrs}^2$$




---

### Exercise D.

**Comment on the suitability of the quadratic model**

* The quadratic model in ***Exercise C*** seems more suitable than the linear model in ***Exercise B***.

* The ***adjusted R-squared*** has increased, indicating a better fit to the data. 

* Additionally, the coefficient of the quadratic term is significantly different from zero, suggesting that the inclusion of the quadratic term captures the non-linear relationship between education years and claim amounts.

---

### Polynomial Curve

```{r eval=FALSE,fig.width=10}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in GBP)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")

### Sorting to remove jaggedness

# ix means index
ix <- sort(EducationYears,index.return=T)$ix

lines(EducationYears[ix],predict(m2)[ix],col="red",type="l")
```

```{r echo=FALSE}
ix <- sort(EducationYears,index.return=T)$ix
```

Sorted Indices

```{r}
ix
```

---

```{r echo=FALSE,fig.width=10}
plot(EducationYears, ClaimAmount, 
     xlab = "Number of years in education", 
     ylab = "Claim amount (in GBP)", 
     main = "Scatterplot of the amount of claims against years in education",
     pch=16,col="blue")

### Sorting to remove jaggedness
ix <- sort(EducationYears,index.return=T)$ix


lines(EducationYears[ix],predict(m2)[ix],col="red",type="l")
```


---



---
