---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

2 A financial consultancy working with large firms wishes to model the relationship
between a firm’s assets and the number of senior management positions in the firm.
The data file firms.Rdata contains the variables:
assets – this is the value of assets (in millions of £).
sn_positions – this is the number of senior managements positions.

You can load the file into R using load("firms.Rdata").

---
(i) Plot the number of senior management positions as a function of assets. 
(ii) Plot the number of senior management positions as a function of log10_assets
where log10_assets is the assets at log10 scale. [4]
(iii) Comment on your plots in parts (i) and (ii). 
An analyst wishes to check if the number of senior management positions follows a
Poisson distribution.

---

(iv) Calculate the mean of the number of senior management positions. 
Use the command set.seed(222) to initialise the random number generator.
(v) Generate a sample of size equal to the number of firms from a Poisson
distribution with parameter equal to the mean calculated in part (iv). 
(vi) Plot a histogram of the sample simulated in part (v) and a histogram of
sn_positions on two separate graphs but on the same scale specifying
appropriate axis limits and labels. 
(vii) Comment on your plots in part (vi). 

Here are the solutions to the given problem, step by step:

---

### (i) Plot the number of senior management positions as a function of assets:

```r
load("firms.RData")
plot(assets, sn_positions, xlab = "Assets", ylab = "Number of senior positions", main = "Firm's assets vs number of senior positions")
```
This code produces a scatterplot showing the relationship between a firm's assets and the number of senior management positions.

---

### (ii) Plot the number of senior management positions as a function of log10_assets:

```r
log10_assets = log10(assets) 
plot(log10_assets, sn_positions, xlab = "Assets transformed at log10 scale", ylab = "Number of senior positions", main = "Firm's assets (log10) vs number of senior positions")
```
This code creates a scatterplot of the number of senior management positions against the log10-transformed assets.

---

### (iii) Comment on your plots in parts (i) and (ii):

In part (i), the variable assets is extremely widespread, making it difficult to discern any clear pattern or relationship. In part (ii), taking the log10 of the assets helps to reveal an increasing relationship between the variables, indicating that the number of senior management positions tends to rise as the log-transformed assets increase.

---

### (iv) Calculate the mean of the number of senior management positions:

```r
mu = mean(sn_positions) 
mu
```
The mean number of senior management positions is calculated as approximately 13.58065.

---

### (v) Generate a sample of size equal to the number of firms from a Poisson distribution:

```r
set.seed(222)
theoretical_samp = rpois(length(sn_positions), mu) 
```
This code generates a sample from a Poisson distribution with the parameter equal to the mean calculated in part (iv), using a seed for reproducibility.

---

### (vi) Plot histograms of the sample and sn_positions:

**Histogram of the sample:**

```r
hist(theoretical_samp, xlim = c(0, 100), ylim = c(0, 100), main = "Distribution of simulated senior positions", xlab = "Number of senior positions")
```

**Histogram of sn_positions:**

```r
hist(sn_positions, breaks = 20, xlim = c(0, 100), ylim = c(0, 100), main = "Distribution of observed senior positions", xlab = "Number of senior positions")
```
These codes create histograms of the simulated sample and the observed sn_positions on the same scale.

---

### (vii) Comment on your plots in part (vi):

The two distributions are very different. The observed counts exhibit an over-representation of zero values compared to the simulated Poisson distribution. This suggests that the number of senior management positions (sn_positions) does not follow a Poisson distribution with parameter 13.58.

---

These steps and comments cover the solution to the given problem. If you have further questions or need additional clarification, feel free to ask!


---
Q2
(i)
plot(assets,sn_positions, xlab="Assets", ylab="Number of senior positions", main="Firm's assets vs number of senior positions") 

(ii)
```{r}
log10_assets = log10(assets) 
plot(log10_assets,sn_positions, xlab="Assets transforms at log10 scale", 
ylab="Number of senior positions", main="Firm's assets (log10) vs number of senior positions") 
```

<p>
---


(iii)
The variable assets is extremely widespread in part (i). 
Taking the log10 of the assets in (ii) helps to show an increasing relationship between the variables. 
(iv)
mu = mean(sn_positions) 
mu
13.58065 
(v)
```{r}
set.seed(222)
theoretical_samp = rpois(length(sn_positions), mu) 
```

(vi)
```{r}
hist(theoretical_samp,xlim=c(0,100), ylim=c(0,100), 
main = “Distribution of simulated senior positions", xlab="Number of senior positions") 
```
<p>
---

```{r}
hist(sn_positions, breaks=20,xlim=c(0,100), ylim=c(0,100), 
	main="Distribution of observed senior positions", xlab="Number of senior positions") 
```

(vii)
We can see that the two distributions are very different. 
<p>
---

For example, we can see an over-representation of zero values in the observed counts. 
Therefore, the number of senior positions sn_positions variable does not follow a Poisson distribution with parameter 13.58. 
[Total 21]
Commentary:
(i), (ii) Well answered overall. There were some answers where the graph annotation was not informative, and some cases with reversed axes.
(iii) Well answered, with most candidates receiving full or partial credit.
(iv), (v) Very well answered.
(vi) Mixed answers. While most candidates produced the two plots, in many cases the axes were not matched. Suboptimal titles and/or axis labels were used in some cases.
(vii) Most candidates correctly observed that the distributions are different, but failed to comment specifically on the data not following the Poisson distribution that we have here.
