---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

2 A financial consultancy working with large firms wishes to model the relationship
between a firm’s assets and the number of senior management positions in the firm.
The data file firms.Rdata contains the variables:
assets – this is the value of assets (in millions of £).
sn_positions – this is the number of senior managements positions.

You can load the file into R using load("firms.Rdata").

---
(i) Plot the number of senior management positions as a function of assets. 
(ii) Plot the number of senior management positions as a function of log10_assets
where log10_assets is the assets at log10 scale. [4]
(iii) Comment on your plots in parts (i) and (ii). 
An analyst wishes to check if the number of senior management positions follows a
Poisson distribution.

---

(iv) Calculate the mean of the number of senior management positions. 
Use the command set.seed(222) to initialise the random number generator.
(v) Generate a sample of size equal to the number of firms from a Poisson
distribution with parameter equal to the mean calculated in part (iv). 
(vi) Plot a histogram of the sample simulated in part (v) and a histogram of
sn_positions on two separate graphs but on the same scale specifying
appropriate axis limits and labels. 
(vii) Comment on your plots in part (vi). 

---
Q2
(i)
plot(assets,sn_positions, xlab="Assets", ylab="Number of senior positions", main="Firm's assets vs number of senior positions") 

(ii)
```{r}
log10_assets = log10(assets) 
plot(log10_assets,sn_positions, xlab="Assets transforms at log10 scale", 
ylab="Number of senior positions", main="Firm's assets (log10) vs number of senior positions") 
```

<p>
---


(iii)
The variable assets is extremely widespread in part (i). 
Taking the log10 of the assets in (ii) helps to show an increasing relationship between the variables. 
(iv)
mu = mean(sn_positions) 
mu
13.58065 
(v)
```{r}
set.seed(222)
theoretical_samp = rpois(length(sn_positions), mu) 
```

(vi)
```{r}
hist(theoretical_samp,xlim=c(0,100), ylim=c(0,100), 
main = “Distribution of simulated senior positions", xlab="Number of senior positions") 
```
<p>
---

```{r}
hist(sn_positions, breaks=20,xlim=c(0,100), ylim=c(0,100), 
	main="Distribution of observed senior positions", xlab="Number of senior positions") 
```

(vii)
We can see that the two distributions are very different. 
<p>
---

For example, we can see an over-representation of zero values in the observed counts. 
Therefore, the number of senior positions sn_positions variable does not follow a Poisson distribution with parameter 13.58. 
[Total 21]
Commentary:
(i), (ii) Well answered overall. There were some answers where the graph annotation was not informative, and some cases with reversed axes.
(iii) Well answered, with most candidates receiving full or partial credit.
(iv), (v) Very well answered.
(vi) Mixed answers. While most candidates produced the two plots, in many cases the axes were not matched. Suboptimal titles and/or axis labels were used in some cases.
(vii) Most candidates correctly observed that the distributions are different, but failed to comment specifically on the data not following the Poisson distribution that we have here.
