---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

1 An insurance company wants to study the association between the number of years
their clients spent in education and their claim amounts.
Data from 25 randomly selected claims are contained in the file
AmountYears.RData in the following two variables:
ClaimAmount – this is the claim amount (in £).
EducationYears – this is the number of years the client spent in education.

---

(i) Plot the claim amounts against the years of education. 
A statistician suggests fitting a simple linear model to the data.
(ii) (a) Fit a linear model to the data. 
(b) Plot the regression line by adding it to the graph in part (i). 
Another statistician looks at the plot in part (i) and suggests a non-linear relationship.
(iii) (a) Fit a model with a quadratic term added to the model fitted in part (ii).

(b) State the equation of the fitted model. 
(iv) Comment on the suitability of the quadratic model in part (iii), compared to
the model in part (ii), based on the output from part (iii). 


---

Q1 load("AmountYears.RData")
(i)
plot(EducationYears, ClaimAmount, xlab = "Number of years in education", ylab = "Claim amount (in £)", main = "Scatterplot of the amount of claims against years in education")


---

(ii)
(a)
```{r}
m1 <-lm(ClaimAmount~EducationYears)  
summary(m1) 
```

Call: lm(formula = ClaimAmount ~ EducationYears) Residuals: Min 1Q Median 3Q Max -43747 -19875 1768 18575 61813 Coefficients: Estimate Std. Error t value Pr(>|t|) (Intercept) -32220 40159 -0.802 0.43058
<p>
---

EducationYears 11101 3124 3.553 0.00169 ** --- Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 Residual standard error: 27450 on 23 degrees of freedom Multiple R-squared: 0.3544, Adjusted R-squared: 0.3263 F-statistic: 12.62 on 1 and 23 DF, p-value: 0.001695 
(b)
abline(m1, col = "red") 

(iii)(a) m2 <- lm(ClaimAmount~EducationYears + I(EducationYears^2))  
summary(m2) Call: lm(formula = ClaimAmount ~ EducationYears + I(EducationYears^2)) Residuals: Min 1Q Median 3Q Max
<p>
---

-51368 -15047 -2360 14008 45625 Coefficients: Estimate Std. Error t value Pr(>|t|) (Intercept) 609272 242570 2.512 0.0199 * EducationYears -95999 40154 -2.391 0.0258 * I(EducationYears^2) 4371 1635 2.674 0.0139 * --- Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 Residual standard error: 24380 on 22 degrees of freedom Multiple R-squared: 0.5127, Adjusted R-squared: 0.4684 F-statistic: 11.57 on 2 and 22 DF, p-value: 0.000368

---

(b)
The fitted model is:
ClaimAmount = 609272−95999×EducationYears + 4371×EducationYears2 

---

(iv)
The quadratic model in part (iii) seems more suitable.

The adjusted R-squared has increased.

The coefficient of the quadratic term is significantly different from 0.

---


Commentary:
(i), (ii): Well answered in general. A number of candidates failed to provide informative annotation on the graphs. A small number of candidates presented the graphs with the axes reversed.
(iii): Mixed answers with some candidates not showing appropriate output.
(iv): Not very well answered. A common error was referring to the unadjusted R-squared measure.
<p>
---


---