---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```5 An insurance company, which currently only sells home insurance, is interested in
entering the car insurance market. An underwriting manager at the company believes
that the age and gender of the policyholder will be the most important factors in
estimating the number of claims made under a car insurance policy.
The underwriting manager has commissioned a survey of its current home insurance
customers who also have car insurance, choosing a male customer and a female
customer for every age from 18 to 65, asking them how many car insurance
claims they have made in the past 3 years. This dataset is saved in the file
ClaimsData.Rdata. After loading this data into R, using the command
load("ClaimsData.Rdata"), the data frame ClaimsData will be available,
which contains the following three variables:
age – this is the age (in years) of the policyholder.
gender – this is either ‘M’ for male or ‘F’ for female.
claim_count – this is the number of car insurance claims reported by the
policyholder over the past 3 years.
(i) Fit a normal linear regression model to the data using claim_count as the
response variable and age as the explanatory variable. Your answer should
include the estimated intercept and slope of the regression line. 
A colleague suggests that the response variable would be better modelled as having a
Poisson distribution.
(ii) Fit a Generalised Linear Model (GLM) to the data using claim_count as
the response variable and age as the explanatory variable, assuming a Poisson
distribution for the response variable. Your answer should include the
estimated coefficients and the Akaike’s Information Criterion (AIC) of the
fitted model. [4]
The underwriting manager wishes to compare the fit of the GLM in part (ii) against
that of the normal linear regression model in part (i).
(iii) Explain why scaled deviances cannot be used to compare the fit of the models
in parts (i) and (ii). 
(iv) Fit, by choosing a suitable argument for family in the glm command, a
GLM to the data that is equivalent to the model fitted in part (i). Your answer
should include the estimated coefficients and the AIC of this fitted model. [4]
(v) Compare the fit of the models fitted in parts (ii) and (iv). 
---–7
The underwriting manager believes the Poisson GLM would be improved by adding
the explanatory variable gender as well as its interaction with age.
(vi) (a) Fit a Poisson GLM to the data of the form age*gender. Your
answer should include the estimated coefficients and the AIC of this
fitted model.
(b) Compare, using scaled deviances, the fit of this model to that in
part (ii).
[7]
[Total 23]
END OF PAPER

Q5
Load the data:
> load("ClaimsData.Rdata")
(i)
> model_lin <- lm(claim_count ~ age) 
> model_lin$coefficients
(Intercept) age
4.29414532 -0.05026053
The intercept is 4.29 and the slope is -0.05. 
(ii)
> model_glm_age <- glm(claim_count ~ age, family = poisson) 
> model_glm_age
# Call: glm(formula = claim_count ~ age, family = poisson)
Coefficients:
(Intercept) age
1.70492 -0.02323
Degrees of Freedom: 95 Total (i.e. Null); 94 Residual
Null Deviance: 149.5
Residual Deviance: 128.2 AIC: 349.8 
<p>
---

(iii)
A comparison using scaled deviances is only valid for nested models. 
The model in (i) assumed a Normal distribution, whilst the model in (ii) assumed a Poisson distribution. Therefore, neither model is a subset of the other. 
(iv)
> model_lin_glm <- glm(claim_count ~ age, family = gaussian) 
> model_lin_glm
# Call: glm(formula = claim_count ~ age, family = gaussian)
Coefficients:
(Intercept) age
4.29415 -0.05026
Degrees of Freedom: 95 Total (i.e. Null); 94 Residual
Null Deviance: 295.8
Residual Deviance: 249.3 AIC: 370 
(v)
The AIC of the Poisson model in part (ii) is lower than that of the Normal model in part (iv).  Therefore, the Poisson GLM provides a better fit. 
(vi)
(a)
> model_glm_full <- glm(claim_count ~ age*gender, family = poisson) 
> model_glm_full
<p>
---

# Call: glm(formula = claim_count ~ age * gender, family = poisson)
Coefficients:
(Intercept) age genderM age:genderM
1.84642 -0.03459 -0.18217 0.01805
Degrees of Freedom: 95 Total (i.e. Null); 92 Residual
Null Deviance: 149.5
Residual Deviance: 113.5 AIC: 339 
(b) H0: the bigger (full) model is NOT a significant improvement over the simpler model; against H1: it is a significant improvement. 

> 

```{r}
anova(model_glm_age, model_glm_full, test="Chisq") 
```

Analysis of Deviance Table
Model 1: claim_count ~ age Model 2: claim_count ~ age * gender
Resid. Df Resid. Dev Df Deviance Pr(>Chi) 1 94 128.25 2 92 113.46 2 14.789 0.0006148 *** --- Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
The p-value is clearly significant (less than 0.1%). 
Therefore, the GLM with age*gender is a significant improvement over that with gender alone. 
<p>
---

[Total 23]
Commentary:
(i) Well answered. A fair number of candidates failed to pull out the intercept and slope estimates explicitly.
(ii) Well answered. As in part (i), a fair number of candidates failed to pull out the coefficient estimates and AIC explicitly.
(iii) Not well answered. Only a relatively small number of candidates referenced nesting.
(iv) Well answered in general. A number of candidates failed to present the coefficient estimates and the value of the AIC explicitly.
(v) Very well answered.
(vi)(a) Well answered in general. A number of candidates failed to pull out the coefficient estimates and AIC value explicitly.
(vi)(b) Not well answered. Only a small number of candidates performed a proper comparison (e.g. anova) with reference to statistical significance in the difference of the deviances.
[Paper Total 100]
END OF EXAMINERS’ REPORT
www.actuaries.org.uk
© 2021 Institute and Faculty of Actuaries