---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```



```{r}
library(markovchain)
```

---
# (i)

```{r}
TransitionName <- c("Pass","Shoot","Dribble")
MarkovData <- c(0.5,0.1,0.4,
              0.7,0.2,0.1,
              0.4,0.3,0.3) 
```

---


#### Part (B)

```{r}
# MarkovData<-c(0.5,0.1,0.4,0.7,0.2,0.1,0.4,0.3,0.3)
# TransitionName<-c("Pass","Shoot","Dribble") 
BallTransition<-matrix(MarkovData,nrow = 3,
                       byrow = TRUE,
                       dimnames = list(TransitionName,TransitionName))
```

---

```{r}
BallTransition


```


---

```{r}
markov2<-new("markovchain",
             transitionMatrix=matrix(data=MarkovData,
 byrow=TRUE,nrow=3,
 dimnames=list(TransitionName,TransitionName)))
```

---


Part (c)

```{r}

plot(markov2) 
```
---



### Part D

```{r}
summary(markov2)

```


---

  
```{r}

steadyStates(markov2)

```

---

#### Part E

```{r}


markov2^2


```

After 2 plays the probability is 0.48

---

After 5 plays the probability is 0.50475

```{r}

markov2^5
```


---



```{r}

set.seed(100)
seq<-markovchainSequence(100,markov2)
#Frequency of the terms
table(seq)

```
---

#### Part G.

```{r}
#(vii)
library(lattice)

barchart(prop.table(table(seq)))

```

#### Part G.

---

```{r}
plot(prop.table(table(seq)),main = "Frequency of different states")
```

---


#### Part (H) 

According to the steady state distribution, the frequency of Pass, Shoot and Dribble needs to be 50.47, 18.09, 31.42 respectively. 

Pass frequency of the simulation is matching with that of the expected. But there is some discrepancy in the shoot and dribble.

---

#### Main Reasons 
- (1) The number of simulations (100) is very small and needs 
to be increased. When the simulations are small, different seed values can produce different results altogether because of sample error and those differences might be significant
--- 