---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

### Question 4

The ‘sunspot’ dataset contains the monthly mean relative sunspot numbers from 1749 to 1983. 
Collected at Swiss Federal Observatory, Zurich until 1960, then Tokyo Astronomical 
Observatory. Please import this dataset from the ‘datasets’ package in R.
Libraries required:
‘tseries’
‘forecast’

---

#### Exercises 
1. Plot the timeseries with frequency = 12, giving appropriate labels for each axis and paste 
the chart into your answer sheet. 

2.  Plot the sample Auto-correlation Function (ACF) and the sample Partial Auto-Correlation 
Function (PACF) of the data, giving appropriate labels for the x and y axes. 

3.   Estimate the ACF and PACF values at each lag. 

4.   Comment on the ACF based on the charts and values computed in (2.  and (iii). 

5.  Calculate the simple seasonal difference with period 12. Print the top 12 values of the 
seasonal difference. 

6.  Decompose the trend, seasonality and randomness of the data. Plot the decomposed values.

7. Comment on the general features of your chart. 

---

Solution 4:
library(tseries)
## Warning: package 'tseries' was built under R version 4.0.5
## Registered S3 method overwritten by 'quantmod':
## method from
## as.zoo.data.frame zoo
library(forecast)
## Warning: package 'forecast' was built under R version 4.0.5

---

# (i)
```{r}
data(sunspots)
a<-ts(sunspots,frequency = 12)
plot(a, main = "Time series of plots")
```
---
```{r}
#(ii)
par(mfrow=c(1,2))
acf(a,main="Sample ACF")
pacf(a,main="Sample PACF")
```

---

# (iii)
```{r}
acf(a,plot = FALSE)

```


## 
## Autocorrelations of series 'a', by lag
## 
## 0.0000 0.0833 0.1667 0.2500 0.3333 0.4167 0.5000 0.5833 0.6667 0.7500 0.83
33 
## 1.000 0.922 0.890 0.875 0.864 0.850 0.836 0.819 0.805 0.797 0.7
78 
## 0.9167 1.0000 1.0833 1.1667 1.2500 1.3333 1.4167 1.5000 1.5833 1.6667 1.75
00 
## 0.756 0.734 0.710 0.686 0.663 0.632 0.602 0.569 0.542 0.512 0.4
80 
## 1.8333 1.9167 2.0000 2.0833 2.1667 2.2500 2.3333 2.4167 2.5000 2.5833 2.66
67 
## 0.451 0.426 0.391 0.367 0.334 0.298 0.264 0.233 0.204 0.173 0.1
47 
## 2.7500 2.8333 
## 0.115 0.083

```
---
```{r}

pacf(a,plot = FALSE)
## 
## Partial autocorrelations of series 'a', by lag
## 
## 0.0833 0.1667 0.2500 0.3333 0.4167 0.5000 0.5833 0.6667 0.7500 0.8333 0.91
67 
## 0.922 0.272 0.189 0.135 0.064 0.044 -0.005 0.014 0.046 -0.045 -0.0
58 
## 1.0000 1.0833 1.1667 1.2500 1.3333 1.4167 1.5000 1.5833 1.6667 1.7500 1.83
33 
## -0.059 -0.077 -0.056 -0.046 -0.099 -0.075 -0.102 -0.037 -0.055 -0.054 -0.0
13 
## 1.9167 2.0000 2.0833 2.1667 2.2500 2.3333 2.4167 2.5000 2.5833 2.6667 2.75
00 
## 0.014 -0.068 0.045 -0.034 -0.050 -0.029 -0.019 0.012 -0.006 0.026 -0.0
17 
## 2.8333 
## -0.039 

```
---


# (iv)
'The ACF indicates that there is correlation between an observation and
the past observations. There is a correlation beyond 2.5 years but it keeps g
etting weaker'
## [1] "The ACF indicates that there is correlation between an observation and the past observations. There is a correlation beyond 2.5 years but it keep
s getting weaker" [2M]
```
---
```{r}
# (v)
s<-diff(a, lag=12,differences=1)
head(s,12) 
```
## Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec
## 2 15.3 13.3 19.2 32.6 5.0 16.5 -9.4 36.7 15.3 -9.8 -95.3 -9.8

---

```{r}
# (vi)
f<-decompose(a)
plot(f)

```

---

# (vii)
'The cyclical trend with more variations at the longer time lags, 
there is seasonality in the dataset'
## [1] "The cyclical trend with more variations at the longer time lags, \nth
ere is seasonality in the dataset" [2M]
```
---
