---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


A random variable follows a Weibull distribution with parameters for the shape and scale as 
follows:
```{r}
Shape = 1.5
Scale = 100000
```
Run the following code in R before beginning your answer:
```{r}
set.seed(1234)
options(scipen=5)
```

---

Use the in-build functions in R to:

### Exercises

1. Simulate 1,000 values from a Weibull distribution with parameters given above, assign 
this to a vector variable called ‘***WeibullSim***’. 
<p>
2.  Calculate the mean and standard deviation of the simulated values. 
<p>
3.   Using the mean and standard deviation estimated from exercise 2, estimate the $\mu$ and $\sigma$ parameters 
of a LogNormal distribution following the method of moments approach. 
<p>
4.   Simulate 1,000 values from the LogNormal distribution using the parameter estimates above and assign this to a vector variable called ‘***LognormalSim***’ and calculate the 
mean and standard deviation of the simulated values. 

---

### Exercises

<ol start="5">
  <li> Plot a histogram of ‘<tt>WeibullSim</tt>’ and ‘<tt>LognormalSim</tt>’ showing the frequencies and paste 
the plot into your answer.</li>
<p>
  <li>Plot the probability density function for ‘<tt>WeibullSim</tt>’ and ‘<tt>LognormalSim</tt>’ as a line plot.</li>
</ol>


 


  * Set the <tt>WeibullSim</tt> line to color ‘red’ and the <tt>LognormalSim</tt> line to color ‘green’.
  * Set the y-axis limits from 0 to 0.00001.

Hint: Use the density function in R to estimate densities of your simulated values. 


---

#### Exercise 1

```{r}
shape <- 1.5
scale <- 100000
WeibullSim <- rweibull(1000,shape,scale) 

```


```{r}
head(WeibullSim)
```

---

```{r}
hist(WeibullSim,breaks=30)
```

---



#### Exercise 2

Calculate the mean and standard deviation of the simulated values. 

```{r}
SimMean <- mean(WeibullSim)
SimMean
```


```{r}
SimSD<-sd(WeibullSim)
SimSD 
```

---


#### Exercise 3

Using the mean and standard deviation estimated from exercise 2, estimate the $\mu$ and $\sigma$ parameters 
of a LogNormal distribution following the method of moments approach.


---



#### Exercise 3

#### Method of Moments	- Mean

$${\displaystyle \mu =\log \left({\frac {\operatorname {E} [X]^{2}}{\sqrt {\operatorname {Var} [X]+\operatorname {E} [X]^{2}}}}\right)}$$




```{r}

mu<-log( (SimMean^2) / (SimSD^2 + SimMean^2)^0.5 )
mu


```

---

#### Method of Moments	- Standard Deviation
$${\displaystyle \sigma ^{2}=\log \left({\frac {\operatorname {Var} [X]}{\operatorname {E} [X]^{2}}}+1\right)}$$

Remark - we are asked to find $\sigma$, so we take the additional step of getting the square root.
```{r}

sigma<-log(SimSD^2/SimMean^2 + 1)^0.5
sigma

```

---


#### Exercise 4

Simulate 1,000 values from the LogNormal distribution using the parameter estimates above and assign this to a vector variable called ‘***LognormalSim***’ and calculate the 
mean and standard deviation of the simulated values. 


```{r}
LogNormalSim <- rlnorm(1000,mu,sigma)
mean(LogNormalSim)
```

---



#### Exercise 4

```{r}
mean(LogNormalSim)
```

```{r}
sd(LogNormalSim)
```

---


#### Exercise 5

Plot a histogram of ‘***WeibullSim***’ and ‘***LognormalSim***’ showing the frequencies

```{r}
SimResults<-as.data.frame(cbind(WeibullSim,LogNormalSim))
```

---

```{r}

hist(WeibullSim, breaks = 30)

```

---

```{r}

hist(LogNormalSim, breaks = 30) 
```



---


#### Exercise 6 
Plot the probability density function for ‘***WeibullSim***’ and ‘***LognormalSim***’ as a line plot.

<pre><code>
plot(density(WeibullSim),type = "l",col="red",ylim = c(0,0.00001))
lines(density(LogNormalSim),col="green") 
</code></pre>

---

```{r}
plot(density(WeibullSim),type = "l",col="red",ylim = c(0,0.00001))
lines(density(LogNormalSim),col="green") 

```

---


---