---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


This exercise involves modeling the ball-by-ball actions of a football player using a Markov Chain. Here's how you can approach each part of the exercise in R using the `markovchain` package:

### **i) Create a vector with the state space of this Markov Chain in R**

```r
states <- c("Pass", "Shoot", "Dribble")
```

### **ii) Create a matrix of transition probabilities in R with names of that states in the matrix. Then create a Markov Chain Object using the ‘new’ function and details from the matrix created.**

```r
# Load the markovchain package
library(markovchain)

# Define the transition matrix
transitionMatrix <- matrix(c(0.5, 0.1, 0.4,
                             0.7, 0.2, 0.1,
                             0.4, 0.3, 0.3),
                           nrow = 3, byrow = TRUE)
rownames(transitionMatrix) <- states
colnames(transitionMatrix) <- states

# Create the Markov Chain object
footballMC <- new("markovchain", states = states, transitionMatrix = transitionMatrix)
```

### **iii) Plot the transition probability matrix**

```r
# Plot the Markov Chain
plot(footballMC)
```

### **iv) Calculate the steady state of this distribution**

```r
# Calculate steady state
steadyState <- steadyStates(footballMC)
steadyState
```

### **v) Calculate the probability that a player who Passes the ball initially, will Pass the ball after:**

#### a) 2 plays have taken place

```r
# Calculate probability after 2 plays
twoPlays <- footballMC^2
initialState <- c(1, 0, 0) # Starts with Pass
probTwoPlays <- initialState %*% twoPlays
probTwoPlays["Pass"]
```

#### b) 5 plays have taken place

```r
# Calculate probability after 5 plays
fivePlays <- footballMC^5
probFivePlays <- initialState %*% fivePlays
probFivePlays["Pass"]
```

### **vi) Use the function set.seed(100). Simulate a sequence of 100 states using markovchainSequence function. Compute the frequency of the states.**

```r
set.seed(100)
sequence <- markovchainSequence(n = 100, markovchain = footballMC, t0 = "Pass")
freqTable <- table(sequence)
freqTable
```

### **vii) Use a bar chart to plot the relative frequency of the states and label the chart appropriately.**

```r
barplot(freqTable / sum(freqTable), main = "Relative Frequency of States", xlab = "State", ylab = "Relative Frequency")
```

### **viii) Comment by comparing the results obtained in (vii) with the steady state distribution obtained in (iv).**

- **Comparison**: By comparing the relative frequencies obtained from the simulation with the steady-state distribution calculated in part (iv), you can observe how closely the simulated states align with the theoretical steady-state distribution. If the number of simulations is sufficiently large, the relative frequencies should converge to the steady-state probabilities.

If you need further clarification or have any questions, feel free to ask!



```{r}
library(markovchain)
```

---
# (i)

```{r}
TransitionName <- c("Pass","Shoot","Dribble")
MarkovData <- c(0.5,0.1,0.4,
              0.7,0.2,0.1,
              0.4,0.3,0.3) 
```

---


#### Part (B)

```{r}
# MarkovData<-c(0.5,0.1,0.4,0.7,0.2,0.1,0.4,0.3,0.3)
# TransitionName<-c("Pass","Shoot","Dribble") 
BallTransition<-matrix(MarkovData,nrow = 3,
                       byrow = TRUE,
                       dimnames = list(TransitionName,TransitionName))
```

---

```{r}
BallTransition


```


---

```{r}
markov2<-new("markovchain",
             transitionMatrix=matrix(data=MarkovData,
 byrow=TRUE,nrow=3,
 dimnames=list(TransitionName,TransitionName)))
```

---


Part (c)

```{r}

plot(markov2) 
```
---



### Part D

```{r}
summary(markov2)

```


---

  
```{r}

steadyStates(markov2)

```

---

#### Part E

```{r}


markov2^2


```

After 2 plays the probability is 0.48

---

After 5 plays the probability is 0.50475

```{r}

markov2^5
```


---



```{r}

set.seed(100)
seq<-markovchainSequence(100,markov2)
#Frequency of the terms
table(seq)

```
---

#### Part G.

```{r}
#(vii)
library(lattice)

barchart(prop.table(table(seq)))

```

#### Part G.

---

```{r}
plot(prop.table(table(seq)),main = "Frequency of different states")
```

---


#### Part (H) 

According to the steady state distribution, the frequency of Pass, Shoot and Dribble needs to be 50.47, 18.09, 31.42 respectively. 

Pass frequency of the simulation is matching with that of the expected. But there is some discrepancy in the shoot and dribble.

---

#### Main Reasons 
- (1) The number of simulations (100) is very small and needs 
to be increased. When the simulations are small, different seed values can produce different results altogether because of sample error and those differences might be significant
--- 