---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


The Logistic function can be defined by the function:

$$ f(𝑥) = \frac{1}{(1 + exp(-x))}$$

1. Create a function in R to calculate f(x). 

2. Create a plot of this function and evaluate it at values for x ranging from -10 to 10. Add appropriate x and y-axis labels. 


---
#### Exercise 1
```{r}
logistic<-function(x){ 1/(1+exp(-x))}
```

```{r}
logistic(4)
```

---

```{r}
X <- seq(-10,10,by=0.25)
```

```{r}
logistic(X)
```


---

#### Exercise 2
```{r eval=FALSE}
plot(logistic,
  xlim = c(-10,10), 
  main = "Logistic Function", 
  ylab = "y=1/(1+exp(-x))")
```
---

```{r echo=FALSE}
plot(logistic,
  xlim = c(-10,10), 
  main = "Logistic Function", 
  ylab = "y=1/(1+exp(-x))")
```
---


---