---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Solution 3:
IAI CS2B-0921
Page 5 of 13


```{r}
library(markovchain)
```

---
# (i)

```{r}
TransitionName <- c("Pass","Shoot","Dribble")
MarkovData <- c(0.5,0.1,0.4,
              0.7,0.2,0.1,
              0.4,0.3,0.3) 
```

---


#### Part (B)

```{r}
# MarkovData<-c(0.5,0.1,0.4,0.7,0.2,0.1,0.4,0.3,0.3)
# TransitionName<-c("Pass","Shoot","Dribble") 
BallTransition<-matrix(MarkovData,nrow = 3,
                       byrow = TRUE,
                       dimnames = list(TransitionName,TransitionName))
```

---

```{r}
BallTransition


```


---

```{r}
markov2<-new("markovchain",
             transitionMatrix=matrix(data=MarkovData,
 byrow=TRUE,nrow=3,
 dimnames=list(TransitionName,TransitionName)))
```

---


Part (c)

```{r}

plot(markov2) 
```
---



#### PArt 4
```{r}

summary(markov2)

```


---

  
```{r}

steadyStates(markov2)
## Pass Shoot Dribble
## [1,] 0.5047619 0.1809524 0.3142857
#(v)
#(a)

```

---

#### Part 5

```{r}


markov2^2
## Unnamed Markov chain^2 
## A 3 - dimensional discrete Markov Chain defined by the following states: 
## Pass, Shoot, Dribble 
## The transition matrix (by rows) is defined as follows: 
## Pass Shoot Dribble
## Pass 0.48 0.19 0.33
## Shoot 0.53 0.14 0.33
## Dribble 0.53 0.19 0.28

```

After 2 plays the probability is 0.48

---

After 5 plays the probability is 0.50475

```{r}
#(b)
markov2^5
## Unnamed Markov chain^5 
## A 3 - dimensional discrete Markov Chain defined by the following states: 
## Pass, Shoot, Dribble 
## The transition matrix (by rows) is defined as follows: 
## Pass Shoot Dribble
## Pass 0.50475 0.18075 0.31450
## Shoot 0.50525 0.18100 0.31375
## Dribble 0.50450 0.18125 0.31425
```

---



```{r}


#(vi)
set.seed(100)
seq<-markovchainSequence(100,markov2)
#Frequency of the terms
table(seq)
## seq
## Dribble Pass Shoot 
## 37 51 12 [4M]
```
---

#### Part (g)

```{r}
#(vii)
library(lattice)
## Warning: package 'lattice' was built under R version 4.0.5
barchart(prop.table(table(seq)))
# OR
```

---

```{r}
plot(prop.table(table(seq)),main = "Frequency of different states")
```

---


#### Part (H) 

According to the steady state distribution, the frequency of Pass, Shoot and Dribble needs to be 50.47, 18.09, 31.42 respectively. 

Pass frequency of the simulation is matching with that of the expected. But there is some discrepancy in the shoot and dribble.

---

#### Main Reasons 
- (1) The number of simulations (100) is very small and needs 
to be increased. When the simulations are small, different seed values can produce different results altogether because of sample error and those differences might be significant
--- 