---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

### Bayesian Analysis of Cloud Kitchen Sales Using R

 
This exercise introduces actuarial students to Bayesian statistical methods using R. 

The task involves analyzing sales data from a cloud kitchen, where total monthly sales are assumed to follow a normal distribution with unknown mean and variance. 

Students will compute the moment-generating function (MGF) for the prior distribution, derive the posterior distribution of the mean sales, and visualize both prior and posterior distributions using provided R code. 

---

### Scenario

Total sales per month on a particular cloud kitchen follow a normal distribution with unknown mean θ and variance $20^2$. 

* Sales $x_1, x_2, \ldots,x_n$ are observed over $n$ months. 

* Prior beliefs that θ follows normal distribution with mean 60 and variance $5^2$.

* Sales for last 5 months are extracted for analysis. Total sales for last 5 months were 340.

---

### Exercise A

Mθ(t) denotes MGF of θ. 

- State and compute M'θ(0) and M''θ(0). 

- Use `m1t0` for M'θ(0) and `m2t0` for M''θ(0). 


---


### Exercise A

```{r}
# m1x0 = E[X] 
# m2x0 = E[x^2] 
# var = E[x^2] - E[x]^2
# Equivalently m2x0 = var + E[x]^2

prior_mean <- 60
prior_sd <- 5

m1x0 <- prior_mean
m2x0 <- prior_sd^2 + prior_mean^2
```

---


### Exercise A

```{r}
m1x0
```

```{r}
m2x0
```


---

### Exercise B

1. State the posterior distribution of θ and compute parameters of the distribution using the sales of last 5 months.

2. After 50 months, the same analysis is performed using the 50 months sales data. Total sales for the last 50 months were 3400.

Compute the parameters of the posterior distribution of θ using 50 months sales data.


### Remark: 

This is the application of Bayesian inference to update the probability distribution for the mean sales of a cloud kitchen based on new data.

The exercise demonstrates how to combine the prior distribution (representing initial beliefs) with the likelihood function (based on observed data) to form the posterior distribution.

---

### Exercise B

* θ follows N(prior mean, prior variance)

* Random Variable X follows N(θ, variance)

* Posterior distribution of θ follows Normal Distribution with following parameters
* Post Mean: $$\text{post mean} = \frac{n \cdot \text{sample mean} / \text{variance} + \text{prior mean} / \text{prior variance}}{n / \text{variance} + 1 / \text{prior variance}}$$

* Post Variance: $$\text{post_var} = \left( \frac{n}{\text{variance}} + \frac{1}{\text{prior_variance}} \right)^{-1}$$
---


### Exercise B 

**Last 5 Months**

```{r}
n <- 5
sample_mean <- 340 / 5
sdev <- 20
 
post_mean <- (n * sample_mean / sdev^2 + prior_mean / prior_sd^2) / (n / sdev^2 + 1 / prior_sd^2)

#post variance = 1/(n/variance + 1/prior variance) 
post_var <- 1 / (n / sdev^2 + 1 / prior_sd^2)
```

---

### Exercise B 


**Last 5 Months**

```{r}
post_mean
post_var
sqrt(post_var)
```


---

### Exercise B 


**Last 50 Months**

```{r}
sample2_mean <- 3400 / 50
n2 <- 50
post2_mean <- (n2 * sample2_mean / sdev^2 + prior_mean / prior_sd^2) / (n2 / sdev^2 + 1 / prior_sd^2)

#post variance = 1/(n/variance + 1/prior variance) 
post2_var <- 1 / (n2 / sdev^2 + 1 / prior_sd^2)
```

---


### Exercise B 

**Last 50 Months**
```{r}
post2_mean
post2_var
sqrt(post2_var)
```

---



### Exercise C

Use the code below to plot the prior distribution of θ:
```{r fig.width=10,eval = FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5 
y <- dnorm(x, mean=60, sd=5) 
plot(x, y, ylim=c(0, .02))
```

* Add a line to overlay the posterior distribution of θ.
* Add another line to overlay the posterior distribution of θ after 50 months.
* Place the final graph and comment on it.

---

## Exercise C

**Solution**


What we are given to start with
```{r fig.width=10,eval=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5

y <- dnorm(x, mean=60, sd=5)

plot(x, y, ylim=c(0, 0.2), 
     pch=18,col="red")
```


---

What we are given to start with

```{r fig.width=10,echo=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5
y <- dnorm(x, mean=60, sd=5)
plot(x, y, ylim=c(0, 0.2), 
     pch=18,col="red")
```


---


## Exercise C

*** Add a line to overlay the posterior distribution of θ**

```{r fig.width=10,eval=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5
y <- dnorm(x, mean=60, sd=5)
plot(x, y, ylim=c(0, 0.2), 
     pch=18,col="red")

px1 <- post_mean + seq(-3, 3, by=0.2) * sqrt(post_var)

py1 <- dnorm(px1, mean=post_mean, sd=sqrt(post_var))

lines(px1, py1, col="green")
```

---

## Exercise C

**Add a line to overlay the posterior distribution of θ**



```{r fig.width=10,echo=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5
y <- dnorm(x, mean=60, sd=5)
plot(x, y, ylim=c(0, 0.2), 
     pch=18,col="red")

px1 <- post_mean + seq(-3, 3, by=0.2) * sqrt(post_var)
py1 <- dnorm(px1, mean=post_mean, sd=sqrt(post_var))
lines(px1, py1, col="green")
```

---

## Exercise C

**Add another line to overlay the posterior distribution of θ after 50 months.**

```{r fig.width=10,eval=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5
y <- dnorm(x, mean=60, sd=5)
plot(x, y, ylim=c(0, 0.2),  pch=18,col="red")

px1 <- post_mean + seq(-3, 3, by=0.2) * sqrt(post_var)
py1 <- dnorm(px1, mean=post_mean, sd=sqrt(post_var))
lines(px1, py1, col="green")

px2 <- post2_mean + seq(-3, 3, by=0.2) * sqrt(post2_var)
py2 <- dnorm(px2, mean=post2_mean, sd=sqrt(post2_var))
lines(px2, py2, col="blue")
```


---

## Exercise C

**Add another line to overlay the posterior distribution of θ after 50 months.**


```{r fig.width=10,echo=FALSE}
x <- 60 + seq(-3, 3, by=0.2) * 5
y <- dnorm(x, mean=60, sd=5)
plot(x, y, ylim=c(0, 0.2),  pch=18,col="red")

px1 <- post_mean + seq(-3, 3, by=0.2) * sqrt(post_var)
py1 <- dnorm(px1, mean=post_mean, sd=sqrt(post_var))
lines(px1, py1, col="green")


px2 <- post2_mean + seq(-3, 3, by=0.2) * sqrt(post2_var)
py2 <- dnorm(px2, mean=post2_mean, sd=sqrt(post2_var))
lines(px2, py2, col="blue")
```

---

### Comments

#### First Scenario
- The posterior distribution with sample size = 5 is close to the prior distribution. 
- There is a slight shift of the mean towards the sample mean and a similar dispersion.

#### Second Scenario
- When the sample size increased, the posterior distribution moves towards the sample mean and dispersion. 
- More weight is given to the sample where the sample size is big. 
- Additionally, the variation is reduced with a larger sample size.

---

---