---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

---

**A.** Generate 150 random numbers from a uniform [0,1] distribution and calculate the mean:

```{r}
set.seed(052023)
u <- runif(150)
mean_u <- mean(u)
mean_u
```

---

**B.** Simulate a sample from the chi-square distribution with 2 degrees of freedom using the random numbers generated in A:

```{r}
sample_chisq <- qchisq(u, df = 2)
```

#### First 10 values

```{r}
sample_chisq[1:10]
```

---


```{r fig.width=10}
hist(sample_chisq, 
     main = "Histogram of Chi-Square Sample", 
     xlab = "Value", ylab = "Frequency",breaks=20)
```

---

**C.** Simulate a sample from the Gamma (1, 1/2) distribution using the same random numbers:

```{r}
sample_gamma <- qgamma(u, shape = 1, rate = 0.5)
```

Samples generated in exercises B and C will be the same because the chi-square distribution with 2 degrees of freedom is a special case of the Gamma distribution with shape parameter 1 and rate parameter 0.5.

---

**D.** Plot histogram of the sample generated in exercise C and comment on the shape:

```{r eval=FALSE}
hist(sample_gamma, 
     main = "Histogram of Gamma Sample", 
     xlab = "Value", ylab = "Frequency",
     breaks=20)
```

---


### Gamma Sample (again)

```{r echo=FALSE,fig.width=10}
hist(sample_gamma, 
     main = "Histogram of Gamma Sample", 
     xlab = "Value", ylab = "Frequency",
     breaks=20)
```


---

### Chi-Square Sample (again)

```{r echo=FALSE,fig.width=10}
hist(sample_chisq, 
     main = "Histogram of Chi-Square Sample", 
     xlab = "Value", ylab = "Frequency",breaks=20)
```

---

**E.** Compute mean and median of the sample and explain why mean is greater than median:

```{r}
mean_sample <- mean(sample_chisq)
median_sample <- median(sample_chisq)
mean_sample
median_sample
```

The mean is greater than the median because the chi-square distribution is skewed to the right, meaning it has a longer tail on the right side.

---

**F.** Simulate 1000 values of the sum of samples of size 150 from the chi-square distribution with 2 degrees of freedom:


#### Single Instance
```{r}
sum(qchisq(runif(150), df = 2))
```


#### Simulate 1000 values
```{r}
set.seed(052023)
sums_chisq <- replicate(1000, sum(qchisq(runif(150), df = 2)))
```

#### First 10 values
```{r}
sums_chisq[1:10]
```
---

**G.** Plot histogram of the 1000 samples sum generated in exercise F and comment on the shape in the context of the central limit theorem:

```{r eval=FALSE}
hist(sums_chisq, 
     main = "Histogram of 1000 Samples Sum", 
     xlab = "Sum", 
     ylab = "Frequency",
     breaks=30)
```

---

```{r echo=FALSE,fig.width=10}
hist(sums_chisq, main = "Histogram of 1000 Samples Sum", xlab = "Sum", ylab = "Frequency",
     breaks=30)
```

---

* The histogram should show a normal distribution shape, which aligns with the central limit theorem.

* This theorem states that the sum (or average) of a large number of independent, identically distributed random variables will be approximately normally distributed, regardless of the original distribution.

---

---

Feel free to run this code in your R environment and let me know if you need further explanations or assistance with any part of it!