---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)
library(scales)
options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

### Compound Poisson Distribution

* The annual aggregate claim amount of an insurer follows a compound Poisson 
distribution with parameter 1,000. 

* Individual claim amounts follow a Gamma 
distribution with shape parameter $\alpha = 750$ and rate parameter $\lambda = 0.25$. 

* The relevant exercises are on the next slide

---

### Exercises

**1**. Generate 20,000 simulated aggregate claim values for this insurer, using a 
random number generator seed of 914. Use the R function, <tt>head()</tt>, to 
display the first seven simulated claim values in your answer script. 

**2**. Determine the mean and the standard deviation of the simulated aggregate 
claim values from part 1. 

**3**. Plot the empirical density function of the simulated aggregate claim values 
from part 1, setting the x-axis range from 2,600,000 to 3,300,000 and the 
y-axis range from 0 to 0.0000045. 

**4**. Suggest a suitable distribution, including its parameters, that approximates the simulated aggregate claim values from part 1.  

**5**. Generate 20,000 values from your suggested distribution in part 4 using a 
random number generator seed of 914. Use the R function, <tt>head()</tt>, to 
display the first seven generated values in your answer script. 

**6**. Plot the empirical density function of the simulated values in part 5. as a 
different coloured line in the chart that was produced in part 3.  


---

### Exercise 1

**Understanding the Problem:**

We're dealing with a compound Poisson distribution, where:

* **Primary process:** Poisson with rate λ = 1000, representing the number of claims.<p>
* **Secondary process:** Gamma with shape α = 750 and rate λ = 0.25, representing individual claim amounts.<p>

**Simulation:**
  - We simulate the number of claims using a Poisson distribution.<p>
  - For each claim, we simulate its amount using a Gamma distribution.<p>
  - The total aggregate claim is the sum of all individual claim amounts.
   
   


---

### Exercise 1
   
```{r}
set.seed(914) 

n = rpois(20000, 1000) 
s = numeric(20000) 

for(i in 1:20000) {
  x = rgamma(n[i], shape = 750, rate = 0.25) 

  s[i] = sum(x)
  } 
head(s, 7) 
```


---

### Exercise 1

#### Alternative Approach

```{r}
set.seed(914) 
s = numeric(20000) 

for(i in 1:20000) {
  n = rpois(1, 1000) 
  x = rgamma(n, shape = 750, rate = 0.25) 
  s[i] = sum(x)
} 

head(s, 7) 
```

---

### Exercise 2. 

**Mean and Standard Deviation:**
   - We calculate the mean and standard deviation of the simulated aggregate claims.

```{r}
mean(s) 
```

The mean of the simulated claim values is `r comma(mean(s))`

```{r}
sd(s) 
``` 

The standard deviation of the simulated claim values is `r comma(sd(s))`

---

### Exercise 3. 

```{r eval = FALSE}
plot( 
  density(s), 
  xlim = c(2600000, 3300000), 
  ylim = c(0, 4.5e-06), 
  xlab = "Simulated Claims", 
  main = "Probability Density Function",
  sub = "Simulated Claims from a Compound Poisson Distribution") 
```
---

### Exercise 3. 

```{r echo = FALSE}
plot( 
  density(s), 
  xlim = c(2600000, 3300000), 
  ylim = c(0, 4.5e-06), 
  xlab = "Simulated Claims", 
  main = "Probability Density Function",
  sub = "Simulated Claims from a Compound Poisson Distribution") 
```
---

### Exercise 4.

Suggest a suitable distribution, including its parameters, that approximates the simulated aggregate claim values

* Given the shape of the Gamma distribution, a Normal distribution might be a good approximation for the aggregate claims.
* We can use the mean and standard deviation calculated in exercise 2 as parameters for the Normal distribution.

#### Normal distribution 

* mean = 3,000,982 
* standard deviation = 93,872.61  
* 93,872.61^2 = 8,812,067,277 
 $$ X \sim N(\mu=3,000,982, \sigma^2 = 8,812,067,277)$$

---

#### Exercise 5.

20,000 values from your suggested distribution in part 4 using a 
random number generator seed of 914.

```{r}
set.seed(914) 
approx_dist = rnorm(20000, mean(s), sd(s)) 
head(approx_dist, 7) 
```

---

### Exercise 6.

* By comparing the two density functions, we can visually assess how well the Normal distribution approximates the compound Poisson distribution. 

* If the two distributions are close, the Normal approximation can be used for further calculations and risk analysis.

---

### Exercise 6.

```{r eval=FALSE}
plot( 
  density(s), 
  xlim = c(2600000, 3300000),   ylim = c(0, 4.5e-06), 
  xlab = "Simulated Claims", 
  main = "Probability Density Functions")

### Super-imposed theoretical distribution 

lines( 
  density(approx_dist), 
  col = "red") 

legend("topright", 
  legend = c("Simulated Claims", "Approximate Normal Distribution"), 
  col = c("black","red"), 
  pch = 7)
```

---

### Exercise 6.



```{r echo=FALSE}
plot( 
  density(s), 
  xlim = c(2600000, 3300000),   ylim = c(0, 4.5e-06), 
  xlab = "Simulated Claims", 
  main = "Probability Density Functions")

lines( 
density(approx_dist), 
col = "red") 
legend("topright", 
legend = c("Simulated Claims", "Approximate Normal 
Distribution"), 
  col = c("black","red"), 
  pch = 7)
```
---


---

---