---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


2 Before answering this question, the data set 𝑥, representing n = 100 observations from 
an ARMA(2,0) model, must be generated in R using the following code: 
```{r}
set.seed(12456) 
x = arima.sim(n = 100, model = list(ar = c(0.7,0.2)))
```

---

### Exercise 1 and 2

1. Plot the sample autocorrelation function (sample ACF) and sample partial 
autocorrelation function (sample PACF) for the data set 𝑥. 

2. Comment on the general features of the two plots in Part A. with emphasis on 
how they compare to the theoretical behaviour of the corresponding functions 
for an ARMA(2,0) process.  

---



```{r}
set.seed(12456)
x = arima.sim(n = 100, model = list(ar = c(0.7,0.2)))

```

```{r}
acf(x, 
main = "Sample ACF of 100 Simulated Values from an ARMA(2,0) 
model") 
```

---

```{r}
pacf(x,  main ="Sample PACF of 100 Simulated Values from anARMA(2,0) model") 
```

---


### Exercise 2

Commenting on ACF and PACF

ACF: The ACF plot shows a significant decay over the first few lags, which is characteristic of an AR process. This decay is due to the autoregressive terms in the ARMA(2,0) model.
PACF: The PACF plot shows significant spikes at lags 1 and 2, and then cuts off. This is consistent with the theoretical behavior of the PACF for an AR(2) process. The spikes at lags 1 and 2 indicate the presence of two autoregressive terms.

---
### Exercise 3

Fit the following three models to the generated data set 𝑥, displaying their 
Akaike Information Criterion (AIC) values in the R output in your answer 
script: 
ARMA(1,0) 
ARMA(1,1) 
ARMA(2,0) 
  
---

#### Exercises
Part D Explain, using the results from part Part C, which of the fitted models is the most 
appropriate for modelling the data set x.  
Part E Comment on your answer to part Part D. 
Part F Explain, with reasons, how the value of n in the R code for generating data set 
x may be changed to ensure that the ARMA(2,0) model becomes the best 
fitting model.  






(ii)
EITHER:
The ACF is decreasing (exponentially) to zero 
which is in line with the theoretical behaviour of an ARMA(2,0) process 
OR:
The ACF decreases to zero more slowly than might be expected from an ARMA(2,0)
process 
THEN:
* But the PACF seems to cut off after lag 1. 
* We might expect the PACF of an ARMA(2,0) process to cut off after lag 2 
However, given the small sample size and the fact that this process is close to being non-stationary, the two plots are not out-of-line with expectations. 

---
Part C

```{r}
fitx10 = arima(x, order = c(1,0,0)) 
fitx10$aic 
 305.4467 
 
```

```{r} 
fitx11 = arima(x, order = c(1,0,1)); fitx11$aic 
 307.4325
```


```{r} 
fitx20 = arima(x, order = c(2,0,0)); fitx20$aic 
 307.4309
```

---
 
#### Part (d)
The preferred model here is ARMA(1,0) 
as the AIC value is the lowest 
---
Part E

We might have expected the ARMA(2,0) model to be the best fit as that is the “true”
model that x was generated from 
However, the number of observations of x is quite small 
and so sampling uncertainty has produced a better fitting / forecasting model that is
not ARMA(2,0) 

---
Part F
We could increase the value of n in the R code (to say 1,000) 
which would reduce the sampling uncertainty of the observations 

---

### Exercise 7
```{r}
set.seed(12456) 
y = arima.sim(n = 1000, model = list(ar = c(0.7,0.2))) 
```

---

#### Exercises

7. Generate a new data set, y, using your suggested change to the value of $n$ in 
part F. Set the same random number generator seed as above before 
generating y. 

8. Fit the three models to the new data set y, displaying their AIC values in the R 
output in your answer script. 

9. Explain, using the results from part (viii), why the ARMA(2,0) model is the 
most appropriate model for modelling data set y.  

10. Comment on the differences between the suggested models fitted for data sets 
x and y in parts Part D and (ix), respectively. 




---

### Exercise 8

```{r}
fity10 = arima(y, order = c(1,0,0)); fity10$aic 
```

```{r}
fity11 = arima(y, order = c(1,0,1)); fity11$aic 
```

```{r} 
fity20 = arima(y, order = c(2,0,0)); fity20$aic 
```


---

### Exercise 9

* The preferred model here is ARMA(2,0) as the AIC value is the lowest 

---
(x)

```{r}
fitx10
```

Call:
arima(x = x, order = c(1, 0, 0))
Coefficients:
 ar1 intercept
 0.9530 -1.0632
s.e. 0.0279 1.9238
sigma^2 estimated as 1.142: log likelihood = -149.72, 
aic = 305.45 
fity20
Call:
arima(x = y, order = c(2, 0, 0))
Coefficients:
 ar1 ar2 intercept
 0.7116 0.2112 0.2409
s.e. 0.0309 0.0309 0.4064
sigma^2 estimated as 1.014: log likelihood = -1426.62,
aic = 2861.24 

---

For data set y, the best fitting model is the “true” model 
Whereas the best fitting model for data set x is not the “true” model 

---

* This is what we expected as the more observations we have the more likely it is that
the “true” model is the best fit 

* The estimated AR parameters for data set y are close to the real values of 0.7 and 0.2 
*The estimated error variance $sigma^2$ is closer to the real value of 1 for data set y
than for data set x. 

* The estimated intercept term is closer to the real value of zero for data set y than
for data set x. 

* These observations on the estimated error variance and intercept term are to be
expected given that the data volume is higher in data set y than in data set x 
and given that the fitted model matches the true model for data set y but not for
data set x 

---