---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
CS2B A2022–4 
3 An international pensions provider is interested in quantifying the force of mortality at 
certain ages for a particular country, for the period from 1 January 2017 through to 
1 January 2020. The ‘CS2B_A22_Qu_3_Data.csv’ file contains mortality data from a 
recent mortality investigation. The deaths and exposure data are all in respect of the 
period 1 January 2020 to 31 December 2020 inclusive. The file contains the following 
six variables: 
Age Age last birthday, x, in single ages from 55 to 95 inclusive 
Deaths Number of observed deaths at age x, Dx
Exposure Central exposed to risk at age x, Ex
c (measured in years) 
Graduation1 Central mortality rate at age x, mx, derived from graduation method 1 
Graduation2 Central mortality rate at age x, mx, derived from graduation method 2 
Graduation3 Central mortality rate at age x, mx, derived from graduation method 3 
The graduation methods are set out below: 
Graduation 1 Graduation by parametric formula 
Graduation 2 Graduation obtained by removing one parameter from the formula 
underlying Graduation 1 
Graduation 3 Graduation obtained by adding one parameter to the formula 
underlying Graduation 1 
Before answering this question, the ‘CS2B_A22_Qu_3_Data.csv’ file should be 
loaded into R and assigned to a data frame called graduation. 

---
(i) Construct three new columns in graduation, called zx1, zx2 and zx3, 
containing the standardised deviations: 
zx = Dx – mxEx
c
 
 √ mxEx
c
for each of the graduations. 
Use the R function, head(), to display the first seven values of columns zx1, 
zx2 and zx3 ONLY in your answer script. [6] 
(ii) Determine, using R, the p-value of the chi-square goodness-of-fit test for each 
of the three graduations given that the numbers of degrees of freedom are as 
follows: 
Graduation 1 2 3 
Degrees of freedom 36 37 35 
--- 
(iii) Comment on your results in part (ii), with reference to the suitability of the 
three graduations for the observed mortality rates. [6] 
(iv) Determine, using R, the numbers of positive and negative deviations for each 
of the three graduations. 

(v) Determine, using R, the numbers of groups of positive deviations for each of 
the three graduations. 
(vi) Determine, using R, the p-value of the grouping of signs test for each of the 
three graduations. 
[Hint: If you are using the binomial coefficients ൫
n
r
൯ in your calculations, 
these may be obtained in R as choose(n, r).] [9] 
(vii) Comment on whether your conclusions in part (iii) have changed in light of 
the results in part (vi). [2] 
 [Total 51] 
END OF PAPER 
---

Q3
(i)
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2022 - Examiners’ report
---

```{r}
graduation = 
read.csv(file = "CS2B_A22_Qu_3_Data.csv", head = TRUE) 
```

---
graduation$zx1 = (graduation$Deaths - graduation$Exposure * 
graduation$Graduation1) / sqrt(graduation$Exposure * 
graduation$Graduation1) [1½]
graduation$zx2 = (graduation$Deaths - graduation$Exposure *
graduation$Graduation2) / sqrt(graduation$Exposure * 
graduation$Graduation2) 

graduation$zx3 = (graduation$Deaths - graduation$Exposure * 
graduation$Graduation3) / sqrt(graduation$Exposure * 
graduation$Graduation3) 

---


head(graduation[, 7:9], 7) 
 zx1 zx2 zx3
1 -0.4532486 2.08008602 0.077224448
2 -0.7304714 1.21170927 -0.453050849
3 0.4301427 1.78813614 0.498299817
4 -0.1565084 0.57642875 -0.250408199
5 0.5517038 0.72246846 0.351202375
6 0.2811644 -0.07163144 0.002057777
7 -0.5559179 -1.34146800 -0.839474197 
---
(ii)
EITHER:
```{r}
chisq = vector(length = 3) 
chisq = sum(graduation$zx1^2) 
chisq[2] = sum(graduation$zx2^2) 
chisq[3] = sum(graduation$zx3^2)
```
--- 
df = c(36, 37, 35) 
1 - pchisq(chisq, df = df) [1½]
 7.002971e-02 5.582300e-08 7.128381e-02 
OR:
chisq1 = sum(graduation$zx1^2) 
chisq2 = sum(graduation$zx2^2) 
chisq3 = sum(graduation$zx3^2) 
1 - pchisq(chisq1, df = 36) [1½]
 0.07002971 
1 - pchisq(chisq2, df = 37) 
 5.5823e-08 
1 - pchisq(chisq3, df = 35) 
 0.07128381 
THEN:
The p-value for graduation 1 is 0.07003
The p-value for graduation 2 is 5.582e-08
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2022 - Examiners’ report
---
The p-value for graduation 3 is 0.07128 

---
(iii)
Based on the p-values, there is enough evidence to reject graduation 2 at the 5% significance 
level but not graduations 1 and 3 
Since graduation 2 does not fit the rates well, the parameter that was removed from 
graduation 1 to give this graduation is statistically significant and should be added
back in [2]
The results do not support the need for the additional parameter in graduation 3
compared with graduation 1. 
For graduation 3 the improvement in fit approximately balances out the lost
degree of freedom. 
Although graduations 1 and 3 appear suitable based on the chi-square test, they
may be unsuitable for other reasons 
Any appropriate example, e.g. small consistent bias, a few large outliers, or clumps
of deviations of the same sign 

---
(iv)
EITHER:
pos = vector(length = 3) 
neg = vector(length = 3) 
pos = length(graduation$zx1[graduation$zx1 0]) 
pos[2] = length(graduation$zx2[graduation$zx2 0]) 
pos[3] = length(graduation$zx3[graduation$zx3 0]) 
neg = length(graduation$zx1[graduation$zx1 < 0]) 
neg[2] = length(graduation$zx2[graduation$zx2 < 0]) 
neg[3] = length(graduation$zx3[graduation$zx3 < 0]) 
pos 
 22 20 21 
neg 
 19 21 20 
OR:
pos1 = length(graduation$zx1[graduation$zx1 0]) 
pos2 = length(graduation$zx2[graduation$zx2 0]) 
pos3 = length(graduation$zx3[graduation$zx3 0]) 
neg1 = length(graduation$zx1[graduation$zx1 < 0]) 
neg2 = length(graduation$zx2[graduation$zx2 < 0]) 
neg3 = length(graduation$zx3[graduation$zx3 < 0]} 
pos1
 22 
pos2
 20 
pos3
 21 
neg1
 19 
neg2

---
 21 
neg3
 20 
THEN:
Graduation 1 has 22 positive and 19 negative deviations.
Graduation 2 has 20 positive and 21 negative deviations.
Graduation 3 has 21 positive and 20 negative deviations. 

---
(v)
EITHER:
groups = vector(length = 3) 
for(j in 1:3){ 
pos_z = (graduation[, j + 6] 0) * 1 
groups[j] = 
sum( 
duplicated(c(which(pos_z == 1) - 1, which(pos_z == 0))) * 1) 
 [3½]
+ pos_z * 1} [2]
groups 
 11 5 11 
OR:
pos_z = (graduation$zx1 0) * 1 
groups1 = 
sum( 
duplicated(c(which(pos_z == 1) - 1, which(pos_z == 0))) * 1) 
 [3½]
+ pos_z * 1 [2]
pos_z = (graduation$zx2 0) * 1
groups2 =
sum(
duplicated(c(which(pos_z == 1) - 1, which(pos_z == 0))) * 1)
+ pos_z * 1 
pos_z = (graduation$zx3 0) * 1
groups3 =
sum(
duplicated(c(which(pos_z == 1) - 1, which(pos_z == 0))) * 1)
+ pos_z * 1 
groups1 
 11 
groups2
 5 
groups3
 11 
THEN:
Graduation 1 has 11 groups of positive deviations.

---
Graduation 2 has 5 groups of positive deviations.
Graduation 3 has 11 groups of positive deviations. 
ALTERNATIVE SOLUTION:
EITHER:
```{r}
groups = vector(length = 3) 
for(j in 1:3){ 
pos_z = (graduation[, j + 6] 0) * 1 
groups[j] = 
sum( 
duplicated(c(which(pos_z == 1) - 1,
which(pos_z == 0))) * 1) [3½]
+ pos_z * 1 [2]
}
groups 
```

 11 5 11 
OR:
pos_z = (graduation$zx1 0) * 1 
groups1 = 
sum( 
duplicated(c(which(pos_z == 1) - 1,
which(pos_z == 0))) * 1) [3½]
+ pos_z * 1 [2]
pos_z = (graduation$zx2 0) * 1
groups2 =
sum(
duplicated(c(which(pos_z == 1) - 1,
which(pos_z == 0))) * 1)
+ pos_z * 1 
pos_z = (graduation$zx3 0) * 1
groups3 =
sum(
duplicated(c(which(pos_z == 1) - 1,
which(pos_z == 0))) * 1)
+ pos_z * 1 
groups1 
 11 
groups2
 5 
groups3
 11 
THEN:
Graduation 1 has 11 groups of positive deviations.
Graduation 2 has 5 groups of positive deviations.
Graduation 3 has 11 groups of positive deviations. 
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2022 - Examiners’ report
---
(vi)
EITHER:
pval = vector(length = 3) 
for(j in 1:3){ 
pval[j] = 0 
for(t in 1:groups[j]){ 
pval[j] = pval[j] + choose(pos[j] - 1, t - 1) *
choose(neg[j] + 1, t) / 
choose(pos[j] + neg[j], pos[j])}} [3]
pval 
 0.6844862501 0.0004065953 0.6787210548 
OR:
pval1 = 0 
for(t in 1:groups1){ 
pval1 = pval1 + choose(pos1 - 1, t - 1)*
choose(neg1 + 1, t) /
choose(pos1 + neg1, pos1)} [3]
pval2 = 0
for(t in 1:groups2){
pval2 = pval2 + choose(pos2 - 1, t - 1)*
choose(neg2 + 1, t) /
choose(pos2 + neg2, pos2)} 
pval3 = 0
for(t in 1:groups3){
pval3 = pval3 + choose(pos3 - 1, t - 1)*
choose(neg3 + 1, t) /
choose(pos3 + neg3, pos3)} 
pval1 
 0.6844862501 
pval2
 0.0004065953 
pval3
 0.6787210548 
THEN:
The p-value for graduation 1 is 0.6845
The p-value for graduation 2 is 0.0004066
The p-value for graduation 3 is 0.6787 
(vii)
The conclusions in part (iii) remain unchanged 
as the p-values calculated under the grouping of signs test support the conclusions
in part (iii) 
