---
title: "Spanning Trees"
subtitle: "DiscreteMaths.github.io"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    font_size: 20pts
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

<style type="text/css">
  body{
  font-size: 20pt;
}
</style>

```{r eval=FALSE,include=FALSE}

# CT6 
# - ASI 2005 April Question B - 1
# - ASI 2006 April Question B - 1

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#4169e1", #)(vvv#TREES
  #background_color = "#E6E6FA",  #TREE
  base_color = "black", #)(
  background_color =  "#dbcccc",
  text_color="black",
  text_bold_color = "black",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "400", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```
<style> 
  /* Hide page numbers */ .
  remark-slide-number { display: none; } 
</style>

## Bayesian Statistics with R
### Bayesian Analysis of Cloud Kitchen Sales
<br><br><br><br>
## <tt>Statistical Computing with R</tt>
###<tt>CS1B (ASI - May 2023) 

---

###  Logical Operations
Let $n \in \{1,2,3,4,5,6,7,8,9\}$ and let p and q be the following propostions concerning 
the integers $n$.


* ***p***: $n$ is event
* ***q***: $n<5$


Find the values of $n$ for which each of the following compound statement is true,

A. $\neg p$

B. $p \wedge q$

C. $\neg p \vee q$ 

D. $p \oplus q$

---


Find the values of $n$ for which each compound statement is true.

#### Given Propositions:

- $p$: $n$ is even.
- $q$: $n < 5$.

#### Possible Values of $n$:

$$n \in \{1, 2, 3, 4, 5,6, 7, 8, 9\}$$
---

### Exercise A: $\neg p$

**($n$ is not even):**

$$
\neg p: n \text{ is odd}
$$

The values of $n$ for which $\neg p$ is true are:
$$\{1, 3, 5, 7, 9\}$$

---

### Exercise B: $p \wedge q$ 

**( $n$ is even and $n < 5$ ):**

$$
p \wedge q: n \text{ is even and } n < 5
$$

The values of $n$ for which $p \wedge q$ is true are:
$$\{2, 4\}$$
