---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


---

Here's a short tutorial on using the `calc_auc` function in R with the `ggRandomForest` package:

### Example Code
```{r}
# Load the necessary libraries
library(ggRandomForests)
library(randomForestSRC)

# Use the built-in iris dataset for demonstration
data(iris)

# Build a random forest model using the iris dataset
rfsrc_iris <- rfsrc(Species ~ ., data = iris)

```


---

```{r}
# Generate ROC curves for the first and second outcomes
gg_dta1 <- gg_roc(rfsrc_iris, which.outcome = 1)
gg_dta2 <- gg_roc(rfsrc_iris, which.outcome = 2)
```


---

```{r}
# Calculate the Area Under the ROC Curve (AUC) for both outcomes
auc1 <- calc_auc(gg_dta1)
auc2 <- calc_auc(gg_dta2)
```


---

```{r}
# Print the AUC values
print(paste("AUC for outcome 1:", auc1))
print(paste("AUC for outcome 2:", auc2))
```

### Explanation:
1. **Load Libraries**: Start by loading the necessary libraries, `ggRandomForests` and `randomForestSRC`.
2. **Build Random Forest Model**: Utilize the built-in `iris` dataset to create a random forest model predicting `Species`.
3. **Generate ROC Curves**: Apply the `gg_roc` function to generate ROC curves for the first and second outcomes.
4. **Calculate AUC**: Use the `calc_auc` function to compute the Area Under the ROC Curve (AUC) for both outcomes.
5. **Print Results**: Finally, print the AUC values for each outcome to the console.

This step-by-step example demonstrates how to calculate and interpret the Area Under the ROC Curve for a random forest model built with the iris dataset.

If you have any further questions or need more assistance, feel free to ask!

```{r}
library(ggRandomForests)
```


---


