---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3
---


Q. 3) An analysis was carried out to investigate the annual average rainfall of two countries. 
The data is as below:
Iran: 128,125,133,104,146,132,125,118,129,124
Belgium:160,128,169,105,151,164,162,177,185,150,182,158,156,123,141,176,162,172
i) Perform a suitable test to determine whether the rainfall in both countries has equal 
variance or not, at the 5% confidence level. (6)
ii) Test whether the mean rainfall in both countries is equal or not, at 5% confidence 
level. 
iii) Calculate the 95% confidence interval for difference in means. (2)
iv) Comment on the results in part (ii) and (iii).
---

```{r}
Iran: 128,125,133,104,146,132,125,118,129,124
Belgium:160,128,169,105,151,164,162,177,185,150,182,158,156,123,141,176,162,172
```

---