---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
lib_dir: libs
nature:
  highlightStyle: github
  countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3
---


Q. 3) An analysis was carried out to investigate the annual average rainfall of two countries. 
The data is as below:
Iran: 128,125,133,104,146,132,125,118,129,124
Belgium:160,128,169,105,151,164,162,177,185,150,182,158,156,123,141,176,162,172

#### Exercises

1.  Perform a suitable test to determine whether the rainfall in both countries has equal 
variance or not, at the 5% confidence level. (6)
2.  Test whether the mean rainfall in both countries is equal or not, at 5% confidence 
level. 
3.   Calculate the 95% confidence interval for difference in means. 
4.  Comment on the results in part (2.  and (3.  .
---

```{r}
Iran <- c(128,125,133,104,146,132,125,118,129,124)
Belgium <- c(160,128,169,105,151,164,162,177,185,150,182,158,156,123,141,176,162,172)
```

---

Solution 3:
i)
H0: Annual average rainfall of Belgium and Iran are same
H1: Annual average rainfall of Belgium and Iran are not same


```{r}
Iran <- c(128,125,133,104,146,132,125,118,129,124)
Belgium <- c(160,128,169,105,151,164,162,177,185,150,182,158,156,123,141,176,162,172)
```

---

```{r}
var.test(Iran, Belgium)
```

---

F test to compare two variances
data: Iran and Belgium
F = 0.25802, num df = 9, denom df = 17, p-value = 0.04385
alternative hypothesis: true ratio of variances is not equal to 1
95 percent confidence interval:
0.0864436 0.9602591
sample estimates:
ratio of variances 
 0.258022
Since p-value < 0.05 we fail the variance test thus we reject the null hypothesis that both have equal 
variance

---

ii)

```{r}
t.test(Iran, Belgium, var.equal = FALSE)
```

OUTPUT
data: Iran and Belgium
t = -4.9984, df = 25.904, p-value = 3.407e-05
alternative hypothesis: true difference in means is not equal to 0

95 percent confidence interval:
-42.79403 -17.85041
sample estimates:
mean of x mean of y 
126.4000 156.7222
---

#### Conclusion

Since, P-value<0.05 we reject the null hypothesis and can conclude that both cities have different amount 
of rainfall with 95% confidence.

---

### Exercise C

iii) Confidence interval can be read from part b
95 percent confidence interval:
-42.79403 -17.85041

---

### Exercise D

* The confidence interval (-42.8,-17.8) does not contain 0. 
* Therefore the assumption of equal means is 
not true. This result is in line with the conclusion in part (b). 

---