---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Q. 1) Identify the probability distribution that best fits the below questions. Calculate the 
following using R functions:
1.  Assume golf balls from the driving range next door lands in your yard at an average 
rate of 3 balls per hour during the day. What is the probability that 10 or fewer golf 
balls will land in your yard during the afternoon, assuming the afternoon is 5 hours 
long? 
2.  You are surveying people exiting from a polling booth and asking them if they voted 
independently. The probability that a person voted independently is 20%. What is 
the probability that 70 people must be asked before you can find 5 people who voted 
independently? 

---
3.   Assume you flip a fair coin 100 times. What is the number N such that, 90% of the 
time, the number of heads is less than or equal to N?
4.  A researcher is waiting outside of a library to ask people if they support a certain 
law. The probability that a given person supports the law is p = 0.2. What is the 
probability that the fourth person the researcher talks to is the first person to support
the law?
5.  Assume that a light bulb has a mean lifetime of 1000 hours. What is the probability 
that the light bulb survives to 2000 hours?
6.   Assume a random variable Z is distributed according to the normal distribution with 
mean 6 and standard deviation 4. What is the probability that Z takes on a value 
between -1 and 3?
---

---
Solution 1:
i) X ~ Poisson (15)
ppois(10, 15)
[1] 0.1184644

---

ii) X ~ NB(5,0.2)

```{r}
dnbinom(65,5,0.2)
```
[1] 0.00013892

---

iii) X ~ Binom (100,0.5)

```{r}
qbinom(0.9, 100, 0.5)
```
[1] 56

---

iv) X ~ Geometric (0.2)

```{r}
dgeom(x=3, prob=0.2)
```

[1] 0.1024


---

v) X ~ Exp(1/1000)

```{r}
1 - pexp(2000, 0.001)
```
[1] 0.1353353

---
vi) X ~ N(6,16)
pnorm(3, 6, 4) - pnorm(-1, 6, 4)
[1] 0.1865682

---