---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


```{r include=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(knitr)
library(magrittr)
library(flextable)
library(kableExtra)
```

In a small country, a political election was held recently to decide on a political party 
to govern the country. A survey was conducted to monitor the approval rating for the 
winning political party. 

A sample of 15 voters were asked to complete the same survey twice: once before the 
election (Approval before) and a second time 6 months after the election (Approval 
after). 

---

The survey asked each of the 15 voters to record their approval for the winning 
political party on a scale from 1 to 10, where ***1 = Strongly disapprove*** and ***10 = Strongly approve***.
<p>
The results are shown in the table below:


```{r echo=FALSE, warning=FALSE}
voters <- as.character(1:15)

approval_before <- c(8, 6, 8, 7, 7, 4, 2, 10, 8, 7, 10, 
8, 8, 9, 6) 
approval_after <- c(5, 6, 2, 3, 4, 1, 4, 7, 4, 2, 10, 1, 
5, 3, 7) 

thisTable <- data.frame(voters,approval_before,approval_after)

thisTable_t <-thisTable %>% data.table::transpose() 

thisTable_t <- bind_cols(varnames=c("Voter","Before","After"),thisTable_t)

thisTable_t %>% kable( col.names = NULL) %>%  
      kable_styling(bootstrap_options = c("bordered", full_width = F, font_size = 16)) %>%
  row_spec(1, background = "AAEEFF", color = "black", font_size = 18)


```
<p>
------------------------------------
<p>
The values can be entered in R using: 
<pre><code>
approval_before <- c(8, 6, 8, 7, 7, 4, 2, 10, 8, 7, 10, 
8, 8, 9, 6) 
approval_after <- c(5, 6, 2, 3, 4, 1, 4, 7, 4, 2, 10, 1, 
5, 3, 7) 
</code></pre>
---


#### Exercises

1. Calculate the means for the Approval before and Approval after results. 

2.  Outline why Kendall’s Tau would be a suitable correlation coefficient to use 
for these results. 

3. Calculate Kendall’s Tau coefficient between the Approval before and 
Approval after results. 

4. Comment on your results in Exercises 1 and 3. 

---

### Exercise 1

#### Sample Estimates

Enter the data into R using the commands provided in the question
```{r}
approval_before <- c(8, 6, 8, 7, 7, 4, 2, 10, 8, 7, 10, 8, 8, 9, 6)
approval_after <- c(5, 6, 2, 3, 4, 1, 4, 7, 4, 2, 10, 1, 5, 3, 7)
```

```{r}
mean(approval_before)

mean(approval_after) 
```

---
### Exercise 2 and 3

#### Kendall’s Rank Correlation Coefficient

* Kendall’s rank correlation coefficient will measure the strength of monotonic relationship between approval before and approval after results
<p>
* It considers only the relative values of the bivariate data, and not their actual values
<p>
* It doesn’t assume linearity between the bivariate data.


```{r}
cor(approval_before,approval_after,method="kendall") 
```

---
### Exercise 3

#### Spearman Correlation Coefficient



```{r}
cor(approval_before,approval_after,method="spearman") 
```


#### Pearson Correlation Coefficient

```{r}
cor(approval_before,approval_after) 
```

---

### Exercise 4

* From Exercise 1, the mean approval has fallen from before (7.2) to after (4.3) 
* Kendall’s Tau in Exercise 3 shows that there is weak correlation between the approval
scores before and after 
* From Exercises 1 and 3, this suggests that generally, the 15 voters’ approval for the
political party has changed, i.e. approval has fallen (or, the shift in opinion has not
been the same across those surveyed)

---

---