---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

4 An actuarial modeller in a health insurance company is constructing a Generalised 
Linear Model (GLM) to analyse claim numbers for its critical illness policies. For 
every policy over the past year, the company has collected the number of reported 
claims (Claim_number) and data on the following covariates:

* Age: Age of policyholder, a number between 18 and 80 
* Gender: Gender of the policyholder (Male or Female) 
* Region: A description of the region 
where the policyholder lives 
* Pre_existing_health_condition: A categorical value representing whether the 
policyholder has a pre-existing health condition, 
0 = no and 1 = yes 

---

The data given in the file named Claims_Experience.RData show the past 
year’s reported claims for this company’s critical illness policies. After loading the 
data into R, using the command load("Claims_Experience.RData"), the 
data frame data_claims with its variables listed above will be available. 

The modeller wants to fit a GLM with the Claim_number as the response variable 
and is deciding whether to fit the GLM using a Poisson or Normal distribution. 

---

#### Exercises 

A. Fit a GLM, using a Normal distribution, that treats age as a numerical variable 
and the remaining three covariates as factors, calling this model_g. Your 
answer should include the estimated value, standard error and p-value of each 
parameter in the model. 

B.  Fit a GLM, using the Poisson distribution, that treats age as a 
numerical variable and the remaining three covariates as factors, 
calling this model_p. 
Your answer should include the estimated value, standard error and 
p-value of each parameter in the model. 

C. Justify which model the modeller should use, using your answers to 
parts (i) and (ii)(a). 
(c) Comment on the dependence of the number of reported claims on 
Pre_existing_health_condition based on your answers to 
parts (ii)(a) and (ii)(b). 

---

#### Exercises 

The modeller has fitted the GLM using a Poisson distribution and now wants to 
investigate which of the three factors should be selected in the model, by using a 
method similar to the backward selection approach. The following models are 
considered: 

* model_p2: removes Gender from model_p
* model_p3: removes Gender and Pre_exisiting_health_condition from model_p
* model_p4: removes Gender, Pre_exisiting_health_condition and Region from <tt>model_p</tt> 
---

D. Compare the values of the Akaike’s Information Criterion for the four models 
<tt>model_p</tt>, <tt>model_p</tt>2, <tt>model_p</tt>3 and <tt>model_p</tt>4. 
---

E. Comment on your answer to part (iii). 
(b) Comment on the differences between the process in part (iii) and the 
full backward selection method. 

---
The modeller suggests refining the fitted GLM (<tt>model_p</tt>) by also including the 
interactions between the variables: Age, Region and 
Pre_existing_health_condition. 

---


(v) (a) Fit an appropriate model, to include the interactions between the 
variables Age, Region and 
Pre_existing_health_condition. 
(b) Justify whether the refined model in part (v)(a) improves the model 
(<tt>model_p</tt>) fitted in part (ii)(a). 
(c) Comment on whether any of the interactions are significantly 
associated with claim numbers. 

---

**Part A: Normal GLM**

```R
# Load the data
load("Claims_Experience.RData")

# Fit the Normal GLM
model_g <- glm(Claim_number ~ Age + Gender + Region + Pre_existing_health_condition, data = data_claims, family = gaussian)

# Summarize the model
summary(model_g)
```

**Part B: Poisson GLM**

```R
# Fit the Poisson GLM
model_p <- glm(Claim_number ~ Age + Gender + Region + Pre_existing_health_condition, data = data_claims, family = poisson)

# Summarize the model
summary(model_p)
```

**Interpreting the Results:**

Both models will provide estimates for the coefficients of the covariates. These coefficients represent the change in the expected value of the response variable (Claim_number) associated with a unit change in the corresponding covariate, holding other covariates constant.

**Choosing the Appropriate Model:**

The choice between a Normal and Poisson GLM depends on the nature of the response variable and the assumptions underlying each model:

* **Normal GLM:** Assumes the response variable is normally distributed. It's suitable for continuous variables.
* **Poisson GLM:** Assumes the response variable is count data (non-negative integers). It's suitable for modeling rates or counts of events.

**In this case, the `Claim_number` is a count variable, so the Poisson GLM is more appropriate.**

**Additional Considerations:**

* **Overdispersion:** If the observed variance in the data is larger than the expected variance under the Poisson model, overdispersion may be present. This can be addressed using quasi-Poisson or negative binomial models.
* **Zero-Inflated Data:** If there are many zero counts in the data, a zero-inflated Poisson or zero-inflated negative binomial model may be considered.
* **Model Diagnostics:** It's important to check the model assumptions (e.g., linearity, constant variance, normality of residuals for the Normal GLM) and consider model selection techniques (e.g., AIC, BIC) to identify the best-fitting model.
