---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 3 - Probability Distributions
* Exam on basis of ***Base R*** 

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

An insurer’s marketing team has developed a new lottery that gives each new 
customer the chance to win a cash prize. The insurer states that any new customer can 
win with probability 0.36 independently of all other customers.

Let $Y$ denote the number of winners in a random sample of 900 new customers. 

#### Exercises
1.  State a suitable distribution for the random variable Y. 
2. Calculate, to four decimal places, the probabilities $P[k < Y \leq k + 20]$ for all 
values $k = 220, 240, 260, \ldots, 400$.

3. It is suggested that the true distribution of Y can be approximated with a Normal 
distribution with expectation $E[Y] = 324$ and a standard deviation of 14.4. Calculate, to four decimal places, the probabilities $P[k < Y \leq k + 20]$ for all 
values $k = 220, 240, 260, \ldots, 400$, based on the suggested Normal distribution. 

4. Plot the values calculated in parts 2. and 3. on the same graph. 
*Hint: you may find the R command lines(…) useful.*

5. Discuss your answers in parts 1 to 4.


---

#### Exercise 1

This is a Binomial(900, 0.36) distribution. 

```{r}
n <- 900
p <- 0.36
```

---

### Cumulative Distribution Function

```{r}

q = 220 + 20*(0:10)
q
```

```{r}
cdfb = pbinom(q, n, p) 

round(cdfb,4)
```

---

### Probability for intervals under Binomial distribution

```{r}
pb = diff(cdfb) 


round(pb,4)



```

### Alternative Approach

```{r}
pb = cdfb[-1] - cdfb[-11]
round(pb,4)
```


---

### Alternative Approach

```{r} 

k <- seq(220,400,by=20)

pb <- pbinom(k+20,900,0.36)-pbinom(k,900,0.36)

round(pb,4)
```

---

### Exercise 4

```{r}

cdfn = pnorm(q, mean=324, sd=14.4) 
pn = diff(cdfn) 
```



```{r}
round(pn, 4)
```

---

### Alternative using continuity correction:

```{r}
k <- seq(220,400,by=20)

pn <- pnorm(k+20.5, mean=324, sd=14.4) -
                pnorm(k+0.5, mean=324, sd=14.4)
round(pn,4)
```

---

### Exercise 4

Plot both sets of probabilities on the same graph to compare the values:

<pre><code>
plot(q[-1], round(pb,4), type="p", xlab="k", ylab="probability",col="black", pch=16) 
# small circles show for Bin(n,p) prob. 

lines(q[-1], round(pn,4),col="green",lwd=2) 

</code></pre>


---


```{r echo=FALSE}
plot(q[-1], round(pb,4), type="p", xlab="k", ylab="probability",col="black") 
# small circles show for Bin(n,p) prob. 

lines(q[-1], round(pn,4),col="green") 

```
---

### Central Limit Theorem

* The plot shows that the suggested $N(324, 14.42)$ distribution is a very good approximation to 
the binomial distribution 

* The reason for that is the CLT since the binomial distribution is the distribution of a sum of $n$ independent random variables and $n = 900$ is large 

* The approximation only works well since the parameters of the normal distribution have been 
chosen to match the first two moments of the normal distribution and the binomial 
distribution, $E(Y) = 900 \times  0.36 = 324$ 
and $$ Var(Y) = 900 \times  0.36 \times  0.64 = (30 \times  0.6 \times  0.8)^2 = 14.42 $$

---


---