---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 2 - Random Number Generation
* Exam on basis of ***Base R*** 

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

### Reproducible Simulations


When you execute any R code in this question, make sure you run the entire R script 
including the line ***set.seed(1234)*** at the start.

---

Use the command <tt>set.seed(1234)</tt> to initialise the random number generator. 


#### Exercises
1.  Generate a sample of size $n = 180$ from an Exponential distribution 
with parameter $\lambda  = 5$. 

2. Calculate the sample median of the sample in part 1.

3. Calculate the median of the Exponential distribution with $\lambda  = 5$. 

4. Comment on your results in parts 2 and 3. 

5.  The maximum likelihood estimator for $\lambda$ is given by $\lambda = 1/X$ where $X$ is the sample 
mean. Estimate the parameter $\lambda$  using the maximum likelihood estimator and 
the sample in part 1

6. Generate another sample of size $n = 180$ from an Exponential 
distribution with parameter $\lambda  = 5$ and estimate the parameter $\lambda$  using 
the maximum likelihood estimator and the new sample. 

7. Comment on the estimated values in parts 5 and 6


---

### Set Up 

```{r}
set.seed(1234) 

lambda <- 5

samplesize <- 180

```

---

#### Exercise 1

Generate a sample of size $n = 180$ from an Exponential distribution 
with parameter $\lambda  = 5$. 
  
```{r}
x = rexp(samplesize,lambda) 

```

```{r}
head(x)
```


#### Alternative Approach

* Using Inverse Transform Sampling

```{r}
x=-(1/5)*log(runif(samplesize))
```

---

#### Exercise 2

Calculate the sample median of the sample in part 1.

```{r}
m = median(x) 
m 
```

---

#### Exercise 3

Calculate the median of the Exponential distribution with $\lambda  = 5$. 

```{r}
mq = qexp(0.5,lambda)
mq 
```

Remark: This is the analytical result, i.e. the population median

---

#### Exercise 4

Comment on your results in parts 2 and 3.

* The sample median in part 2 is an estimator for the true median in part 3 based
on the sample in part 1 

* The results in parts 2 and 3 are not equal but similar 

---

#### Exercise 5

* The maximum likelihood estimator for $\lambda$ is given by $\lambda = 1/X$ where $X$ is the sample 
mean. 

* Estimate the parameter $\lambda$  using the maximum likelihood estimator and 
the sample in part 1



* For the parameter of an exponential distribution the MLE is the inverse of the
sample mean

```{r}
l = 1/mean(x) 
l 
```

Compare to $\lambda$

```{r}
lambda
```
---

#### Exercise 6

Generate another sample of size n = 180 from an Exponential distribution with parameter $\lambda$  = 5 and estimate the parameter $\lambda$ using the maximum likelihood estimator and the new sample. 

```{r}
x2 <- rexp(samplesize,lambda)

1/mean(x2) 
```

---

#### Exercise 7

Comment on the estimated values in parts 5 and 6

* The estimated values are both close to the true value ( $\lambda = 5$ ) but not identical. 

* The actual value of the estimator is a random variable and the estimated value depends on the actual sample


---