---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Q. 4) Consider the Dataset “fraud.csv”. There are two feature variables (1) State and (2) 
Sum_insured indicating the geography of the insured and the size of the sum insured for 
1000 different life insurance policy claims. 

You are asked to perform the following 
computations:

---

### Exercises

Part A.  Compute the number of fraudulent claims as a proportion of total number of claims. 
Part B. Compute the proportion of total claims associated with a medium sum insured policy 
from state C3, based on the data provided. (3)
Part C . Given the claim is a fraudulent claim compute the proportion of claims to be associated 
with a medium sum insured policy from state C3, based on the data provided. (4)

Part D Using the results in (Part A, (Part Band (Part Ccompute the probability of a fraudulent claim given 
that a new claim comes from state C3 and has a medium sum insured. 

---

Part E Compute the proportion of fraudulent claims from medium sum insured policies from 
C3 as a proportion of total claims from medium sum insured policies from C3. (3)
Part F Compare the results obtained in (Part D and (Part E with suitable explanation. (2)
Part GProvide two major limitations of computing the probability of a fraudulent claim using 
this method. (2)

---

Solution.4
```{r}
fraud <- read.csv("fraud.csv", stringsAsFactors=TRUE)
```

### Part A 
```{r}
x<-prop.table(table(fraud$Fraudulent))[2]
x
```

 Yes 
0.057 

---

Part B
y<-sum(fraud$State == "C3"& fraud$Sum_insured =="Medium")/nrow(fraud)
y
[1] 0.199 


---

Part C
```{r}
fraud_claims<-fraud[fraud$Fraudulent=="Yes",]

z<-sum(fraud_claims$State=="C3"&fraud_claims$Sum_insured =="Medium")/nrow(fraud_claims)
 z
```

 [1] 0.4035088 


---
Part D
prob<-x*z/y
prob
 Yes 
0.1155779 


---

Part E 
subset2<-fraud[fraud$Sum_insured=="Medium"&fraud$State == "C3",]
sum(subset2$Fraudulent == "Yes")/nrow(subset2)
[1] 0.1155779 
(3)
Part F
'The values need to be equal as this is an application of Bayes Theorem' 
(2)
vPart B
'(a) Assumes that all the variables are independent' 
'(b) If your test data set has a categorical variable of a category that wasn not present in the training 
data set, the Naive Bayes model will assign it zero probability and will not be able to make any 
predictions in this regard'

