---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Survival Analysis

Load the dataset “ovarian.csv” and save this as a data frame called ‘ovarian’. 
The columns in the dataset can be defined as:

* ***futime***: survival or censoring time
* ***fustat***: censoring status
* ***age***: in years
* ***resid.ds***: residual disease present (1=no,2=yes)
* ***rx***: treatment group
* ***ecog.ps***: ECOG performance status

Carry out the following analysis:

```{r message=FALSE,warning=FALSE}
library(tidyverse)
library(survival)
ovarian <- read_csv("ovarian.csv")
```
---

```{r}
ovarian %% filter(rx==1) %% arrange(futime)
```

---

```{r}
ovarian %% filter(rx==2) %% arrange(futime)
```

---

### Exercises

1. Compute the Cox regression using the column <ttrx</tt as the covariate. 
2. Summarise the results of the regression. 
3. Comment on the hazard ratio and statistical significance of the <ttrx</tt 
4. Compute the Kaplan-Meier estimators for the two groups of <ttrx</tt and save the results 
to an object called ‘KM’. 
5. Summarise the Kaplan-Meier estimator results. 
6. Plot the ‘KM’ results and set the colour for group 1 of <ttrx</tt to ‘red’ and group 2 to 
‘green’. Add a plot label and x and y axes labels to your plot. 
7. Comment on the KM analysis using the summary and plot results. 

---

### Part 1

```{r}

library(survival)

cph1r <- coxph(Surv(futime,fustat)~rx,
               data=ovarian)

```

---

### Part 2

```{r}
summary(cph1r)
```

---

### Part 3

* The p-value of rx is 0.31 which is greater than a 
significance level of 0.05 and hence is not a siginificant predictor

* The hazard ratio indicates that an rx value of 2 has a 0.5508 or ~50%
lower hazard rate than rx value of 1'

---

### Part 4

```{r}
KM <- survfit(Surv(futime,fustat)~rx,ovarian)
KM
```


---

### Part 5

```{r}
summary(KM)
```

---

### Part 6

```{r eval=FALSE}
plot(KM,col=c("red","green"),
        main="Kaplan-Meier Survival Curve",
        xlab="Time",ylab="Survival Probability") 
```

---

### Part 6

```{r echo=FALSE}
plot(KM,col=c("red","green"),
        main="Kaplan-Meier Survival Curve",
        xlab="Time",ylab="Survival Probability") 
```

---

### Part 7

* <ttrx 2</tt seems to perform better than <ttrx 1</tt initially upto time 353 post which 
there is a decline in the Survival Probability. 

* There is a steep decline in the numbers at risk between times 365 and 464 for rx at 2' 



---