---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 4

---

Q. 3) Set the simulation seed to 1234 prior to carrying out the analysis.

Part A Create a vector of values from 1 to 1000 and store in a variable ‘x’. Create another 
vector ‘y’ by using the rule: y = sin(x/40)+ε, where ε is a normal random variable with 
a mean of 0 and a standard deviation of 0.1. 

Part BPlot the x and y as a line and add grey colour to the line. Add gridlines to the plot. 

Part CCreate a vector to capture the ‘y’ values with a lag of the past 40 values (including the 
current value). Plot the lag as a red line added to the existing plot. 

Part D Create a vector to capture the ‘y’ values with a lag of the past 20 values, the current 
value and 20 future values. Plot the lag as a blue line added to the existing plot. 

---

Solution.3
Part A
```{r}
#Generate time series
set.seed(1234)
x <- 1:1000
y <- sin(x/40) + rnorm(1000,sd=.1) 

#(3 
#w (4)
```
---

iPart A
```{r}
# Plot the unsmoothed data (gray)
plot(x, y, type="l", col=grey(.5))
# add gridlines
grid() 
```

---

### Part C

Part CCreate a vector to capture the ‘y’ values with a lag of the past 40 values (including the 
current value). Plot the lag as a red line added to the existing plot. 


* Applies linear filtering to a univariate time series or to each series separately of a multivariate time series.
```{r}
# Smoothed with lag:
# average of current sample and 39 previous samples (red)
f40 <- rep(1/40, 40)
f40
```
---

```{r echo=FALSE}
# Plot the unsmoothed data (gray)
plot(x, y, type="l", col=grey(0.5))
# add gridlines
grid() 
```

---

```{r echo=FALSE}
# Plot the unsmoothed data (gray)
plot(x, y, type="l", col=grey(0.5))
y_lag <- stats::filter(y, filter= f40, sides=1)
lines(x, y_lag, col="red") 
# add gridlines
grid() 
```

---

```{r eval=FALSE}
y_lag <- stats::filter(y, filter= f40, sides=1)
lines(x, y_lag, col="red") 
```


```

---


w (3)
Part D



---

Part D Create a vector to capture the ‘y’ values with a lag of the past 20 values, the current 
value and 20 future values. Plot the lag as a blue line added to the existing plot. 

```{r}
# Smoothed symmetrically:
# average of current sample, 20 future samples, and 20 past samples (blue)
f41 <- rep(1/41,41)
f41
```


---


```{r echo=FALSE}
# Plot the unsmoothed data (gray)
plot(x, y, type="l", col=grey(0.5))
y_sym <- stats::filter(y, f41, sides=2)
lines(x, y_sym, col="blue") 
# add gridlines
grid() 
```

```{r eval=FALSE}
y_sym <- stats::filter(y, f41, sides=2)
lines(x, y_sym, col="blue") 

```

---
