---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


---

### Investigating the Proportion of Electric Cars in Two Populations

#### Description:

This exercise involves simulating data, estimating confidence intervals, and performing hypothesis testing to investigate the proportion of electric cars among registered cars in two large populations (A and B).


This exercise demonstrates how to simulate data, estimate confidence intervals, and perform hypothesis testing using R. 
---

A researcher wants to investigate the proportion of electric cars among all registered
cars in two large populations (denoted as populations A and B). 

* Two samples are
considered, one from each population. 

* The first sample consists of $n_A = 900$ registered
cars, while the second sample consists of $n_B = 1,200$ registered cars. 

* The type of each
car (electric or not) is assumed to be independent of the type of other cars in the
samples.

---

### Exercises 

**A.** Simulate the two samples in R assuming that the proportion of electric cars in
population A is 0.02, while in population B it is 0.025, and save the simulated
samples for later use.

<i>You must use the command <tt>set.seed(12345)</tt> to initialise the random
number generator, before you start the simulation.</i>

**B.** The true proportions of electric cars in the two populations are unknown to the
researcher, and the researcher wants to estimate them.

Determine an equal-tailed 99% confidence interval for the difference in the
true proportions of electric cars in the two populations A and B, using your
sample data. 

---

**C.** Based on previous knowledge, the researcher believes that the proportion of electric
cars in population A is lower than the proportion of electric cars in population B.

Perform a hypothesis test to investigate this belief using your sample data. 

In doing so, you should include the following steps, in addition to any other
necessary steps:
*  Compute the value of a suitable test statistic under an appropriate normal
approximation.
*  Compute the p-value of the test.
*  State your conclusion based on the calculated p-value.



---

### Exercise A.

**Simulate the Two Samples**

* Simulate two samples assuming the proportion of electric cars in population A is 0.02, while in population B it is 0.025. 

* Initialize the random number generator with `set.seed(12345)`.


```{r}
set.seed(12345)
data.A = rbinom(900, 1, 0.020)
data.B = rbinom(1200, 1, 0.025)
```

---

### First 150 outcomes for both groups

```{r}
data.A[1:150]
```


```{r fig.width=10}
data.B[1:150]
```

---



### Exercise B. 

**Determine a 99% Confidence Interval for the Difference in Proportions**

Using the sample data, determine an equal-tailed 99% confidence interval for the difference in the true proportions of electric cars in populations A and B.

```{r}
n.A = length(data.A)
n.B = length(data.B)

p.hat.A = sum(data.A) / n.A
p.hat.B = sum(data.B) / n.B
```

---

***Point Estimates***
```{r}
print(p.hat.A)
print(p.hat.B)
```


---

***Standard Error***

```{r}
ese.diff = sqrt(p.hat.A * (1 - p.hat.A) / n.A + p.hat.B * (1 - p.hat.B) / n.B)
```


***Quantiles***

```{r}
Zs = c(-1, 1) * qnorm(0.995)
Zs
```

***Confidence Intervals***

```{r}
ci.p.diff = (p.hat.A - p.hat.B) + (Zs * ese.diff)
ci.p.diff

```


The 99% CI for the difference in proportions is [-0.025631365, 0.009520254].

---


### Exercise C.

**Perform a Hypothesis Test**

Perform a hypothesis test to investigate whether the proportion of electric cars in population A is lower than in population B.

**Hypotheses:**
- $H_0: p_A = p_B$
- $H_1: p_A < p_B$

---

***Aggregate Proportion***
```{r}
p.hat.common = (sum(data.A) + sum(data.B)) / (n.A + n.B)
```

***Standard Error***
```{r}
ese.diff.H0 = sqrt(p.hat.common * (1 - p.hat.common) * (1 / n.A + 1 / n.B))
```

***Test Statistic***
```{r}
stat.diff = (p.hat.A - p.hat.B) / ese.diff.H0
stat.diff
```

---

***p-value***
```{r}
pvalue.diff = pnorm(stat.diff)
pvalue.diff
```


---

**Conclusion:**

* The value of the test statistic is -1.154163, and the p-value is 0.1242166. 

* Since the p-value is greater than the typical significance level (e.g., 0.05), we do not reject the null hypothesis. 

* Therefore, we do not have sufficient evidence to conclude that the proportion of electric cars in population A is lower than in population B.

---



---

Q2
(i)
Simulate samples:
```{r}
set.seed(12345)
data.A = rbinom(900,1,0.020)
data.B = rbinom(1200,1,0.025)
#data.A; data.B 
```

---

(ii)
```{r}
n.A = length(data.A); n.B = length(data.B)
p.hat.A = sum(data.A)/n.A
p.hat.B = sum(data.B)/n.B
p.hat.A; p.hat.B
# 0.02111111
# 0.02916667 

```

---

Parts (i)-(iii) were well answered in general. A common error in part (i) was not using properly the inverse transform method as the question asked.
Part (iv) was not well answered, with comments often being vague or not consistent with the graphs plotted earlier.

---

```{r}
ese.diff = sqrt( p.hat.A*(1-p.hat.A)/n.A + p.hat.B*(1-p.hat.B)/n.B )
ese.diff
# 0.006823359 
ci.p.diff = (p.hat.A-p.hat.B) + c(-1,1)*qnorm(0.995)*ese.diff
ci.p.diff
# -0.025631365 0.009520254 

```

The 99% CI for the difference in proportions is [-0.025631365, 0.009520254] 
---

### Part C

#### Hypotheses:
H0: p_A = p_B 
v. 
H1: p_A < p_B 

#### Statistic:

```{r}
p.hat.common = (sum(data.A)+sum(data.B))/(n.A+n.B)
p.hat.common
```


---

#### Standard Error

```{r}
ese.diff.H0 = sqrt( p.hat.common*(1-p.hat.common) * (1/n.A + 1/n.B))
ese.diff.H0
# 0.006979562 
```

---


```{r}
stat.diff = (p.hat.A-p.hat.B)/ese.diff.H0
stat.diff

```

---

The value of the statistic is -1.154163 
P-value:

```{r}
pvalue.diff = pnorm(stat.diff)
pvalue.diff
```


The p-value of the test is 0.1242166 

#### Conclusion:
We do not have evidence against the null hypothesis of equal proportions 
at level 12.4% 
so we do not reject it. 

---


---


