---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


---

### Simulation and Analysis of Exponential Distribution Using Inverse Transform Method


This exercise involves simulating samples from an exponential distribution, analyzing their means, and examining the application of the Central Limit Theorem (CLT).

---

### Exercises

**A.** Simulate 10,000 samples from an Exp(3) distribution, each of size n = 10,
using the inverse transform method. 

* You should save the generated values in
R for later use.

* You must use the command <tt>set.seed(3202)</tt> to initialise the random
number generator, before you start the simulation. 

**B.** Plot a histogram of the means of the samples generated in exercise A, using an
appropriate option in R for plotting the histogram on the probability density
scale. 

**C.** Plot the Probability Density Function (PDF) of the sampling distribution of the
sample mean, under the Central Limit Theorem (CLT), corresponding to the
samples generated in exercise A. 

The graph of the PDF should be superimposed on the histogram produced in exercise B. 

**D.** Comment on the sampling distribution of the sample mean and the application
of the CLT in this case based on your answers to exercise B and C. 

---

To simulate exponential random variables using the inverse transform method, you can use the following formula:

$$X = -\frac{1}{\lambda} \ln(1 - U)$$

Here's a brief explanation of the components:
- $X$ is the exponential random variable you want to simulate.
- $\lambda$ is the rate parameter of the exponential distribution.
- $U$ is a uniform random variable on the interval (0, 1).

By generating a uniform random variable $U$ and applying the formula, you can obtain a random variable $X$ that follows an exponential distribution with rate parameter $\lambda$.

---

### Exercise A.

**Simulate 10,000 samples from an Exp(3) distribution**

* Simulate 10,000 samples of size $n = 10$ from an exponential distribution with rate parameter $\lambda = 3$ using the inverse transform method. 

* Initialize the random number generator with `set.seed(3202)` before the simulation.


```{r}
set.seed(3202)
n = 10
lambda = 3

# Initialize matrix, each sample in a row
m1 = matrix(0, nrow = 10^4, ncol = n)  
for (i in 1:10^4) {
  u = runif(n, 0, 1)
  m = -log(1 - u) / lambda  # Generate samples in rows
  m1[i, ] = m
}
```

---


```{r}
head(m1) %>% round(3)
```

```{r}
tail(m1) %>% round(3)
```

---

### Exercise B. 

**Plot a histogram of the means of the samples**

Plot a histogram of the means of the samples generated in exercise A, using the probability density scale.


```{r eval=FALSE}

# compute mean of each row
xbar.a = apply(m1, 1, mean)

hist(xbar.a, 
     freq = FALSE, 
     main = "Sampling distribution of Xbar")
```

---

### Exercise B. 

```{r echo=FALSE, fig.width=11}
xbar.a = apply(m1, 1, mean)
hist(xbar.a, freq = FALSE, main = "Sampling distribution of Xbar")
```

---

### The Exponential Distribution - Key Characteristics.

1. **Mean (μ)**: This is the average value of the distribution, and for an exponential random variable, it's given by $\mu = \frac{1}{\lambda}$, where $\lambda$ is the rate parameter.

2. **Standard Deviation (σ)**: This measures the spread of the distribution, and for an exponential random variable, it's the same as the mean, so $\sigma = \frac{1}{\lambda}$.

3. **Standard Error (SE)**: This measures the accuracy of the sample mean as an estimate of the population mean. For an exponential distribution, it is defined as the standard deviation of the sampling distribution of the sample mean. 

For a sample size $n$, the standard error is $SE = \frac{\sigma}{\sqrt{n}} = \frac{1}{\lambda\sqrt{n}}$.


---

### Exercise C. 

**Plot the PDF of the sampling distribution under the CLT**

Using the CLT, plot the Probability Density Function (PDF) of the sampling distribution of the sample mean. 
Superimpose this PDF on the histogram from part B..

```{r eval=FALSE}
hist(xbar.a, freq = FALSE, 
	main = "Sampling distribution of Xbar")

x.grid = seq(0, 1, length = 1000)
lines(x.grid, 
      dnorm(x.grid, 1/3, sqrt(1/90)),
      col="red")
```



---

### Exercise C. 

```{r echo=FALSE, fig.width=11}
hist(xbar.a, freq = FALSE, main = "Sampling distribution of Xbar")
x.grid = seq(0, 1, length = 1000)
lines(x.grid, dnorm(x.grid, mean = 1/3, sd= sqrt(1/90)),col="red")
```

---

### Exercise D.

**Comment on the sampling distribution and the CLT application**

Based on the histogram and the superimposed PDF, comment on the sampling distribution of the sample mean and the application of the CLT in this case.

**Comments:**
- The sampling distribution of the sample mean is relatively asymmetrical compared to the normal distribution.
- The Central Limit Theorem does not work very well here as the sample size is small ($n = 10$).

---

This exercise illustrates the process of simulating data from an exponential distribution, analyzing sample means, and understanding the limitations of the CLT with small sample sizes.

---

---
---
Q1
(i)
R code:

```{r}
set.seed(3202)
n = 10; lambda = 3
m1 = matrix(0,nrow=10^4,ncol=n) # Initialise matrix, each sample in a row
for (i in 1:10^4){ 
u = runif(n,0,1) 
m= - log(1-u)/lambda # Generate samples in rows 
m1[i,] = m 
}

```

---

(ii)
R Code:

```{r}
xbar.a = apply(m1,1,mean) 
hist(xbar.a,freq=F,main="Sampling distribution of Xbar")
```

---
(iii)
Using the CLT, the sampling distribution should be N(1/3, 1/90). 
R code:

```{r}
hist(xbar.a,freq=F,main="Sampling distribution of Xbar")
x.grid = seq(0,1,length=1000) 
lines(x.grid,dnorm(x.grid,1/3,sqrt(1/90))) 

```

---

The curve shows the exact pdf of the sampling distribution. 
(iv)


---

* The sampling distribution is relatively asymmetrical compared to the normal. 
* The CLT does not work very well here 
as the sample size is small (n = 10).

---

---

---
