---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 4

---

2 An insurance company designed a new product and wanted to assess its clients’ 
responses to the product. A survey was carried out giving an opportunity to each 
participating client to give a positive or negative response to the product, 
independently of other clients. Let X be the random variable representing the positive 
responses to the new product. 

---

### Exercises

(i) Identify the distribution of X and its parameters.  
Out of 160 clients who responded independently to the survey, 101 gave a positive 
response for the new product. 

---

Q2
(i)
𝑋𝑋 follows a Binomial (𝑛𝑛, 𝜃𝜃) distribution, where 𝑛𝑛 is the number of clients surveyed 
and 𝜃𝜃 is the probability of getting a positive approval of the product from a client 
---


The probability of obtaining a positive response for the product is denoted by θ and a 
Beta prior distribution with parameters ሺα, βሻ is assumed for θ. The posterior 
distribution of θ is proportional to: 
fሺθ|xሻ ∝ θx + α –1ሺ1 – θሻn – x + β – 1, 
where x is the number of positive responses obtained out of n clients surveyed. 
(ii) Specify the posterior distribution of θ with its parameters. 
(iii) Comment on the prior distribution of θ in relation to the posterior distribution. 
  
---

(ii)
The posterior distribution of 𝜃𝜃 is Beta with parameters
𝛼𝛼 + 𝑥𝑥 = 𝛼𝛼 + 101 and 𝑛𝑛 − 𝑥𝑥 + 𝛽𝛽 = 59 + 𝛽𝛽 . 
(iii)
The prior and posterior are of the same family. So, we have a conjugate prior 
  
---
  
(iv) State the parameter values for which the prior is a Uniform(0, 1) distribution. 
  
(v) (a) Plot the prior density of θ with the parameters obtained in part (iv). Set 
the maximum limit of the y axis to 12. 
(b) Plot the posterior distribution of θ on the same graph as above. 
[Hint: you may find the lines function useful.] 

---


An Analyst consulted by the company suggests that based on previous experience, a 
Beta prior with parameters ሺ40, 24ሻ is more appropriate. 
(vi) Plot the new prior and posterior distributions of θ on the same graph from 
part (v). 

---

(vii) Comment on the plots obtained in parts (v) and (vi). 
The company will put the new product on the market only if there is a high 
probability that θ is higher than 60%. 
(viii) (a) Calculate the probability Pሺθ > 0.6 | Xሻ in the case of both priors; that 
is, Uniform(0, 1) and Beta with parameters (40, 24). [4] 
(b) Comment on your answer to part (viii)(a). [2] 



---

(iv)
Beta(1, 1) is the same as a Uniform(0,1) distribution: 𝛼𝛼 = 1, 𝛽𝛽 = 1 
(v)(a)
```{r}
n = 160
x = 101
alpha = 1
beta = 1
p = seq(0,1, by=0.01) 
plot(p, dbeta(p, alpha, beta), ylab="density", type ="l",
ylim = c(0,12), main = "Plot for the prior density of θ") 
```
---
(b) 

```{r}
plot(p, dbeta(p, alpha, beta), ylab="density", type ="l",
ylim = c(0,12), main = "Plot for the prior density of θ") 
lines(p, dbeta(p, x+alpha, n-x+beta), type ="l", col="red") 
legend(0.,12, c("Prior U(0,1)", 
"Posterior"),lty=c(1,1),col=c("black", "red")) 
```

---

### Part (6)

```{r eval = FALSE}
alpha1 = 40
beta1 = 24
p = seq(0,1, by=0.01)
plot(p, dbeta(p, alpha, beta), ylab="density", type ="l",
ylim = c(0,12), main = "Plot for the prior density of θ") 

lines(p, dbeta(p, alpha1, beta1), col="blue") 
lines(p, dbeta(p, x+alpha1, n-x+beta1), type ="l", 
col="brown") 
legend(0.,12, c("U(0,1)", "Beta(102,60)", "Beta(40, 24)", 
"Beta(141,83)"),lty=c(1,1),col=c("black", "red", "blue", 
"brown"))
```

---


```{echo=FALSE}
alpha1 = 40
beta1 = 24
p = seq(0,1, by=0.01)
plot(p, dbeta(p, alpha, beta), ylab="density", type ="l",
ylim = c(0,12), main = "Plot for the prior density of θ") 

lines(p, dbeta(p, alpha1, beta1), col="blue") 
lines(p, dbeta(p, x+alpha1, n-x+beta1), type ="l", 
col="brown") 
legend(0.,12, c("U(0,1)", "Beta(102,60)", "Beta(40, 24)", 
"Beta(141,83)"),lty=c(1,1),col=c("black", "red", "blue", 
"brown"))
```

---

### Part (7)

(vii)
The Beta(40, 24) prior contains more information than the uniform (0, 1) prior 
and this is reflected in the posterior densities obtained as the posterior in (vi) is narrower 
than the one in (v) 

---

(viii)(a) 
```{r}
#Using the Uniform(0, 1) prior:
1 - pbeta(3/5, x+alpha, n-x+beta) 
0.7840908 
```
---


Using the Beta(40, 24) prior:
```{r}
1 - pbeta(3/5, x+alpha1, n-x+beta1) 
```

#0.8199189 

---
(b) 

* The Beta(40, 24) prior provides a slightly higher posterior probability than the proportion of 
clients positively perceiving the new product 
* As the difference between the two priors is considerable and the difference between the 
two posteriors is small, we can conclude that most information is coming from the data rather 
than the priors  

---

BLANK
---
This question was not well answered, with many candidates not attempting parts of it. 
In parts requiring comments, credit was awarded for valid comments that are not shown 
here. 
In parts where a legend is shown on the graph, credit was given when candidates 
provided a relevant description or comment in the answer instead of the legend. 
In part (iv), a common error was to give (0,1) as the parameters of the distribution. 
In part (viii)(a), a common error was using the prior instead of the posterior distribution.
