---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Q. 3)
(7)
[27]
(4)
[22]
The number of claims in a year on an individual policy follows a Poisson distribution with
a parameter of 0.75. Individual claim amounts follow a Gamma distribution with parameters
(shape parameter α = 2 and rate parameter \lambda = 20,000). Set seed value of 100.


---

i)
Simulate claims for 10,000 policies and create a frequency distribution of the number
of claims made on a policy per year.
(4)

ii) Simulate the claim amounts for each claim and compute the aggregate claim amounts
for each policy. Create a histogram of the aggregated claim amounts with proper
labelling and legends.
iii) Compute the mean and variance of the aggregated claim amounts by using the method
of moments of the compound Poisson distribution. 


---

Part (b)
An insurer plans to go for an aggregate excess of loss reinsurance policy. The insurer wants
to evaluate the expected cost for the retention limits from 50,000 to 100,000 in steps of
5,000.

iv) Compute expected costs for the insurer and the reinsurer for each of the retention limits. 
v) The insurer decides to set a maximum retention limit where the mean cost to the insurer
is not more than 75% of the mean aggregated claims. Identify the maximum possible
retention limit based on the computations of (iv) and the proportion of total claims to
be handled by the reinsurer. 
vi) Based on that limit, compute the standard deviation and coefficient of skewness of the
costs to the reinsurer.

---


Solution 3:
i)

```{r}
set.seed(100)
freq<-rpois(10000,0.75)
table(freq)
```


---

# There was a typo in the question. It should have been rate parameter is
1/20000 or scale parameter should have been 20000
# In case the students follow either one of the approaches, full marks will be
awarded.
#Solution Assuming rate parameter = 20000

---

## Part (b)

```{r}
AggclaimAmount<-c()

for (i in 1:10000) {
  claimAmount<-sum(rgamma(freq[i],shape = 2, rate = 20000))
  AggclaimAmount<-c(AggclaimAmount,claimAmount)
}
```

---

```{r}
hist(AggclaimAmount, breaks =30, 
  main = "Histogram of Aggregate Claim Amount", 
  xlab = "AggregateClaim Amount", 
  ylab = "Frequencies")

```

---

## Part (c)

```{r}
mean_poisson<-0.75
mean_gamma<-2/(20000)
var_poisson<-0.75
var_gamma<-2/((20000)^2)
mean_aggregate<-mean_poisson*mean_gamma
```

mean_aggregate
 7.5e-05

```{r}
var_aggregate<-mean_poisson*var_gamma+var_poisson*mean_gamma^2
var_aggregate
```
 1.125e-08

---


## Part (d)

```{r}
mean_claims_I<-c()
mean_claims_R<-c()
for (i in seq(50000,100000,5000)) {
mean_claims_R<-c(mean_claims_R,mean(pmax(AggclaimAmount-i,0)))
mean_claims_I<-c(mean_claims_I,mean(pmin(AggclaimAmount,i)))
}
```

---
# Mean Costs to the Insurers

```{r}
mean_claims_I
```

 7.395992e-05 7.395992e-05 7.395992e-05 7.395992e-05 7.395992e-05 7.395992e-
05
[7] 7.395992e-05 7.395992e-05 7.395992e-05 7.395992e-05 7.395992e-05
# Mean Costs to the Reinsurer
mean_claims_R
 0 0 0 0 0 0 0 0 0 0 0


---

v)
mean_agg_cost<-mean(AggclaimAmount)
# 75% of the Aggregate claims cost
mean_Cost_Insurer<-mean_agg_cost*0.75
# 75% of the Aggregate claims cost = 5.546994e-05
# Retention limits should be much lesser than the limits specified in part (iv)
# Can be recalculated by considering a different range from 0.0001 to 0.0002
# If the student does not compute this range but mentions that no values from the range are applicable,
then full marks should be awarded.

---

```{r}
mean_claims_I<-c()
mean_claims_R<-c()
for (i in seq(0.0001,0.0002,0.00001)) {
mean_claims_R<-c(mean_claims_R,mean(pmax(AggclaimAmount-i,0)))
mean_claims_I<-c(mean_claims_I,mean(pmin(AggclaimAmount,i)))
}
```

---

```{r}
mean_claims_I
```
 4.201910e-05 4.482882e-05 4.740955e-05 4.979510e-05 5.198747e-05 5.400000e-
05
[7] 5.584620e-05 5.751952e-05 5.904900e-05 6.044518e-05 6.170202e-05
mean_claims_R
 3.194081e-05 2.913110e-05 2.655037e-05 2.416482e-05 2.197244e-05 1.995992e-
05
[7] 1.811371e-05 1.644040e-05 1.491091e-05 1.351474e-05 1.225790e-05


---

```{r}
Retention_Limit<-0.00015

Reinsurer_Claims<-pmax(AggclaimAmount-Retention_Limit,0)
#Proportion of Claims to be taken up by the reinsurer
sum(Reinsurer_Claims>0)/10000
```

---


## Part (4)
vi)
If for the part (v), the student identifies that no values from the range are applicable and stops here, 

full marks should be awarded

```{r}
SD_Retention<-sd(Reinsurer_Claims)
SD_Retention

```

> SD_Retention
 5.991976e-05

---
```{r}
Skew_Retention<-mean((Reinsurer_Claims-mean(Reinsurer_Claims))^3)/(sd(Reinsurer_Claims))^3
Skew_Retention
```
