---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Mortality Investigation

* Use the data file “Mortality_Investigation.csv” to answer the below questions. 

* This table
contains the details of mortality investigation performed on 100 different male lives
between 1-Jan-2016 and 31-Dec-2018.

---

### Part 1

Print first 6 rows of the data ensuring correct data formats for all the columns and
calculate the proportion of the total sample died, survived and withdrew from the
observation period.


```{r}
Mort_Inv <- read.csv("Mortality_Investigation.csv")

Mort_Inv$DoB<-as.Date(Mort_Inv$DoB)
Mort_Inv$DoJ<-as.Date(Mort_Inv$DoJ)
Mort_Inv$DoE<-as.Date(Mort_Inv$DoE)

```

---


```{r}
head(Mort_Inv)
```
```{r}
prop.table(table(Mort_Inv$Exit_Reason)) 

```




---

### Part 2

Compute two new columns ‘Age_At_Entry’ and ‘Age_At_Exit’ for each person (you
can assume a completed year as 365.25 days and you can compute the age as difference
between the dates in decimals like 35.234 years) and print the last 6 rows of the newly
formed table.

---
### Part 2

```{r}
Mort_Inv$Age_At_Entry<-round((Mort_Inv$DoJ-Mort_Inv$DoB)/365.25,4)
Mort_Inv$Age_At_Exit<-round((Mort_Inv$DoE-Mort_Inv$DoB)/365.25,4)
tail(Mort_Inv)
```


---

### Part 3

Compute the average age at entry and the average age at exit of the people for whom
“Death” was the reason of Exit. 


```{r}
mean(Mort_Inv$Age_At_Entry[Mort_Inv$Exit_Reason == "Death"])
mean(Mort_Inv$Age_At_Exit[Mort_Inv$Exit_Reason == "Death"])
```



---


### Part 4

If the age label used is “Age last Birthday”, then calculate the number of lives out of total 100 lives who have contributed the full year towards the
central exposed to risk $E^{C}_{37}$



```{r}
sum((Mort_Inv$Age_At_Entry)<37 & Mort_Inv$Age_At_Exit>38 )
```

---


### Part 5

If the age label used is “Age last Birthday”, then calculate
the number of lives who did not contribute at all towards the central exposed to risk $E^{C}_{37}$




```{r}
sum((Mort_Inv$Age_At_Entry)>38|Mort_Inv$Age_At_Exit<37)
```



---

### Part 6

If the age label used is “Age last Birthday”, then calculate the total number of years of contribution to $E^{C}_{37}$
by all the lives together (Round your
answer to 2 digits after the decimal).

```{r}
Mort_Inv$Contribution37 <- ifelse(
  (Mort_Inv$Age_At_Exit<37 |Mort_Inv$Age_At_Entry> 38),
  "No","Yes" )

Mort_Inv$contribution37_Period <- ifelse(
  Mort_Inv$Contribution37 == "Yes",
  (pmin(38,Mort_Inv$Age_At_Exit)- pmax(37,Mort_Inv$Age_At_Entry)),0)

sum(Mort_Inv$contribution37_Period)
```

---
