---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Q. 4)
(4)
[30]
Consider the data set ‘Covid_2019.csv’ (name it covid19) where the first row of the csv file
contains the headings for the columns.


Exercises
===================================================================================================
i)
Print the number of missing values in each of the columns and create a new data set
‘covid19_1’ by removing all the missing values.
(4)
From covid19_1, use the columns from Population Density (8th column) to Life Expectancy
(17th column) to answer the following questions.
ii) Create a new data frame “Covid_Cluster” containing only the above mentioned
columns. Normalize all the columns of the data frame using the scale function. (3)
iii) Classify the countries into five groups by using the values obtained from (ii) applying
K-Means clustering algorithm. It is mandatory to set a seed value of 100 before
executing the algorithm. Print the number of countries in each cluster. 

Exercises
===================================================================================================
iv) What proportion of total countries in each cluster are severe with respect to COVID-
19? You can use the “Severe” column from the original dataset. (5)
v) Print the total number of cases and total number of deaths for each cluster.

http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS2B_COVID_2019.csv
CS2B_Mortality_Investigation
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS2B_Mortality_

===================================================================================================
Solution 4:
i)

```{r}
covid19 <- read.csv("Covid_2019.csv")
missingvalues<-sapply(covid19,FUN = function(x)sum(is.na(x)))
missingvalues

```

> missingvalues
Continent Country
total_case
0 0 total_deaths total_cases_per_million 0 0 population population_density median_ag
0 11 2
aged_65_older gdp_per_capita cardiovasc_death_rat
27 27 2
diabetes_prevalence female_smokers male_smoker
17
1
hospital_beds_per_thousand
e
45
0 69 7
life_expectancy Sever
s
0
total_deaths_per_millio
n
0
e
4
e
4
s

```{r}
covid19_1<-na.omit(covid19)
```

---

Part 2
====================================================================================
ii)

```{r}
Covid_Cluster<-covid19_1[,c("population_density","median_age","aged_65_older","gdp_per_capita",
"cardiovasc_death_rate","diabetes_prevalence","female_smokers",
"male_smokers","hospital_beds_per_thousand","life_expectancy")]
Covid_Cluster<-scale(Covid_Cluster)
```

---

Part 3
====================================================================================
iii)
Page 10 of 11IAI

```{r}
set.seed(100)
cluster1<-kmeans(Covid_Cluster,centers = 5)
cluster1$size

```

---

Part 4
====================================================================================


```{r}
covid19_1$cluster<-cluster1$cluster
```

```{r}
table(covid19_1$cluster,covid19_1$Severe)

```

---

Part 4
====================================================================================


```{r}
prop.table(table(covid19_1$cluster,covid19_1$Severe),margin = 1)
```
> table(covid19_1$cluster,covid19_1$Severe)
No Yes
1 16
5
2 27
1
3 10 16
4 15
6
5 15 15
> prop.table(table(covid19_1$cluster,covid19_1$Severe),margin = 1)
1
2
3
4
5
No
0.76190476
0.96428571
0.38461538
0.71428571
0.50000000
Yes
0.23809524
0.03571429
0.61538462
0.28571429
0.50000000



---


Part 5
====================================================================================
v)

```{r}
aggregate(total_cases~cluster,data = covid19_1, FUN = "sum")
aggregate(total_deaths~cluster,data = covid19_1, FUN = "sum")

```
> aggregate(total_cases~cluster,data = covid19_1, FUN = "sum")
cluster total_cases
1
1
1467904
2
2
658083
3

```{r}
aggregate(total_deaths~cluster,data = covid19_1, FUN = "sum")

```