---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Exercise 1
===============================
i)
Simulate a time series of length 200 observations in a line chart using the following R
function and comment on stationarity. (Set a seed value of 100)

```{r}
set.seed(100)
observations <- arima.sim(list(order = c(1,1,1), 
          ar = 0.7, 
          ma = 0.3), 
          n = 200)
```

---


```{r}

observations
```

---

#### Exercise 2:

```{r}

plot(observations, 
     main = "Line chart of the time series observations")


```


---

#### Stationarity

Stationarity means that the statistical properties of a process that creates a time series are constant over time. 

This statistical consistency makes predictable distributions and is an assumption of many time series forecasting models.



* The data is not stationary as we observe that the values are changing with time

* Mean and Standard Deviation are different at different points in time

---

#### Exercise 2

Plot the auto correlation function and partial auto correlation function of the data if the data
is stationarity, otherwise take the first difference of the observations and plot the above
graphs. Comment on the possible ARMA model based on the graphs. 


---

As the data is not stationary, we take the first difference of the observations

```{r}
diff_obs<-diff(observations)
acf(diff_obs)
```

---


```{r}

pacf(diff_obs)


```

As both ACF and PACF are seen to have spikes only for the first two lags and they appear to tail off after that,
ARMA(2,2) model appears to be the most appropriate model based on the ACF and PACF plots.


---

#### Exercise 3

Fit AR(1), AR(2), MA(1) and ARMA(1,1) models to the observations from (ii) and
comment on the best fitting model using AIC or any other metric.

---

Exercise 3
```{r}
AR1model <- arima(diff_obs,order = c(1,0,0))
AR1model
```


---

```{r}

AR2model <- arima(diff_obs,order = c(2,0,0)) 
AR2model
```


---

```{r}

MA1model <- arima(diff_obs,order = c(0,0,1))
MA1model
```

---
```{r}
ARMA1.1model <- arima(diff_obs,order = c(1,0,1))
ARMA1.1model
```
---

### Akaike Information Criterion

```{r}
AIC(AR1model)
```

```{r}
AIC(AR2model)
```

```{r}
AIC(MA1model)
```

```{r}
AIC(ARMA1.1model)
```





AIC is appearing the least for AR(2) model. 
The same is being by PACF graph also.

---

#### Exercise 4

Using the best fitted model from exercise 3 create a forecast of the observations for the next
3 periods.




```{r}
AR2model <- arima(diff_obs,order = c(2,0,0))

predict(AR2model,n.ahead = 3)
```
---

---