---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Before answering this question, the R package for Markov chains should be loaded 
into R using the following code: 

<pre><code>
install.packages("markovchain") 
library(markovchain)
</code></pre>

---

```{r echo = FALSE}
library(markovchain)
```

$X_t$ is a Markov chain on the state space {1,2,3} with the following transition matrix: 

$$
\left(\begin{array}{ccc}
p &  1 – p & 0\\
q & 0 & 1 – q\\
0 & r & 1 – r\\
\end{array}\right)
$$
where $0 < p, q, r < 1$. 

---

### Exercises

1. Construct an R function, with arguments p, q and r, that generates a Markov 
chain object for Xt. 

2.


---

#### Exercise 1


```{r}

p <- 0.4
q <- 0.2
r <- 0.25

P <- matrix(data = c(p,1-p,0,q,0,1-q,0,r,1-r), nrow =3, byrow = TRUE) 
P
```

---


```{r}
markovchain = function(p,q,r){ 
 P = matrix(data = c(p,1-p,0,q,0,1-q,0,r,1-r), nrow =3, byrow = TRUE) 
 mc = new("markovchain", transitionMatrix = P) 
 mc
} 
 
```

---

```{r}
test_mc <- markovchain(0.5,0.3,0.2)
```
```{r}
steadyStates(test_mc)
```

---

#### Exercise 2

Assume that p = 0.75 and r = 0.25. 
Construct R code that calculates the stationary distribution of Xt for values of 
q from 0.1 to 0.9 inclusive, at intervals of 0.1.



```{r}
statdist = matrix(data = 0, nrow = 3, ncol = 9) 
 

for(i in 1:9){ 
  q = 0.1*i 
  mc = markovchain(p = 0.75, q, r = 0.25) 
  statdist[,i] = steadyStates(mc)} 

statdist 
 
```


--- 

 Assume that p = 0.75 and r = 0.25. 
(ii) Construct R code that calculates the stationary distribution of Xt for values of 
q from 0.1 to 0.9 inclusive, at intervals of 0.1, pasting your results into your 
answer script. 

---


(iii) Plot a graph showing the stationary distributions of Xt calculated in part (ii) as 
a function of q. You should include three separately coloured lines on your 
graph, each line representing the stationary probabilities of each state. 

---

(iii)

```{r}
q = seq(from = 0.1, to = 0.9, by = 0.1) 
 

plot( 
 

q, 
 

statdist[1,], 
 

type = "l", 
 

main = "Stationary Distributions of X_t as a
function of q", 
 

ylab = "Probability", 
 

col = "red") 
 

lines(q, statdist[2,], col = "black") 
 

lines(q, statdist[3,], col = "blue") 
 

legend("top", 
legend = c("State 1", "State 2", "State 3"), 
 

col = c("red", "black", "blue"), 
 

pch=7)

```

```{r}


```

 

[Total 20]
Part (i) was very well answered although some candidates lost marks for not creating 
a function, or for creating a function outputting a matrix rather than a Markov chain 
object. Candidates who did not create a function or who created a function outputting 
a matrix rather than a Markov chain object needed to perform additional work in part 
(ii). 
Part (ii) was well answered. Candidates who were less familiar with loops in R 
included more extensive R code than necessary. Candidates who answered part (ii) by 
raising the transition matrices to a large power were required to raise the matrices to 
a second large power to demonstrate convergence otherwise only partial credit was 
awarded.
Part (iii) was poorly answered. Many candidates lost marks for missing or 
inappropriate titles, axis labels and/or legends. Where candidates were unfamiliar 
with the legend function in R, a manually applied legend was accepted provided that it 
correctly represented the line colours and was situated next to the chart.


