---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


Part 6

(vi) Estimate a Cox proportional hazards model with death as the event of interest 
using the two covariates, drug and gender, with no interaction term, pasting 
your results into your answer script. You should use the Breslow method for 
tie handling. 

(vii) Comment on the results produced in part (vi) with reference to the effects of 
the two covariates, drug and gender, on the mortality rate. [5]
(viii) Update the Cox proportional hazards model in part (vi) to include an 
interaction term between drug and gender, pasting your results into your 
answer script. 
(ix) Analyse the effectiveness of MediCo, commenting on any differences between 
males and females. 

(vi)


```{r}
coxph( 
 

Surv(mortalitydata$Time, mortalitydata$Status) ~ Drug + 
Gender, 
data = mortalitydata, 
 

ties = "breslow") 
 
```

---
Call:
coxph(formula = Surv(mortalitydata$Time, mortalitydata$Status) 
~ 
 Drug + Gender, data = mortalitydata, ties = "breslow")
Coef exp(coef) se(coef) z p
Drug -0.35174 0.70347 0.03439 -10.23 <2e-16
Gender -0.93263 0.39352 0.03575 -26.09 <2e-16
Likelihood ratio test=820.3 on 2 df, p=< 2.2e-16
n= 4400, number of events= 3441 
 
---
OR:

```{r}
coxph( 
 

Surv(mortalitydata$Time, mortalitydata$Status) ~ 
mortalitydata$Drug + mortalitydata$Gender, [2½]
ties = "breslow") 
```

---

Call:
coxph(formula = Surv(mortalitydata$Time, mortalitydata$Status) ~ 
 mortalitydata$Drug + mortalitydata$Gender, ties = "breslow")
 coef exp(coef) se(coef) z p
mortalitydata$Drug -0.35174 0.70347 0.03439 -10.23 <2e-16
mortalitydata$Gender -0.93263 0.39352 0.03575 -26.09 <2e-16
Likelihood ratio test=820.3 on 2 df, p=< 2.2e-16
n= 4400, number of events= 3441 
 
---


#### Exercise 7

* The results suggest that MediCo reduces the mortality rate of patients by around 30% 
 
* The results suggest that female lives’ mortality rates are around 40% of that of males 
 

* However, the graphs from part (iv) suggest that an interaction term may be required in the 
Cox model to fully analyse the effects of MediCo and Gender on the mortality rate, and 
hence the above analysis is potentially misleading 
 

* The results suggest that the Gender covariate appears to have a greater effect on the mortality 
rate than the Drug covariate 
 

* Both p-values indicate that the coefficients are unlikely to be 0 
 
---


#### Exercise 8


```{r}
coxph( 
 

Surv(mortalitydata$Time, mortalitydata$Status) ~ Drug*Gender, 

 

data = mortalitydata, 
 

ties = "breslow") 
 
```

---

Call:
coxph(formula = Surv(mortalitydata$Time, mortalitydata$Status) ~ 
 Drug * Gender, data = mortalitydata, ties = "breslow")
Coef exp(coef) se(coef) z p
Drug -0.07143 0.93106 0.04515 -1.582 0.114
Gender -0.63564 0.52960 0.04711 -13.494 <2e-16
Drug:Gender -0.65874 0.51750 0.06977 -9.442 <2e-16
Likelihood ratio test=910.2 on 3 df, p=< 2.2e-16
n= 4400, number of events= 3441 
 

---

OR:
coxph( 
 

Surv(mortalitydata$Time, mortalitydata$Status) ~ 
mortalitydata$Drug * mortalitydata$Gender, [1½]
ties = "breslow") 
 

Call:
coxph(formula = Surv(mortalitydata$Time, mortalitydata$Status) ~ 
 mortalitydata$Drug * mortalitydata$Gender, ties = "breslow")
 coef exp(coef) se(coef) z p
mortalitydata$Drug -0.07143 0.93106 0.04515 -1.582 0.114
mortalitydata$Gender -0.63564 0.52960 0.04711 -13.494 <2e-16
mortalitydata$Drug:mortalitydata$Gender -0.65874 0.51750 0.06977 -9.442 <2e-16
Likelihood ratio test=910.2 on 3 df, p=< 2.2e-16
n= 4400, number of events= 3441 
 
---


(ix)
MediCo reduces male mortality rates by 1 – 0.93106 = 6.9% 
 

which is not statistically significant 
 

and reduces female mortality rates by 1 – 0.93106 * 0.51750 = 51.8% 
 

which is statistically significant 
 

The results from the Cox analysis are consistent with the Kaplan–Meier plots in part (iv)

---

The likelihood ratio test statistic for the model with the interaction term compared with the 
model without the interaction term is: 
910.2 – 820.3 = 89.9 
OR 
L1 = m1$loglik
 

L2 = m2$loglik
 

-2 *(L1 - L2)

 
 89.8897
(where m1 = Cox model fitted in part (vi) and m2 = Cox model fitted in part (viii))
which is highly significant under the chi-squared distribution with one degree of freedom.

 

---

---