---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Page 2 of 5
Q. 1)
The data set “Monthly_Corn.csv” contains monthly average closing prices and the volume of corn traded on the Chicago Mercantile Exchange from January 2017 to March 2023. Load the data into R.
i) Create a time series of the closing price by using an appropriate R function.
(1)
ii) Plot this series, labelling each axis appropriately.

iii) Based on visual inspection of the series, comment on whether the series is stationary or not.
(1)
iv) Plot the Auto-Correlation Function (ACF) and Partial Auto-Correlation Function (PACF) of the closing prices, labelling each axis appropriately.

v) Comment on the stationarity of the closing price by observing the plots in part (iv).

vi) Create another time series “monthly_returns” by using the formula log(Pt/Pt-1), where Pt and Pt-1 correspond to the closing prices of month t and t-1 respectively.

vii) Plot the ACF and PACF of the monthly returns series, labelling each axis appropriately.

viii) Comment on the stationarity of the monthly returns by observing the plots in part (vii).

ix) Based on the ACF and PACF of the returns series obtained in part (vii), identify the most appropriate ARMA model to be fitted to the series.
(1)
x) Fit the following four ARMA models to the returns series:
a) ARMA(0,0),
b) ARMA (0,1),
c) ARMA (1,0) and
d) ARMA (1,1)
and identify the model with the least Akaike Information Criteria (AIC) value.

xi) Compare the result in part (ix) and part (x) and give a suitable explanation for any deviation observed.

[24]

---


Solution 1:

```{r}
Corn_Prices<-read.csv("Monthly_Corn.csv")
```

```{r}

Close<-ts(Corn_Prices$Close,start = c(2017,1),frequency=12)
```

```{r}
plot(Close, ylab = "Corn Monthly prices", 
	main = "Monthly plot of corn prices")

```


---

### Part C

The series is not stationary as it appears to be trending upwards with 
time. The mean at different periods appears not to be constant For periods be
tween 2017 to 2020, it appeared to be stationary.

---

### Part D


```{r eval=FALSE}
acf(Close, main = "ACF", xlab = "Lags", 
    ylab = "Serial Correlation", lag.max=60)

```

---

```{r echo=FALSE}
pacf(Close, main = "PACF", xlab = "Lags", 
     ylab = "Serial Correlation", lag.max=60)
```


---

(v) 

* If a series is stationary, the ACF should decay to zero quickly and sho
uld not display any oscillation.
* The series is not stationary because auto cor
relation function is not decaying to zero quickly. 
* Also, if the number of lag
s is increased, oscillation is also observed, hence cannot be stationary

---

(vi) Create another time series "monthly_returns" by using the formula log(P
t/Pt-1), where Pt and Pt-1 correspond to the closing prices of month t and t1 respectively.
```{r}
returns<-diff(log(Close))
```

---

(vii) 

```{r}
acf(returns, main = "ACF", xlab = "Lags", ylab = "Serial Correlation", 
lag.max=60)
```


```{r}
pacf(returns, main = "PACF", xlab = "Lags", ylab = "Serial Correlation", lag.max=60)
```

---

(viii) 

* The monthly returns appear stationary. If a series is stationary, the A
CF should decay to zero quickly and should not display any oscillation.
 
---

(ix) The most appropriate ARMA will be ARMA (0,0) as the ACF and PACF do not 
show significance at any of the lags 
(x) 

```{r}
library(forecast)
```


---

```{r}
m1<-arima(returns,order = c(0,0,0))
m2<-arima(returns,order = c(1,0,0))
m3<-arima(returns,order = c(0,0,1))
m4<-arima(returns,order = c(1,0,1))

```

---

```{r}
m1$aic
##  -164.6838
m2$aic
##  -163.5281
m3$aic
##  -163.3079
m4$aic
##  -165.2896
```

--- 

```{r}
auto.arima(returns)
## Series: returns 
## ARIMA(0,0,0)(1,0,0)[12] with zero mean 
## 
## Coefficients:
## sar1
## 0.2548
## s.e. 0.1120
## 
## sigma^2 estimated as 0.005687: log likelihood=86.37
## AIC=-168.74 AICc=-168.58 BIC=-164.14
# ARMA (1,1) has the lowest AIC.
```

---

(xi) The deviation is observed because seasonality is not considered in the 
model. ACF is actually demonstrating a small significant serial correlation at lag 12, which might give rise to seasonal ARMA. 
# If we are forcefully fitting non seasonal models to a seasonal data, the di
screpancy is possible.
 [1 mark for each valid observation, Max 2]

---