---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


Before answering this question, generate the vector, X, in R using the following code:
```{r}
set.seed(1027); 
X = rexp(n=1000, rate=0.01)
```

The vector X represents the gross claim sizes of 1,000 claims. 

---


### Part 1. 
The payments are to be split between an insurance company and its reinsurer under an Excess of Loss reinsurance arrangement with a retention level M = 400. 

Determine the proportion of the claims that are fully covered by the insurer. 

```{r}
length(X[X<=400])/1000
```

```{r}
sum(X<=400)/1000

```

---


### Part 2.

Generate an additional vector, Y, which is of the same length as X, such that Y represents the amounts to be paid by the insurer for each component of X.  

```{r}
M = 400
Y = pmin(X, M)
```

```{r}
#Inspect first 20 values of Y

Y[1:20]
```
---

### Part 3.

Generate an additional vector, Z, which is of the same length as X, such that Z
represents the amounts to be paid by the reinsurer for each component of X.

An actuary assumes that the underlying gross claims distribution follows an exponential distribution of some unknown rate $\lambda$. 

```{r}
Z = pmax(0, X-M)

#Inspect Non-Zero Values of Z
Z[Z>0]
```

```{r}
z = X-Y

#Inspect Non-Zero Values of Z
Z[Z>0]
```


---

### Part 4.

The actuary needs to estimate $\lambda$ using only the claim amounts recorded in vector Y. 

Construct R code that calculates the log-likelihood, as a function of the parameter $\lambda$, given the claim amounts data in vector Y. 

```{r}
S = sum(X[X<=400])
logLik = function(lambda){
  -5200 * lambda + 987 * log(lambda) - lambda * S
}
```

---

### Part 4.

```{r}
nz = length(Y[Y==M])
Y_exc_M = Y[Y<M]
flnL = function(parameter){
  nz*pexp(M, rate=parameter, lower.tail=FALSE,
  log.p=TRUE)+
  sum(dexp(Y_exc_M, rate=parameter, log=TRUE))
}
```

---

### Part 5.

Determine the value of $\lambda$ at which the log-likelihood function reaches its maximum.  

```{r}
987 / (5200 + sum(X[X<=400]))
```

---

```{r}
nlm(f = function(x) - logLik(x), p = 0.01)$estimate
```

Hence, the maximum likelihood estimate of lambda = 0.01023209.
