---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

### Life Insurance Portfolios 
* An analyst is investigating a life insurance portfolio data set that comprises two variables, x1 and x2, for 200 policyholders. 

* The analyst is exploring whether the 200 policyholders can be divided into two clusters (labelled A and B) based on the two variables, x1 and x2.  

* Before answering this question, generate the data set, portfolio, in R using the following code: 

```{r}
set.seed(2019);  
portfolio <-  data.frame(
  x1=rnorm(200,3,1),  
  x2=scale(c(rnorm(70,4,1), rnorm(130,10,1)))) 
```
---

```{r}
head(portfolio)
```

### Exercises

In the first stage of the investigation, the analyst decides to assign the first 100 policyholders in the data set to cluster A, and the remaining policyholders to cluster B. 

1. Construct a new column in the data set, portfolio, called <tt>group_label_stage1</tt>, containing the policyholder cluster labels, defined above.

2. Determine the coordinates (x1A, x2A) of the centre of cluster A and the coordinates (x1B, x2B) of the centre of cluster B. 

3. Construct a new column in the data set, portfolio, called <tt>dist_A</tt> containing the Euclidean distances between the policyholders and the centre of cluster A.   

--- 
<ol>
<li>Construct a new column in the data set, portfolio, called dist_B containing the Euclidean distances between the policyholders and the centre of cluster B.  </li> 
</ol>
The analyst decides to update the cluster labels by assigning to each policyholder the label of the cluster whose centre is nearest, according to the distances calculated in parts 3 and 4.. 
--- 

5. Construct a new column in the data set, portfolio, called <tt>group_label_stage2</tt>, containing the updated policyholder cluster labels, defined above.

--- 

6. Generate a 2x2 matrix showing the number of policyholders with each possible combination of values of <tt>group_label_stage1</tt> and <tt>group_label_stage2</tt>. 


7. Comment on the matrix generated in part (vi) with reference to how the cluster labels have changed between <tt>group_label_stage1</tt> and <tt>group_label_stage2</tt>.

8. Plot the column, x1,of data set, portfolio, against column, x2, (with x1 on the  x-axis and x2 on the y-axis), using two distinct colours to identify clusters Aand B according to the label, <tt>group_label_stage2</tt>. 

9. The analyst decides to stop at this stage and to report <tt>group_label_stage2</tt> as the final set of cluster labels. 
Comment on this decision. 

---


3
(i) 

```{r}
portfolio$group_label_stage1 <- c(rep("A", length=100),
rep("B", length=100))
```

```{r}
x1_A <- mean(portfolio$x1[portfolio$group_label_stage1=="A"]); x1_A


x2_A <- mean(portfolio$x2[portfolio$group_label_stage1=="A"]); x2_A

x1_B <- mean(portfolio$x1[portfolio$group_label_stage1=="B"]); x1_B

x2_B <- mean(portfolio$x2[portfolio$group_label_stage1=="B"]); x2_B
 
```
```{r}
head(portfolio)
```

```{r}
library(dplyr)
portfolio %>% group_by(group_label_stage1) %>% 
  summarize(meanx1=mean(x1),
            meanx2=mean(x2)
            )
```

```{r}
model1 = kmeans(portfolio[1:100,1:2],1)
model1$centers
```
```{r}
x1_A <- model1$centers[1,1]; x1_A
x2_A <- model1$centers[1,2]; x2_A
```

```{r}
model2 = kmeans(portfolio[101:200,1:2],1)
model2$centers
```

```{r}
x1_B <- model2$centers[1,1]; x1_B
x2_B <- model2$centers[1,2]; x2_B
```

Therefore:
the coordinates of the centre of cluster “A”, (x1_A, x2_A) = (2.926666, -0.7054048)

and:
the coordinates of the centre of cluster “B”, (x1_B, x2_B) = (2.829781, 0.7054048)

---

The analyst decides to update the cluster labels by assigning to each policyholder the
label of the cluster whose centre is nearest, according to the distances calculated in
parts (iii) and (iv).
(v) Construct a new column in the data set, portfolio, called group_label_stage2,
containing the updated policyholder cluster labels, defined above.

(vi) Generate a 2x2 matrix showing the number of policyholders with each
possible combination of values of group_label_stage1 and
group_label_stage2.
(vii)

Comment on the matrix generated in part (vi) with reference to how the cluster
labels have changed between group_label_stage1 and group_label_stage2. 
(viii) Plot the column, x1, of data set, portfolio, against column, x2, (with x1 on the
x-axis and x2 on the y-axis), using two distinct colours to identify clusters A
and B according to the label, group_label_stage2.

The analyst decides to stop at this stage and to report group_label_stage2 as the final
set of cluster labels.
(ix)
Comment on this decision.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

```{r}
portfolio$dist_A <- sqrt((portfolio$x1 - x1_A)^2 +
(portfolio$x2 - x2_A)^2) 

portfolio$dist_B <- sqrt((portfolio$x1 - x1_B)^2 +
(portfolio$x2 - x2_B)^2) 

portfolio$group_label_stage2 <- rep("A", 200)
portfolio$group_label_stage2[portfolio$dist_B <portfolio$dist_A] <- "B" 

table(portfolio$group_label_stage1,portfolio$group_label_stage2)
```
---

(vii)

There are 70 policyholders whose cluster labels were originally assigned to “A” in the
first stage of the investigation which remained unchanged after the update.

There are 97 policyholders whose cluster labels were originally assigned to “B” in the
first stage of the investigation which remained unchanged after the update.

However, the labels of 30 policyholders were updated from “A” to “B”... 
...whereas, the labels of 3 policyholders were updated from “B” to “A”. 

---
(viii) 

```{r}
col_vec <- rep("blue", 200)
col_vec[portfolio$group_label_stage2=="B"] <- "red"
```




col=col_vec)
```{r}
plot(
  portfolio$x1,
  portfolio$x2,
  main="Plot of x1 against x2 with group_label_stage2 policyholder cluster labels",
legend("topright", legend=c("group A", "group B"),
col=c("blue", "red"), 
pch=1))

```

(ix)
The graph in part (viii) shows that the analyst was able to reasonably identify two sets
of clusters.

However, the analyst did not check the convergence of the clustering algorithm... 
... although, inspection of the table might suggest that the clusters are unlikely to
change much.

---

* The top three group A policyholders might possibly be better assigned to group B. 
* The analyst could potentially improve the results by updating the centres of the
clusters, re-calculating the distances, updating the labels and repeating this process
until convergence (i.e. until the labels remain constant).

* The analyst could implement the full kmeans algorithm to ensure convergence of the
final clusters.

---

The analyst may want to apply feature scaling / data normalisation to the values of
x_1 and x_2 so that each of them contributes approximately proportionately to the
Euclidean distances and then re-run the analysis.


---