---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Solution 3: 
i) 
#a > Sports <- read.csv(Sports.csv") 
> Model1<-lm(Sports$Points~Sports$X100m+Sports$X400m+Sports$X110m.hurdle+Sports
 $High.jump+Sports$Long.jump+Sports$Pole.vault+Sports$Shot.put+Sports$Javeline+Sports
 $Discus) 
> summary(Model1)  
(3) 
(2) 
(3) 
(2) 
[15] 
Page 5 of 11 
IAI                             
CS1B-1124 
Call: 
lm(formula = Sports$Points ~ Sports$X100m + Sports$X400m + Sports$X110m.hurdle +  
Sports$High.jump + Sports$Long.jump + Sports$Pole.vault +  
Sports$Shot.put + Sports$Javeline + Sports$Discus) 
Residuals: 
Min      1Q  Median      3Q     Max  -97.106 -25.043  -7.748  33.856 119.528  
Coefficients: 
Estimate Std. Error t value Pr(>|t|)     
(Intercept)         
8632.632   1368.083   6.310 7.83e-06 *** 
Sports$X100m        
Sports$X400m         -163.414     76.208  -2.144 0.046756 *   -78.141     16.269  -4.803 0.000166 *** 
Sports$X110m.hurdle -120.901     41.818  -2.891 0.010152 *   
Sports$High.jump     810.196    211.532   3.830 0.001340 **  
Sports$Long.jump     209.187     63.472   3.296 0.004269 **  
Sports$Pole.vault    183.345     66.876   2.742 0.013912 *   
Sports$Shot.put       
90.349     28.200   3.204 0.005204 **  
Sports$Javeline       
Sports$Discus         
17.757      
10.986      
2.975   5.969 1.52e-05 *** 
5.781   1.900 0.074491 .   --- 
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
Residual standard error: 60.74 on 17 degrees of freedom 
Multiple R-squared:  0.9794, 
Adjusted R-squared:  0.9685  
F-statistic: 89.91 on 9 and 17 DF,  p-value: 1.491e-12 
#b Plot of residuals for the model  
ii) 
Using a general rule if the p-value is less than  0.05, then the concerned explanatory variable 
is considered to be significant. This is shown by *,  **, *** signs in the R summary output . 
So, based on the R-output, only points for Discus Throw are not significant. All other 
explanatory variables are considered to be significant.    
residuals(Model1)
 0 50-100
 0
 5
 10
 15
 Index
 20
 25
 > Model2_poisson<-glm(Sports$Points~Sports$X100m+Sports$X400m+Sports$X110m.hurd
 le+Sports$High.jump+Sports$Long.jump+Sports$Pole.vault+Sports$Shot.put+Sports$Javelin
 e+Sports$Discus,family="poisson") 
> summary(Model2_poisson)  
Call: 
glm(formula = Sports$Points ~ Sports$X100m + Sports$X400m + Sports$X110m.hurdle +  
Sports$High.jump + Sports$Long.jump + Sports$Pole.vault +  
Sports$Shot.put + Sports$Javeline + Sports$Discus, family = "poisson") 
Deviance Residuals:   
(6) 
(2) 
Page 6 of 11 
IAI                             
CS1B-1124 
Min        
1Q    Median        
3Q       Max   -1.02319  -0.25491   0.03473   0.36449   1.31328   
Coefficients: 
(Intercept)          
Estimate Std. Error z value Pr(>|z|)     
9.0988649  0.2501151  36.379  < 2e-16 *** 
Sports$X100m        
Sports$X400m        -0.0211126  0.0139153  -1.517  0.12921     -0.0093734  0.0029665  -3.160  0.00158 **  
Sports$X110m.hurdle -0.0161147  0.0076612  -2.103  0.03543 *   
Sports$High.jump     0.1009034  0.0387525   2.604  0.00922 **  
Sports$Long.jump     0.0239972  0.0116951   2.052  0.04018 *   
Sports$Pole.vault    0.0225249  0.0122296   1.842  0.06550 .   
Sports$Shot.put      
0.0111537  0.0051471   2.167  0.03024 *   
Sports$Javeline      
Sports$Discus        
0.0021487  0.0005379   3.995 6.48e-05 *** 
0.0012340  0.0010555   1.169  0.24236     --- 
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
(Dispersion parameter for poisson family taken to be 1) 
Null deviance: 374.4158  on 26  degrees of freedom 
Residual deviance:   8.3483  on 17  degrees of freedom 
AIC: 321 
Number of Fisher Scoring iterations: 3 
iii) 
iv) 
Scaled deviance can be used to compare only nested models.  
Since, Model 1 and Model2_Poisson are models with different distributional assumptions. 
Model 1 assumes normal distribution and Model 2 assumes Poisson distribution.  
They are not nested models. 
Hence, scaled deviance cannot be used to compare the models in parts (i) and (ii). 
An equivalent model to a linear regression model will be a GLM with normal distribution.   
> Model2_normal<-glm(Sports$Points~Sports$X100m+Sports$X400m+Sports$X110m.hurdl
 e+Sports$High.jump+Sports$Long.jump+Sports$Pole.vault+Sports$Shot.put+Sports$Javelin
 e+Sports$Discus,family=gaussian()) 
> summary(Model2_normal)  
Call: 
glm(formula = Sports$Points ~ Sports$X100m + Sports$X400m + Sports$X110m.hurdle +  
Sports$High.jump + Sports$Long.jump + Sports$Pole.vault +  
Sports$Shot.put + Sports$Javeline + Sports$Discus, family = gaussian()) 
Deviance Residuals:  
Min       
1Q   Median       3Q      Max   -97.106  -25.043   -7.748   33.856  119.528   
Coefficients: 
Estimate Std. Error t value Pr(>|t|)     
(Intercept)         
8632.632   1368.083   6.310 7.83e-06 ***  
(4) 
(2) 
Page 7 of 11 
IAI                             
CS1B-1124 
Sports$X100m        
Sports$X400m         -163.414     76.208  -2.144 0.046756 *   -78.141     16.269  -4.803 0.000166 *** 
Sports$X110m.hurdle -120.901     41.818  -2.891 0.010152 *   
Sports$High.jump     810.196    211.532   3.830 0.001340 **  
Sports$Long.jump     209.187     63.472   3.296 0.004269 **  
Sports$Pole.vault    183.345     66.876   2.742 0.013912 *   
Sports$Shot.put       
90.349     28.200   3.204 0.005204 **  
Sports$Javeline       
Sports$Discus         
17.757      
10.986      
2.975   5.969 1.52e-05 *** 
5.781   1.900 0.074491 .   --- 
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
(Dispersion parameter for gaussian family taken to be 3689.4) 
Null deviance: 3048179  on 26  degrees of freedom 
Residual deviance:   62720  on 17  degrees of freedom 
AIC: 307.89 
Number of Fisher Scoring iterations: 2 
v) 
vi) 
Model 1 in part (i) and Model 2 (Normal) in part (iv) are exact equivalents of each other. The 
same can be seen from the estimates of coefficients in the R summary output. 
So, for comparing the fit of Model 1 and Model 2 (Poisson), Model 2 (Normal) can be used as 
a proxy for Model 1. And then, the AIC of Model 2 (Poisson) can be compared with the AIC 
of Model 2 (Normal). 
AIC (Model 2 Normal) = 307.89 
AIC (Model 2 Poisson) = 321 
Smaller the AIC, better is the fit. So, Model 2 Normal is a better fit as compared to Model 2 
Poisson.  
Consequentially, the linear multiple regression model in part (i) is a better fit to the data as 
compared to the GLM fitted in part (ii). 
> run<-data.frame(Sports$X100m,Sports$X400m,Sports$X110m.hurdle) 
> pr_run<-prcomp(run,scale. = TRUE) 
>  
> summary(pr_run) 
Importance of components: 
PC1    PC2    PC3 
Standard deviation     1.492 0.6680 0.5726 
Proportion of Variance 0.742 0.1487 0.1093 
Cumulative Proportion  0.742 0.8907 1.0000 
>  
> jump<-data.frame(Sports$High.jump,Sports$Long.jump,Sports$Pole.vault) 
> pr_jump<-prcomp(jump,scale. = TRUE) 
>  
> summary(pr_jump) 
Importance of components: 
PC1    PC2    PC3  
(4) 
(2) 
Page 8 of 11 
IAI                             CS1B-1124 
     Page 9 of 11 
 
Standard deviation     1.2588 1.0307 0.5941 
Proportion of Variance 0.5282 0.3541 0.1176 
Cumulative Proportion  0.5282 0.8824 1.0000 
  
>  
> throw<-data.frame(Sports$Shot.put,Sports$Javeline,Sports$Discus) 
> pr_throw<-prcomp(throw,scale. = TRUE) 
>  
> summary(pr_throw) 
Importance of components: 
                          PC1    PC2     PC3 
Standard deviation     1.4141 0.8727 0.48853 
Proportion of Variance 0.6665 0.2539 0.07955 
Cumulative Proportion  0.6665 0.9204 1.00000  
  
For Run Sports Category, 74.2% variance is captured by PC1. 
For Jumping Sports Category, 52.82% variance is captured by PC1. 
For Throw Sports Category, 66.65% variance is captured by PC1.  (7) 
   
vii) Considering the summary R output generated in part (vi),  
 
In case of Run Sports Category, all three PCs cumulatively capture at least 90% of the total 
variance of the data. 
 
In case of Jump Sports Category, all three PCs cumulatively capture at least 90% of the total 
variance of the data. 
 
In case of Throw Sports Category, first two PCs cumulatively capture 92.04% (at least 90%) 
of the total variance of the data. 
 
So, if PCs capturing at least 90% of the variance is a criterion for reducing the dimensionality 
of the data set, then Throw Sports Category satisfies it. For Throw Sports, PC1 and PC2 can 
be retained and PC3 can be dropped thus reducing the dimensionality of the data set. In case 
of Run Sports and Jumping Sports, all three PCs need to be retained and thus the 
dimensionality of the data set will not be reduced.  
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
(3) 
   
viii) > Model3<-lm(Sports$Points~Sports$Pole.vault)  
  
H0: Beta coefficient is equal to 0 
H1: Beta coefficient is not equal to 0.   
   
 > confint(Model3,level=0.95) 
                      2.5 %     97.5 % 
(Intercept)       5654.3838 10895.1124 
Sports$Pole.vault -573.3559   508.9227  
   
 As the 95% confidence interval for beta (-573.3559, 508.9277) contains the value 0, we do 
not have sufficient evidence to reject the null hypothesis at 5% level of significance. Hence, 
based on this test, one can conclude that there is no correlation between pole vault score and 
winning points.  (5) 
   
ix) > Model4<-lm(Sports$X110m.hurdle~Sports$X100m) 
> summary(Model4) 
  
IAI                             CS1B-1124 
     Page 10 of 11 
 
Call: 
lm(formula = Sports$X110m.hurdle ~ Sports$X100m) 
 
Residuals: 
     Min       1Q   Median       3Q      Max  -0.48075 -0.28936 -0.05353  0.21428  0.76203  
 
Coefficients: 
             Estimate Std. Error t value Pr(>|t|)     
(Intercept)    2.2036     2.7252   0.809 0.426379     
Sports$X100m   1.1183     0.2478   4.512 0.000132 *** --- 
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
 
Residual standard error: 0.356 on 25 degrees of freedom 
Multiple R-squared:  0.4489, Adjusted R-squared:  0.4268  
F-statistic: 20.36 on 1 and 25 DF,  p-value: 0.0001319 
  
> score<-c(10.68,10.42,11.68,11.62,10.54) 
> hurdle_score<-2.2036+1.1183*score 
> hurdle_score 
[1] 14.14704 13.85629 15.26534 15.19825 13.99048  
  (5) 
  [40] 
   