---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---
Solution 4:  
   
i)  > exam.success = matrix(c(132,120,27,51),ncol=2,nrow=2) 
> exam.success 
     [,1] [,2] 
[1,]  132   27 
[2,]  120   51 
 
Data is properly getting displayed.  
   
 H0: tutorial attendance and exam success are independent, against  
H1: tutorial attendance and exam success are not independent (1) 
   
ii) > chisq.test(exam.success)$expected 
         [,1]     [,2] 
[1,] 121.4182 37.58182 
[2,] 130.5818 40.41818 (2) 
   
iii) > chisq.test(exam.success) 
 
 Pearson's Chi-squared test with Yates' continuity correction 
 
data:  exam.success 
X-squared = 6.8349, df = 1, p-value = 0.008939  
   
 
 
The p-value is significant (e.g. at the 1% level), since 0.008939 < 0.01 – therefore there is evi
 dence to reject the null hypothesis and we conclude that tutorial attendance and exam success 
are not independent. 
 
 
(3) 
   
IAI                             
CS1B-1124 
iv) 
v) 
vi) 
> fisher.test(exam.success) 
Fisher's Exact Test for Count Data 
data:  exam.success 
p-value = 0.006544 
alternative hypothesis: true odds ratio is not equal to 1 
95 percent confidence interval: 
1.190372 3.671876 
sample estimates: 
odds ratio  
2.073216   
The p-value is significant (e.g. at the 1% level), 0.006544 < 0.01 – therefore there is evidence 
to reject the null hypothesis and we conclude that tutorial attendance and exam success are not 
independent. Conclusion under  Fisher’s exact test is similar to conclusion under contingency 
table test. 
(a) Fisher’s test is an exact test whereas chi-square test is an approximation  
(b) Fisher’s test is suitable for 2×2 datasets whereas chi-square test can be used for N×N datas
 ets. 
H0: the proportion of students passing the exam is 60% (p = 0.60) 
H1: the proportion of students passing the exam is not equal to 60% (p <> 0.60)  
> x=132+120 
> n=x+27+51 
> binom.test(x,n,conf.level = 0.95) 
Exact binomial test 
data:  x and n 
number of successes = 252, number of trials = 330, p-value < 
2.2e-16 
alternative hypothesis: true probability of success is not equal to 0.5 
95 percent confidence interval: 
0.7140288 0.8084419 
sample estimates: 
probability of success  
0.7636364   
The p-value is < 2.2e-16 which is definitely less than 5% and hence we have sufficient eviden
 ce to reject the null hypothesis and hence we can conclude that the proportion of students pass
 ing the examination is not equal to 60%.  
************* 
(3) 
(2) 
(4) 
[15] 
Page 11 of 11 