---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Solution 2:  
   
i) A. Pibar represents the total number of claims for every insurer (total of no of claims across 
all the years) 
B. Pstar is a quantity calculated by taking sum over all insurers of (total number of claims per 
insurer multiplied by proportion of total no of claims of that insurer to total) and divided 
by total number of cells minus 1. This quantity is used in the denominator while 
determining V(m(θ)). 
C. m or E(m(θ)) is the expected value of average claims per policy based on the collateral 
data. 
D. s or E(s2(θ)) is the expected value of the variance of claims per policy based on the 
collateral data. 
E. v or V(m(θ)) is the variance of average claims per policy based on the collateral data 
 
(5) 
   
ii) > claims<-data.frame( 
+   Year1 = c(14.2,58.6,123), 
+   Year2 = c(15.8,63.1,132), 
+   Year3 = c(22.7,81.0,161), 
+   Year4 = c(19,64.2,133) 
+ ) 
>  
> nopols<-data.frame( 
+   Year1 = c(163,4435,16184), 
+   Year2 = c(189,4761,17443), 
 
 
 
 
IAI                             
CS1B-1124 
+   Year3 = c(252,5576,20102), 
+   Year4 = c(199,4581,18000) 
+ ) 
>  
> n <- ncol(claims) ## This stands for n 
> N <-nrow(claims) ## This stands for N 
> X <- claims/nopols ## This stands for Xij 
> Xibar <-rowSums(claims) / rowSums(nopols) # Xibar   
> Pibar <- rowSums(nopols) # Pibar  
> Pbar <-sum(Pibar) # Pbar 
> Pstar <-sum(Pibar * (1-Pibar/Pbar))/(N*n-1) # Pstar 
> m <-sum(claims) / Pbar # E[m(θ)] 
> print(m) 
[1] 0.009659901 
>  
> s <- mean(rowSums(nopols *(X-Xibar)^2)/(n-1)) # E[s2(θ)]  
> print(s) 
[1] 0.002749923 
> v <- (sum(rowSums(nopols*(X-m)^2))/(n*N-1)-s)/Pstar 
> print(v) 
[1] 0.0001793695 
iii) 
iv) 
v) 
> Zi <- Pibar / (Pibar + s/v) 
> print(Zi) 
[1] 0.9812655 0.9992084 0.9997863 
> Premi <- Zi * Xibar + (1-Zi) * m 
> print(Premi) 
[1] 0.087798326 0.013787873 0.007654237 
> nopols_y5 <- c(5000,4800,4200) 
> claims_y5 <- Premi * nopols_y5 
> print(claims_y5) 
[1] 438.99163  66.18179  32.14779 
Since the credibility premiums are high and close to 1 in case of all insurers, we are giving 
more importance to the direct data (related to that particular insurer) and ignoring the 
collateral data (data related to other insurers) 
This is reasonable as there is wide variation in the claims amount for various insurers and 
hence more emphasis should be given on direct data. 
