---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


* This exercise provides a practical application of statistical modeling in the field of actuarial science, specifically focusing on extreme value analysis. 

* Using R, students learn to process raw claims data, transform it using the block maxima method to isolate extreme events, and then fit appropriate distributions like the **Weibull** and **Generalized Extreme Value (GEV)** distributions. 

* The purpose is to equip students with the skills to analyze tail risk, understand the characteristics of large claims through functions like the hazard and mean residual life, and apply computational tools for actuarial problem-solving.


---

### Exercises

Q. 2)
Consider the data file “Claims.csv” of claim amounts in 000’s in respect of a commercial property portfolio over a period of 10 years.

#### Exercises 

i) Calculate the block maxima for these claims using block sizes of 5 claims and paste the output.

ii) Plot a histogram of the block maxima and label the diagram.

iii) Plot an empirical density function in the same graph as in part (ii) above.

---

iv) Fit a Weibull distribution on the block maxima and calculate the estimates of the parameters of the Weibull distribution.

v) Calculate the mean, standard deviation and skewness of the block maxima.

vi) Write a function in R to compute the log-likelihood of the above distribution.

Hint : Use the following PDF function to compute the log likelihood.

---



vii) Find the maximum likelihood estimate of the parameters of the above distribution. Use mean, standard deviation and skewness calculated above as the initial estimates of the parameters.

viii) Superimpose a Generalized Extreme Value distribution on the block maxima using the estimates of the parameters calculated above in part (vii).

ix) Calculate the hazard rate for the block maxima and plot the graph of Hazard rate against block maxima and comment on the graph.

x) Calculate the mean residual life for the block maxima and plot the graph of mean residual life against block maxima. Comment on the graph.


---

Solution 2:
i)
<code for reading the input data file> e.g. 

```{r}
Claims <- read.csv("Claims.csv")
Claims$block <-(Claims$Claim_number-1) %/% 5 +1
blockmax <- aggregate(Claims ~ block, Claims, max)
blockmax 
```

block Claims 1 1 104 2 2 94 3 3 218 4 4 235 5 5 140 6 6 84 7 7 213 8 8 222 9 9 128 10 10 247 11 11 152 12 12 202 13 13 193 14 14 201 15 15 180 16 16 291 17 17 243 18 18 163 19 19 267 20 20 203
[2]
---

### Part B 

```{r}
hist(blockmax$Claims, 
     xlab = "Claims", main = "Histogram of block maxima")
```

---

### Part C

```{r}
hist(blockmax$Claims, xlab = "Claims", 
main = "Histogram of block maxima", freq = FALSE) 
lines(density(blockmax$Claims), col = "blue")

```
[2]

---
iv) 

```{r}
library(MASS) 
est <- fitdistr(blockmax$Claims, "weibull", lower = 0) 
est 
```

shape scale 3.8973242 209.3505841 ( 0.7069667) ( 12.6163461)
c = est$estimate["scale"]^ (-est$estimate["shape"]) > c scale 9.011579e-10 > g = est$estimate["shape"]

--- 

> g shape 3.897324
[3]

---

v) 


```{r}
alpha = mean(blockmax$Claims) 
alpha 
```

[1] 189 

```{r}
beta = sd(blockmax$Claims)
beta 

```

[1] 57.78271 gamma = skewness(blockmax$Claims) gamma [1] -0.273219


---

### Part F

**vi)**

```r
MLE = function(x){
  f <- 1/x[2]*(1+x[3]*(blockmax$Claims - x[1])/x[2])^(-1-1/x[3])*exp(-(1+x[3]*(blockmax$Claims- x[1])/x[2])^(-1/x[3]))
  lnf <- log(f)
  sum(-lnf)
}
```

**vii)**

```r
p = c(alpha,beta,gamma)
MLE(p) 
# [1] 110.3667

f_MLE <- nlm(MLE,p)
```

**Results of nlm() for MLE:**
`$minimum` [1] 108.4947
`$estimate` [1] 173.0482986 59.5496685 -0.4257482
`$gradient` [1] -2.207406e-07 -7.863145e-07 -5.439915e-05
`$code` [1] 1
`$iterations` [1] 33

**viii)**

```r
GEV <- function(x,a,b,c){
  f = 1/b * (1+c*(x-a)/b)^-(1+1/c)*exp(-((1+c*(x-a)/b)^(-1/c)))
}

fit = GEV(blockmax$Claims,f_MLE$estimate[1],f_MLE$estimate[2],f_MLE$estimate[3])
fit
```

**Output of fit:**
[1] 0.002216593 0.001752716 0.006657504 0.005922606 0.004311508 0.001356100 0.006776331
[8] 0.006529717 0.003563970 0.005131517 0.005053503 0.006876667 0.006798624 0.006874941
[15] 0.006456144 0.001361035 0.005416809 0.005683342 0.003474704 0.006876611

-----


The R code is implementing the negative log-likelihood function for the Generalized Extreme Value (GEV) distribution. The function takes a vector of parameters `x` where `x[1]` is the location parameter, `x[2]` is the scale parameter, and `x[3]` is the shape parameter. The code first calculates the probability density function (PDF) `f` for each claim in the `blockmax$Claims` dataset, then takes the negative sum of the log of these values. Minimizing this negative log-likelihood function is equivalent to maximizing the log-likelihood function, which is the standard approach for finding the Maximum Likelihood Estimates (MLE) of the GEV parameters.

---

### Part F
vi) 

```{r}
MLE = function(x){
    f <- 1/x[2]*(1+x[3]*(blockmax$Claims - x[1])/x[2])^(-1-1/x[3])*exp(-(1+x[3]*(blockmax$Claims- x[1])/x[2])^(-1/x[3])) 

    lnf <- log(f) 

sum(-lnf) #
}

```
----
vii) 

```{r}p = c(alpha,beta,gamma) MLE(p) [1] 110.3667
f_MLE <- nlm(MLE,p)

```

f_MLE $minimum [1] 108.4947 $estimate [1] 173.0482986 59.5496685 -0.4257482 $gradient 
[1] -2.207406e-07 -7.863145e-07 -5.439915e-05 $code [1] 1 $iterations [1] 33

---

viii) 

```{r}
GEV <- function(x,a,b,c){f = 1/b * (1+c*(x-a)/b)^-(1+1/c)*exp(-((1+c*(x-a)/b)^(-1/c)))} 
```


```{r}
fit = GEV(blockmax$Claims,f_MLE$estimate[1],f_MLE$estimate[2],f_MLE$estimate[3]) 
fit 

```

[1] 0.002216593 0.001752716 0.006657504 0.005922606 0.004311508 0.001356100 0.006776331 
[8] 0.006529717 0.003563970 0.005131517 0.005053503 0.006876667 0.006798624 0.006874941 
[15] 0.006456144 0.001361035 0.005416809 0.005683342 0.003474704 0.006876611

---

### Part G

ix) h = dweibull(blockmax$Claims,g,est$estimate["scale"])/(1-pweibull(blockmax$Claims,g,est$estimate["scale"])) h 

[1] 0.002452262 0.001829610 0.020933123 0.026020746 0.005802246 0.001320774 0.019572126 
[8] 0.022065445 0.004475462 0.030059810 0.007363354 0.016784861 0.014708485 0.016545242 
[15] 0.012017747 0.048335279 0.028670960 0.009015548 0.037666671 0.017026742

plot(m,blockmax$Claims)

### Part G

**ix)**

```r
h = dweibull(blockmax$Claims,g,est$estimate["scale"])/(1-pweibull(blockmax$Claims,g,est$estimate["scale"]))
h
```

**Output of h:**
[1] 0.002452262 0.001829610 0.020933123 0.026020746 0.005802246 0.001320774 0.019572126
[8] 0.022065445 0.004475462 0.030059810 0.007363354 0.016784861 0.014708485 0.016545242
[15] 0.012017747 0.048335279 0.028670960 0.009015548 0.037666671 0.017026742

```r
plot(m,blockmax$Claims)
```

**Conclusion:**
The hazard function is an increasing function of x. An increasing hazard function indicates lighter tail.

-----

---
The hazard function is an increasing function of x. An increasing hazard function indicates lighter tail.
[4]

---

### Part J
x) Sy = function(y,g,b){(1-pweibull(y,g,b))} int = integrate(Sy,0, Inf, g, est$estimate[“scale”]) ex = int$value/Sy(blockmax$Claims,g,est$estimate[“scale”]) plot(ex,blockmax$Claims)
The mean residual life function is an increasing function of x. An increasing mean residual function indicates a lighter tail.

