---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


#### AutoClaims

Description of the given data sets

AutoClaims.csv

An insurance company is trying to analyse the paid and closed claims data of the past quarter.

The columns carry the following explanation

* *STATE*: The state of registration of the vehicle
* *CLASS*: Rating class of the insured based on marital status, use of vehicle and other attributes
* *GENDER*: Gender of the Insured (Same as the Driver)
* *AGE*: Age of the Insured (Same as the Driver)
* *PAID*: Amount paid to settle and close a claim

---


```{r}
getwd()
```


'/home/nbuser/library'



```{r}
list.files()
```


<ol class="list-inline">
	<li>'AutoClaims - 1 - Inspecting the Data.ipynb'</li>
	<li>'AutoClaims.csv'</li>
	<li>'README.md'</li>
</ol>




```{r}
AutoClaims <- read.csv("Data/AutoClaims.csv")
```


---

```{r}
summary(AutoClaims)
```




```{r}
library(readr)
AutoClaims <- read_csv("Data/AutoClaims.csv")

```




```{r}
head(AutoClaims)
```

---



```{r}
names(AutoClaims)
```


```{r}
dim(AutoClaims)
```


---



```{r}
summary(AutoClaims)
```

---



```{r}
AutoClaims$GENDER[1:20]
```



--- 


```{r}
# install.packages("rlang")
# install.packages("magrittr")
# install.packages("janitor")
library(janitor)
library(magrittr)

```

---

#### Tabulations: STATE

```{r}
AutoClaims %>% tabyl(STATE)

```


---


#### Tabulations: GENDER

```{r}
AutoClaims %>% tabyl(GENDER)
```

---


#### Cross-Tabulations: CLASS and GENDER

```{r}
AutoClaims %>% tabyl(CLASS,GENDER)
```




