
\large
# sessionInfo()
### AutoClaims
​
library(readr)
AutoClaims <- read_csv("AutoClaims.csv")
​
Parsed with column specification:
cols(
  STATE = col_character(),
  CLASS = col_character(),
  GENDER = col_character(),
  AGE = col_double(),
  PAID = col_double()
)


Exercise 1
Prepare a table with mean, standard deviation and coefficient of variance (CV) of the claims paid

for each STATE and identify the one with the least and the highest CV
for each CLASS and sort the table in the ascending order of the CV

Hint: Coefficient of Variance (CV) can be computed as the ratio of standard deviation to the mean.


Implementation using Base R:
```{r}
state_mean<-aggregate(PAID~STATE,data = AutoClaims,FUN = mean)
names(state_mean)<-c("State","Mean")
```

​
state_mean 
A data.frame: 11 x 2
State	Mean
<chr>	<dbl>
STATE 01	10235.800
STATE 02	7055.078
STATE 03	8714.932
STATE 04	8152.759
STATE 06	8786.739
STATE 07	4960.479
STATE 10	12340.643
STATE 12	6893.705
STATE 14	10399.313
STATE 15	3321.449
STATE 17	7886.282
state_sd<-aggregate(PAID~STATE,data = AutoClaims,FUN = sd)
names(state_sd)<-c("State","SD")
​
state_sd
A data.frame: 11 x 2
State	SD
<chr>	<dbl>
STATE 01	10932.877
STATE 02	6327.473
STATE 03	6494.346
STATE 04	6985.210
STATE 06	10749.517
STATE 07	3065.092
STATE 10	14599.291
STATE 12	8634.955
STATE 14	8406.388
STATE 15	3364.269
STATE 17	7831.913
​
state_summary<-merge(state_mean,state_sd)
​
state_summary
state_summary$CV <- state_summary$SD / state_summary$Mean
​
#Mean, Standard Deviation and Coefficient of Variance for each state
​
state_summary
​
​
The State with minimum coefficient of variance is
​
```{r}
state_summary$State[state_summary$CV==min(state_summary$CV)]
``` 
​
​
​
The State with maximum coefficient of variance is
​
```{r}
state_summary$State[state_summary$CV==max(state_summary$CV)]
``` 
​
The State with minimum coefficient of variance is
​
```{r}
state_summary$State[state_summary$CV==min(state_summary$CV)]
``` 
​
​
The State with maximum coefficient of variance is
​
```{r}
state_summary$State[state_summary$CV==max(state_summary$CV)]
``` 
​
#### State Summarys (Tidyverse)
​
```{r}
library(dplyr)
library(magrittr)
​
state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(Mean = mean(PAID),
  * *    SD   = sd(PAID))
​
``` 
```{r}
head(state_summary_tidy)
``` 
​
```{r}
state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(Mean = mean(PAID),
  * *    SD   = sd(PAID)) %>%
   mutate(CV = Mean/SD) %>%
   arrange(CV)
​
```   
​
​
### Claims (Base R)
​
```{r}
Class_mean<-aggregate(PAID~CLASS,data = AutoClaims,FUN = mean)
names(Class_mean)<-c("Class","Mean")
Class_sd<-aggregate(PAID~CLASS,data = AutoClaims,FUN = sd)
names(Class_sd)<-c("Class","SD")
Class_summary<-merge(Class_mean,Class_sd)
Class_summary$CV<-Class_summary$SD/Class_summary$Mean
#Mean, Standard Deviation and Coefficient of variance for each rating class
Class_summary
``` 
​
​
```{r}
Class_summary[order(Class_summary$CV),]
``` 
​
```{r}
Class_summary_tidy <- AutoClaims %>% group_by(CLASS) %>%
   summarize(Mean = mean(PAID),
  * *    SD   = sd(PAID)) %>%
   mutate(CV = Mean/SD) %>%
   arrange(CV)
​
```   
​
```{r}
Class_summary_tidy
``` 
​
​
​
Class_mean<-aggregate(PAID~CLASS,data = AutoClaims,FUN = mean)
names(Class_mean)<-c("Class","Mean")
Class_sd<-aggregate(PAID~CLASS,data = AutoClaims,FUN = sd)
names(Class_sd)<-c("Class","SD")
Class_summary<-merge(Class_mean,Class_sd)
Class_summary$CV<-Class_summary$SD/Class_summary$Mean
#Mean, Standard Deviation and Coefficient of variance for each rating class
Class_summary
​
```{r}
Class_summary[order(Class_summary$CV),]
``` 
​
​
Exercise 2
By using a box plot, identify the STATE(s) and CLASS(es) which have no outlier values in terms of the claims paid.

options(repr.plot.width=12, repr.plot.height=8)
​
boxplot(PAID~STATE,data = AutoClaims)
​

States with no outliers are State 14.

boxplot(PAID~CLASS,data = AutoClaims)
​

Class with no outlier is C11.

​
library(ggplot2)
​
p <- ggplot(data=AutoClaims,aes(x=STATE,y=PAID)) + geom_boxplot()
​
p
​

​
p <- ggplot(data=AutoClaims,aes(x=CLASS,y=PAID)) + geom_boxplot() 
​
p + theme_bw()
​
​

​
