---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---




### AutoClaims

```{r}
library(readr)
AutoClaims <- read_csv("data/AutoClaims.csv")
```

---

Exercise 1
==============================================
Prepare a table with mean, standard deviation and coefficient of variance (CV) of the claims paid

for each STATE and identify the one with the least and the highest CV
for each CLASS and sort the table in the ascending order of the CV

Hint: Coefficient of Variance (CV) can be computed as the ratio of standard deviation to the mean.

---

Implementation using Base R:

```{r}
state_mean<-aggregate(PAID~STATE,data = AutoClaims,FUN = mean)
names(state_mean)<-c("State","Mean")
```

---

```{r}
state_sd<-aggregate(PAID~STATE,data = AutoClaims,FUN = sd)
names(state_sd)<-c("State","SD")

state_sd
```


---

```{r}

state_summary<-merge(state_mean,state_sd)

state_summary

```

---

```{r}


state_summary$CV <- state_summary$SD / state_summary$Mean
```

---

```{r}
#Mean, Standard Deviation and Coefficient of Variance for each state

state_summary
```

---


The State with minimum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==min(state_summary$CV)]
``` 

The State with maximum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==max(state_summary$CV)]
``` 

The State with minimum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==min(state_summary$CV)]
``` 


---

The State with maximum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==max(state_summary$CV)]
``` 

#### State Summarys (Tidyverse)

```{r}
library(dplyr)
library(magrittr)

state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(Mean = mean(PAID),
   SD   = sd(PAID))

``` 

---

```{r}
head(state_summary_tidy)
``` 


```{r}
state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(Mean = mean(PAID),
   SD   = sd(PAID)) %>%
   mutate(CV = Mean/SD) %>%
   arrange(CV)

```   

---

### Claims (Base R)

```{r}
Class_mean<-aggregate(PAID~CLASS,data = AutoClaims,FUN = mean)
names(Class_mean)<-c("Class","Mean")
Class_sd<-aggregate(PAID~CLASS,data = AutoClaims,FUN = sd)
names(Class_sd)<-c("Class","SD")
Class_summary<-merge(Class_mean,Class_sd)
Class_summary$CV<-Class_summary$SD/Class_summary$Mean
#Mean, Standard Deviation and Coefficient of variance for each rating class
Class_summary
``` 

---


```{r}
Class_summary[order(Class_summary$CV),]
``` 

```{r}
Class_summary_tidy <- AutoClaims %>% group_by(CLASS) %>%
   summarize(Mean = mean(PAID),
   SD   = sd(PAID)) %>%
   mutate(CV = Mean/SD) %>%
   arrange(CV)

```   

```{r}
Class_summary_tidy
``` 

---

```{r}
Class_mean<-aggregate(PAID~CLASS,data = AutoClaims,FUN = mean)
names(Class_mean)<-c("Class","Mean")
Class_sd<-aggregate(PAID~CLASS,data = AutoClaims,FUN = sd)
names(Class_sd)<-c("Class","SD")
Class_summary<-merge(Class_mean,Class_sd)
Class_summary$CV<-Class_summary$SD/Class_summary$Mean
#Mean, Standard Deviation and Coefficient of variance for each rating class
Class_summary


Class_summary[order(Class_summary$CV),]
``` 

---

Exercise 2
=====================
By using a box plot, identify the STATE(s) and CLASS(es) which have no outlier values in terms of the claims paid.

options(repr.plot.width=12, repr.plot.height=8)

boxplot(PAID~STATE,data = AutoClaims)


States with no outliers are State 14.

boxplot(PAID~CLASS,data = AutoClaims)
​

Class with no outlier is C11.


library(ggplot2)
​
p <- ggplot(data=AutoClaims,aes(x=STATE,y=PAID)) + geom_boxplot()
​
p

p <- ggplot(data=AutoClaims,aes(x=CLASS,y=PAID)) + geom_boxplot() 
​
p + theme_bw()
