---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

### AutoClaims

```{r}

AutoClaims <- read.csv("data/AutoClaims.csv")
```


```{r}

head(AutoClaims)
```

---

### Exercise 1

Prepare a table with mean, standard deviation and coefficient of variance (CV) of the claims paid

* for each STATE and identify the one with the least and the highest CV
* for each CLASS and sort the table in the ascending order of the CV

Hint: Coefficient of Variance (CV) can be computed as the ratio of standard deviation to the mean.

---

### Implementation using Base R:

```{r}
state_mean<-aggregate(PAID~STATE,data = AutoClaims,FUN = mean)
names(state_mean)<-c("State","Mean")
```


```{r}

head(state_mean)
```

---

### Implementation using Base R:

```{r}
state_sd<-aggregate(PAID~STATE,data = AutoClaims,FUN = sd)

names(state_sd)<-c("State","SD")

head(state_sd)
```


---

### State Summary

```{r}

state_summary<-merge(state_mean,state_sd)

state_summary

```

---

```{r}


state_summary$CV <- state_summary$SD / state_summary$Mean



```

Mean, Standard Deviation and Coefficient of Variance for each state

```{r}

state_summary
```

---

### Coefficients of Variance

The State with minimum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==min(state_summary$CV)]
``` 

The State with maximum coefficient of variance is

```{r}
state_summary$State[state_summary$CV==max(state_summary$CV)]
``` 



---



#### State Summarys (Tidyverse)

```{r echo=FALSE, include=FALSE, warning=FALSE}
library(dplyr)
library(magrittr)
```

<pre><code>
library(dplyr)
library(magrittr)
</code></pre>


### using ***group_by*** and ***summarize***
```{r}



state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(Mean = mean(PAID),
             SD = sd(PAID))

``` 

---

```{r}
head(state_summary_tidy)
``` 

---

### using ***mutate***

```{r}
state_summary_tidy <- AutoClaims %>% group_by(STATE) %>%
   summarize(
      Mean = mean(PAID),
      SD = sd(PAID)) %>%
   mutate(CV = SD/Mean) %>%
   arrange(CV)

```   


---

```{r}
head(state_summary_tidy)
``` 

---

### Class (Base R)


Same as before, but just using Class rather than States

```{r}
Class_mean<-aggregate(PAID~CLASS,data = AutoClaims,FUN = mean)

names(Class_mean)<-c("Class","Mean")

Class_sd<-aggregate(PAID~CLASS,data = AutoClaims,FUN = sd)

names(Class_sd)<-c("Class","SD")
```

---

```{r}
Class_summary<-merge(Class_mean,Class_sd)
Class_summary$CV<-Class_summary$SD/Class_summary$Mean
#Mean, Standard Deviation and Coefficient of variance for each rating class
Class_summary
``` 

---


```{r}
Class_summary[order(Class_summary$CV),]
``` 


---

```{r}
Class_summary_tidy <- AutoClaims %>% group_by(CLASS) %>%
   summarize(Mean = mean(PAID),
             SD = sd(PAID)) %>%
   mutate(CV = SD/Mean) %>%
   arrange(CV)

```   
---

```{r}
Class_summary_tidy
``` 



---

### Exercise 2

By using a box plot, identify the STATE(s) and CLASS(es) which have no outlier values in terms of the claims paid.

<pre><code>
options(repr.plot.width=12, repr.plot.height=8)

boxplot(PAID~STATE,data = AutoClaims)
</code></pre>

---

````{r echo=FALSE}
options(repr.plot.width=12, repr.plot.height=8)

boxplot(PAID~STATE,data = AutoClaims)
```


States with no outliers are State 14.
---

```{r}
boxplot(PAID~CLASS,data = AutoClaims)

```

Class with no outlier is C11.

--- 

### Exercise 2 (ggplot2)

<pre><code>

library(ggplot2)

p <- ggplot(data=AutoClaims,aes(x=STATE,y=PAID)) + geom_boxplot()

p
</code></pre>

---

```{r echo=FALSE}

library(ggplot2)

p <- ggplot(data=AutoClaims,aes(x=STATE,y=PAID)) + geom_boxplot()

p
```

---

### Exercise 2 (ggplot2)

<pre><code>

p <- ggplot(data=AutoClaims,aes(x=CLASS,y=PAID)) + geom_boxplot() 

p + theme_bw()

</code></pre>

---

```{r echo=FALSE}

library(ggplot2)

p <- ggplot(data=AutoClaims,aes(x=STATE,y=PAID)) + geom_boxplot()

p
```

---
