---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

### AutoClaims

```{r}

AutoClaims <- read.csv("data/AutoClaims.csv")
```


```{r}

head(AutoClaims)
```

---
### Exercise 3

* Prepare  a  contingency  table  identifying  the  proportion  of  insured  across  differentCLASS(es) for each gender.

* Test the null hypothesis “**There is no relationship between the  GENDER  and  the  CLASS  of  the  insured**”  against  an  alternate  hypothesis  of existence of significant relation at 95% confidence level.


* Perform an appropriate test and comment.

---

#### Contingency Table for the distribution of Gender across rating classes

```{r}

State_Class_Freq <-table(AutoClaims$GENDER,AutoClaims$CLASS) 
State_Class_Freq
```

```{r}

addmargins(State_Class_Freq)
```

---

#### Contingency Table in terms of proportion of Rating classes

(overall proportion)
```{r}

State_Class_RowProp<-prop.table(State_Class_Freq)

State_Class_RowProp 
```

```{r}

addmargins(State_Class_RowProp)
```

---

#### Contingency Table in terms of proportion of Rating classes for each gender

(proportions inside gender)

```{r}

State_Class_RowProp<-prop.table(State_Class_Freq, margin=1)

State_Class_RowProp 
```

```{r}

addmargins(State_Class_RowProp)
```


---

### Chi-Square test for independence of two variables 

```{r}
chisq.test(AutoClaims$GENDER,AutoClaims$CLASS) 
```
---

#### Interpretation: 

* Chi-Squared values is 13.704 with 3 degrees of freedom. 
* P-Value  for the chi  square test is 0.0033 < 0.05. 
* Hence the null hypothesis of independence between the gender and rating class is rejected at 95% confidence level. 

Looking at the proportions in the table, it is evident that more than 55% of the females are 
in rating class C6 whereas as only 46% of the males are in that rating class. 

Similarly 34.6% 
of the males are in C11 as against only 30% of females in that class.

Same is the case with F6. Hence the difference in proportions

---

#### Exercise 4 


Use  an  appropriate test to  test  if  the  mean  claim  amount  paid  and  variance  of  claim amounts paid is uniform between males and females.

---


### Testing for mean amount paid between Male and Female


```{r}
t.test(AutoClaims$PAID[AutoClaims$GENDER=="M"],AutoClaims$PAID[AutoClaims$GENDER=="F"]) 
```

---

* P-value = 0.28 > 0.05 
* Null hypothesis of “**Mean claim paid is same between males and 
females**” cannot be rejected at 95% confidence level 

---

#### ***var.test()***

Testing for variance of amount paid between Male and Female

```{r}
var.test(AutoClaims$PAID[AutoClaims$GENDER=="M"],AutoClaims$PAID[AutoClaims$GENDER=="F"]) 
```
---


* P-value = 0.0013 < 0.05 
*  Null hypothesis of “Variance of claim paid is same between males and females” is rejected at 95% confidence level.

* Claims paid among females is more  volatile compared to that of males

---
---