---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---



```{r}
library(readr)
AutoClaims <- read_csv("Data/AutoClaims.csv")
``` 

Your actuarial friend has suggested you to use natural logarithm of “PAID” claims instead of the actual “PAID” Claim amount because the log e (PAID) is more closer to
normal distribution than “PAID” Claims. 

--- 
#### Exercises
1. Verify the statement made by your friend by comparing the Skewness and Excess Kurtosis of both the PAID claims as well as
log e (PAID). Write appropriate custom functions to compute both of them.

2. Repeat the model in (i) above by considering the suggestion in (iii). 

3. Identify and comment on the key differences between both the models.

4. Your Manager has suggested that the model can be improved by adding interaction effects between STATE and CLASS, STATE and GENDER, CLASS and GENDER
as additional variables to the set of independent variables taken in (i). Evaluate the worthiness of this suggestion.



---


#### Exercise 1
* Reason for better model
* Checking for the normality of Auto Claims Paid vs. Logarithm of Auto Claims Paid
* Writing Functions for Skewness and Kurtosis

```{r}

skew<-function(x)mean((x-mean(x))^3)/sd(x)^3
kurt<-function(x)(mean((x-mean(x))^4)/sd(x)^4)-3

``` 


---

```{r}
skew(AutoClaims$PAID)
```

```{r}

skew(log(AutoClaims$PAID))
``` 

#### Remark

* The closer to Zero, the better


---

```{r}

kurt(AutoClaims$PAID)
``` 




```{r}

kurt(log(AutoClaims$PAID))
``` 

#### Remark

* The closer to Zero, the better

---


#### Conclusion

Skewness and Kurtosis of Log (Claims) are more close to Zero compared to those of actual
claims paid, thus indicating the possibility of using linear regression with this dependent
variable

---

#### Model with untransformed "PAID" variable


```{r}

model1 <- lm(PAID~.,data = AutoClaims)
summary(model1)
```

---

Using Natural Logarithm of the claims paid

```{r}

model2 <- lm(log(PAID)~.,data = AutoClaims)
summary(model2)
```


---


```{r}

anova(model2)
```

---


#### Key Differences

1. R-Squared and Adjusted R-Squared improved fit compared to the initial model and hence the model is a better
2. While a the significance level of a few factor coefficients when compared with the base categories changed, the overall significant variables did not change which can be inferred from the ANOVA table

---

#### Exercise 3

```{r eval=FALSE}
# Using Interaction effects in the model
model3<-lm(PAID~.+STATE:CLASS+STATE:GENDER+CLASS:GENDER,data = AutoClaims)
summary(model3)
```



---

```{r echo=FALSE}
# Using Interaction effects in the model
model3<-lm(PAID~.+STATE:CLASS+STATE:GENDER+CLASS:GENDER,data = AutoClaims)
summary(model3)
```


---
#### Exercise 2
Interaction effect between a few classes and states emerged out to be very
significant (State 6 and Class F6 came out to be significantly negative).

* Though State 15 came out to be significantly negative when main effects alone were considered,
the interaction effects compensated that negative significantly when interacted with class C6 and Class 11 whereas the interaction coefficient is not significant between State 15 and Class F6 indicating that the claims paid is significantly lesser when the
state is 15 and class is F6 compared to other rating classes. 
* Digging deeper into the relationships is possible with the interaction effect. 
* Similarly main effect of Gender is significantly negative compared to females but that is offset to some extent for some
states (2,3,7,15) and for some rating classes (C6) whereas it is further negative in case of F6. 
* So the differences can be magnified by considering the interaction effects, improving the predictability of the model


---

#### Exercise 3
ANOVA table for the model suggests that except the age all the main effects and
their interaction effects are significant at 5% significance level
indicating their
contribution to the predictability of the model

---

---

Part (i) and Part (ii) were attempted by maximum number of students. Many of them
were successful in writing the code to fit a linear regression model but only successful
students were able to provide a good interpretation of the results. Likewise many
students were able to write the code for plotting but very few of them ended up in
writing the interpretations based on the graphs. 

---

Part (iii) was not attempted by many
students and those who attempted also ended up in writing wrong functions. Overall,
this question was very poorly done. 

Part (iv) was attempted well, code was written well
but a few ended up in writing correct interpretations. 

Part (v) was attempted by a few
students only and many of them failed to provide appropriate interpretation.
---

---