---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


Part 1
===========================
The following amounts are the sizes of claims (in £) on house insurance policies for a 
certain type of repair. 

1990, 2400, 2150, 2090, 2300, 
2100, 2180, 2150, 2030, 2100, 
2180, 2010, 2060, 2160, 2120 




---

Part 1
==============

Enter data in R. 

```{r} 
mydata <- c(1990,2400,2150,2090,2300,
           2100,2180,2150,2030,2100,
           2180,2010,2060,2160,2120)
```

---

Part 1
==============

 Calculate Q1, Q2, Q3 and Inter-quartile range. (4)

```{r}
quantile(mydata) 

```

```{r}
fivenum(mydata)
```


```{r}
IQR(mydata) 
```

```{r}
# Q3 - Q1
2170 - 2075
```





---

Part 3
=======================

Determine the sample mean and variance of the data. 

```{r}
mean(mydata) 

var(mydata) 

sd(mydata)

```

*** Save For Later ***

```{r}
xbar <- mean(mydata) 
Ssquare <- var(mydata)
```

---


Test the hypothesis whether the mean claim amount is equal to £2000 and comment 
on the results. 

<h4>Hypotheses</h4>

Ho: The mean claim amount is £ 2000 
H1: Mean claim amount is not equal to £ 2000 

```{r}
t.test(mydata,mean=2000) 
```

Since the p-value is $2.2\times 10^{-16}$ is less than 5%, there is sufficient evidence to reject $H_o$ of mean equal to £ 2000. 

---

Part 5
===============================

Assuming the data to be normally distributed, calculate the probability of a claim amount 
exceeding £2300. 
 

 
 
```{r}
pnorm(2300,mean = xbar , sd = sqrt(Ssquare) ,lower.tail = FALSE) 
```

---

Part 6
========================

Calculate the revised mean and median after removing the largest two values from the 
dataset. Comment on the result.
 
```{r}
max(mydata)

sort(mydata)
```

```{r}
rev(sort(mydata))

```


```{r}
rev(sort(mydata))[1:2]

```

```{r}
mydata1 = mydata[mydata < max(mydata)] 
```



```{r}
mydata1 = mydata[mydata < max(mydata)] 
mydata1 
```


```{r}
mydata2 = mydata1[mydata1 < max(mydata1)] 
mydata2
```

```{r}
mean(mydata2) 
median(mydata2) 
```


Impact of removing outliers from the data has led to the mean and median being almost equal. Earlier the 
mean was higher than median, which shows that the mean is more likely to be affected by outliers. 
