---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Probability Distributions
* Fundamentals of Statistical Analysis

* Question 2

---

* The prior and posterior distribution for values of systolic blood pressure follows Normal 
distribution. 

* Prior distribution of systolic blood pressure (x) has a mean of 120 and standard 
deviation of 10. 

```{r}
priormean <- 120 
priorsd <- 10 
```


---

### Part 1


Generate range of values of x in the interval [80,160] and using `len = 100`. 


```{r}
x <- seq(80,160,len = 100) 
```


<h4> Inspecting the data </h4>

```{r}
head(x)

max(x)
```


---

### Part 2

Plot the posterior probability density function of x using answer from Part 1. 


```{r eval=FLASE}
y<-dnorm(x,mean = priormean,sd=priorsd)
plot(x,y,type = "l")
```

---

```{r echo=FLASE}
y<-dnorm(x,mean = priormean,sd=priorsd)
plot(x,y,type = "l")
```

---

### Part 3


Perform a simulation of 1000 posterior samples for the parameter x.      



```{r}
z <- rnorm(1000,priormean,priorsd)

head(z)
```

---

### Part 4


Plot a histogram of the posterior distribution of x. 


```{r echo=FALSE} 
hist(z,main = "Posterior distribution of x")
```
---


```{r echo=FALSE} 
hist(z,main = "Posterior distribution of x")
```

---

#### Part 5


What is the mean and standard deviation of posterior distribution of x?      


```{r}
mean(z) 

sd(z) 

```

---

### Part 6
#

Calculate the 95% confidence interval for systolic blood pressure using the posterior 
distribution. 


```{r}
sbp <- mean(z) + qnorm(c(0.025,0.975)) * sd(z) 
sbp
```

---
 
 
--- 
