---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3
---

An agency has collected data on the number of COVID19 cases of two cities in order to 
analyse the similarities & differences between them. Below is the data for two cities on 
monthly basis. 



| Month |  City A |  City B |  
|----- |---- |------ |
| 1 |  9150 |  8919 |  
| 2 |  9418 |  9095 |  
| 3 |  9218 |  9046 |  
| 4 |  9539 |  9321 |  
| 5 |  9179 |  9719 |  
| 6 |  8907 |  9704 |  
| 7 |  9472 |  9107 |  
| 8 |  8921 |  9275 |  


---

Part 1
====================
Enter data in R. 


```{r}
city1 <- c(9150,9418,9218,9539,9179,8907,9472,8921) 

city2 <- c(8919,9095,9046,9321,9719,9704,9107,9275)
```

---

Part 2
====================
Test at 5% level with clearly mentioning the hypothesis, if there is a difference in the 
mean of the two sample data assuming equal & unknown variances. 

* ***Ho:*** There is no difference in the average number of monthly COVID19 cases between two cities. 
* ***H1:*** There is a difference in the average number of monthly COVID19 cases between two cities. 

```{r}
t.test(x=city1,y=city2,var.equal = TRUE,conf.level = 0.95) 
```

---

* Since the p-value is 0.7289 is significantly greater than 5%, there is insufficient evidence to reject Ho. 

* Thus, we have no evidence to suggest that the means are different between the two samples. 

---

Suppose we dont assume equal & unknown variances

```{r}
t.test(x=city1, y=city2, conf.level = 0.95) 
```

---
Part 3
====================

Test whether the variances are equal at 5% level and comment on the results. 


```{r}
var.test(x=city1,y=city2,conf.level = 0.95) 

```

The p-value is 0.5691 > 5%, so there is insufficient evidence to reject the assumption of equal variance. 

---

Part 4
====================
Calculate the 95% confidence interval for the difference in means. 



Confidence interval can be read from Part 2 or can be derived as below: 
```{r}
t.test(x=city1,y=city2,var.equal = TRUE,conf.level = 0.95)$conf.int 
```
i.e. 95% CI is (-337.39, 241.89) 

---

Part 5
====================

Comment on your findings in part 2 and part 4. 

The confidence interval (-337,241) contains 0, therefore the assumption of equal means holds. 
