---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Probability Distributions
* Fundamentals of Statistical Analysis

* Question 4

---

Datasets
=================

* Five years of marketing expenditure and company sales by month 

```{r}
budget = read.csv("marketingbudget.csv")
```


---

```{r}
head(budget)
```

---

Part 1
=================

Five years of marketing expenditure and company sales by month 


Construct a scatterplot of the data. Comment on the relationship between the Sales & 
Spend based on the plot. 

<pre><code>
plot(budget$Spend,budget$Sales)
</code></pre>

---

```{r, echo =FALSE} 
plot(budget$Spend,budget$Sales)
```

The above scatter plot shows a positive linear relationship between marketing Spend and Sales data. 


---

```{r} 
plot(budget$Spend,budget$Sales, pch=18, col="blue")
```

---

Part 2
=================

Calculate Pearson’s correlation coefficient between Sales and Spend of the company. 


```{r} 
cor(budget$Sales,budget$Spend) 

```

---

Part 3
=====================================

Perform a hypothesis test for the null hypothesis that Pearson’s population correlation 
coefficient is equal to zero, against the alternative that it is positive. You should report the 
p-value of the test and a clear conclusion. 


<pre><code>
cor.test(budget$Spend,budget$Sales,
         method="pearson",
         alternative = "greater")
</code></pre>

---

```{r}
cor.test(budget$Spend,budget$Sales,
         method="pearson",
         alternative = "greater")
```

* The p-value is $2.2 \times 10^{-16}$, showing very strong evidence against the null hypothesis. 

* Thus, we reject that the Pearson’s correlation coefficient is equal to 0 and conclude that it is positive. 

---

Part 4
=====================================

Perform a simple linear regression analysis on the data. 

Your answer should report the estimate of parameter sigma - The Residual standard error. 

```{r include = FALSE}
reg = lm(Sales ~ Spend, data = budget) 
summary(reg) 
```

---

```{r}
reg = lm(Sales ~ Spend, data = budget) 
summary(reg) 
```

From the output, the estimate of parameter sigma is 10,650. 

---

Part 5
====================================

Plot the fitted line on the data scatterplot. 

<pre><code>
plot(budget$Spend,budget$Sales)
abline(reg) 
</code></pre>



---

```{r}
plot(budget$Spend,budget$Sales)
abline(reg) 
```


---

Part 6
====================================
State the proportion of the total variability of the responses explained by the model based on your output in Part 4. 

From the R output, the proportion of total variability of the responses explained by the model is 94.12%. 

---

Part 7
====================================
Plot a graph of the residuals of the model fitted in Part 6 against the explanatory variable. 


```{r} 
plot(budget$Spend,residuals(reg))
```

---

Part 8
====================================

Obtain a 99% confidence interval for parameter sigma. (4)

```{r}
es = resid(reg)
```


From the above, the confidence interval for parameter sigma is (-3630.15, 3630.15) 

---

Part 9
============================
Comment on the validity of the model based on results in part (vii) and part (viii). 


ix)
Based on the results in both part (vii) and part (viii), the errors seem to be close to zero and the 
confidence interval of residuals also contains 0. Hence the model seems to be a good fit. 

---


Part 10
============================

Calculate the p-value of a hypothesis test for this suggestion (slope equal to 10), by 
creating a suitable test statistic. 

Let Ho: Beta = 10 and H1: Beta not equal to 10 

```{r}
b1 = (coef(reg))[['Spend']]
n = 60 
s = sqrt(sum(es^2)/(n-2))
```

---

```{r}
SE = s/sqrt(sum((budget$Spend-mean(budget$Spend))^2))
t = (b1-10)/SE 

#p-value
pt(t,58,lower.tail = FALSE)
```

```{r}
pvalue = 2*pt(t,58,lower.tail = FALSE)
pvalue 
```

---

Part 11
==================================

Comment on the suggestion in Part 11. 

* There is insufficient evidence to reject the null hypothesis at 5% level of significance. 
* The slope can be assumed to be equal to 10 for this data. 

---


Part 12
==================================
Calculate the predicted amount of sales when the marketing spend is £4500. 


```{r} 
y = 3431.5592 + (b1*4500) 
y 

```

With a marketing spend of $4,500, the Sales would be $50,821. 

---