---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

Q. 4)
The table below shows the total claim number (cancellations) per year, Xij, for 4 travel companies over last 4 years.
Years,j
1
2
3
4
Travel Companies, i
Make
455
458
587
531
Ease
251
322
292
340
Go
309
246
217
120
Clear
400
426
470
547
i) Using Empirical Bayes Credibility Theory (EBCT) Model 1, compute the following
a) Copy the below code to load the data:
q4<-matrix(c(455,251,309,400,
458,322,246,426,
587,292,217,470,
531,340,120,547),
ncol=4,nrow=4)
(1)
b) E[m(θ)]
(2)
c) E[s2(θ)]
(2)
d) Var[m(θ)]
(3)
e) Z
(2)

ii) Using part (i), calculate the expected claim number for Go and Clear.

iii) What additional information is required to use EBCT Model 2.

---

iv) Travel company “Ease” launched a membership program last year providing full refund on cancellations. Number of cancellations believed to follow binomial distribution with parameters n=3 and 0.20.
Number of cancellations in last year on 150 memberships are as follows:
Cancellations
0
1
2
3
Members
61
71
15
3
Carry out goodness of fit test for the binomial model specified for number of cancellations on each membership.
(5)

---


Empirical Bayes Credibility Theory (EBCT) Model 1

The Empirical Bayes Credibility Theory (EBCT) Model 1, formulated by Hans Bühlmann in 1967, is a foundational model in actuarial science. It is used to estimate the credibility premium, which is a weighted average of the individual risk's past experience and the overall average experience of a group of similar risks³.

### Key Features of EBCT Model 1:
1. **Assumptions**:
   - The model assumes that the observed data for each risk follows a certain distribution.
   - The parameters of this distribution are considered random variables with their own prior distribution.

2. **Credibility Factor (Z)**:
   - The credibility factor \( Z \) determines the weight given to the individual risk's past experience versus the overall average.
   - It ranges between 0 and 1, where a higher \( Z \) indicates more reliance on the individual risk's data.

3. **Calculation**:
   - The credibility premium is calculated as:
     $$
     \text{Credibility Premium} = Z \cdot \text{Individual Risk's Mean} + (1 - Z) \cdot \text{Overall Mean}
     $$
   - This formula balances the specific risk's historical data with the collective experience of the group.

4. **Applications**:
   - EBCT Model 1 is widely used in insurance to set premiums based on past claims data.
   - It helps in making more accurate predictions by incorporating both individual and collective risk information.

Would you like to know more about how this model is applied in practice or any specific examples?



[20 Marks]
Solution 4:
i)

```{r}
q4<-matrix(c(455,251,309,400,
458,322,246,426,
587,292,217,470,
531,340,120,547),
ncol=4,nrow=4)
```


```{r}
q4
```

```{r}
n<-ncol(q4)
n
```


```{r}
m<-mean(rowMeans(q4))
m
```


```{r}
s<-mean(apply(q4,1,var))
s
```


```{r}
v<-var(rowMeans(q4)) - mean(apply(q4,1,var))/n
Z<- n/(n+s/v)
n
```


Z
[1] 0.9443816
[1+2+2+3+2]
ii)
Z*rowMeans(q4[3:4,])+(1-Z)*m
[1] 231.3532 455.8799


---

iii)
Risk Volumes are required to apply EBCT2


---

iv)

```{r}
obs<-c(61,71,15,3)
```


* Combine 2 and 3 since value of 3 is less than 5

```{r}
obs.comb<-c(61,71,15 + 3)
```


```{r}
p<-0.2
exp<-dbinom(c(0:1),3,p)
exp[3]<-1-pbinom(1,3,p)
sum(exp)
```



---

```{r}
chisq.test(x=obs.comb,p=exp)
```


* Since p-value <0.5, we have sufficient evidence to reject the null hypothesis that cancellation follows bin(3,0.2)

---

Source: Conversation with Copilot, 5/10/2024
(1) MTH5131 Actuarial Statistics - Queen Mary University of London. https://qmplus.qmul.ac.uk/mod/resource/view.php?id=2522684.
(2) Risk Modelling in General Insurance - Cambridge University Press .... https://assets.cambridge.org/97805218/63940/index/9780521863940_index.pdf.
(3) 5 Credibility Theory - Lecture notes 1 - 5 Theory (Unit 6) 5 ... - Studocu. https://www.studocu.com/in/document/universal-business-school/statistics/5-credibility-theory-lecture-notes-1/16670923.
(4) Credibility Premium Estimation of Insurance Claims in Nigeria - IDEAS/RePEc. https://ideas.repec.org/p/hal/wpaper/hal-01855972.html.


