---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

Q. 4)
The table below shows the total claim number (cancellations) per year, Xij, for 4 travel companies over last 4 years.
Years,j
1
2
3
4
Travel Companies, i
Make
455
458
587
531
Ease
251
322
292
340
Go
309
246
217
120
Clear
400
426
470
547
i) Using Empirical Bayes Credibility Theory (EBCT) Model 1, compute the following
a) Copy the below code to load the data:
q4<-matrix(c(455,251,309,400,
458,322,246,426,
587,292,217,470,
531,340,120,547),
ncol=4,nrow=4)
(1)
b) E[m(θ)]
(2)
c) E[s2(θ)]
(2)
d) Var[m(θ)]
(3)
e) Z
(2)

ii) Using part (i), calculate the expected claim number for Go and Clear.

iii) What additional information is required to use EBCT Model 2.

---

iv) Travel company “Ease” launched a membership program last year providing full refund on cancellations. Number of cancellations believed to follow binomial distribution with parameters n=3 and 0.20.
Number of cancellations in last year on 150 memberships are as follows:
Cancellations
0
1
2
3
Members
61
71
15
3
Carry out goodness of fit test for the binomial model specified for number of cancellations on each membership.
(5)

---

[20 Marks]
Solution 4:
i)

```{r}
q4<-matrix(c(455,251,309,400,
458,322,246,426,
587,292,217,470,
531,340,120,547),
ncol=4,nrow=4)
```


```{r}
q4
```

```{r}
n<-ncol(q4)
n
```


```{r}
m<-mean(rowMeans(q4))
m
```


```{r}
s<-mean(apply(q4,1,var))
s
```


```{r}
v<-var(rowMeans(q4)) - mean(apply(q4,1,var))/n
Z<- n/(n+s/v)
n
```


Z
[1] 0.9443816
[1+2+2+3+2]
ii)
Z*rowMeans(q4[3:4,])+(1-Z)*m
[1] 231.3532 455.8799


---

iii)
Risk Volumes are required to apply EBCT2


---

iv)

```{r}
obs<-c(61,71,15,3)
```


* Combine 2 and 3 since value of 3 is less than 5

```{r}
obs.comb<-c(61,71,15 + 3)
```


```{r}
p<-0.2
exp<-dbinom(c(0:1),3,p)
exp[3]<-1-pbinom(1,3,p)
sum(exp)
```



---

```{r}
chisq.test(x=obs.comb,p=exp)
```


* Since p-value <0.5, we have sufficient evidence to reject the null hypothesis that cancellation follows bin(3,0.2)

---