---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---
Q. 1)
i) Hotel room prices in Mumbai are normally distributed with a mean of Rs. 5400 and standard deviation of Rs. 900, whereas prices in Kolkata are normally distributed with a mean of Rs. 3600 and standard deviation of Rs. 1500. Compute the probability that a hotel room price in Mumbai is atleast twice the price in Kolkata. Note that the hotel prices in Mumbai and Kolkata are independent of each other.
Store mean and sd values as x.mean, x.sd, y.mean,y.sd,etc.
Hint: p**** is a useful function to ascertain CDF.
(5)
ii)
a) Show that the population mean and standard deviation of “differences in hotel prices between Mumbai and Kolkata” (i.e., Mumbai hotel prices – Kolkata hotel Prices) are 1800 and 1749.86.
(3)
b) Generate a sample of size 50 for the ” differences in hotel prices between Mumbai and Kolkata”.
(2)
c) Draw qqplot and qqline for the sample generated and comment on the results.
(4)
Store population mean and standard deviation of differences as dif.mean and dif.sd
Make sure to set the seed value to 1234 i.e use set.seed(1234) before generating the sample.
Store the sample as dif.sample
Don’t paste the sample

---

iii) Test using 5% significance level whether the mean of “difference in hotel prices” is less than 1375 based on the sample generated in part (ii).
a) Perform z-test and comment.
(5)
b) Perform t-test (assuming population is not known) and comment.
(3)
Please ensure to include test conclusion.
iv)
a) Generate another sample but with size=1000 and store as dif.sample2.
(1)
b) Draw qqplot and qqline for the sample generated in (iv.a) and comment on the results.
(3)
[26]


Page 2 of 8
Solution 1:
i)

```{r}
#X~N(5400,900^2)
x.mean=5400
x.sd=900
#Y~N(3600,1500^2)
y.mean= 3600
y.sd=1500
#P(X>2Y) = P(X-2Y>0)
# X-2Y ~ N(x.mean - 2y.mean, x.sd^2+4y.sd^2)
```


---

```{r}
f.mean<-x.mean - 2*y.mean
f.var <- x.sd^2+4*y.sd^2
>
1-pnorm(0,mean=f.mean,sd=sqrt(f.var))
[1] 0.2827485
```

----
ii)
a)
b)
dif.mean<-x.mean - y.mean
dif.mean
[1] 1800
>
dif.sd <-sqrt(x.sd^2+y.sd^2)
dif.sd
[1] 1749.286

---

set.seed(1234)
dif.sample<-rnorm(50,dif.mean,dif.sd)
[3]


--- 

c)
qnorm(dif.sample)
qqline(dif.sample)

---

#### Conclusion 

Above results indicates that despite sample generated from normal distribution it doesn’t seem to

align with normality probably due to low sample size and high variability.

---

iii)
a)
sample.mean<-mean(dif.sample)
sample.mean
[1] 1007.481
z<- (sample.mean - 1375)/(dif.sd/sqrt(50))
pnorm(z)
[1] 0.06869143

---

Since p-value is >0.05 we cannot reject the null hypothesis and thus, don’t have sufficient evidence to say that mean is less than 1375.

---

b)

```{r}
t.test(dif.sample, mu=1375, alternative = "less")
```

---

One Sample t-test
data: dif.sample
t = -1.6786, df = 49, p-value = 0.0498
alternative hypothesis: true mean is less than 1375
95 percent confidence interval:
-Inf 1374.558
sample estimates:
mean of x
1007.481


* Since p-value is < 0.05 we can reject the null hypothesis and can imply that mean is less than 1375.

---

iv)
a)
b)

```{r}
dif.sample2<-rnorm(1000,dif.mean,dif.sd)
qqnorm(dif.sample2)
qqline(dif.sample2)
```

With larger sample size, it indicates normality.

---