---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

Q. 3)
An analyst fitted various models on the data containing claims information of 20 policies and shared the output to conduct analysis.
Output 1:
Call:
glm(formula = Claim ~ 1, family = poisson(lin = "log"), data = q3)
Deviance Residuals:
Min 1Q Median 3Q Max
-1.8439 -0.8975 -0.1791 0.3925 2.5561
Coefficients:
Estimate Std. Error z value Pr(>|z|)
(Intercept) 0.5306 0.1715 3.094 0.00197 **
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
Null deviance: 30.147 on 19 degrees of freedom
Residual deviance: 30.147 on 19 degrees of freedom
AIC: 71.114
Number of Fisher Scoring iterations: 5

---

i) Write the equation of the fitted model using Output 1 as above and specify which distribution is used to model the response variable.
(2)
ii) Using output 1, show that the sample mean of the response variable is 1.7 (when rounded to one decimal place).
(2)
Output 2:
Call:
glm(formula = Claim ~ Gender + Health - 1, family = poisson(lin = "log"),
data = q3)
Deviance Residuals:
Min 1Q Median 3Q Max
-1.22474 -0.81754 -0.07119 0.27453 1.44149
Coefficients:
Estimate Std. Error z value Pr(>|z|)
GenderF 1.1394 0.3490 3.265 0.001096 **
GenderM 1.1394 0.2216 5.143 2.71e-07 ***
HealthNonDiabetic -1.4271 0.3939 -3.623 0.000291 ***
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
Null deviance: 38.229 on 20 degrees of freedom
Residual deviance: 14.436 on 17 degrees of freedom
AIC: 59.403
Number of Fisher Scoring iterations: 5

---

iii) Write the equation of the fitted model using Output 2.
Make sure to define the explanatory (categorical) variables.
(4)
iv) Output 2 doesn’t have an intercept as a coefficient unlike output1. Please provide the reason for the same by comparing glm R formulas.
(2)
v) Compare output 1 and output 2, and suggest which model is better using suitable statistics.
(2)
vi) Compute log likelihood of the model given in output 2.
(3)
vii) Claim follows Poisson distribution. Test for the parameter to be equal to 1.5 at 1% level of significance.
Hint: Use outputs and various sub-parts to determine x for poisson test.
(5)
[20]


Solution 3:
i)
```{r}
log(y) = 0.5306
```

poisson distribution is used to model response variable.

ii)
```{r}
Claim.mean<-round(exp(.5306),1)
Claim.mean
```

[1] 1.7

---

### Part C

iii)

```{r}
log(y) = 1.1394 * GenderF + 1.1394 * GenderM -1.4271 *HealthNonDiabetic
```

where GenderF =1 if Gender = F else 0
GenderM =1 if Gender = M else 0
HealthNonDiabetic =1 if Health= NonDiabetic else 0

--- 

### Part D

iv)

* For model 2, “-1” is used in Glm R formula to not take the intercepts while fitting the model.
* Thus, no intercept exits.


---


v)
AIC (=59.4) of model 2 which is lower than AIC (71.1) of model 1. This indicates that model 2 is better.

---

### Part F
vi)
#AIC = - 2LogL(Model) + 2*Parameters
#LogL(Model) = Parameters - AIC/2
aic<-59.403
par<-3
>
L<- par - aic/2
L
[1] -26.7015

--- 

vii)
#Total claims = Mean claims * Total policies
x<-Claim.mean*20
poisson.test(x=X,T=20,r=1.5,conf.level = 0.99)
Exact Poisson test

data: X time base: 20
number of events = 34, time base = 20, p-value = 0.4639
alternative hypothesis: true event rate is not equal to 1.5
99 percent confidence interval:
1.042837 2.605372
sample estimates:
event rate
1.7
>
# We cannot reject the null hypothesis that parameter is equal to 1.5.
[Max 5]

---