---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}

/* Style for <pre> tags, commonly used for code blocks */
pre {
  background-color:  #ADD8E6;  /* Light grey background */
  padding: 10px;              /* Some padding for spacing */
  border-radius: 5px;         /* Rounded corners */
}

/* Style for <code> tags, commonly used for inline code */
code {
  background-color: #ADD8E6; /* Slightly darker grey for inline code */
  padding: 2px 4px;           /* Padding for inline code */
  border-radius: 3px;         /* Slightly rounded corners */
}
```

### The Beta distribution

* The Beta distribution is a continuous probability distribution defined on the interval $[0, 1]$ and is parameterized by two shape parameters $a$ and $b$ (often denoted as $\alpha$ and $\beta$ in some literature). 

* It is often used to model random variables that represent proportions or probabilities. 

---

### Exercise 

A random variable X with values in the interval [0,1] has the following Cumulative
Distribution Function (CDF):
$$F(X) = 1 - (1-x^a)^b, x \in[0,1]$$
and inverse CDF:


$$F^{-1}(u) = (1 - (1-u)^{\frac{1}{b}})^{\frac{1}{a}}$$

where $a > 0$ and $b > 0$ are parameters. 

The Probability Density Function (PDF) is then given by:

$$f(x) = abx^{a-1}(1-x^a)^{b-1}, x \in[0,1]$$

You do not need to check the form of the inverse CDF or PDF.

---

### Exercise

A. Plot the inverse CDF $F^{-1}(u)$ for an appropriate range of values of $u$ when $a = 0.7$ and $b = 0.5$. 

B. Simulate 1,000 values of $X$ for $a = 0.7$ and $b = 0.5$ using the inverse transform method and store the 1,000 simulated values in a vector in <tt>R</tt> for later use.

You must use the command <tt>set.seed(123)</tt> to initialise the random
number generator before you start the simulation. 

C. Plot a histogram of the 1,000 simulated values obtained in exercise B using
relative frequencies. 

D. Plot the PDF $f(x)$ of $X$ for $a = 0.7$ and $b = 0.5$ for values of $x \in [0,1]$. 

<i>The graph of the PDF should be superimposed on the histogram produced in ***Exercise C***. </i>


---

### Exercises 

E.  Comment on the plot in ***Exercise D*** comparing the simulated values of $X$ with the
PDF, also taking into account the size of the simulated sample. 

F. Plot the PDF $f(x)$ of $X$ for $a = 0.7$ and $b = 0.5$ for values of $x \in [0,1]$ in a new
plot, and add the PDF $f(x)$ for $a = 3$ and $b = 2$ to the same plot. 

G.  A statistician wants to use one of the PDFs in ***Exercise F*** as a prior density for a
Bayesian analysis.

Comment on the prior beliefs about the underlying parameter that the two
PDFs represent. 

---


### Exercise A.

**Plot the inverse CDF**

```{r eval=FALSE}
a <- 0.7
b <- 0.5

u <- seq(0, 1, length.out = 100)

inv_cdf <- (1 - (1-u)^(1/b))^(1/a)

plot(u, inv_cdf, type = "l", 
     xlab = "u", 
     ylab = "F^-1(u)", 
     main = "Inverse CDF")
```

---

### Exercise A.

```{r echo=FALSE}
a <- 0.7
b <- 0.5

u <- seq(0, 1, length.out = 100)
inv_cdf <- (1 - (1-u)^(1/b))^(1/a)
plot(u, inv_cdf, type = "l", xlab = "u", ylab = "F^-1(u)", main = "Inverse CDF")
```

---

### Exercise B.

**Simulate 1000 values**

```{r}
set.seed(123)
simulated_values <- (1 - runif(1000)^(1/b))^(1/a)
```

```{r}
#First 10 values

simulated_values[1:10]
```

---

### Exercise C.

**Plot histogram**
```{r eval=FALSE}

hist(simulated_values, prob = TRUE, 
     main = "Histogram of Simulated Values")
```

---

### Exercise C.

**Plot histogram**
```{r echo=FALSE,fig.width=10}

hist(simulated_values, prob = TRUE, 
     main = "Histogram of Simulated Values")
```


---

### Exercise D.

**Plot PDF and histogram**

```{r eval = FALSE}

hist(simulated_values, prob = TRUE, 
     main = "Histogram of Simulated Values")

x <- seq(0, 1, length.out = 100)

pdf <- a*b*x^(a-1)*(1-x^a)^(b-1)

lines(x, pdf, col = "red")
```

---

### Exercise D.

**Plot PDF and histogram**

```{r echo = FALSE}

hist(simulated_values, prob = TRUE, 
     main = "Histogram of Simulated Values")

x <- seq(0, 1, length.out = 100)
pdf <- a*b*x^(a-1)*(1-x^a)^(b-1)
lines(x, pdf, col = "red")
```

---

### Exercise E.

**Comment on the Plot**

* The histogram and the superimposed PDF show a good match. This indicates that the simulated values are indeed drawn from the specified distribution. 

* The simulated distribution closely follows the theoretical distribution, especially in the central region. 

* However, there are some minor discrepancies, particularly in the tails, which are likely due to the finite sample size of 1000.

* As the sample size increases, the histogram would converge more closely to the theoretical PDF, reducing the variability in the tails.



---


### Exercise F. 

**Plot two PDFs**

```{r eval=FALSE}

pdf1 <- a*b*x^(a-1)*(1-x^a)^(b-1)

pdf2 <- 3*2*x^(3-1)*(1-x^3)^(2-1)

plot(x, pdf1, type = "l", ylim = c(0, 3), 
     main = "PDFs for Different Parameter Values")

lines(x, pdf2, col = "red")
```

---



### Exercise F. 

**Plot two PDFs**

```{r echo=FALSE}

pdf1 <- a*b*x^(a-1)*(1-x^a)^(b-1)

pdf2 <- 3*2*x^(3-1)*(1-x^3)^(2-1)

plot(x, pdf1, type = "l", ylim = c(0, 3), 
     main = "PDFs for Different Parameter Values")

lines(x, pdf2, col = "red")
```

---


### Exercise G.

* The PDF with $a = 0.7$ and $b = 0.5$ is more uniform, suggesting a less informative prior. It is more concentrated towards 1, indicating a prior belief that the underlying parameter (i.e. proportion) is likely to be high. 

* The PDF with a = 3 and b = 2 peaks in the interval between 0.6 and 0.8, suggesting a more informative prior, i.e. a prior that places more weight on high values of the parameter.

---


**Key Points:**

- The inverse CDF is used to generate random numbers from the distribution.

- The histogram visualizes the empirical distribution of the simulated data.

- The PDF is the theoretical probability density function.

- The two PDFs in ***Exercise F*** represent different prior beliefs about the parameter.


---



**Part (iv): Superimposing PDF on Histogram**

```{r echo=FALSE}
# ... (previous code)

# Plot histogram with superimposed PDF
hist(simulated_values, prob = TRUE, main = "Histogram and PDF")
lines(x, pdf, col = "red", lwd = 2)
```




**Understanding the Problem**

We're tasked with:
1. **Visualizing the Inverse CDF:** Plotting the inverse CDF for specific parameter values.
2. **Simulating Data:** Generating random numbers from the distribution using the inverse transform method.
3. **Visualizing the Distribution:** Plotting a histogram of the simulated data and overlaying the theoretical PDF.
4. **Comparing PDFs:** Plotting two PDFs with different parameter values and interpreting their implications for Bayesian analysis.

---


---


### Exercise E.

**Comment**

* The histogram closely matches the theoretical PDF, indicating that the inverse transform method effectively generates samples from the desired distribution.

* However, with a sample size of 1000, there's still some variability in the histogram, especially in the tails.

---
**R Code Implementation**

```R
# (i) Plot the inverse CDF
a <- 0.7
b <- 0.5

u <- seq(0, 1, length.out = 100)
inv_cdf <- (1 - (1-u)^(1/b))^(1/a)
plot(u, inv_cdf, type = "l", xlab = "u", ylab = "F^-1(u)", main = "Inverse CDF")

---

# (ii) Simulate 1000 values
set.seed(123)
simulated_values <- (1 - runif(1000)^(1/b))^(1/a)


# (iii) Plot histogram
hist(simulated_values, prob = TRUE, main = "Histogram of Simulated Values")

# (iv) Plot PDF and histogram
x <- seq(0, 1, length.out = 100)
pdf <- a*b*x^(a-1)*(1-x^a)^(b-1)
lines(x, pdf, col = "red")

---

# (v) Comment:
# The histogram closely matches the theoretical PDF, indicating that the inverse transform method effectively generates samples from the desired distribution.
# However, with a sample size of 1000, there's still some variability in the histogram, especially in the tails.

# (vi) Plot two PDFs
pdf1 <- a*b*x^(a-1)*(1-x^a)^(b-1)
pdf2 <- 3*2*x^(3-1)*(1-x^3)^(2-1)
plot(x, pdf1, type = "l", ylim = c(0, 3), main = "PDFs for Different Parameter Values")
lines(x, pdf2, col = "red")


