---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

###  Exercises

A. Simulate 10,000 values from a N(0,1) distribution using an appropriate R 
command. 
<br>
<i>You should save the generated values in R for later use.</i> 

B Simulate 10,000 values from a χ2(df=4) distribution using an appropriate R 
command. <br>
<i>You should save the generated values in R for later use.</i> 

C. Discuss the main disadvantage of using the inverse transform method for 
simulating values from a chi-square distribution. 

D. Simulate 10,000 values from a $t_{4}$ distribution using your answers to exercise A 
and B, explaining your answer. <br>
<i>You should save the generated values in R for later use.</i>   

E. Simulate 10,000 values from a $t_{20}$ distribution by modifying your R code from 
exercises A, B and D where necessary. 
<br>
<i>You should save the generated values in R for later use.</i>  

F. Demonstrate an important property of the $t$ distribution by presenting and 
comparing two appropriate plots of the values simulated in exercise E and F. 

---

**1. Simulation and Analysis**

**1(i) Simulating N(0,1) values:**

```{r}
set.seed(123)  # For reproducibility
n <- 10000
z <- rnorm(n, mean = 0, sd = 1)
```

**1(ii) Simulating χ²₄ values:**

```{r}
chi_squared <- rchisq(n, df = 4)
```


---

**1(iii) Disadvantage of Inverse Transform Method for χ²:**

The inverse transform method requires finding the inverse cumulative distribution function (CDF) of the chi-square distribution. For the chi-square distribution, this inverse CDF is not analytically tractable, making it computationally inefficient and potentially less accurate, especially for large degrees of freedom.

---

### Exercise D.

**Simulating t₄ values:**


To simulate values from a $t_{4}$ distribution, we can use the fact that the $t_{4}$ distribution can be represented as the ratio of a standard normal variable to the square root of a chi-square variable divided by its degrees of freedom.

We can simulate t-distributed values using the formula:

$$t = \frac{z}{ \sqrt{(\chi^{2}/df)} }$$

where `z` is a standard normal random variable and `chi_squared` is a chi-square random variable with `df` degrees of freedom.

```{r}
t4 <- z / sqrt(chi_squared/4)
```

---

**1(v) Simulating t₂₀ values:**

```{r}
t20 <- z / sqrt(chi_squared/20)
```


---

### Exercise F.

**Demonstrating a Property of the t-distribution:**

As the degrees of freedom increase, the t-distribution approaches a standard normal distribution. We can visualize this by plotting the densities of the t₄ and t₂₀ distributions:


---

```{r}
library(ggplot2)

# Create a data frame to hold the simulated values
df <- data.frame(t4 = t4, t20 = t20)

# Melt the data frame for easier plotting
df_melted <- reshape2::melt(df)
```

---

```{r}
# Plot the densities
ggplot(df_melted, aes(x = value, color = variable)) +
  geom_density() +
  labs(title = "Density Plots of t4 and t20 Distributions",
       x = "Value", y = "Density")
```

This plot will show that the t₂₀ distribution is more concentrated around 0 and has a shape closer to a normal distribution compared to the t₄ distribution.
