---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


4 The table below shows the total claim amounts (£1,000s) per year, Xij, over a 5-year 
period for five insurance companies’ critical illness book of business. 
Year, j 
2016 2017 2018 2019 2020 
Insurance 
company, i
A 5,000 2,720 3,170 2,950 6,300 
B 3,680 3,360 3,900 2,910 8,020 
C 880 800 550 620 1,890 
D 6,150 3,880 5,780 5,220 7,100 
E 1,100 970 1,900 2,300 3,430 
---

The data can be entered into R in matrix form using the following code: 

```{r}
claims <-  matrix(  
        c(5000,3680,880,6150,1100,
          2720,3360,800,3880,970, 
          3170,3900,550,5780,1900,
		      2950,2910,620,5220,2300,
		      6300, 8020,1890,7100,3430),nrow=5,ncol=5) 
```

---

## Exercises

(i) Calculate, using ***Empirical Bayes Creditability Theory (EBCT)*** Model 1, the 
following: 
(a) E(mሺϑ)  
(b) Var(mሺϑ)  
(c) E(s2ሺϑ)  

---

Part B Calculate the expected claim amount for Company E, using your answers from 
part (i).  
The table below shows additional information that has been provided showing the 
number of claims, Pij, over a 5-year period for the five insurance companies. 
Year, j 
2016 2017 2018 2019 2020 
Insurance 
company, i
A 580 540 490 500 650 
B 440 460 360 390 510 
C 180 160 170 120 220 
D 990 1,000 880 930 1,100 
E 210 200 175 170 130 
The data can be entered into R in matrix form using the following code: 
```{r}
volumes <- 
matrix(c(580,440,180,990,210,540,460,160,1000,200,490, 
360,170,880,175,500,390,120,930,170,650,510,220, 
1100,130),nrow=5,ncol=5) 
```

---

Part C Calculate, using EBCT Model 2, the following: 
(a) E(mሺϑ)  
(b) Var(mሺϑ)  
(c) E(s2ሺϑ)  
Part D Calculate the expected claim amount for Company E, using your answers from 
part Part C, assuming that the number of claims for Company E in the following 
year remains at 130.  
Part E Comment on the difference between models EBCT Model 1 and EBCT 
Model 2 applied in earlier parts, by comparing your answers in parts Part B 
and Part D.  

---

Q4
Note the answers shown here are in units of 1,000s (or 1,000,000s for variances).

### Part A
```{r}
n = ncol(claims)
```

(i)(a)
```{r}
m <- mean(rowMeans(claims)) 
```

 3,383.2 
(i)(b)
```{r}
v <- var(rowMeans(claims)) - mean(apply(claims,1,var))/n 
```

 3,232,667 
(i)(c)
```{r}
s <- mean(apply(claims,1,var)) 
```
 1,887,450 
 
---

Part B
Z <- n/(n+s/v) 
 0.8954367 or 89.54%
Z * rowMeans(claims)+(1-Z)*m 
 3960.578 4270.399 1202.633 5391.485 2090.906
Estimated claims amount for company E is 2,090.91. 

---

Part C(a)
```{r}
n <- ncol(claims)
N <- nrow(claims) 
X <- claims/volumes 
Xi_bar <- rowSums(claims)/rowSums(volumes) 
Pi <- rowSums(volumes) 
P <- sum(Pi) 
```

```{r}
P_star <- sum(Pi*(1-Pi/P))/(N*n-1) 
m2 <- sum(claims)/P 
```

---

# 7.319775 
(b)
```{r}
v2 <- (sum(rowSums(volumes*(X-m2)^2))/(N*n-1) - mean(rowSums(volumes*(X-Xi_bar)^2)/(n-1)))/P_star 

```

 3.151102 

---

Part C(c)


```{r}
s2 <- mean(rowSums(volumes*(X-Xi_bar)^2)/(n-1)) 
```
 4356.445 

---

Part D
```{r}
Zi <- Pi/(Pi+s2/v2)
```

 0.6662620 0.6097364 0.3807366 0.7799425 0.3902952
or 39.03% credibility factor for company E.

```{r}
Zi*Xi_bar + (1-Zi)*m2 
```
 7.304668 9.030223 6.656035 6.088278 8.740714

* Therefore the premium (i.e. expected claims) for company E in the coming year will be: 8.740714 * 130 = 1,136.29 

---
Part E
The expected claim amount for company E is lower using EBCT model 2. 
EBCT model 2 allows for the number of claims (volumes), 
which has been declining 

---

---

Part (i): Very well answered overall. In (b) and (c) a small number of candidates gave confused answers between v and s.
Part Part B: Very well answered overall. A number of candidates did not mention explicitly the figure for company E.
Part Part C: Very well answered overall. As in part (i), a small number of candidates showed confused answers between v and s.
Part Part D: Well answered generally, with some small calculation errors. As in part Part B, some candidates did not mention explicitly the figure for company E.
Part Part E: This part was not well answered. A number of candidates did not attempt it, but most candidates received partial credit for providing some valid comments.

---