---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


1 (i) Simulate 10,000 values from a N(0,1) distribution using an appropriate R 
command. You should save the generated values in R for later use.  
Part B Simulate 10,000 values from a χ4
2 distribution using an appropriate R 
command. You should save the generated values in R for later use.  
Part C Discuss the main disadvantage of using the inverse transform method for 
simulating values from a chi-square distribution.  
Part D Simulate 10,000 values from a t4 distribution using your answers to parts (i) 
and Part B, explaining your answer. 
You should save the generated values in R for later use.  
Part E Simulate 10,000 values from a t20 distribution by modifying your R code from 
parts (i), Part B and Part D where necessary. 
You should save the generated values in R for later use.  
(vi) Demonstrate an important property of the t distribution by presenting and 
comparing two appropriate plots of the values simulated in parts Part D and Part E. 
 [6] 
 [Total 22] 
CS1B A2023–3 


--

Q1
(i) R code:
```{r}
norm.val = rnorm(10000) 
```

---

### Part B 

R code:

```{r}
chisq4.val = rchisq(10000,4) 
```

---

Part C 

The main disadvantage of the inverse transform method is the necessity to have an explicit expression for the inverse of the distribution function, which is not available for the chi-square distribution. 
The use of the method would require numerical methods for inverting the cdf. 

---

### Part D 

A random variable having the t_n distribution can be simulated as

$$Z / \sqrt(Y/n) $$
where Z ~ N(0,1) and Y ~ 𝜒𝜒𝑛𝑛2 
independently. 
R code:

```{r}
t4.val = norm.val/sqrt(chisq4.val/4) 
```

---
Part E R code:
```{r}
norm.val = rnorm(10000) 
```

[Or, use same values from (i)]

```{r}
chisq20.val = rchisq(10000,20) 
t20.val = norm.val/sqrt(chisq20.val/20) 
```

---

(vi) R code:
(vii)

```{r}
par(mfrow=c(1,2))
hist(t4.val) 
hist(t20.val) 
```

Histograms are shown below: 

#### Commentary

The histograms illustrate that the t distribution approaches the standard normal distribution as the number of the degrees of freedom increases. 

---

END

---


END

---

Parts (i), Part B: Very well answered. A small number of candidates used runif() in (i).
Part Part C: Candidates’ comments were often unclear. Many candidates failed to mention the need to use numerical methods.
Part Part D: Mostly well answered. A common error was failing to mention the required independence.
Part Part E: Mostly well answered. There were some errors concerning the degrees of freedom of the involved chi square distribution.
Part (vi): The answers here were of mixed quality. A range of different plots were presented and received credit as appropriate.
