---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


4 The table below shows the total claim amounts (£1,000s) per year, Xij, over a 5-year 
period for five insurance companies’ critical illness book of business. 
Year, j 
2016 2017 2018 2019 2020 
Insurance 
company, i
A 5,000 2,720 3,170 2,950 6,300 
B 3,680 3,360 3,900 2,910 8,020 
C 880 800 550 620 1,890 
D 6,150 3,880 5,780 5,220 7,100 
E 1,100 970 1,900 2,300 3,430 
---

The data can be entered into R in matrix form using the following code: 
claims <- 
matrix(c(5000,3680,880,6150,1100,2720,3360,800,3880,970, 
3170,3900,550,5780,1900,2950,2910,620,5220,2300,6300, 
8020,1890,7100,3430),nrow=5,ncol=5) 


(i) Calculate, using Empirical Bayes Creditability Theory (EBCT) Model 1, the 
following: 
(a) Eሾmሺϑሻሿ [2] 
(b) Varሾmሺϑሻሿ [3] 
(c) Eሾs2ሺϑሻሿ [2] 

---

(ii) Calculate the expected claim amount for Company E, using your answers from 
part (i). [3] 
The table below shows additional information that has been provided showing the 
number of claims, Pij, over a 5-year period for the five insurance companies. 
Year, j 
2016 2017 2018 2019 2020 
Insurance 
company, i
A 580 540 490 500 650 
B 440 460 360 390 510 
C 180 160 170 120 220 
D 990 1,000 880 930 1,100 
E 210 200 175 170 130 
The data can be entered into R in matrix form using the following code: 
```{r}
volumes <- 
matrix(c(580,440,180,990,210,540,460,160,1000,200,490, 
360,170,880,175,500,390,120,930,170,650,510,220, 
1100,130),nrow=5,ncol=5) 
```

---

(iii) Calculate, using EBCT Model 2, the following: 
(a) Eሾmሺϑሻሿ [7] 
(b) Varሾmሺϑሻሿ [4] 
(c) Eሾs2ሺϑሻሿ [3] 
(iv) Calculate the expected claim amount for Company E, using your answers from 
part (iii), assuming that the number of claims for Company E in the following 
year remains at 130. [3] 
(v) Comment on the difference between models EBCT Model 1 and EBCT 
Model 2 applied in earlier parts, by comparing your answers in parts (ii) 
and (iv). [2] 
 [Total 29] 
END OF PAPER 