---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


2 Consider a sample of 1,000 insurance policies. For each policy, there is a record of 
whether the policyholder has submitted at least one claim during the last 6-month 
period, the age of the policyholder and the duration of the policy. The number of 
policies, X, with one or more submitted claims is modelled as a random variable with 
a Binomial distribution X ~ Binomial(1000, p). 

---
(i) State the underlying assumptions that lead to the choice of a Binomial 
distribution for X.  
Use the command <tt>load("CS1policies.Rdata")</tt> to load the observed data into 
R. This will create a data frame called CS1policies with three entries in your R 
workspace. You can check the names of the entries with names(CS1policies): 
*  age: age of the policyholder in years 
*  duration: duration of the policy in months 
*  claimed: takes the values 0 (no claim in 6-month period) or 1 (at least one claim 
in 6-month period). 

---

An Actuary wishes to fit different Generalised Linear Models (GLMs) to the data, 
assuming that the number of policies with one or more submitted claims has a 
Binomial distribution and the link function of the GLM is the logit function. 

B. Fit a GLM to the data such that p depends on the age of the policyholder and 
report the estimated parameter values.  

C. Fit a GLM to the data such that $p$ depends on the duration of the policy and 
report the estimated parameter values.  

D. Compare the fit of the models in Part B and Part C using the deviance and 
the ***Akaike’s Information Criterion (AIC)***. 
 
E. Fit a GLM to the data, such that p depends on both the age of the policyholder 
and the duration of the policy, and report the estimated parameter values.  

F. Comment on which of the three models in parts Part B, Part C and Part E is most 
appropriate, justifying your answer.  

---

```{r}
load("CS1policies.Rdata")
```

```{r}
head(CS1policies)
```

---

### Exercise  B 


We need to assume that the 1,000 policies are independent 
and that the probability to claim in a six-month period is the same for all policyholders 

---

### Exercise  B 

```{r}
model1 = glm(claimed ~ age, data = CS1policies, family = binomial) 
```
#OR

```{r}
model1 <- glm(claimed~age, family = "binomial" (link = "logit"))
summary(model1) 
```
intercept = -2.336974, 
slope (age) = 0.024023 
Histogram of t4.valt4.valFrequency-10-5051015010003000Histogram of t20.vat20.valFrequency-4-202405001000

---

### Exercise  C

Part C

```{r}
model2 = glm(claimed ~ duration, data = CS1policies, family = binomial) 
summary(model2 ) 
```
intercept = 0.390062, 
slope (duration) = -0.027608 

---

### Exercise D

* Residual deviance(model1) = 1051.8 , deviance(model2) = 948.98 
* AIC(model1) = 1055.8 , AIC(model2) = 952.98 
* The residual deviance and the AIC are both smaller for the model in part Part C. 
* So duration seems to be a better predictor than age, and we would choose the model in part Part C.

---

### Exercise D

Alternative answer:
The accuracy of the AIC depends on which R command is used: model1, summary(model1) or AIC(model1), which give 1056, 1055.8 and 1055.78, respectively.
For model2 they give 953, 952.98 and 952.9805, respectively.

---

### Exercise E

```{r}
model3 = glm(claimed ~ age + duration, data = CS1policies, family = binomial) 
summary(model3) 
```

* Intercept = -0.909944

* Slope(age) = 0.029434

* Slope(duration) = -0.028040 

---

### Exercise F

* The AIC (AIC (model3)= 945.51) has decreased compared to the model in part Part C (which is already better than the model in part Part B) 

* We would therefore use both, age and duration, in a linear predictor, and choose the model in part Part E. 

### Alternative answer:

* The accuracy of the AIC depends on which R command is used: model3, summary(model3) or AIC(model3), which give 945.5, 945.51 and 945.5106, respectively.

---


---
Parts (i)-Part C were well answered generally. In part Part B (and later parts) a number of candidates failed to report clearly the estimated parameter values.
CS1B - Actuarial Statistics - Core Principles - April 2023 - Examiners’ report
CS1B A2023 © Institute and Faculty of Actuaries
Part Part D: Well answered. A fair number of candidates failed to state the statistics (or the conclusion) explicitly. Some candidates considered the AIC only and not the deviance.
Part Part E: Candidates performed well here and consistently with their answers in earlier parts Part B and Part C. Answers including the interaction term received full marks as appropriate (and later).
Part (vi): Well answered generally. Some candidates used an alternative answer involving the deviances and anova() to compare models, and received credit as appropriate.

