---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
lib_dir: libs
nature:
  highlightStyle: github
  countIncrementalSlides: false
---


In a very small empirical study a sample from a random variable $X$ is observed. 
The data can be entered into R using the following code:


```{r}
x <- c(0.22, 0.38, 1.28, 0.54, 0.56, 1.36,
   0.55, 0.37, 0.43, 0.46, 0.62, 0.54,
   0.54, 0.51, 0.44, 0.68, 0.55, 0.30)
```

---

\subsection*{Exercise 1} 
Estimate the expected value of $X$.


```{r}
mean(x)
```

\begin{verbatim}
0.573888
\end{verbatim}


---

Exercise 2
============================
Calculate a 95\% confidence interval for the expected value of $X$, assuming that $X$ is Normally distributed.

```{r}
t.test(x, conf=0.95)
```
\begin{verbatim}
	One Sample t-test

data:  x

t = 8.2805, df = 17, p-value = 2.276e-07
alternative hypothesis: true mean is not equal to 0

95 percent confidence interval:
 0.4276654 0.7201123

sample estimates:
mean of x 
0.5738889 

\end{verbatim}

---

\textbf{Shapiro-Wilk normality test}

```{r}
shapiro.test(x)

```

---

```{r}

qqnorm(x, cex=1.5, pch=18, col="black" )

qqline(x ,lwd=2,col="red")

```

\includegraphics[scale=0.45]{00-C1/images/00-C1-Q1-qqnorm.png}

---

Tidyverse
==========================

```{r}

# install.packages(c("magrittr","broom"))

library(magrittr)
library(broom)

t.test(x, conf=0.95) %>% tidy()


```

---

Exercise 3
============================

Construct a confidence interval for the expected value of $X$ using the bootstrap
method with 10,000 bootstrap replications.

```{r}
1:4

1:5

1:6
```

---


> sample(1:4,replace =TRUE)
[1] 1 1 4 1
>
> sample(1:5,replace =TRUE)
[1] 5 3 1 5 4
>
> sample(1:6,replace =TRUE)
[1] 4 2 1 2 2 4



```{r}
replicate(10, mean(sample(x,replace =TRUE)))
```


\begin{verbatim}
 [1] 0.5927778 0.4955556 0.5144444 0.5444444 0.6272222
 [6] 0.5577778 0.5438889 0.4916667 0.6233333 0.6266667
\end{verbatim}




```{r}
y = replicate(10000,mean(sample(x,replace =TRUE))) 

quantile(y,prob=c(0.025,0.975)) 

```


\begin{verbatim}
 2.5% 97.5% 
0.4544444 0.7222222 
\end{verbatim}

---


Exercise  4
=============================

Comment on the differences between the confidence intervals in exercises 2 and 3.


* The CIs are almost identical with the CI in exercise 3 being narrower. 
* The reason might be that the distribution of $X$ is not normal, and therefore the
distribution of the mean is not normal for the small sample size we have in this
question. 
* 


The bootstrap method provides a good approximation of the distribution of the mean
independently of the type of distribution.

---