---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```



The dataset “***Interest_rates.csv***” contains a series of returns on bonds of maturities 
1 year, 5 years, 10 years, 15 years, 20 years and 30 years (i.e. bonds that provide a
return of the principal investment after 1, 5, 10, 15, 20 and 30 years respectively).



```{r}
interest_rates <- read.csv("Interest_rates.csv")
``` 

```{r}
head(interest_rates)
```


---

### Exercise 1

Calculate the Pearson correlation coefficient matrix for this data.
Comment on the correlations between the data using the matrix from
Exercise 1



```{r}

# read the data
interest_rates<-read.csv("Interest_rates.csv")
# calculate the Pearson correlation coefficients

``` 

---

```{r}
cor(interest_rates, method="pearson")
``` 

---

The correlation matrix shows that there is strong (positive) correlation between returns on bonds of similar maturity.

It also shows that the correlation between returns is weaker as the length of maturity between bonds increases.

---


### ***ggcorrplot()***

<pre><code>
library(ggcorrplot)

cor(interest_rates, method="pearson") %>%ggcorrplot::ggcorrplot()

</code></pre>

---

### ***ggcorrplot()***

```{r echo=FALSE, warning=FALSE}
library(ggcorrplot)

cor(interest_rates, method="pearson") %>%ggcorrplot::ggcorrplot()

```

---

### Part 2

Perform a reduction of the dimension of the data using principal component analysis with the method of singular value decomposition.

Your answer should include a summary of the principal component analysis.

Suggest with reasons, using the output of the R analysis, how many 
components of the transformed data should be retained.

---
 
### Principal component analysis
 
```{r}
# carry out principal component analysis using SVD method

pca <- prcomp(interest_rates)
``` 


```{r}
# review the results of the principal component analysis
summary(pca)

```


---

### Conclusion



The R-output shows that the proportion of variance explained by the first two principal
components is approx 98%, and by the first three components approx 99%.

Therefore it would be reasonable to reduce the dimensions of the dataset by using the
first two (or three) principal components.

---


---
Candidates performed generally well in this question. Part (i) was very well answered,
with some partial answers in (i)(b) where many candidates observed a relationship in
individual years without drawing out the overall trend. Answers in part (ii) were also
satisfactory. Note that in part (ii) the princomp() function can alternatively be used in R.


