---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---



*   Use the command \texttt{set.seed(2019)} to initialise the random number generator.

*   When you execute any R code, make sure you run the entire R script including the line \texttt{set.seed(2019)} .
```{r}
sample(1:6,10,replace=TRUE)

set.seed(2019)
sample(1:6,10,replace=TRUE)
```

\begin{center}
\includegraphics[scale=0.5]{00-C1/images/00-C1-Q3-Setseed.png}* 
\end{center}


---

Exercise 1
===============================

Consider a random sample $X_1 , \ldots , X_n$ from an Exponential distribution with parameter $\lambda$ and define 
$$ Y = \sum^{n}_{i=1} X_i $$

State the distribution of $Y$, giving all the parameters of the distribution.


---

Theory

*   The sum of $n$ exponential ($\lambda$) random variables is a gamma ($n, \lambda$) random variable.
*   The rate parameter is usually denoted as $\beta$ for the Gamma Distribution, i.e. gamma ($n, \beta$) 

   Mean for Gamma Distribution	 ${\displaystyle {\frac {\alpha }{\beta }}} $

  Variance	for Gamma Distribution $ {\displaystyle {\frac {\alpha }{\beta ^{2}}}}$

*   If the exponential random variables have a common rate parameter, their sum has an Erlang distribution, a special case of the gamma distribution.


The distribution of $Y$ is a Gamma distribution,
$$Y \sim \mbox{Gamma}(n, \lambda)$$


---


```{r}
set.seed(2019);rexp(5,0.5)

set.seed(2019);sum(rexp(5,0.5))

set.seed(2019)
X = replicate(1000,sum(rexp(20,0.5)))

mean(X)

var(X)

```

---

\includegraphics[scale=0.5]{00-C1/images/00-C1-Q3-2.png}

---

Exercise 2
=========================

Perform simulation of a sample $x_1 , ..., x_n$ with sample size n = 15 from an
exponential distribution with parameter $\lambda = 2$.


```{r}
n = 15
lambda = 2
set.seed(2019);X = rexp(n, lambda)

X 

mean(X)

var(X)

sum(X) 


```


---




Exercise 3
===================================

Calculate the value of Y for the sample in part 2.

```{r}
Y = sum(X)
```


---

Exercise 4
=======================
Perform 1,000 repetitions of parts 2 and 3 to obtain a Bootstrap sample
$y_1 , ..., y_B$ from the random variable Y with B = 1,000.


Using \texttt{for} loops.
```{r}
y = 0*(1:1000) # generate a vector of size 1000

for (i in 1:1000){

   y[i] = sum(rexp(n, lambda))

} 
```

Using \texttt{replicate()}
```{r}

y = replicate(1000, sum(rexp(n, lambda)))

```

---

Exercise  5
=======================

Plot a histogram showing the relative frequencies of the sample $y_1 , ..., y_B$  from
part 4.

```{r}
hist(y, prob=TRUE) 
```



---


Exercise 6
=================================
Compare graphically the histogram in part 4 to the density of a
suitable Normal distribution. You might find the following R command
useful:

<pre><code>
curve(dnorm(x,mean= ...., sd= ... , 
 add=TRUE, 
 lwd=2,
 col="red"))
<code></pre>


```{r}
hist(y,prob=TRUE,
  xlim=c(0,20),
  ylim=c(0,0.25))

curve(dnorm(x,
 mean=n/lambda,
 sd=sqrt(n/(lambda^2))), 
   add=TRUE, lwd=2, col="red")

```

\includegraphics[scale=0.6]{00-C1/images/00-C1-Q3-Curve.png}


--- 


```{r}
hist(y,prob=TRUE,
  xlim=c(0,20),
  ylim=c(0,0.225),
  col=c("pink","lightblue"))

curve(dnorm(x,
 mean=n/lambda,
 sd=sqrt(n/(lambda^2))), 
  add=TRUE, lwd=2.75, 
   col="red")


```

\includegraphics[scale=0.6]{00-C1/images/00-C1-Q3-Curve-2.png}

---

Comment on your findings in the context of the Central
Limit Theorem.


*   In contrast to a normal distribution, the histogram is clearly not symmetrical.

  This comes from the fact that $Y$ can take only positive values.
For a larger sample size of $n$ of $x_1 , ... , x_n$ (not a larger B) the CLT ensures that the
distribution of $Y$ becomes approximately normal.



```{r}
n = 75
lambda = 2
set.seed(2019)
y = replicate(1000, sum(rexp(n, lambda)))

hist(y,prob=TRUE,
  xlim=c(20,60),
  ylim=c(0,0.10),
  col=c("pink","lightblue"))

curve(dnorm(x,
 mean=n/lambda,
 sd=sqrt(n/(lambda^2))), 
   add=TRUE, lwd=2.75, 
   col="red")

```


\includegraphics[scale=0.5]{00-C1/images/00-C1-Q3-Curve-Q3.png}

---

Parts (i)-(iii) were generally very well answered. In part (iv) there was wide variation in
the quality of the answers with various errors in the details of the code. 

A common error in part (v) was omitting the (prob = TRUE) part of the code, which is required for relative frequencies. 

Part (vi)(a) was reasonably well attempted, while many candidates did not
attempt part (vi)(b), with some giving incomplete comments. Part (vi)(a) can alternatively
be answered using R code based on the lines() command.

