---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 4 - Regression Analysis
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```




A recent study suggests that the maximum heart rate of a person, related to age in
years, is given by the equation:

$$ \mbox{Max Rate} = 220 – \mbox{Age}$$

Suppose this is to be empirically proven and 15 people of varying ages are tested for
their maximum heart rate. 

---

The following data are collected:

| | | | | | | | | | | | |
|:-------------:|------|------|------|------|------|------|------|------|------|------|
|***Age (years)***| 18| 23| 25| 35| 65| 54| 34| 56| 72| 19| 23| 42| 18| 39| 37|
|***Max Rate ***| 202| 186| 187| 180| 156| 169| 174| 172| 153| 199| 193| 174| 198| 183| 178|

The data can be entered into R using the following commands:

<pre><code>
x <- c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)
y <- c(202,186,187,180,156,169,174,172,153,199,193,174, 198,183,178)
</code></pre>

---

## Exercise 1

Plot the fitted line for the regression of Max Rate on Age.

<pre><code>
x <- c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)

y <- c(202,186,187,180,156,169,174,172,153,199,193,174,198,183,178)

plot(x,y, pch=18,col="red",cex=2)
</code></pre>

---
### Scatterplot

```{r echo=FALSE,warning=FALSE}
x <- c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)

y <- c(202,186,187,180,156,169,174,172,153,199,193,174,198,183,178)

plot(x,y, pch=18,col="red",cex=2)
```


---

### Regression model

```{r}
lm.result = lm(y ~ x)
```


```{r}
# Obtain the basic values of the regression analysis

lm.result
```

---

<pre><code>
# make a plot
plot(x,y,pch=18,cex=2,col="blue")

# plot the regression line
abline(lm(y ~ x), col="red") 

</code></pre>

---

```{r echo=FALSE,warning=FALSE}
# make a plot
plot(x,y,pch=18,cex=2,col="blue")

# plot the regression line
abline(lm(y ~ x), col="red")

```


---


```{r}
summary(lm.result)
```

---


Max heart rate reduces as age increases. The fit of the model seems good.

---


```{r}
coef(lm.result)
```


```{r}
confint(lm.result)
```


---



## Exercise 2

Comment on the results.

A researcher reviews the plot in ***Exercise 1*** and suggests the slope should be equal to –1.


```{r}
confint(lm.result) %>% round(2)
```




---

## Exercise 3

Calculate the p-value of a hypothesis test for this suggestion, by creating a suitable test statistic.


* Can do a test to see if the slope of -1 is correct. 


* Let $H_0$ be that $\beta = -1$, and $H_A$ be that $\beta \neq -1$. 


* Then we can create the test statistic and the p-value as follows:

---

```{r}
es <- resid(lm.result)
# the residuals lm.result

es %>% round(3)
```


```{r}
b1 <- (coef(lm.result))[['x']] #the x part of the coefficients
n  <-  15

b1 %>% round(3)

n
```


---

### Residual Standard Error

Under the assumption that the population error term has a constant variance, the mean square error (MSE) of the regression is the estimate of that variance, and is given by:

$${\displaystyle {\hat {\sigma }}_{\varepsilon }^{2}={\frac {SSR}{n-2}}.\,}$$

The residual standard error is the square root of the MSE.

```{r}
#calculate residual standard error
s  <-  sqrt( sum( es^2 ) / (n-2) )
s

```

```{r}

sqrt(deviance(lm.result)/df.residual(lm.result))

```
---
The standard errors of the parameter estimates are given by

$${\displaystyle {\hat {\sigma }}_{\beta _{1}}={\hat {\sigma }}_{\varepsilon }{\sqrt {\frac {1}{\sum (x_{i}-{\bar {x}})^{2}}}}}$$

$${\displaystyle {\hat {\sigma }}_{\beta _{0}}={\hat {\sigma }}_{\varepsilon }{\sqrt {{\frac {1}{n}}+{\frac {{\bar {x}}^{2}}{\sum (x_{i}-{\bar {x}})^{2}}}}}={\hat {\sigma }}_{\beta _{1}}{\sqrt {\frac {\sum x_{i}^{2}}{n}}}.}$$

```{r}


SE <-  s/sqrt(sum((x-mean(x))^2))
SE
```


---

### Test Statistic

```{r}
t  <-  (b1 - (-1) )/SE
t
```


---

* find the right tail for this value of t with 15-2 d.f.


```{r}
pt(t,13,lower.tail=FALSE)
```



---


### P-value


The p-value is twice this as the problem is two-sided,




```{r}
2 * pt(t,13,lower.tail=FALSE)
```

---

## Exercise 4
Comment on the researcher’s suggestion, using your answer to part 3.




The null hypothesis is rejected at the 5% level of significance. The slope may not be
equal to -1 for these data. (Which is the slope predicted by the original formula 220 -
Age.)


---