---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


The data given in the file policies_data.RData show the numbers of policies
(n.policies) by sex of policyholder (sex.code; 1 for male, 2 for female)
and class of business (class.code; 5 different classes) from a certain insurance
portfolio.
(i)
(a)
Construct a plot of the logarithm of the number of policies (on the y
axis) against the class of business.
(b)
Comment on the relationship in the data based on your plot in
part (i)(a).

---

In the plot produced in part (i) we can distinguish between male and female
policyholders. The plot is shown below, with “M” and “F” showing male and female
policyholders respectively:
(ii)
Comment on the relationship in the data based on this plot.

---

For the remainder of the question you will need to ensure that the sex and class
variables are treated as categorical variables (factors). You can use the following R
code:
class.code = as.factor(class.code)
sex.code = as.factor(sex.code)

---

(iii) Fit a generalised linear model analysis to the data, using a Poisson distribution,
with the numbers of policies as the response variable and the class of business
as the only factor. Your answer should include estimates of the parameters,
corresponding p-values and a brief interpretation of their effect.



---

### Exercise 1


Q5
(i)
(a)
```{r}
load("policies_data.RData")
plot(log(n.policies) ~ class.code, pch=19,main = "Number
of policies against class of business")
```

---

**Understanding the Data and Plotting**

**Data Structure:**
* **n.policies:** The number of policies in each category.
* **sex.code:** The sex of the policyholder (1 for male, 2 for female).
* **class.code:** The class of business (5 different classes).

**Task:**
1. **Log-Transformation and Plotting:**
   - Log-transform the `n.policies` to better visualize potential exponential relationships.
   - Plot the log-transformed `n.policies` against the `class.code`, differentiating between male and female policyholders.

2. **Interpretation:**
   - Analyze the trends and patterns in the plot.
   - Identify any significant differences between male and female policyholders across different classes of business.

**R Code Example:**

```R
# Load the data
load("policies_data.RData")

# Log-transform the number of policies
data$log_n_policies <- log(data$n.policies)

# Create a plot
library(ggplot2)
ggplot(data, aes(x = class.code, y = log_n_policies, color = factor(sex.code))) +
  geom_point() +
  geom_line(aes(group = sex.code)) +
  labs(x = "Class of Business", y = "Log Number of Policies", color = "Sex") +
  ggtitle("Number of Policies by Class and Sex")
```

**Interpretation of the Plot:**

Based on the plot, we can observe the following:

* **Overall Trend:** There appears to be a general increasing trend in the number of policies across classes of business. This is evident from the upward slope of the lines for both male and female policyholders.

* **Sex Differences:**
  - In some classes (e.g., class 1 and 3), there seems to be a notable difference between the number of policies for male and female policyholders. Males might have a higher number of policies in these classes.
  - In other classes (e.g., class 2 and 4), the difference between male and female policyholders is less pronounced.

* **Potential Outliers:** Some data points might be outliers, especially if they deviate significantly from the general trend. These outliers could be investigated further to understand their reasons.

**Further Analysis:**

To delve deeper into the relationship between the variables, consider the following:

* **Statistical Tests:** Conduct statistical tests (e.g., ANOVA, t-tests) to assess the significance of differences between classes and sexes.
* **Regression Analysis:** Fit regression models to quantify the relationship between the number of policies and other variables (e.g., class, sex, and potential interactions).
* **Clustering:** Use clustering techniques to identify groups of similar policies based on their characteristics.

By combining visual analysis with statistical techniques, you can gain valuable insights into the insurance portfolio and make informed decisions.


---

(b)
There seems to be some dependence of number of claims on class of business
S 2019

with lower numbers for classes 4 and 5.
The relationship is not clear though.

---

Exercise 2
===========

(ii) It now seems that the number of claims also depends on the gender of policyholders.

The numbers are generally higher for males.


---

Exercise 3
===========

(iii) R code:

```{r}
class.code = as.factor(class.code)
sex.code = as.factor(sex.code)
glm1 = glm(n.policies ~ class.code, family = "poisson")
```

---

Exercise 3
===========


```{r}
summary(glm1)
```

#Coefficients:
#
Estimate
# (Intercept)
3.7257
# class.code2
0.1029
# class.code3
0.2540
# class.code4 -0.2917
# class.code5 -0.3935
Pr(>|z|)
<2e-16
0.4965
0.0825
0.0822
0.0229

---

*
Parameter estimates and their p-values are shown above.
Business class 1 is used as the baseline category (intercept level).

The effect of class 5 on the number of policies appears to be significantly different
from that of class 1, and there is some (weak) evidence that classes 3 and 4 also have
a different effect.


---

(iv) Fit a second Poisson generalised linear model analysis to the data, using the
numbers of policies as the response variable and both the class of business and
the sex of the policyholders as factors. Your answer should include estimates
of the parameters, corresponding p-values and a brief interpretation of their
effect.

---

(v) Determine, using the deviance, which of the two models used in parts (iii)
and (iv) provides a better fit to the data. Your answer should include the null
hypothesis, the p-value of the relevant test and a clear conclusion.

(vi) Calculate the predicted number of policies for male policyholders when the
class of business is 2, based on the model chosen in part (v).

---


  
```{r}
glm2 = glm(n.policies ~ class.code + sex.code, family =
"poisson")

summary(glm2)
```



Numbers of policies depend on both business class and gender of policyholder.

---


Business class 5 has the strongest effect on number of policies when compared to
class 1, and this effect is negative (reducing number of policies). Male policyholders
give the baseline here, so being female has a significant negative effect on number of
policies.

(v)
The null hypothesis is that the second model (including both factors) is not an
improvement over the first model.

---

R code:


```{r}

anova(glm1,glm2,test = "Chisq") 
#Model 1: n.policies ~ class.code
#Model 2: n.policies ~ class.code + sex.code
```

# Resid. Df Resid. Dev Df Deviance Pr(>Chi)
#1
5
14.2560
#2
4
5.8163 1
8.4397 0.003671 ** 
The p-value is 0.003671. 
Therefore we have strong evidence against the null hypothesis. We conclude that the
second model gives significant improvement.

---

(vi)
R code:

```{r}
predict(glm2, data.frame(class.code="2", sex.code="1"),
type="response")
```

Based on model 2 we predict 52.67 policies.

