---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

---

## Exercise 1

The following data represent the average total number of marks obtained for a particular exam, observed over seven exam sessions that had been administered by a professional examination body:

\[87, 53, 72, 90, 78, 85, 83 \]

Enter these data into R and compute their sample mean and variance.

```{r}
y <- c(87, 53, 72, 90, 78, 85, 83)
mean(y)
c(mean = mean(y), variance = var(y))
```



---

## Exercise 2

Investigate whether the Poisson model is appropriate for these data by calculating the sample mean and sample variance of 10 Poisson samples having the same size and mean as the sample given above.

### Recall:
For the Poisson distribution: $$E(X) = \text{Var}(X)$$

```{r}
xbar <- numeric(10)
s2 <- numeric(10)
xbar
s2
```

---
Use `rpois()` to generate 7 random numbers.

```{r}
x <- rpois(7, lambda = 78.29)
x
```

---

```{r}
set.seed(1234)
for (j in 1:10) {
    x <- rpois(7, 78.29)
    xbar[j] <- mean(x)
    s2[j] <- var(x)
}
```
#
It is unusual to get as large a difference between the mean and the variance as that observed for these data, making it doubtful that these data are from a Poisson distribution.

---

## Exercise 3

Investigate whether the Poisson model is appropriate for these data by calculating the sample mean and sample variance of 10,000 Poisson samples having the same size and mean as the sample given above.

```{r}
set.seed(1234)
for (j in 1:10000) {
    x <- rpois(7, 78.29)
    xbar[j] <- mean(x)
    s2[j] <- var(x)
}
```


---

```{r}
hist(xbar, breaks = 200, col = "lightblue")
```


---

```{r}
hist(s2, breaks = 200, col = "pink")
```

---

```{r}
mean(s2 > 159)
```

---
