---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


In a small empirical study, data are recorded on the number of waves per hour and the
average wave height per hour at a location just off the coast of Scotland. 

The data are
given in the file named CS1waves.Rdata. 

Loading the data into R will create two
vectors in your R workspace, called \textbf{\textit{Wn}} (number of waves per hour) and \textbf{\textit{Wheight}}
(average wave height in cm during the hour).

Generate an appropriate plot to visually inspect the relationship between wave
height and number of waves per hour.


---

### Exercise  1


Generate an appropriate plot to visually inspect the relationship between wave
height and number of waves per hour.



```{r}
# load data
load("CS1waves.Rdata")

#plot data
plot(Wn, Wheight)
```


---

### Exercise  2

Comment on the plot in Exercise 1


#### Comments 

There seems to be a linear relationship between wave height and number of waves.
The more waves per hour, the smaller the waves (negative association).

---

### Exercise 3

Calculate Pearson’s correlation coefficient between the number of waves per
hour and the average wave height.


```{r}
cor(Wn, Wheight,method = "pearson")

```


---

### Exercise 4

Calculate Spearman’s rank correlation coefficient between the number of
waves per hour and the average wave height.

Comment on your findings in parts (iii) and (iv).

```{r}
cor(Wn, Wheight, method = "spearman")

```

---


### Exercise  5
We now model the number of waves per hour, X, as a random variable with a Poisson
distribution with unknown parameter l. The log likelihood function for estimating l
is given by l(l) = log(l)
n
x i ) – ln where n is the number of observations.
( Â i=1


---

### Exercise 5


* Both correlation coefficients confirm the negative relationship that is
already obvious in the plot.
* The rank correlation is lower than the Pearson correlation, indicating that the relationship is stronger when we take the magnitude of observations into account rather than just their relative rank.
* In other words, for observations with similar magnitude, the ranks are not always ordered.



---

### Exercise  6

Plot the log likelihood function for values of $l = 220, 221, ..., 280$.

```{r}
l = 220:280
ll = log(l)*sum(Wn)-168*l
plot(l,ll) [1]
```


---

### Exercise  7

Determine an approximate maximum likelihood estimate for l using the plot in Exercise 6

#### Solution


* By inspection, the maximum of $ll(\hat{\lambda})$ is at about 250.
* The maximum likelihood estimate is $\hat{\lambda} \approx 250$ waves per hour.


---

### Exercise  8

Calculate the exact maximum likelihood estimate of $l$.


#### Solution

The exact MLE is the mean, that is, \hat{\lambda}̂ = 168 ∑ 168
ii=1 XX ii
mean(Wn)

$\hat{\lambda}̂ =248.8579$

A common error in part (vii) was to provide the value of the log likelihood
instead of lambda.



---


---