---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


---

In a small empirical study, data are recorded on the number of waves per hour and the average wave height per hour at a location just off the coast of Scotland. The data are given in the file named `CS1waves.Rdata`. Loading the data into R will create two vectors in your R workspace, called **Wn** (number of waves per hour) and **Wheight** (average wave height in cm during the hour).

---
### Exercise 1
Generate an appropriate plot to visually inspect the relationship between wave height and the number of waves per hour.

```{r echo=FALSE}
# load data

Waves <- readr::read_csv("WavesData.csv", show_col_types = FALSE)
attach(Waves)
#load("CS1waves.Rdata")
```

---

```{r}
head(Waves)
```


```{r eval=FALSE}
# plot data
plot(Wn, Wheight,
     pch=18,col="red")
```

---

```{r echo=FALSE}
# plot data
plot(Wn, Wheight,
     pch=18,col="red")
```



---

### Exercise 2
**Comment on the plot in Exercise 1.**


#### Comments
There seems to be a linear relationship between wave height and the number of waves. The more waves per hour, the smaller the waves (negative association).

---

### Exercise 3
Calculate Pearson’s correlation coefficient between the number of waves per hour and the average wave height.

```{r}
cor(Wn, Wheight, method = "pearson")
```

---

### Exercise 4

Calculate Spearman’s rank correlation coefficient between the number of waves per hour and the average wave height. Comment on your findings in parts (iii) and (iv).

```{r}
cor(Wn, Wheight, method = "spearman")
```

---

### Exercise 5

We now model the number of waves per hour, X, as a random variable with a Poisson distribution with unknown parameter λ. The log likelihood function for estimating λ is given by:

$$l(\lambda) = \log(\lambda) \cdot \sum_{i=1}^{n} x_i - n \cdot \lambda$$

where $n$ is the number of observations.

---


### Exercise 5

#### Comments

* Both correlation coefficients confirm the negative relationship that is already obvious in the plot.

* The rank correlation is lower than the Pearson correlation, indicating that the relationship is stronger when we take the magnitude of observations into account rather than just their relative rank.

* In other words, for observations with similar magnitude, the ranks are not always ordered.

---
### Exercise 6
Plot the log likelihood function for values of $\lambda = 220, 221, ..., 280$.

```{r}
λ = 220:280
ll = log(λ) * sum(Wn) - 168 * λ
plot(λ, ll)
```

---
### Exercise 7
Determine an approximate maximum likelihood estimate for λ using the plot in Exercise 6.


#### Solution
* By inspection, the maximum of $ll(\hat{\lambda})$ is at about 250.
* The maximum likelihood estimate is $\hat{\lambda} \approx 250$ waves per hour.

---
### Exercise 8
Calculate the exact maximum likelihood estimate of $\lambda$.


#### Solution
The exact MLE is the mean, that is,

$$\hat{\lambda} = \frac{\sum_{i=1}^{168} X_i}{168}$$
$$\hat{\lambda} = 248.8579$$

A common error in part (vii) was to provide the value of the log likelihood instead of λ.

---