---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

A statistician is carrying out an exercise to analyse a dataset that describes the failure
times of outdoor telephone cables, with respect to the cable material quality (graded
1 to 4) and level of rainfall in centimetres that the cable is exposed to.


The data given in the file “***CablesDataset.csv” show failure times in years for 20
different cables.

---

### Import Data 

```{r}
# read the data
cables_data<-read.csv("Data/CablesDataset.csv")
head(cables_data)
```


---


#### Exercise 1

Fit a linear model to the data with the ***failure time*** as the response, including
both cable material quality and level of rainfall as the two covariates.

Your answer should include a summary of the fitted model.





```{r}
# fit the linear model

linear_model<-lm(Failure.Time ~ Material.Type + Rainfall,data =
cables_data)


```

---

```{r}
# output the linear model fit summary statistics
summary(linear_model)
```   
 

---

#### Exercise 2


State the formula of the model fitted in Exercise 1, clearly explaining the
notation that you use.
Comment on the significance of the parameters of the model fitted.


```{r}
coef(linear_model)
```

```{r}
confint(linear_model)
```

---

#### Model Summary 
From analysing the R output, we see that the fitted linear model is:
$$ \hat{y} = 0.081086 - 0.014499x_1 + 0.005591x_2$$

where $x_1$ is the ‘material type’ variable, and $x_2$ is the ‘rainfall’ variable. $\hat{y}$ is the predicted failure time.


* The R output shows that the ‘material type’ parameter is significantly
different to zero (at the 0.1% level).
* The ‘rainfall’ parameter is not significantly different to zero - this is
indicated by the t-tests shown in the fourth column of the R output.
* The intercept is also significantly different to zero



---

#### Exercise 3


Plot the residuals of the model. Comment on the plot.



```{r}
# obtain the residuals

residuals(linear_model)

```


---

```{r}
# obtain the largest residuals 

which.max( residuals(linear_model) ) 

```



```{r}
# obtain the largest residuals

which.max( abs ( residuals(linear_model) ) )

```

---

```{r eval = FALSE}
# plot the residuals

plot(residuals(linear_model),
      pch=16,col="red",cex=1.4)

abline(h=0)

```


---

```{r echo=FALSE}
# plot the residuals

plot(residuals(linear_model),
      pch=16,col="red",cex=1.4)

abline(h=0)

```

---



The residuals exhibit a fairly random scatter around zero
apart from the 6th point.

---

#### Exercise 4

An analyst suggests that the 6th row of the original data should be removed. 
Construct a new data set from the original data ***CablesDataset.csv*** with the 6th row removed. 
Justify the removal of the 6th row from the original data.


```{r}
# remove data point 6 and redefine the dataset

cables_data2 <- cables_data[-6,]

```

Remark: Previously the residuals plot indicates that data point 6 is an outlier.


---

  
#### Exercise 5

Fit a linear model to the new data set constructed in Exercise 4. 
Comment on the fit of the new model compared to the model fitted previously, by comparing suitable statistics from the R outputs.

```{r}
# refit the linear model

linear_model2 <- lm(Failure.Time~Material.Type + Rainfall,
    data = cables_data2)


```


---

```{r}
summary(linear_model2)
```

---

* The adjusted $R^2$ statistic for the model fitted to the data with the outlier
removed is 0.7475. 

* This shows an improved fit relative to the model fitted to all
20 data points, which had an adjusted $R^2$ statistic of 0.4776.



---

#### Exercise 6

Fit a generalised linear model (GLM) to the data set constructed in Exercise 4 using a Gamma distribution.
State the formula of the model fitted, clearly explaining the notation that you use.
Comment on the significance of the parameters of the fitted model.

---

```{r}
# fit a Gamma GLM
gfit<-glm(Failure.Time~Material.Type+Rainfall,
        family= Gamma(link=inverse), 
        data = cables_data2)
```

---

```{r}
summary(gfit)

```

---


```{r}
coef(gfit)
```






\[
\hat{\eta} = \frac{1}{\mu} 
= 6.1920558 + 6.7286452x_1 + 0.4814427x_2
\]

where $x_1$ is the ‘material type’ variable, and $x_2$ is the ‘rainfall’ variable.

---

```{r}
# review the model fit
summary(gfit)
```



---

Reviewing the model fit output from R, the ‘rainfall’ parameter is not
significantly different to zero, whereas the ‘material type’ parameter is
significant at the 0.1% level.
