---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

In a very small empirical study a sample from a random variable $X$ is observed. 
The data can be entered into R using the following code:


```{r}
x <- c(0.22, 0.38, 1.28, 0.54, 0.56, 1.36,
   0.55, 0.37, 0.43, 0.46, 0.62, 0.54,
   0.54, 0.51, 0.44, 0.68, 0.55, 0.30)
```

---

### Exercise 1
Estimate the expected value of $X$.


```{r}
mean(x)
```



---

### Exercise 2

Calculate a 95% confidence interval for the expected value of $X$, assuming that $X$ is Normally distributed.

```{r}
t.test(x, conf=0.95)
```

---

#### Shapiro-Wilk normality test

```{r}
shapiro.test(x)

```

---

```{r}

qqnorm(x, cex=1.5, pch=18, col="black" )

qqline(x ,lwd=2,col="red")

```

---

#### Tidyverse


```{r warning=FALSE}

# install.packages(c("magrittr","broom"))

library(magrittr)
library(broom)
```


```{r}
t.test(x, conf=0.95) %>% tidy()


```

---

### Exercise 3


Construct a confidence interval for the expected value of $X$ using the bootstrap
method with 10,000 bootstrap replications.

```{r}
1:4

1:5

1:6
```

---

```{r}
sample(1:4, 3, replace =TRUE)

```

```{r}
sample(1:5, 3, replace =TRUE)
```

```{r}
sample(1:6, 3, replace =TRUE)
```
---


---

```{r}
sample(1:4,  replace =TRUE)

```

```{r}
sample(1:5,  replace =TRUE)
```

```{r}
sample(1:6, replace =TRUE)
```

---

### ***replicate()***

```{r}
replicate(5, mean(sample(x,replace =TRUE)))
```


```{r}
replicate(5, mean(sample(x,replace =TRUE)))
```


```{r}
replicate(5, mean(sample(x,replace =TRUE)))
```
---



```{r}
y = replicate(10000,mean(sample(x,replace =TRUE))) 

quantile(y,prob=c(0.025,0.975)) 

```


---


### Exercise  4


Comment on the differences between the confidence intervals in exercises 2 and 3.


* The CIs are almost identical with the CI in exercise 3 being narrower. 
<p>
* The reason might be that the distribution of $X$ is not normal, and therefore the
distribution of the mean is not normal for the small sample size we have in this
question. 
<p>


* The bootstrap method provides a good approximation of the distribution of the mean
independently of the type of distribution.

---