---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 1 - Binomial Probability Distribution
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```



The probability that India will win a cricket match against South Africa is 0.7


#### Exercises

1. Prepare a probability distribution table of number of wins if Indians are going to play 10
matches against the South Africans. 
2. Plot a bar chart of the probabilities of number of wins from 0 to 10.

3.Find the mean and median number of wins for India against South Africa.

---


#### Exercise Solution 1:


$$ X \sim Bin(n=10,p=0.7)$$

```{r}
p<-0.7
```
probability distribution table of number of number of wins

```{r}
prob<-dbinom(0:10,size = 10,prob = 0.7)
probability_Densities<-data.frame(No_Successes = 0:10,prob)
```

---

```{r}
probability_Densities
```


---

### Exercsie 2

Bar chart of the probabilities of number of wins

<pre><code>
barplot(prob,main = "Bar Chart of Probability of Successes", 
    xlab = "Number of Successes", 
    names.arg = 0:10)
</code></pre>

---


```{r echo=FALSE}
barplot(prob,main = "Bar Chart of Probability of Successes", 
    xlab = "Number of Successes", 
    names.arg = 0:10)
```

---
### Exercise 3

Mean and median number of wins for India against South Africa


#### Mean

```{r}
#Definition of Binomial Mean
mean<-10*0.7 #or
mean
```

```{r}

#Definition of Expectation of Discrete Random Variable
X <- probability_Densities$No_Successes
Prob.X <- probability_Densities$prob
mean <- sum(X * Prob.X)

mean

```

---


#### Median

```{r}
median<-qbinom(0.5,size = 10,prob = 0.7)
median
```


---



