---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 4 - Categorical Variables
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Refer to the data file ***Indices_Returns.csv*** and answer the following questions:


```{r}
# Load the data file

indices<-read.csv("Indices_Returns.csv")
```

```{r}
head(indices)
```

---

### Exercise 1

Express the number of months with negative Sensex returns as a proportion of total 
number of months. 

---

number of months with negative Sensex returns as a proportion of total number of months

```{r}
negative_Sensex <- sum(indices$Sensex<0) 
negative_Sensex 
```

```{r}
proportion_neg <- negative_Sensex / nrow(indices) 
proportion_neg 
```

```{r}
mean(indices$Sensex<0)
```

---

### Exercise 2

Test whether the proportion of months with negative Sensex returns is less than 50% at 
95% confidence level as well as at 99% confidence level. 

---

```{r}
binom.test(negative_Sensex,nrow(indices),
           p=0.5,
           alternative = "less") 
```

---

```{r}
test <- binom.test(negative_Sensex,nrow(indices),
           p=0.5,
           alternative = "less") 

```

```{r}
library(broom)

tidy(test)

```

---

#### Interpretation

* The p-value corresponding to the test is 0.02529. 

* So the null hypothesis of Proportion of Negatives is at least 50% is rejected at 
95% Confidence level but is failed to be rejected at 99% Confidence level 
 
---

### Exercise 3

Classify the monthly returns of FI and IT sectors as follows and prepare a contingency 
table of counts. 


|**Monthly Return** |**Classification**|
|:----:|:----:|
|≤ First Quartile value |Low|
|> Third Quartile value |High|
|All others| Medium|


---

### Exercise 3


Classify the monthly returns of FI and IT 

```{r}
FI<-ifelse(indices$FI <= quantile(indices$FI,0.25),"Low",
           ifelse(indices$FI > quantile(indices$FI,0.75),"High",
                  "Medium")) 

```

```{r}
IT<-ifelse(indices$IT <= quantile(indices$IT,0.25),"Low",
           ifelse(indices$IT > quantile(indices$IT,0.75),"High",
                  "Medium")) 
```

---

#### Categorical Variable

```{r}
FI
```

---

#### Categorical Variable

```{r}
class(FI)
```

```{r}
table(FI)
```


---

```{r}

cut(x = indices$FI, 
    breaks = c(min(indices$FI),
               quantile(indices$FI,0.25),
               quantile(indices$FI,0.75),
               max(indices$FI)), 
    labels = c("Low","Medium","High"))

```

---

```{r}
table(FI,IT) 
```

---

### Exercise 4

Use the contingency table provided in Exercise 3 and test for the independence of 
monthly returns between FI and IT sector returns using an appropriate test and provide an 
interpretation of the test results. 

---

Test if the returns of FI and IT sectors are independent of each other

```{r}
chisq.test(FI,IT) 
```

---

### Interpretation
*  p-value < 0.05 indicating the rejection of null hypothesis of independe
nce of returns of both the sectors

* There is a lot of interdependence in the sectoral returns

* There were very few instances where one sector's returns were below the Q1 and the other sector's returns were above Q3

* The numbers in the diagonals are much higher to the ones in the off 
diagonals indicating the strength of the relationship

---

### Exercise 5

Test whether the returns of FI sector are significantly higher compared to that of IT Sector 
at 95% confidence level using appropriate test. 

---

### Exercise 5

Test whether the returns of FI sector are significantly higher compared to 
that of IT Sector

```{r}
t.test(indices$FI,indices$IT, alternative = "greater") 
```

---

#### Interpretation  

*  p-value =0.4094 > 0.05 indicates failure to reject the null hypothesis 
of the return of FI not greater than that of IT sector at 95% confidence level

*  There is no sufficient evidence to infer that the returns of FI sector 
are significantly higher than that of IT Sector

---

---