---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 1 - Binomial Probability Distribution
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Refer to the data file ***Indices_Returns.csv*** and answer the following questions:


```{r}
# Load the data file

indices<-read.csv("Indices_Returns.csv")
```

```{r}
head(indices)
```





---

### Exercise 1

Load the csv file into R and create a new column called “***Sensex_Direction***”. 

The value of this column will be “Positive” when the Sensex returns are positive and “Negative” when they are negative and convert the variable as a factor variable. 

---


#### New Categorical Variable

```{r}
# Creation of new column called Sensex_Direction

indices$Sensex_Direction <- ifelse(indices$Sensex>0,
                                   "Positive",
                                   "Negative")

indices$Sensex_Direction <- as.factor(indices$Sensex_Direction)

```

---


```{r}
head(indices)
```


---

### Exercise 2


Fit an appropriate generalized linear model (GLM) to with a ‘logit’ link function to relate the “Sensex_Direction” with the returns of 10 sectors as a multivariate model and display
the summary of the model. 



```{r}
# Fit the model and display the summary
model1 <- glm( Sensex_Direction ~ BM+CD+EN+FI+FM+HC+IN+IT+TE+UT,
               data = indices, 
               family = binomial(link = "logit"))
```

---

```{r}

summary(model1)

```

---

```{r}
library(broom)

tidy(model1)

```
---

## Exercise 

Identify which sectors have significantly impacted the direction of Sensex returns at 95% and 99% confidence level.

Verify the relationship between residual deviance of the model and the Akaike Information Criteria (AIC). 


---



### Exercise 4

Sectors significantly impacted

* Sectors which have significantly impacted the direction of Sensex retur
ns are CD, EN, FI, FM, IT and TE at 95% Confidence level

* But only FI has impacted the Sensex direction at 99% Confidence level


Relationship between the residual deviance and AIC
* all.equal(AIC(model1), model1$deviance+2*11)
* (11 number of model parameters (the number of variables in the model pl
us the intercept))


---


---

