---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Question 5
========================

Refer to the data file “Indices_Returns.csv” and answer the following questions:

Indices_Returns.csv file is provided in the system.

```{r}
indices<-read.csv("D:\\Indices_Returns.csv")
```


---

```{r}

head(indices)

```




---

i) Load the csv file into R and create a new column called “Sensex_Direction”. 

The value of this column will be “Positive” when the Sensex returns are positive and “Negative” when they are negative and convert the variable as a factor variable. 

---

```{r}
# Creation of new column called Sensex_Direction

indices$Sensex_Direction <-     ifelse(indices$Sensex>0,"Positive","Negative")

indices$Sensex_Direction <- as.factor(indices$Sensex_Direction)

```

---

Exercise 2
==================================

Fit an appropriate generalized linear model (GLM) to with a ‘logit’ link function to relate the “Sensex_Direction” with the returns of 10 sectors as a multivariate model and display
the summary of the model. 



```{r}
# Fit the model and display the summary
model1 <- glm( Sensex_Direction ~ BM+CD+EN+FI+FM+HC+IN+IT+TE+UT,
               data = indices, 
               family = binomial(link = "logit"))
```

---

## Exercise 

Identify which sectors have significantly impacted the direction of Sensex returns at 95% and 99% confidence level.

Verify the relationship between residual deviance of the model and the Akaike Information Criteria (AIC). 


---

\subsection*{Exercise 3}


---

Exercise 4
=================================

Sectors significantly impacted

* Sectors which have significantly impacted the direction of Sensex retur
ns are CD, EN, FI, FM, IT and TE at 95% Confidence level

* But only FI has impacted the Sensex direction at 99% Confidence level


Relationship between the residual deviance and AIC
* all.equal(AIC(model1), model1$deviance+2*11)
* (11 number of model parameters (the number of variables in the model pl
us the intercept))


