---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 1 - Binomial Probability Distribution
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Refer to the data file ***Indices_Returns.csv*** and answer the following questions:


```{r}
# Load the data file

indices<-read.csv("Indices_Returns.csv")
```

```{r}
head(indices)
```

---


```{r}
v) Plot the residuals of the fitted model and identify which month is the most significant
outlier in the residuals. (4)

vi) Comment on the appropriateness of the model fitted. (2)


Your actuarial friend has suggested that the current model can be improved by removing the
variables which do not impact the direction of Sensex returns at 95% confidence level and
refitting the GLM with ‘logit’ link function.

```


```{r}
##  TRUE
v)
Plot the residuals of Model
plot(model1$residuals)

Page 11 of 13IAI

which(model1$residuals==min(model1$residuals))
## 41
## 41
indices$Month[which(model1$residuals==min(model1$residuals))]

##  Jun-09
## 164 Levels: Apr-06 Apr-07 Apr-08 Apr-09 Apr-10 Apr-11 Apr-12 ... Sep
-19



---

vi)
Interpretation

 As the residual deviance came down significantly from Null Deviance of
223.21 to 32.90, the variables are able to classify the direction appropriately


*  One huge outlier (Jun-09) can impact the accuracy of the result (Removing this may reduce the residual deviance further)
*   The independent variables are not independent and they are interdependent (Correlations are very high among the sectors). 
Hence the standard errors may not be appropriate

*  As this data is a time series data serial correlation between the observations need to be considered and the model may have to be fitted after removing serial correlation.



```

----

vii) Update the model fitted in (ii) above, as suggested by your friend and display the summary
of the model.


viii) Compare the models in (ii) and (vii) using an appropriate test and comment on the
difference in the residual deviances between the two models.

---

vii)
Remove the variables that do not impact the model

```{r}


model2<-update(model1,~.-BM-HC-IN-UT)

## Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
summary(model2)


```
---
  
viii)
Call:
glm(formula = Sensex_Direction ~ CD + EN + FI + FM + IT + TE,
family = binomial(link = "logit"), data = indices)
Deviance Residuals:
Min
1Q
-2.08262 -0.00132
Median
0.00000
3Q
0.01757
Max
1.84630
Coefficients:
Estimate Std. Error z value Pr(>|z|)
(Intercept) -0.9623
0.6803 -1.414 0.15722
CD
-98.1771
37.5952 -2.611 0.00902 **
EN
94.2632
33.9930
2.773 0.00555 **
FI
174.4732
53.3659
3.269 0.00108 **
FM
37.7105
15.1808
2.484 0.01299 *
IT
78.7460
26.1203
3.015 0.00257 **
TE
30.7688
11.8875
2.588 0.00964 **
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
(Dispersion parameter for binomial family taken to be 1)
Null deviance: 223.213
Residual deviance: 33.646
AIC: 47.646
on 163
on 157
degrees of freedom
degrees of freedom
Number of Fisher Scoring iterations: 10

anova(model1,model2,test = "Chisq")
##
##
##
##
##
##
##
##

Analysis of Deviance Table
Model 1: Sensex_Direction
UT
Model 2: Sensex_Direction
Resid. Df Resid. Dev Df
1
153
32.905
2
157
33.646 -4
~ BM + CD + EN + FI + FM + HC + IN + IT + TE+
~ CD + EN + FI + FM + IT + TE
Deviance Pr(>Chi)
-0.74102
0.9462
Interpretation





---


p-value of the comparison is 0.94 > 0.05 thus not rejecting the null hypothesis of no significant difference between the two models. 

So the model did not improve significantly based on the friend's suggestion

\end{document}
