---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


```{r echo=FALSE}
library(magrittr)

options(repr.plot.width=8, repr.plot.height=5)
```

Suppose the transition probabilities are a function of the age of a person. The transition
probability of a person aged x moving:

* from Healthy to Sick state is 0.007x
* from Healthy to Death state is 0.001x
* from Sick to Death state is 0.002(100-x)
* from Sick to Healthy is 0.006(100-x).

#### Exercises

Assuming 100 is the maximum age. Calculate the probability of:

1.  Healthy person aged 30 will be in Sick state after 4 years.

2.  Sick person aged 25 will be in Death state after 7 years.

---

The transition matrix can be characterized as follows:
$$
\begin{pmatrix}
1-H2S(x)-H2D(x) & H2S(x) & H2D(x) \\
S2H(x) & 1-S2H(x)-S2D(x) & S2D(x) \\
0 & 0 & 1 \\
\end{pmatrix}
$$

N.B. This matrix varies for differing values of $x$, and hence the process is not a Markov Process.

---


```{r}

H2S<-function(x){ 0.007*x}
H2D<-function(x){ 0.001*x}
S2H<-function(x){ 0.006*(100-x)}
S2D<-function(x){ 0.002*(100-x)}



```


---

```{r}

transmat<-function(x){
  M<-matrix(0,nrow=3,ncol=3)
  M[1,1] <- 1 - H2S(x) - H2D(x)
  M[1,2] <- H2S(x)
  M[1,3] <- H2D(x)
  M[2,1] <- S2H(x)
  M[2,2] <- 1 - S2H(x) - S2D(x)
  M[2,3] <- S2D(x) 
  M[3,1] <- 0
  M[3,2] <- 0
  M[3,3] <- 1
  M
}


```

---

```{r}
transmat(30)
```

---

```{r}
transmat(25)
```

---

#### Exercise 1

```{r}

n <- 30
B <- c(1,0,0)

```

```{r}
i<- 1:4
n +i-1
```

---

```{r}
for (i in 1:4){
  B=B%*%transmat(n+i-1)
  }

```


```{r}
B
```
Hence the probability of Healthy person aged 30 will be in Sick state after 4 years is 0.2608.


---

#### Exercise 2.

```{r}
n <- 25
C <- c(0,1,0)
for (i in 1:7){
  C=C%*%transmat(n+i-1)
  }

```


```{r}
C
```

Hence the probability of sick person aged 25 will be in Death state after 7 years is 0.4333.

---

---