---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


```{r echo=FALSE}
library(magrittr)

options(repr.plot.width=8, repr.plot.height=5)
```

Solution 7: 
i)

```{r}
patients<-c(1:15) 
time<-c(6,6,12,18,27,27,30,36,39,39,54,57,60,60,60) 
censtat<-c(1,0,1,0,1,0,0,0,1,0,0,1,0,0,0) 
data1<-data.frame(patients,time,censtat) 
data1 
library(survival) 
model1<-survfit(Surv(data1$time,data1$censtat)~1) 
```

---


```{r}
data<-summary(model1) 
data 
```

> data 
Call: survfit(formula = Surv(data1$time, data1$censtat) ~ 1) 
 time n.risk n.event survival std.err lower 95% CI upper 95% CI 
 6 15 1 0.933 0.0644 0.815 1 
 12 13 1 0.862 0.0911 0.700 1 
 27 11 1 0.783 0.1115 0.593 1 
 39 7 1 0.671 0.1409 0.445 1 
 57 4 1 0.503 0.1797 0.250 1 

---
ii)

```{r}
tj=c(6,12,27,39,57) 
dj=c(1,1,1,1,1) 
nj=c(15,13,11,7,4) 
1-dj/nj 
```
---

Lambda=cumsum(dj/nj) 
SNA<-exp(-Lambda) 
SNA 
> 1-dj/nj 
[1] 0.9333333 0.9230769 0.9090909 0.8571429 0.7500000 
> SNA 
[1] 0.9355070 0.8662431 0.7909672 0.6856719 0.5340018 

--- 
iii)
survdata<-data.frame(tj,nj,dj,data$surv,SNA) 
survdata 
> survdata 
 tj nj dj data.surv SNA 
1 6 15 1 0.9333333 0.9355070 
2 12 13 1 0.8615385 0.8662431 
3 27 11 1 0.7832168 0.7909672 
4 39 7 1 0.6713287 0.6856719 
5 57 4 1 0.5034965 0.5340018 
