---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


```{r echo=FALSE}
library(magrittr)

options(repr.plot.width=8, repr.plot.height=5)
```


Consider a set of data generated by tossing a fair coin, letting $x_t = 1$ when a head is obtained
and x t = −1 when a tail is obtained. Then, construct y t as
$$y_t = 2 + 1.5x_t − 0.5x_{t−1}$$
Simulate data for three cases, one with a small sample size (n = 10), another one with a
moderate sample size (n = 100) and lastly with large sample size ($n = 1000$).


## Exercise 1
1.  Plot the series for all three cases.



```{r}

set.seed(101010)
x1 = 2*rbinom(11, 1, .5) - 1 # simulated sequence of coin tosses
x2 = 2*rbinom(101, 1, .5) - 1
x3 = 2*rbinom(1001, 1, .5) - 1
```


```{r}
y1 = 2 + filter(x1, sides=1, filter=c(1.5,-.5))[-1]
y2 = 2 + filter(x2, sides=1, filter=c(1.5,-.5))[-1]
y3 = 2 + filter(x3, sides=1, filter=c(1.5,-.5))[-1]

```
---

```{r}
plot.ts(y1, type='s')
```

---

```{r}
plot.ts(y2, type='s')
```

---

```{r}
plot.ts(y3, type='s')
```

---

## Exercise 2 

Find the mean and variance and comment on results.

```{r}
c(mean(y1), mean(y2),mean(y3)) # the sample mean
c(var(y1), var(y2),var(y3)) # the variance

```

---

```{r}
c(mean(y1), mean(y2),mean(y3)) # the sample mean
```

---

```{r}
c(var(y1), var(y2),var(y3)) # the variance

```

As the sample size increases, volatility reduces.

---

---