---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 2 - Correlation Analysis
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A student is performing a study to understand the correlation of temperature among 
different days of the week. 
Temperature on weekdays is recorded for 15 weeks in the file ***Temperature_data.csv***.

Generate the following correlation matrices

1. Pearson’s correlation matrix 

2. Kendall’s rank correlation matrix

3. Spearman’s rank correlation matrix

Comment on the outcomes of the results in parts 1 to 3.


---

#### Exercise 1: Pearson's Correlation

```{r}
Data = read.csv("Temperature_data.csv")
cor (Data, method = "pearson")
```

---

#### Exercise 1: Kendall's Tau

```{r}
cor (Data, method = "kendall")
```

---

#### Exercise 1

```{r}
cor (Data, method = "spearman")
```

---

#### Exercise 2

* The outcome of Pearson method is based on the values of the data whereas Kendall and 
Spearman correlation matrix is based on the rank of the values in the dataset.

* Diagonal in each matrix represents the correlation of temperature with respect to the individual 
days, for example Monday with Monday.

* Pearson shows strong positive correlation between temperatures on Monday, Tuesday, 
Thursday and Friday.

* Kendall and Spearman shows lesser positive correlation as compared to the Pearson’s result 

---

<pre><code>
library(corrplot)

M = cor(Data)
corrplot(M) 
</code></pre>

---

```{r echo =FALSE}
library(corrplot)
M = cor(Data)
corrplot(M) 
```

---

<pre><code>
library(corrplot)

M = cor(Data)
corrplot(M, method = 'number') 
</code></pre>

---


```{r echo =FALSE}
library(corrplot)

M = cor(Data)
corrplot(M, method = 'number') 
```


---

```{r}
library(faraway)
data(punting)

Data = punting 
```


---


```{r echo =FALSE}
library(corrplot)

M = cor(Data)
corrplot(M) 
```
---


```{r echo =FALSE}
library(corrplot)

M = cor(Data)
corrplot(M, method = 'number') 
```


---

### mtcars


```{r echo =FALSE}
library(corrplot)

M = cor(mtcars)
corrplot(M) 
```
---


```{r echo =FALSE}
library(corrplot)

M = cor(mtcars)
corrplot(M, method = 'number') 
```

