---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 2 - Correlation Analysis
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A child tosses $n$ coins and the outcome of heads and tails are recorded in $n$ samples as
$X_1, X_2, \ldots X_n$, where $X_i$’s are independent Bernoulli variables with $p = 0.5$. 

The total outcome of $n$ variables is $Y = X_1 +\ldots + Xn$

Use $n=10$ unless told otherwise

#### Exercises 

A. Specify the distribution of $Y$ 

B. Simulate a sample of 10 values for $Y$ 

C. Assess the value of $Y$ for the sample created in exercise B.  

D. What is the probability of $Y = 4$ 

---

### Exercise A.

* If $X_1, X_2, \ldots X_n$ follows Bernoulli distribution with $p = 0.5$ 

* Then $Y$ follows binomial distribution with $Y \sim \text{Bin}(n,p)$ as sum of independent Bernoulli distribution 
results in binomial distribution with $n$ trials and $p$ is the probability of success. 

---

### Exercise B.  


* The answer below is calculated using <tt>n = 10</tt>

#### R Code

```{r}
n = 10
p = 0.5
```

---

### Exercise B.  

```{r}
Y = rbinom(10, n, p)

Y
```




---

### Exercise C.
 
* The sample simulated in Exercise B are representative value of variable $Y$ itself. 

* Hence, no further 
calculation required to assess the value of variable Y. 
---

### Exercise D. 

```{r}
dbinom (4,n,p)
```


---


---