---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Assume X is a random variable which follows Poi (2) where $X = 1,2,3,4,5,6,7,8,9,10$ i.e. 
values from 1 to 10


#### Exercises

1. Calculate the probability for each of the values of $X$. 
2. Calculate the cumulative probability for each of the values of $X$.
3. Plot a graph of the distribution function of $X$. 


---

#### Exercise 1

```{r}
X  <-  c(1:10)
Lambda  <-  2 
```

#### Probability Mass Function $P(X=x)$

```{r}
dpois(X, Lambda)
```



---

### Exercise 2 

#### Cumulative Distribution Function $P(X \leq x)$

```{r}
ppois(X, Lambda)
```



---

### Exercise 2 

#### $P(X > x)$

```{r}
ppois(X, Lambda, lower.tail = FALSE)
```

```{r}
ppois(X, Lambda, lower.tail = FALSE) %>% round(4)
```

Remember: This is a discrete probability distribution


---

#### Exercise 3 


```{r}
plot(ppois(X,Lambda), pch=18,col="red",cex=2)
```

---


#### Exercise 3 



```{r}
plot(dpois(X,Lambda), pch=16,col="blue",cex=2)
```


---

#### Exercise 3 


```{r}
plot(ppois(X,Lambda,lower.tail = FALSE), pch=17,col="green",cex=2)
```

