---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Q. 5)
(Students can copy the R code as provided file ‘Reference_RCode’)

Claim payment data (Year 1 to Year 4) in INR Crores
Year 1
Year 2
Year 3
Year 4
Insurer A
112
130
178
150
Insurer B
38
50
80
68
Insurer C
89
127
210
150
Insurer D
70
75
77
80

---

#### Exercises

Part A Analyse the data using EBCT Model 1 and calculate the expected total claim payment to be made by each insurer (A, B, C and D) in the Year 5 by replacing question marks‘?’ with appropriate function in the R Code shared in file “Reference_RCode”.

B.  You are now required to calculate the expected total claim payment to be made by each insurer (A, B, C, D and E) in year 5, when additional data of insurer E is made available to you (By using EBCT Model 1).
Additional claim payment data for Insurer E in INR Crores
Year 1
Year 2
Year 3
Year 4
Insurer E
73
87
113
112
a) Provide the new R code so that ‘data’ includes insurer E and calculate the expected claim payment to be paid by each insurer in year 5.

b) Comment on the values of Z and ‘s/v’ ratio in part B  compared to Part A.

c) Comment on the change in expected claim payment for insurers A, B C and D in year 5 in part B compared to Part A.

C. Your manager is not happy with results based on EBCT 1 model. He has now provided you additional information about volume measure for these 5 insurers over Year 1 to Year 5.
Volume measure (Year 1 to Year 5)
Year 1
Year 2
Year 3
Year 4
Year 5
Insurer A
165
186
198
200
210
Insurer B
51
63
78
83
91
Insurer C
119
159
219
188
192
Insurer D
108
112
122
133
144
Insurer E
103
116
126
151
181

---

Analyse the data using EBCT Model 2 to calculate the expected total claim payment to be made by each insurer (A, B, C, D and E) in Year 5.

a) Complete the reference R Code (by replacing question marks‘?’ with appropriate function for data and zi, where data is as used in part (B.  to arrive at the ‘claims5i’ which represents the estimated claim payment for each insurer in Year 5.

b) Comment on change in expected claim payment for insurer A, B, C, D and E in year 5 in part (C. compared to part (B. .



---

Year 1 * Year 2 * Year 3 * Year 4 * 
Insurer A * 112 * 130 * 178 * 150 * 
Insurer  B * 38 * 50 * 80 * 68 * 
Insurer  C * 89 * 127 * 210 * 150* 
Insurer  D * 70 * 75 * 77 * 80 *


---
Solution 5:

```{r}
data=matrix(c(112,38,89,70,130,50,127,75,178,80,210,77,150,68,150,80),4,4)
 
n <- ncol(data)
m <- mean(rowMeans(data))
s <- mean(apply(data,1,var))
v <- var(rowMeans(data))-mean(apply(data,1,var))/n
Z <- n/(n+s/Part E

```

```{r}
Z*rowMeans(data)+(1-Z)*m
```


Expected total claim payment from 

--

*  insurer A is INR 138.09 Crores
*  insurer B is INR 64.48 Crores
*  insurer C is INR 139.41 Crores
*  insurer D is INR 79.02 Crores

As required for other parts
---

```{r}
n
```

```{r}
m
```

```{r}
s
```

```{r}
v
```

```{r}
Z
# [1] 0.8815584 
```
---

### Part (B)

B.  a) R Code
```{r}
data=matrix(c(112,38,89,70,73,130,50,127,75,87,178,80,210,77,113,150,68,150,80,112),5,4)

#no change in other code from part i

n <- ncol(data)
m <- mean(rowMeans(data))
s <- mean(apply(data,1,var))
v <- var(rowMeans(data))-mean(apply(data,1,var))/n
Z <- n/(n+s/Part E

Z*rowMeans(data)+(1-Z)*m
```


--- 


Expected total claim payment from
*  insurer A is INR 137.12 Crores (reduced compared to part Part A
*  insurer B is INR 65.13 Crores (increased compared to part Part A
*  insurer C is USD 138.41 Million (reduced compared to part Part A
*  insurer D is USD 79.35 Million (increased compared to part Part A
*  insurer E is USD 97.24 Million (newly added insurer)

required for other parts

```{r}
n
# [1] 4
m
# [1] 103.45
s
# [1] 824.05
v
# [1] 1288.5
Z
# [1] 0.862154
```

---


Provide the new R code so that ‘data’ includes insurer E and calculate the expected 
claim payment to be paid by each insurer in year 5. 

b) Comment on the values of Z and ‘s/v’ ratio in part (B.  compared to Part A. 

c) Comment on the change in expected claim payment for insurers A, B C and D in 
year 5 in part (B.  compared to Part A. 

C. Your manager is not happy with results based on EBCT 1 model. 
He has now provided you additional information about volume measure for these 5 insurers over Year 1 to Year 5.

---

b) Comparison of calculated quantities
*  With the introduction of insurer E, credibility factor Z has reduced. 

*  E[s2(Theta)] (i.e. ‘s’ in the R code used above) has reduced marginally – this has an effect of 
increasing Z.
* var[m(theta)] has reduced significantly (i.e. ‘v’ in the R code used above) - this has an effect 
of reducing Z.
*  Overall the ratio - ‘s’ / ‘v’ is 0.54 in part i; ‘s’ / ‘v’ increases to 0.64 in part ii – as value of 
denominator (in the calculation of Z = n/ (n + (s/Part E)) has increased in part ii compared to part 
i – hence, credibility factor Z has reduced in part ii compared to part i. 


---

c) Change in expected claim payment for each insurer
Expected claim payment for insurer is based on credibility value and equal to 
Z*individual average + (1-Z) * total average 
(0.5 mark for comment on formula)

---

*  Total average ‘m’ has reduced in part ii compared to part i as average of insurer E is lower 
than total average (m = 105.25) in part i

*  Expected claim payment for insurers A and C has reduced in part ii; as lower weight to individual experience compared to part i (A and C have experienced higher claim payments in 
the past and their average is higher than the total average of 5 insurers 103.45 - value of ‘m’ as calculated above) brings down the total expected value for these insurers towards 103.45

*  Expected claim payment for insurers B and D has increased in part ii; as lower weight to individual experience compared to part i (B and D’s individual average values are lower than 
103.45) and higher weight to total average (103.45) pulls up the total expected value for these 
insurers towards 103.45.

---

#### Exercise 3
C. a) Use data as used in part ii


```{r}
data=matrix(c(112,38,89,70,73,130,50,127,75,87,178,80,210,77,113,150,68,150,80,112),5,4)
#all other code lines as provide in part iii reference R code
volume <- 
matrix(c(165,51,119,108,103,186,63,159,112,116,198,78,219,122,126,200,83,188,133,151),5,4)
n <- ncol(data)
N <- nrow(data)
X <- data/volume #claim payment per unit of volume measure
```

```{r}
Xibar <- rowSums(data)/rowSums(volume)
Pi <- rowSums(volume) #volume measure for each insurer
P <- sum(PPart A
Pstar <- sum(Pi*(1-Pi/P))/(N*n-1)
m <- sum(data)/P #average claim payment per unit volume measure across all insurers
s <- mean(rowSums(volume*(X-Xibar)^2)/(n-1))
v <- (sum(rowSums(volume*(X-m)^2))/(n*N-1)-s)/Pstar
zi <- Pi/(Pi+s/Part E #credibility factor for each insurer
```

```{r}
credibilityi <- zi*Xibar+(1-zPart A*m #expected claim payment per unit of volume measure
volume5i <- matrix(c(210,91,192,144,181),5,1) #’volume5i’ is Year 5 volume measure 
#okay to have 1,5 instead of 5,1 in volume5i (row , column interchange)
claims5i <- credibilityi*volume5i #claims5i is the expected claim payment in year 5 
```

---

Additional output (marks not to be deducted in case if other parts are not provided by student)


```{r}
n
# [1] 4
N
# [1] 5
```

```{r}
X
```

---

```{r}
#Xibar
# [1] 0.7610147 0.8581818 0.8408759 0.6357895 0.7762097
Pi
# [1] 749 275 685 475 496
P
# [1] 2680
Pstar
# [1] 110.0728
m
# [1] 0.7720149
s
# [1] 1.095221
v
# [1] 0.00469698
zi
# [1] 0.7625928 0.5411516 0.7460447 0.6707376 0.6802203
credibility
# [1] 0.7636262 0.8186443 0.8233883 0.6806434 0.7748683
volume5
```


---

b)
claims5
 [,1]
[1,] 160.36151
[2,] 74.49663
[3,] 158.09055
[4,] 98.01265
[5,] 140.25116
Claims5 gives expected claim payment based on EBCT 2 
*  USD 160.36 Million for insurer A
*  USD 74.50 Million for insurer B
*  USD 158.09 Million for insurer C
*  USD 98.01 Million for insurer D
*  USD 140.25 Million for insurer E
(1 mark for identifying expected claim payment)
Expected claim payments for all insurers has increased compared to part ii as it makes use of 
Year 5 volume which is higher for all insurers.
(1 mark for comment on increase and 1 mark for comment on Year 5 volume) (3)

---


