---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 2 - Correlation Analysis
* Exam on basis of ***Base R***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A child tosses n coins and the outcome of heads and tails are recorded in n samples as
$X_1, X_2, \ldots X_n$, where, Xi’s are independent Bernoulli variables with $p = 0.5$. 
The total outcome of $n$ variables is $Y = X_1 +\ldots + Xn$


1. Specify the distribution of Y 

2. Simulate a sample of 10 values for Y 

3.  Assess the value of Y for the sample created in (B.  
4.  what is the probability of Y = 4 

---

#### Solution 3:

Part A If $X_1, X_2, \ldots X_n$ follows Bernoulli distribution with p = 0.5 

Then Y follows binomial distribution ~ Bin(n,p) as sum of independent Bernoulli distribution 
results in binomial distribution with n trials and p is the probability of success. 

---

### Part B.  

Since ‘n’ is not specified in the question, sample simulations of Y can be generated assuming any 
value of ‘n’. 

The answer below is calculated using <tt>n = 10</tt>, however marks are allotted to 
simulations of Y, calculated using any value of ‘n’.


#### R Code

```{r}
n = 10
p = 0.5
```

```{r}
Y = rbinom(10, n, p)

Y
```

Values of y from R (since these are random values, answers for each individual may vary from 
the numbers stated below, please evaluate accordingly)
5 4 4 5 2 6 5 4 2 5 
---

### Part C.
 
* The sample simulated in part (B.  are representative value of variable Y itself. 

* Hence, no further 
calculation required to assess the value of variable Y. 
---

### Part D. 

<pre><code>
dbinom (4,n,p)
</code></pre>

The solution is based on assumption that n = 10 (as assumed in part(B. ). Marks are allotted to 
solutions calculated using any value for “n”.
0.2050781 

---