---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

A child tosses n coins and the outcome of heads and tails are recorded in n samples as
X1, X2, …… Xn, where, Xi’s are independent Bernoulli variables with p = 0.5. The total 
outcome of n variables is Y = X1 +……. + Xn


1. Specify the distribution of Y (2)

2. Simulate a sample of 10 values for Y 

3.  Assess the value of Y for the sample created in (ii) (2)
4.  what is the probability of Y = 4 

---

#### Solution 3:

i) If X1, X2, …… Xn follows Bernoulli distribution with p = 0.5 (1)

Then Y follows binomial distribution ~ Bin(n,p) as sum of independent Bernoulli distribution 
results in binomial distribution with n trials and p is the probability of success. (1)

---

ii) Since ‘n’ is not specified in the question, sample simulations of Y can be generated assuming any 
value of ‘n’. 

The answer below is calculated using <tt>n = 10</tt>, however marks are allotted to 
simulations of Y, calculated using any value of ‘n’.
R Code

```{r}
n = 10
p = 0.5
```

```{r}
Y = rbinom(10, n, p)

Y
```

Values of y from R (since these are random values, answers for each individual may vary from 
the numbers stated below, please evaluate accordingly)
5 4 4 5 2 6 5 4 2 5 (2)
---

iii) 
* The sample simulated in part (ii) are representative value of variable Y itself. 

* Hence, no further 
calculation required to assess the value of variable Y. 
---

iv) dbinom (4,n,p)
The solution is based on assumption that n = 10 (as assumed in part(ii)). Marks are allotted to 
solutions calculated using any value for “n”.
0.2050781 

