---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3 - Hypothesis Testing with ***t.test()***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```
Q. 6) 

You are investigating the level of premium charged by two companies for certain group. 
Random samples of 10 policies from Company1 and Company2 are compared. 

The following Below R 
code provides the premiums charged by Company1 and Company2 in current year for 
10 sample policies.

```{r}
 Company1 <- c(1350,1790,1500,1150,2100,2350,1550,1800,1650,1450)
 Company2 <- c(1500,1200,1300,1700,1800,2400,1450,1950,1850,2100)
 
```


Students are required to provide hypothesis, R code, Output and conclusion

---

A. Assuming that the premiums are normally distributed, carry out a statistical test to 
check equal variance assumption so that it is appropriate to apply a two-sample t 
test to these data. 
 Use R code – var.test 
B. Test whether the level of premiums charged by Company1 and Company2 are 
same.
 Use R code – t.test (use var.equal =TRUE) 
C. The average premium charged by Company2 in the previous year was INR 1500. 
Test whether Company2 appears to have increased its premiums since the previous 
year.
 Use R code -t.test (#one sided) 




---

#### Exercise A

Part A H0 – Company1 and Company2 having equal variance (i.e., ratio of variance =1)
H1 – There is a difference in variance between Company1 and Company2.

```{r}
Company1 <- c(1350,1790,1500,1150,2100,2350,1550,1800,1650,1450)
Company2 <- c(1500,1200,1300,1700,1800,2400,1450,1950,1850,2100)
```

---

#### Test for Equality of Population Variances

```{r}
var.test(Company1,Company2)
```

---

#### Conclusion 

* As the confidence interval for the ratio between variance includes 1, we do not 
have enough evidence to reject H0 at 5% significance level. 

* Hence, we can conclude that ratio of variance between Company1 and Company2 is 1 i.e., Company1 and Company2 having equal 
variance at 5% significance level. 

---
B.  

### Part B

H0 – No difference in mean premium charged by Company1 and Company2. 
H1 – difference in mean premium charged by Company1 and Company2.

```{r}
t.test(Company1, Company2, var.equal = TRUE, conf=0.95)
```
 
---
#### Conclusion 

* As the confidence interval for the difference in mean premium includes 0, we do not have enough evidence to reject H0 at 5% significance level.
* Hence, we can conclude that there is no difference between mean premium charged by Company1 and Company2 at 5% significance level. 


---

C. H0 – mean premium charged by Company2 is INR 1500. 
H1 – mean premium charged by Company2 is greater than 1500.

```{r}
t.test(Company2, mu=1500, alternative= c("greater"), conf=0.95)
```
 One Sample t-test
data: Company2
t = 1.9082, df = 9, p-value = 0.04436
alternative hypothesis: true mean is greater than 1500
95 percent confidence interval:
1508.858 Inf
sample estimates:
mean of x 
 1725

---
 
* Conclusion – As the confidence interval for the mean premium does not include 1500, at 5% 
significance level we have enough evidence to reject null hypothesis. 
(p-value less than 5% also confirms that we have enough evidence to reject null hypothesis at 
5% significance level).
* Hence, we can conclude that the mean premium charged by Company2 is greater than INR 1500 
at 5% significance level. 

---

#### Exercise D

* It was found that out of sample of 200 policies of Company1 and 100 policies of Company2 sold at the start of the year, 60 policies of Company1 and 50 policies of Company2 resulted in claim. 

* Carry out a hypothesis test for the difference in proportions.

* Use R code – prop.test (use <tt>correct=FALSE</tt>)

 
---

#### Part 4

H0 – There is no difference in proportion of (policies that result in) claim between Company1 and 
Company2.

H1 – There is difference in proportion of (policies that result in) claim between Company1 and 
Company2.

```{r}
prop.test(c(60,50),c(200,100),correct=FALSE)
```
 2-sample test for equality of proportions without continuity correction
data: c(60, 50) out of c(200, 100)
X-squared = 11.483, df = 1, p-value = 0.0007023
alternative hypothesis: two.sided
95 percent confidence interval:
-0.31677833 -0.08322167
sample estimates:
prop 1 prop 2 
 0.3 0.5
 
---
 
#### Conclusion 

* As the confidence interval does not include 0 (or as p-value is less than 5%), we have sufficient evidence to reject null hypothesis at 5% significance level. 

* Hence, we can conclude that there is difference in proportion of (policies that result in) claims between  Company1 and Company2 at 5% significance level. 


--- 
#### Exercise E

Part E Company2 wants to study the claim frequency between group1 policies having 
premium less than INR 1500 and remaining policyholders. There were 65 claims 
out of 250 policies of group1 and 45 claims out of 110 other policies in a year. 

Assuming number of claims having Poisson distribution, test at 2.5% level whether 
the ratio of claim frequency between group1 and other policyholders is less than 1.
 Use R code – poisson.test (#one sided)
 
---

Part E 

H0 – There is no difference in ratio of claim frequency between group1 and other groups in 
Company2. 
H1 – Claim frequency of group1 is less than other groups in Company2.

```{r echo=FALSE}
poisson.test(c(65,45),c(250,110), conf=.975, alternative="less")
```

---

```{r include= FALSE}
poisson.test(c(65,45),c(250,110), conf=.975, alternative="less")
```

 Comparison of Poisson rates
data: c(65, 45) time base: c(250, 110)
count1 = 65, expected count1 = 76.389, p-value = 0.01358
alternative hypothesis: true rate ratio is less than 1
97.5 percent confidence interval:
0.0000000 0.9511753
sample estimates:
rate ratio 
0.6355556

---

* Conclusion – As the confidence interval for the ratio does not include 1, at 2.5% significance level 
we have enough evidence to reject null hypothesis. 
(p-value less than 2.5% also confirms that we have enough evidence to reject null hypothesis at 
2.5% significance level).
* Hence, we can conclude that the claim frequency of group1 (having premium less than INR 1500) 
is less than other groups in Company2. 
