---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

For a sample data X =\{ 0.5820, 0.04981, 0.1552, 0.1555, 
0.9036, 0.8501, 0.9288, 0.4408, 
0.9688, 0.6300\}


### Exercises

1. Find the mean value of X 

2. Calculate the standard deviation of X 

3. Find the median of X 

---



```{r}
X <- c(0.5820, 0.04981, 0.1552, 0.1555, 0.9036, 0.8501, 0.9288, 0.4408, 0.9688, 0.6300)
```

---

#### Summary Statistics

```{r}
mean(X)
```

```{r}
sd(X)
```

```{r} 
median(X)
```

---

#### Sort

```{r}
sort(X)
```

#### Reverse the order

```{r}
rev(X)
```



---


```{r}
X
```

#### which,max and which.min()

```{r}
which.min(X)
```

```{r}
which.max(X)
```

---

#### Order

```{r}
order(X)
```


#### Rank and Order

```{r}
cbind(X, rank(X) )
```

---

---