---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

Q. 1) An Actuarial student fits following simple regression model to the data
yi = alpha + beta * xi + ei ; i =1 to 12
where ei are independent normal random variables with mean 0 and variance sigma2
Use following 12 data points for x and y 
where y is response variable while x is explanatory variable

```{r]
x = c(5,10,15,20,25,30,35,40,45,50,55,60)
y = c(15,12,25,23,35,36,33,38,43,45,50,53)
```

Note: Do not use standard ‘model fitting related’ R codes - lm, glm, fitted, residuals, 
predict, anova - to answer parts of this question.

---

i) Calculate Sxx, Sxy and Syy. (7)
ii) Calculate alpha, beta and sigma2
using results in part (i). (3)
iii) Calculate fitted values of y using results in part (ii). (1)
iv) Calculate residuals using results of earlier parts and Calculate mean and variance of 
residuals and comment on the result. (3)
v) Calculate 95% confidence interval for beta and comment if we can conclude that beta 
is not zero stating the Null and alternate hypothesis. Calculate 95% confidence interval 
for sigma2
.

--- 
 
Note: Candidates are expected to find tabulated value using R. (7)
vi) State SSTOT and Calculate SSREG and SSRES. Hence calculate the proportion of 
variability explained by the model using SSREG and SSRES and comment on the result. 
Calculate adjusted R2
and compare it with R^2 to explain result. (5)
vii)Calculate mean predicted response when x=52 and 95% confidence interval for the 
same. Note: Candidates are expected to find tabulated value using R. 

---

Institute of Actuaries of India
Subject CS1-Actuarial Statistics (Paper B)
March 2022 Examination

---


### Solution 1:

#### PArt i) 

```{r}
x= c(5,10,15,20,25,30,35,40,45,50,55,60)
y=c(15,12,25,23,35,36,33,38,43,45,50,53)
```

```{r}
meanx = mean(x)
meany=mean(y)
meanx
meany
```

---

```{r}
x_sq=x*x
x_sq
```

[1] 25 100 225 400 625 900 1225 1600 2025 2500 3025 3600
```{r}
y_sq=y*y
xy=x*y
xy
```

---

```{r}

sumx_sq=sum(x_sq)
sumy_sq=sum(y_sq)
sumxy=sum(xy)
sumx_sq
#[1] 16250
sumy_sq
#[1] 15760
sumxy
#[1] 15775
```

---

```{r}
Sxx=sumx_sq-12*meanx^2
Sxx
```

---

```{r}
Sxy=sumxy-12*meanx*meany
Sxy
```

---

```{r}
Syy=sumy_sq-12*meany^2
Syy

```

---


```{r}

beta=Sxy/Sxx
beta
# [1] 0.7034965

alpha=meany-beta*meanx
alpha
# [1] 11.13636

```

---


```{r}

sigmasq=(1/(12-2))*(Syy-Sxy^2/Sxx)
sigmasq
#[1] 11.87063


```

---


```{r}

expectedy=alpha+beta*x
expectedy

```

#[1] 14.65385 18.17133 21.68881 25.20629 28.72378 32.24126 35.75874 39.27622 42.79371 
#46.31119 49.82867 53.34615

---

```{r}

e=y-alpha-beta*x
e

```

[1] 0.3461538 -6.1713287 3.3111888 -2.2062937 6.2762238 3.7587413
[7] -2.7587413 -1.2762238 0.2062937 -1.3111888 0.1713287 -0.3461538


```{r}
meane=mean(e)
meane
 var(e)
#[1] 10.79148
```

---

Mean value of residuals is close to zero as expected as e~N(0,sigma^2)
(Otherwise, “e” could be calculated as e = y-expectedy)
Var of e is slightly lower than sigma square as calculated in part ii – as denominator is not adjusted
When denominator of 10 gets used instead of 11 we see that var of residuals = sigma^2 

```{r}
var(e)*11/10
```


```
---

```{r}

v) 95% confidence interval for beta
Ho: Beta is zero (i.e. no linear relationship between x and y)
H1: Beta is not equal to zero
(Beta_cap-0)/ sqrt(sigma^2_cap/ Sxx) ~ t10
We use t distribution with n-2 i.e. 10 degrees of freedom
> qt(p=0.025, lower.tail = T, df=10)
[1] -2.228139
Being symmetric distribution, 97.5% point would be 2.228139
> sqrt(sigmasq/Sxx)
[1] 0.0576234
Hence, endpoints of CI would be
> end1=beta+sqrt(sigmasq/Sxx)*qt(p=0.025, lower.tail = T, df=10)
> end1
[1] 0.5751036

```
---

```{r}

> end2=beta-sqrt(sigmasq/Sxx)*qt(p=0.025, lower.tail = T, df=10)
> end2
[1] 0.8318894
Hence 95% Confidence interval for beta is (0.5751, 0.8319)
As confidence interval for beta does not include zero, we can reject null hypothesis (viz. beta=0) and 
Hence, can conclude that beta is not equal to zero at 5% level.
95% CI for sigma^2
(n-2)sigmacap^2/ sigma^2 ~ Chi sq distribution with 10 degrees of freedom

---

Tabulated values of Chi square having 10 df can be obtained as
> chitenend1=qchisq(df=10, p=0.025)
> chitenend2=qchisq(df=10,p=0.975)
> chitenend1
[1] 3.246973
> chitenend2
[1] 20.48318
End points of CI would be
> sigmasqend1=(12-2)*sigmasq/chitenend1
> sigmasqend2=(12-2)*sigmasq/chitenend2
> sigmasqend1
[1] 36.55907
> sigmasqend2
[1] 5.795307
Hence 95% Confidence interval for sigma^2 is (5.795,36.559)

```
---

```{r}

vi)
SSTOT = Syy = 1888 (as calculated in part i)
SSREG = Sxy^2/Sxx 
> ss_reg=Sxy^2/Sxx
> ss_reg
[1] 1769.294
SSRES = SSTOT - SSREG
> ss_res=Syy-ss_reg
> ss_res
[1] 118.7063
R^2 denotes the % of variability explained by the model
R^2 = SSREG / (SSREG + SSRES)
> Rsq = ss_reg/(ss_reg+ss_res)
> Rsq
[1] 0.9371259
Model is a good fit as 93.7% of the variability is explained by the model.

---

```{r}

> adj_Rsq = 1-((12-1)/(12-1-1))*(1-Rsq)
> adj_Rsq
[1] 0.9308385
Adjusted R^2 (93.08%) is lower than R^2 (93.71%) as adjusted R square penalises for extra predictors 
and 
hence is better suited to assess the adequacy of the model (or for comparison between models) 
compared to just using R^2 for model comparison as 
R^2 cannot decrease on addition of more explanatory variables which can be undesirable (as it may 
promote too many explanatory variables though not adding significant improvement in the predicted 
value)
```
---

```{r}
vii)
using results from earlier parts mean predicted response is calculated (using regression line 
equation)
> Emean52=alpha+beta*52
> Emean52
[1] 47.71818

---
  
Expected value of mean predicted response is 47.718 when x=52
varofmean52=((1/12)+(52-meanx)^2/Sxx)*sigmasq
> varofmean52
[1] 2.251822
> mean52end1=Emean52+qt(p=0.025, lower.tail = T, df=10)*sqrt(varofmean52)
> mean52end2=Emean52-qt(0.025,lower.tail = T, df=10)*sqrt(varofmean52)
> mean52end1
[1] 44.37462
> mean52end2
[1] 51.06174
Hence 95% confidence interval for the mean predicted response is (44.3746,51.0617)

---

