---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 2

---

Q. 2) Policy and claims information (PolicyData.csv) of 650 policies is provided to you. The 
data contains following fields:

* Policy : Policy Number
* Claim : Number of claims corresponding to each policy
* Cust_Exp: Policyholder’s experience (VS = Very satisfied, SA= Satisfied, DS= 
Disappointed, VD= Very Disappointed) at the end of the policy tenure.

Amount: Claim amount per policy. Note that amount is set 0 if there is no claim.

---
i) Create a frequency table of claim and share how many policies don’t have any claim. 
ii) Plot a histogram of claim count and suggest 2 distributions that can be a good fit. 
iii) Given that claim count follows Poisson distribution with following two possible values 
for Poisson parameter:
a) 0.35 and 
b) 0.30 
---

Compute confidence interval at 95% confidence level to assess which value is more 
suitable for the given data. 
iv) Compute mean, variance and median of log of claim amount. (Name it: log amount). 
Make sure to exclude policies with no claim. (4)
v) Obtain histogram and Normal QQ Plot of log amount. Add a line to the QQ plot for 
normal distribution 
vi) Indicate which distribution the claim amount might be following using evidence from 
(iv) and (v). 
---
vii)Assuming the log amount (i.e. log of claim amount) follows a Normal distribution, test 
if mean of log amount is greater than 10 at 90% level of confidence. State the 
hypothesis and conclusion clearly. 
viii) Assess whether the policyholder experience (i.e. Cust_Exp) changes with more 
number of claims. Create contingency table and perform test to check the above 
assertion. State the hypothesis clearly. 
---
ix) Please explain why warning message appears while performing the above test. 
x) Again, perform the above test by combining 
a. Very Satisfied and Satisfied customers
b. Disappointed and very disappointed customers
and 
c. 2 or more claims. 
---
Please provide your conclusion on association of policyholder’s experience with 
number of claims. (5)
xi) Amount is defined to be large if the amount is greater than 100,000. Calculate 95% 
confidence interval for proportion of large claim, and comment on the likelihood if 
more than 25% of claims are large. 
---
Solution 2:
i)


```{r}
policydata <- read.csv("PolicyData.csv")
```

---

```{r}
library(dplyr)
str(policydata)

```

```{r}
#a
table(policydata$Claim)
# 0 1 2 3 
# 458 149 36 7 
```

---
#Alternative, if dplyr installed
#count(policydata,Claim)
print("458 Policies don't have any claim")
[1] "458 Policies don't have any claim"
(2)
ii)
```{r}
hist(policydata$Claim)
#poisson and negative binomial distribution
```

---

iii)

```{r}
poisson.test(x=sum(policydata$Claim),T=length(policydata$Policy))
```

---

Exact Poisson test
data: sum(policydata$Claim) time base: length(policydata$Policy)
number of events = 242, time base = 650, p-value < 2.2e-16
alternative hypothesis: true event rate is not equal to 1
95 percent confidence interval:
0.3268739 0.4222903
sample estimates:
event rate 
0.3723077 
#0.35 is more suitable value of parameter since it lies between confidence interval.

---
iv)

```{r}
lx=log(policydata$Amount[policydata$Amount>0])
#Alternative, if dplyr installed
#lx=log(filter(policydata,Amount >0)$Amount)
mean(lx)
```
[1] 9.835205
median(lx)
[1] 9.774659
sd(lx)^2
[1] 3.425705

---
Histogram of policydata$Claim
policydata$Claim
Frequency
0.0 0.5 1.0 1.5 2.0 2.5 3.0
0 200 400

---

v)
```{r}
par(mfrow=c(2,1))
hist(lx)
qqnorm(lx)
qqline(lx)
```

---
vi)

From Histogram and QQPlot it seems log amount closely follows normal distribution.
To add, the mean and median are very close indicating symmtery. One of the characterstics of Z. 
Hence,Claim amount might be following log normal distribution. 

---
vii)
#Null Hypothesis : mu = 10 , alternate hypothesis mu >10
t.test(lx,mu=10,alternative="greater", conf.level = .9)
One Sample t-test
data: lx
t = -1.2337, df = 191, p-value = 0.8906
alternative hypothesis: true mean is greater than 10
90 percent confidence interval:
9.663428 Inf
sample estimates:
mean of x 
Histogram of lx
lx Frequency
6 8 10 12 14
0 10 20 30 40
-3 -2 -1 0 1 2 3
6 8 10 14
Normal Q-Q Plot
Theoretical Quantiles
Sample Quantiles
---
9.835205 
#Given p-value greater than 10% null hypothesis can not be rejected.


---
viii)
ct=table(policydata$Claim,policydata$Cust_Exp)
ct
 
 DS SA VD VS
 0 63 306 20 69
 1 36 90 9 14
 2 16 14 6 0
 3 3 3 1 0
# Null Hypothesis: No association between Policyholder's experience and Claim
chisq.test(ct)
Pearson's Chi-squared test
data: ct
X-squared = 47.749, df = 9, p-value = 2.846e-07
Warning message:
In chisq.test(ct) : Chi-squared approximation may be incorrect

---
ix)
#There are cells where the number of observations are less than 5. (1)
x)
policydata$Claim2=ifelse(policydata$Claim >2,2,policydata$Claim) 
policydata$Cust_Exp2=ifelse(policydata$Cust_Exp %in% c("DS","VD"),"DS","SA") 
ct2=table(policydata$Claim2,policydata$Cust_Exp2)
ct2
 
 DS SA
 0 83 375
 1 45 104
 2 26 17

---
chisq.test(ct2)
Pearson's Chi-squared test
data: ct2
X-squared = 43.514, df = 2, p-value = 3.557e-10
# There is a strong reason to reject null hypothesis.
# Hence, it can concluded that policyholder's experience gets worse as 
claim count increases
---
(5)
xi)
summary(policydata$Amount)
########################
Page 9 of 13
 Min. 1st Qu. Median Mean 3rd Qu. Max. 
 0 0 0 29501 3232 1848069 
---
policydata$large= ifelse(policydata$Amount >100000,1,0)
x = sum(policydata$large)
n = length(policydata$Amount[policydata$Amount>0])
#Alternative, if dplyr installed
#n = length(filter(policydata,Amount >0)$Amount)
---
binom.test(x,n)
Exact binomial test
data: x and n
number of successes = 35, number of trials = 192, p-value < 2.2e-16
alternative hypothesis: true probability of success is not equal to 0.5
95 percent confidence interval:
0.1303796 0.2442928
sample estimates:
probability of success 
 0.1822917 

---

# Since upper bound of c.i is less that .25, it is unlikely that more 
that 
#25% claims are large
(5)
---