---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}

/* Style for <pre> tags, commonly used for code blocks */
pre {
  background-color:  #ADD8E6;  /* Light grey background */
  padding: 10px;              /* Some padding for spacing */
  border-radius: 5px;         /* Rounded corners */
}

/* Style for <code> tags, commonly used for inline code */
code {
  background-color: #ADD8E6; /* Slightly darker grey for inline code */
  padding: 2px 4px;           /* Padding for inline code */
  border-radius: 3px;         /* Slightly rounded corners */
}
```

An Actuarial student fits following simple regression model to the data 
yi = alpha + beta * xi + ei  ;  i =1 to 12 
where ei are independent normal random variables with mean 0 and variance sigma2 
 
Use following 12 data points for x and y  
where y is response variable while x is explanatory variable 
 
```{r} 
x = c(5,10,15,20,25,30,35,40,45,50,55,60) 
y = c(15,12,25,23,35,36,33,38,43,45,50,53) 
```
 
Note: Do not use standard ‘model fitting related’ R codes - <t>lm</tt>, <ttglm</tt>, fitted, residuals, 
predict, anova - to answer parts of this question.  
   
---

### Exercises
   
A. Calculate Sxx, Sxy and Syy. 
   
B. Calculate alpha, beta and sigma2 using results in exercise A. 
   
C. Calculate fitted values of y using results in exercise B. 
   
D. Calculate residuals using results of earlier parts and Calculate mean and variance of 
residuals and comment on the result. 
   
E. Calculate 95% confidence interval for beta and comment if we can conclude that beta 
is not zero stating the Null and alternate hypothesis. Calculate 95% confidence interval 
for sigma2.                                      
(Note: Candidates are expected to find tabulated value using R.)
   
F. State SSTOT and Calculate SSREG and SSRES. Hence calculate the proportion of 
variability explained by the model using SSREG and SSRES and comment on the result.  
Calculate adjusted R2 and compare it with R^2 to explain result.           
   
G. Calculate mean predicted response when x=52 and 95% confidence interval for the 
same. Note: Candidates are expected to find tabulated value using R. 

---
### ExerciseA. Calculate Sxx, Sxy, and Syy

Given the data points:
- \(x = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60)\)
- \(y = c(15, 12, 25, 23, 35, 36, 33, 38, 43, 45, 50, 53)\)

First, we calculate the means of x and y:
$$\bar{x} = \frac{\sum{x}}{n} = \frac{5+10+\ldots+55+60}{12} = 32.5$$
$$\bar{y} = \frac{\sum{y}}{n} = \frac{15+12+\ldots+50+53}{12} = 34.42$$

---

Now, we calculate \(Sxx\), \(Sxy\), and \(Syy\):
$$Sxx = \sum{(x_i - \bar{x})^2} = \sum{(x_i - 32.5)^2}$$
$$Sxy = \sum{(x_i - \bar{x})(y_i - \bar{y})} = \sum{(x_i - 32.5)(y_i - 34.42)}$$
$$Syy = \sum{(y_i - \bar{y})^2} = \sum{(y_i - 34.42)^2}$$

Here are the actual values:
$$Sxx = 2475$$
$$Sxy = 1994.5$$
$$Syy = 2473.5$$


---
### Exercise A. Calculate Sxx, Sxy, and Syy

```{r}
# Data
x <- c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60)
y <- c(15, 12, 25, 23, 35, 36, 33, 38, 43, 45, 50, 53)

# Means
x_mean <- mean(x)
y_mean <- mean(y)
```

---
  
### Exercise A. Calculate Sxx, Sxy, and Syy

```{r}

# Sxx, Sxy, Syy
Sxx <- sum((x - x_mean)^2)
Sxy <- sum((x - x_mean) * (y - y_mean))
Syy <- sum((y - y_mean)^2)

Sxx
Sxy
Syy
```

---
### ExerciseB. Calculate α, β, and σ²

The formulas for α (intercept), β (slope), and σ² (variance of residuals) are as follows:
$$\beta = \frac{Sxy}{Sxx} = \frac{1994.5}{2475} \approx 0.806$$
$$\alpha = \bar{y} - \beta \bar{x} = 34.42 - (0.806 \cdot 32.5) \approx 8.187$$
$$\sigma^2 = \frac{1}{n-2} \sum{e_i^2} = \frac{1}{10} \sum{(y_i - \hat{y_i})^2}$$

Using these:
$$\hat{y_i} = \alpha + \beta x_i = 8.187 + 0.806 \cdot x_i$$
$$\sum{(y_i - \hat{y_i})^2} \approx 697.83$$
$$\sigma^2 \approx 69.78$$


---
### ExerciseB. Calculate α, β, and σ²

```{r}
# Beta (slope)
beta <- Sxy / Sxx

# Alpha (intercept)
alpha <- y_mean - beta * x_mean

# Residuals and sigma2
fitted_values <- alpha + beta * x
residuals <- y - fitted_values
sigma2 <- sum(residuals^2) / (length(x) - 2)

alpha
beta
sigma2
```





---
### ExerciseC. Calculate fitted values of y

Using the fitted model \( \hat{y_i} = \alpha + \beta x_i \):
$$\hat{y} = 8.187 + 0.806 \cdot x_i$$

Calculated fitted values:
$$ \hat{y} = c(12.217, 16.247, 20.277, 24.307, 28.337, 32.367, 36.397, 40.427, 44.457, 48.487, 52.517, 56.547) $$

---
### ExerciseC. Calculate fitted values of y

```{r}
fitted_values
```


---
### ExerciseD. Calculate residuals and their mean and variance

Residuals:
$$ e_i = y_i - \hat{y_i} = c(2.783, -4.247, 4.723, -1.307, 6.663, 3.633, -3.397, -2.427, -1.457, -3.487, -2.517, -3.547)$$
Mean of residuals: 
$$ \frac{\sum e_i}{n} \approx 0$$
Variance of residuals:
$$ \frac{\sum (e_i - \bar{e})^2}{n-1} \approx 69.78$$
Comment: Residuals have zero mean and show a high variance, indicating some deviations in model fitting.



---
### ExerciseD. Calculate residuals and their mean and variance

```{r}
mean_residuals <- mean(residuals)
var_residuals <- var(residuals)

mean_residuals
var_residuals
```

---
### ExerciseE. 95% confidence interval for β and σ²

For β:
$$ CI_{\beta} = \beta \pm t_{\alpha/2} \cdot SE_{\beta}$$
Using \(t_{\alpha/2}\) from t-distribution for df = 10 and SE of β:
$$ SE_{\beta} = \sqrt{\frac{\sigma^2}{Sxx}} \approx 0.168$$
$$ CI = 0.806 \pm 2.228 \cdot 0.168 \approx (0.433, 1.179)$$
Null hypothesis: \(H_0: \beta = 0\)
Since CI does not contain 0, we reject \(H_0\).

For σ²:
$$ CI_{\sigma^2} = \frac{(n-2)\sigma^2}{\chi_{(1-\alpha/2)}^2}, \frac{(n-2)\sigma^2}{\chi_{\alpha/2}^2}$$
For df=10, 95% CI using chi-square distribution values:
$$ CI = (27.20, 193.51)$$


---
### ExerciseE. 95% confidence interval for β and σ²

```{r}
# Standard error of beta
se_beta <- sqrt(sigma2 / Sxx)

# t value for 95% confidence interval
t_value <- qt(0.975, df = length(x) - 2)

# Confidence interval for beta
beta_CI <- beta + c(-1, 1) * t_value * se_beta

# Chi-square values for 95% confidence interval for sigma2
chi2_low <- qchisq(0.025, df = length(x) - 2)
chi2_high <- qchisq(0.975, df = length(x) - 2)

# Confidence interval for sigma2
sigma2_CI <- (length(x) - 2) * sigma2 / c(chi2_high, chi2_low)

beta_CI
sigma2_CI
```




---
### ExerciseF. SSTOT, SSREG, SSRES, R², and adjusted R²

$$ SSTOT = Syy = 2473.5$$
$$ SSREG = \beta \cdot Sxy = 1608.48$$
$$ SSRES = SSTOT - SSREG = 865.02$$
Proportion explained:
$$ R^2 = \frac{SSREG}{SSTOT} \approx 0.65$$

Adjusted \( R^2 \):
$$ R_{adj}^2 = 1 - \frac{SSRES/(n-2)}{SSTOT/(n-1)} \approx 0.61$$

---
### ExerciseF. SSTOT, SSREG, SSRES, R², and adjusted R²

```{r}
# Total Sum of Squares
SSTOT <- Syy

# Regression Sum of Squares
SSREG <- beta * Sxy

# Residual Sum of Squares
SSRES <- SSTOT - SSREG

# R²
R2 <- SSREG / SSTOT

# Adjusted R²
R2_adj <- 1 - ((SSRES / (length(x) - 2)) / (SSTOT / (length(x) - 1)))

SSTOT
SSREG
SSRES
R2
R2_adj
```

---
### ExerciseG. Mean predicted response when x=52 and 95% confidence interval

Using fitted model:
$$ \hat{y_{52}} = \alpha + \beta \cdot 52 = 49.31$$
95% CI:
$$ \hat{y_{52}} \pm t_{\alpha/2} \cdot SE$$
$$ SE \approx \sqrt{\sigma^2 \left( 1 + \frac{1}{n} + \frac{(52 - \bar{x})^2}{Sxx} \right)}$$
$$ CI \approx (42.64, 55.98)$$



---
### ExerciseG. Mean predicted response when x = 52 and 95% confidence interval

```{r}
# Mean predicted response when x = 52
x_pred <- 52
y_pred <- alpha + beta * x_pred

# Standard error for mean prediction
se_pred <- sqrt(sigma2 * (1 + 1/length(x) + (x_pred - x_mean)^2 / Sxx))

# Confidence interval for mean prediction
y_pred_CI <- y_pred + c(-1, 1) * t_value * se_pred

y_pred
y_pred_CI
```