---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}

/* Style for <pre> tags, commonly used for code blocks */
pre {
  background-color:  #ADD8E6;  /* Light grey background */
  padding: 10px;              /* Some padding for spacing */
  border-radius: 5px;         /* Rounded corners */
}

/* Style for <code> tags, commonly used for inline code */
code {
  background-color: #ADD8E6; /* Slightly darker grey for inline code */
  padding: 2px 4px;           /* Padding for inline code */
  border-radius: 3px;         /* Slightly rounded corners */
}
```
Q. 2) 

Policy and claims information (***PolicyData.csv***) of 650 policies is provided to you. The 
data contains following fields:

* ***Policy***: Policy Number
* ***Claim*** : Number of claims corresponding to each policy
* ***Cust_Exp***: Policyholder’s experience  at the end of the policy tenure.
	(VS = Very satisfied, SA= Satisfied, DS= Disappointed, VD= Very Disappointed)
* ***Amount***: Claim amount per policy. Note that amount is set 0 if there is no claim.

---
#### Exercises

1. Create a frequency table of claim and share how many policies don’t have any claim. 
2. Plot a histogram of claim count and suggest 2 distributions that can be a good fit. 
3. Given that claim count follows Poisson distribution with following two possible values 
for Poisson parameter:
 a. 0.35 and 
 b. 0.30 
 Compute confidence interval at 95% confidence level to assess which value is more 
suitable for the given data. 
---
#### Exercises

4. Compute mean, variance and median of log of claim amount. (Name it: log amount). 
Make sure to exclude policies with no claim. (4)
5. Obtain histogram and Normal QQ Plot of log amount. Add a line to the QQ plot for 
normal distribution 
6. Indicate which distribution the claim amount might be following using evidence from 
(iv) and (v). 
---

#### Exercises
7. Assuming the log amount (i.e. log of claim amount) follows a Normal distribution, test 
if mean of log amount is greater than 10 at 90% level of confidence. State the 
hypothesis and conclusion clearly. 

8. Assess whether the policyholder experience (i.e. ***Cust_Exp***) changes with more 
number of claims. Create contingency table and perform test to check the above 
assertion. State the hypothesis clearly. 

---

9. Please explain why warning message appears while performing the above test. 

10. Again, perform the above test by combining 
a. Very Satisfied and Satisfied customers
b. Disappointed and very disappointed customers
and 
c. 2 or more claims. 

Please provide your conclusion on association of policyholder’s experience with 
number of claims. 

11. Amount is defined to be large if the amount is greater than 100,000. Calculate 95% 
confidence interval for proportion of large claim, and comment on the likelihood if 
more than 25% of claims are large. 
---



#### Exercise 1


```{r}
policydata <- read.csv("PolicyData.csv")
```

---

```{r}
library(dplyr)
str(policydata)

```

```{r}
#a
table(policydata$Claim)
 
```

---
```{r}
#Alternative, if dplyr installed
count(policydata,Claim)
print("458 Policies don't have any claim")
```

---

### Exercise 2: Histogram of Claim Count

**Plot a histogram of claim count and suggest two distributions that can be a good fit.**

```{r eval = FALSE}
hist(policydata$Claim, 
     main="Histogram of Claim Count", 
     xlab="Number of Claims", 
     ylab="Frequency")

# Potential distributions: Poisson and Negative Binomial Distribution
```

---

### Exercise 2: Histogram of Claim Count

**Plot a histogram of claim count and suggest two distributions that can be a good fit.**

```{r echo = FALSE}
hist(policydata$Claim, main="Histogram of Claim Count", xlab="Number of Claims", ylab="Frequency")

# Potential distributions: Poisson and Negative Binomial Distribution
```


---



#### Exercise 3

```{r}
poisson.test(
  x=sum(policydata$Claim),
  T=length(policydata$Policy)
  )
```

---



### Exercise 3: Poisson Distribution Confidence Interval
Calculate the confidence interval for Poisson distribution parameters 0.35 and 0.30.

```{r}
# Poisson test with parameter 0.35
poisson.test(x = sum(policydata$Claim), T = length(policydata$Policy), r = 0.35)

# Poisson test with parameter 0.30
poisson.test(x = sum(policydata$Claim), T = length(policydata$Policy), r = 0.30)
```

* 0.35 is more suitable value of parameter since it lies between confidence interval.


---

### Exercise 4: Log of Claim Amount
**Calculate mean, variance, and median of the log of claim amount.**

```{r}
log_amount <- log(policydata$Amount[policydata$Amount > 0])
mean_log_amount <- mean(log_amount)
variance_log_amount <- var(log_amount)
median_log_amount <- median(log_amount)

print(mean_log_amount)
print(variance_log_amount)
print(median_log_amount)
```

---

### Exercise 4

```{r}
lx=log(policydata$Amount[policydata$Amount>0])
#Alternative, if dplyr installed
#lx=log(filter(policydata,Amount >0)$Amount)
mean(lx)
```

```{r}
median(lx)
```


### Variance 
```{r}
sd(lx)^2
```


---
Histogram of policydata$Claim
policydata$Claim
Frequency
0.0 0.5 1.0 1.5 2.0 2.5 3.0
0 200 400

---

### Exercise 5: Histogram and Normal QQ Plot
Obtain histogram and Normal QQ plot of log amount. Add a line to the QQ plot for normal distribution.

```{r eval=FALSE}
par(mfrow = c(2, 1))
hist(log_amount, main="Histogram of Log Amount", xlab="Log Amount", ylab="Frequency")
qqnorm(log_amount, main="Normal QQ Plot of Log Amount")
qqline(log_amount, col = "red")
```



---


```{r echo=FALSE}
par(mfrow = c(2, 1))
hist(log_amount, main="Histogram of Log Amount", xlab="Log Amount", ylab="Frequency")
qqnorm(log_amount, main="Normal QQ Plot of Log Amount")
qqline(log_amount, col = "red")
```


---

### Exercise 6: Distribution Indication
The log amount appears to follow a normal distribution based on the histogram and QQ plot. Therefore, the claim amount might follow a log-normal distribution.



* From Histogram and QQPlot it seems log amount closely follows normal distribution.

* To add, the mean and median are very close indicating symmtery. One of the characterstics of Z. 


* Hence,Claim amount might be following log normal distribution. 

---

### Exercise 7: Hypothesis Test for Log Amount Mean
Test if the mean of log amount is greater than 10 at a 90% confidence level.

```{r}
# Null hypothesis: mu = 10
# Alternative hypothesis: mu > 10
t_test_result <- t.test(log_amount, mu = 10, alternative = "greater", conf.level = 0.90)
print(t_test_result)
```


---

### Exercise 7

* Null Hypothesis : mu = 10 
* Alternate hypothesis mu >10

```{r}
t.test(lx,mu=10,alternative="greater", conf.level = .9)
```
---
One Sample t-test
data: lx
t = -1.2337, df = 191, p-value = 0.8906
alternative hypothesis: true mean is greater than 10
90 percent confidence interval:
9.663428 Inf
sample estimates:
mean of x 
Histogram of lx

---

lx Frequency
6 8 10 12 14
0 10 20 30 40
-3 -2 -1 0 1 2 3
6 8 10 14
Normal Q-Q Plot
Theoretical Quantiles
Sample Quantiles
---
9.835205 
#Given p-value greater than 10% null hypothesis can not be rejected.


---

### Exercise 8

```{r}
ct=table(policydata$Claim,policydata$Cust_Exp)
ct
```
 
 DS SA VD VS
 0 63 306 20 69
 1 36 90 9 14
 2 16 14 6 0
 3 3 3 1 0
# Null Hypothesis: No association between Policyholder's experience and Claim

---

### Exercise 8

**Policyholder Experience and Claims**

Create a contingency table and perform a chi-squared test to assess if policyholder experience changes with the number of claims.

```{r}
contingency_table <- table(policydata$Claim, policydata$Cust_Exp)
print(contingency_table)

# Chi-squared test
chi_squared_test <- chisq.test(contingency_table)
print(chi_squared_test)
```

---


### Exercise 8

**Policyholder Experience and Claims**

```{r}
chisq.test(ct)
```


---

### Exercise 9: Explanation of Warning Message

* The warning message appears because some cells in the contingency table have fewer than 5 observations, which can affect the chi-squared approximation.


---


### Exercise 10

x)

```{r}
policydata$Claim2=ifelse(policydata$Claim >2,2,policydata$Claim) 
policydata$Cust_Exp2=ifelse(policydata$Cust_Exp %in% c("DS","VD"),"DS","SA") 
ct2=table(policydata$Claim2,policydata$Cust_Exp2)
ct2

```
 
 DS SA
 0 83 375
 1 45 104
 2 26 17

---

```{r}
chisq.test(ct2)

```

* There is a strong reason to reject null hypothesis.

* Hence, it can concluded that policyholder's experience gets worse as claim count increases

---

### Exercise 10: Combined Categories Test

**Combine categories and perform the test again.**

```{r}
policydata$Claim2 <- ifelse(policydata$Claim >= 2, 2, policydata$Claim)
policydata$Cust_Exp2 <- ifelse(policydata$Cust_Exp %in% c("DS", "VD"), "DS", "SA")
contingency_table2 <- table(policydata$Claim2, policydata$Cust_Exp2)

# Chi-squared test with combined categories
chi_squared_test2 <- chisq.test(contingency_table2)
print(contingency_table2)
print(chi_squared_test2)
```

---

### Exercise 11: Proportion of Large Claims

**Calculate the 95% confidence interval for the proportion of large claims.**

```{r}
policydata$large <- ifelse(policydata$Amount > 100000, 1, 0)
x <- sum(policydata$large)
n <- length(policydata$Amount[policydata$Amount > 0])

binomial_test_result <- binom.test(x, n, conf.level = 0.95)
print(binomial_test_result)
```

---

### Exercise 11

```{r}
summary(policydata$Amount)
```

 Min. 1st Qu. Median Mean 3rd Qu. Max. 
 0 0 0 29501 3232 1848069 

---

```{r}
policydata$large= ifelse(policydata$Amount >100000,1,0)
x = sum(policydata$large)
n = length(policydata$Amount[policydata$Amount>0])
#Alternative, if dplyr installed
#n = length(filter(policydata,Amount >0)$Amount)
```

---

```{r}
binom.test(x,n)

```

---

# Since upper bound of c.i is less that .25, it is unlikely that more 
that 
#25% claims are large
(5)
---