---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

[20]
Q. 4) A random variable Z which represents claim amount in insurance is lognormally distributed 
with parameters μ=10 and σ2= 4.
i) Compute
a) Likelihood that the claim will be of 5000. (to 2 significant figures) 
b) Probability that the claim pay-out will be greater than 5000 (to 2 significant figures)
c) Maximum claim pay-out in the confidence interval [0.9,0.99] 
d) Median of Z 
e) Interquartile range of Z 
---

ii) Generate a random sample of 100 observed claim amounts using seed value equal to 50 
from this distribution. 
a) Calculate the mean, median, standard deviation and skewness coefficient of the 
sample. 
b) Plot a histogram of the sample vector showing the probability densities, setting the 
y-axis range from 0 to 10^-5 and x axis range from 0 to 10^5 for this graph and paste 
the plot into your answer sheet. 



---

Solution 4:
# i)
# a)
```{r}
mu = 10
var = 4
```

---

```{r}
dlnorm(5000,mu,sqrt(var))
## [1] 3.030741e-05
```

The likelihood that the claim will be of 5000 is 0.000030


```{r}
# b)
1-plnorm(5000,mu,sqrt(var))
## [1] 0.7707756
# probability that the claim payout will be greater than 5000 is 0.77
```

---

```{r}
# c)
qlnorm(0.9,mu,sqrt(var))
## [1] 285815.9
qlnorm(0.99,mu,sqrt(var))
## [1] 2309856

```

maximum claim payout in the confidence interval [0.9,0.99] is [285815.9, 23
09856]

---

# d) 
qlnorm(0.5,mu,sqrt(var))
## [1] 22026.47
#Median of Z is 22026.47


```{r}
# e)
qlnorm(0.75,mu,sqrt(var))-qlnorm(0.25,mu,sqrt(var))
## [1] 79162.81
# Interquartile range for Z is 79162.81
```

---

### Part 2

Generate a random sample of 100 observed claim amounts using seed value equal to 50 
from this distribution. 

```{r}
set.seed(50)
a=rlnorm(100,mu,sqrt(var))
```

---


#### Mean and Median 

```{r}
mean(a)
## [1] 121227.3
median(a)
## [1] 22785.05
```
---

* Standard Deviation and Skewness

```{r}
sd(a)
## [1] 481295.3

#skewness(a)
## [1] 8.163985
```


```{r}
sk = function(x) sum((x-mean(x))^3)/(100*(sd(x))^3)
sk(a)
## [1] 8.041831
```

---


```{r}
# (b)
hist(a, freq = FALSE,xlim = c(0,100000),ylim = c(0,0.00001))
#hist(a,freq = FALSE)
```

---