---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
A study aimed at estimating hazard rates and survival probabilities has produced the results of
Nj=Number at risk and Dj=Decrements for different time intervals tj as shown in 
CS2B_Mar22_Dataset3.csv. 

The researchers analysed the results using Nelson-Aalen model 
of comparison.

---

#### Exercises

A. Write down the formulae for the estimated integrated hazard and the estimated variance of 
this estimator, using the Nelson-Aalen model. 

B. Find the estimated integrated hazard and the estimated variance of this estimator for the 
given data for each time $t_j$. 

C. Produce a scatter plot in R along with corresponding 95% confidence intervals for the 
Nelson-Aalen model. Paste the plot in the answer.

D. Compute the survival function according to the Kaplan Meier estimate. 

E. Demonstrate numerically the validity of the inequality relationship between the survival 
function using the Nelson Aalen model and the Kaplan Meier model using the data in this 
study. 


---

#### Part (A)
For the Nelson-Aalen model the estimated integrated hazard is calculated as:

---

The estimated variance of the estimator of the integrated hazard is 
calculated as: 
 
 
--- 

#### Part (B)

```{r}
data = read.csv("CS2B_Mar22_Dataset3.csv")
names(data) = c("j","tj","nj","dj")
```


```{r}
data$lambda = cumsum(data$dj/data$nj) 
data$sdlambda=sqrt(cumsum(data$dj*(data$nj-data$dj)/data$nj^3)) 
data

data$varlambda=cumsum(data$dj*(data$nj-data$dj)/data$nj^3) 

```


---

```{r}
head(data)


```

---

#### Part (iii)

```{r}
plot(c(0,data$tj),c(0,data$lambda),
  type="s",
  main="Nelson-Aalen estimator of the integrated hazard",
  xlab="Time t",
  ylab="Integrated hazard") 

```

---

```{r eval=FALSE}
plot(c(0,data$tj),c(0,data$lambda),type="s",
main="Nelson-Aalen estimator of the integrated hazard",
xlab="Time t",ylab="Integrated hazard")

lines(data$tj,data$lambda-1.96*data$sdlambda,type="s",lty=2, col="blue")
lines(data$tj,data$lambda+1.96*data$sdlambda,type="s",lty=2, col="blue")
```


---

```{r echo=FALSE}
plot(c(0,data$tj),c(0,data$lambda),type="s",
main="Nelson-Aalen estimator of the integrated hazard",
xlab="Time t",ylab="Integrated hazard")

lines(data$tj,data$lambda-1.96*data$sdlambda,type="s",lty=2, col="blue")
lines(data$tj,data$lambda+1.96*data$sdlambda,type="s",lty=2, col="blue")
```
---

#### Part (iv)
```{r}
data$Survival_KM = cumprod(1-data$dj/data$nj)
data$Survival_KM
```

---

#### Part (v)

The inequality states that:
 
Survival function of Kaplan Meir SKM(t) < survival function of Nelson Aalen SNA(t) 
To demonstrate the inequality:


---

```{r}
data$Survival_NA = exp(-data$lambda)
sum(data$Survival_KM<data$Survival_NA)
```

---

Since all the values are true, the inequality is proved


---

---

**Part A: Formulas for Nelson-Aalen Estimator and Variance**

The Nelson-Aalen estimator for the cumulative hazard function, $\hat{H}(t)$, is given by:

$$\hat{H}(t_j) = \sum_{i=1}^{j} \frac{d_i}{n_i}$$

where:
* $d_i$ is the number of events at time $t_i$
* $n_i$ is the number of individuals at risk at time $t_i$

The variance of the Nelson-Aalen estimator can be estimated using Greenwood's formula:

$$\widehat{Var}(\hat{H}(t_j)) = \sum_{i=1}^{j} \frac{d_i}{n_i^2}$$

---

**Part B: Calculating Estimated Integrated Hazard and Variance**

To perform these calculations, we'll need to load the data and iterate through the time intervals, calculating the estimates at each step.

---

**Part C: Plotting Nelson-Aalen Estimator with Confidence Intervals**

```R
# Load the data
data <- read.csv("CS2B_Mar22_Dataset3.csv")

# Calculate Nelson-Aalen estimator and variance
data$H <- cumsum(data$Dj / data$Nj)
data$Var_H <- cumsum(data$Dj / data$Nj^2)

# Calculate 95% confidence intervals
data$lower <- data$H - 1.96 * sqrt(data$Var_H)
data$upper <- data$H + 1.96 * sqrt(data$Var_H)

# Plot the Nelson-Aalen estimator with confidence intervals
plot(data$tj, data$H, type = "s", xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimator")
lines(data$tj, data$lower, col = "red", lty = 2)
lines(data$tj, data$upper, col = "red", lty = 2)
```

---

**Part D: Kaplan-Meier Estimator**

The Kaplan-Meier estimator for the survival function, $\hat{S}(t)$, is given by:

$$\hat{S}(t_j) = \prod_{i=1}^{j} (1 - \frac{d_i}{n_i})$$

**Part E: Comparing Nelson-Aalen and Kaplan-Meier**

We can numerically verify the relationship between the two estimators:

$$\hat{S}(t) = e^{-\hat{H}(t)}$$

By calculating both estimators for each time point and comparing the results, we can confirm this relationship.

**Note:**
To provide more specific calculations and visualizations, the actual data from `CS2B_Mar22_Dataset3.csv` would be needed. However, the general approach outlined above can be applied to any dataset.
 
Remember to install and load necessary packages like `survival` for survival analysis in R.
