---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
A study aimed at estimating hazard rates and survival probabilities has produced the results of
Nj=Number at risk and Dj=Decrements for different time intervals tj as shown in 
CS2B_Mar22_Dataset3.csv. 

The researchers analysed the results using Nelson-Aalen model 
of comparison.

---

#### Exercises

A. Write down the formulae for the estimated integrated hazard and the estimated variance of 
this estimator, using the Nelson-Aalen model. 

B. Find the estimated integrated hazard and the estimated variance of this estimator for the 
given data for each time $t_j$. 

C. Produce a scatter plot in R along with corresponding 95% confidence intervals for the 
Nelson-Aalen model. Paste the plot in the answer.

D. Compute the survival function according to the Kaplan Meier estimate. 

E. Demonstrate numerically the validity of the inequality relationship between the survival 
function using the Nelson Aalen model and the Kaplan Meier model using the data in this 
study. 


---

#### Part (A)
For the Nelson-Aalen model the estimated integrated hazard is calculated as:

---

The estimated variance of the estimator of the integrated hazard is 
calculated as: 
 
 
--- 

#### Part (B)

```{r}
data = read.csv("CS2B_Mar22_Dataset3.csv")
names(data) = c("j","tj","nj","dj")
```


```{r}
data$lambda = cumsum(data$dj/data$nj) 
data$sdlambda=sqrt(cumsum(data$dj*(data$nj-data$dj)/data$nj^3)) 
data

data$varlambda=cumsum(data$dj*(data$nj-data$dj)/data$nj^3) 

```


---

```{r}
head(data)


```

---

#### Part (iii)

```{r}
plot(c(0,data$tj),c(0,data$lambda),
  type="s",
  main="Nelson-Aalen estimator of the integrated hazard",
  xlab="Time t",
  ylab="Integrated hazard") 

```

---

```{r eval=FALSE}
plot(c(0,data$tj),c(0,data$lambda),type="s",
main="Nelson-Aalen estimator of the integrated hazard",
xlab="Time t",ylab="Integrated hazard")

lines(data$tj,data$lambda-1.96*data$sdlambda,type="s",lty=2, col="blue")
lines(data$tj,data$lambda+1.96*data$sdlambda,type="s",lty=2, col="blue")
```


---

```{r echo=FALSE}
plot(c(0,data$tj),c(0,data$lambda),type="s",
main="Nelson-Aalen estimator of the integrated hazard",
xlab="Time t",ylab="Integrated hazard")

lines(data$tj,data$lambda-1.96*data$sdlambda,type="s",lty=2, col="blue")
lines(data$tj,data$lambda+1.96*data$sdlambda,type="s",lty=2, col="blue")
```
---

#### Part (iv)
```{r}
data$Survival_KM = cumprod(1-data$dj/data$nj)
data$Survival_KM
```

---

#### Part (v)

The inequality states that:
 
Survival function of Kaplan Meir SKM(t) < survival function of Nelson Aalen SNA(t) 
To demonstrate the inequality:


---

```{r}
data$Survival_NA = exp(-data$lambda)
sum(data$Survival_KM<data$Survival_NA)
```

---

Since all the values are true, the inequality is proved


---

---
