---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
[15]
Q. 2) 

An Actuary tried to predict the COVID-19 evolution by using Markov chain which involved 
modelling transitions from one state to another according to a transition probability matrix. 
These predictions are expected to help the authorities to set up adequate protocols for 
managing the post-confinement due to COVID-19.
There are five states in the model namely Infected, Quarantined, Hospitalised, Recovered and 
Death which are measured at the end of each week.
---

#### Data Set 

* The <tt>‘CS2B_Mar22_Dataset2.csv’</tt> file contains the recorded weekly transitions between 
different states. 
* Each row contains a pair of State_From and State_To where State_From 
denotes the starting state and State_To denotes the state after a week’s transition. 
* The following notation has been used:
 I-Infected, 
 Q-Quarantined, 
 H-Hospitalised, 
 R-Recovered, 
 D-Death
* Those states which have no transition to any another state can be assumed to be absorbing.
---

Load ‘CS2B_Mar22_Dataset2.csv’ file into R and name it as coviddata.

#### Exercises 

A. Using transition pair data, compute the transition probabilities between each state. (6)
B. Find out the absorbing states in the model. 
C. Create Markov chain model named “covid19” in R with the respective state space and 
transition probability matrix. Plot this model and paste the plot into your answer sheet. 
(Use library-markovchain) 

---

Parameter 1>150 and Parameter 2 >50
Parameter 
3>300
Surgery Not 
Required
Surgery 
Required
Surgery not
required
---

iv) Using the Markov chain, given that the individual is infected, compute the probability that:
a) He will be in quarantine after a week 
b) He will either be in hospital or in quarantine after two weeks 
c) He will recover after 3 weeks 


v) COVID-19 pandemic is expected to end in a quarter’s time. 
Estimate the number of people who will be (a) Quarantined (b) Hospitalized (c) Recovered 
and (d) Died out of the current 20,000 quarantined and 10,000 hospitalised individuals. 
(Answer to nearest whole number, Assume 13 weeks in a quarter)
vi) Comment on your results obtained in (v). 

---
Solution 2:
# i)



```{r}
covid_data=read.csv("CS2B_Mar22_Dataset2.csv")
```

```{r}
head(covid_data)
```

---


```{r}
covid_data$State_From = as.factor(covid_data$State_From)
covid_data$State_To = as.factor(covid_data$State_To)
```

```{r}
States_From = levels(covid_data$State_From)
States_To = levels(covid_data$State_To)
States = unique(c(States_From,States_To))
```

---

```{r}
# Transition Probabilities for each pair
transitions_Master = c()

for (i in States) {
 for (j in States) {
 
 transition = sum(covid_data$StateFrom == i & covid_data$State.To == j)/sum(covid_data$StateFrom == i)
 
 transition = ifelse(sum(covid_data$StateFrom == i) ==0, ifelse(i ==j, 1,0),transition)
 
 transitions_Master = c(transitions_Master,transition)
 
 }
 
}

```

---

```{r}
transition_matrix = matrix(transitions_Master, nrow=5, 
                           byrow = T, 
                           dimnames=list(States, States))

transition_matrix

```

---

The two absorbing states are "Recovered" and "Death"


```{r}
library(markovchain)

```


```{r}
covid19 <- new("markovchain",
               states=c("H","I","Q", "D","R"),
    transitionMatrix=transition_matrix)

covid19

```


---

```{r}
## Unnamed Markov chain 
## A 5 - dimensional discrete Markov Chain defined by the following states: 
## H, I, Q, D, R 
## The transition matrix (by rows) is defined as follows: 
## H I Q D R
## H 0.2 0 0.0 0.2 0.6
## I 0.4 0 0.6 0.0 0.0
## Q 0.2 0 0.3 0.0 0.5
## D 0.0 0 0.0 1.0 0.0
## R 0.0 0 0.0 0.0 1.0
plot(covid19)
```


# (a)

---

```{r}
initialstate=c(0,1,0,0,0)
afterweek=initialstate*(covid19)
afterweek
## H I Q D R
## [1,] 0.4 0 0.6 0 0
## [1] 0.6
```

---

(1)
# (b)

```{r}
aftertwoweeks=initialstate*(covid19*covid19)
aftertwoweeks
## H I Q D R
## [1,] 0.2 0 0.18 0.08 0.54
# Probability of either hospital or quarantine after 2 weeks
p1 = aftertwoweeks[1]+aftertwoweeks[3]
p1
```
---

```{r}
## [1] 0.38 (2)
# (c)
afterthreeweeks=initialstate*(covid19*covid19*covid19)
afterthreeweeks
## H I Q D R
## [1,] 0.076 0 0.054 0.12 0.75
#Probability of recovery after 3 weeks
p2 = afterthreeweeks[5]
p2
## [1] 0.75 (2)
# (v)
# Long term probability (after 13 weeks)
# Quarantined People

```

---

```{r}
initial_Q = c(0,0,1,0,0)
after_quarter = initial_Q*(covid19^13)
initial_H = c(1,0,0,0,0)
after_quarter1 = initial_H*(covid19^13)
Recovered = 20000*after_quarter[5]+10000*after_quarter1[5]
Died = 20000*after_quarter[4]+10000*after_quarter1[4]
Quarantined = 20000*after_quarter[3]+10000*after_quarter1[3]
Hospitalized = 20000*after_quarter[1]+10000*after_quarter1[1]

```
---

```{r}
round(Recovered,0)
## [1] 26071
round(Died,0)
## [1] 3929
round(Hospitalized,0)
## [1] 0
round(Quarantined,0)
## [1] 0
```

---

#### Part F.

* In the long run, everyone needs to reach a steady state probabilities. 
* As there are two absorbing states, one of the two is the destination
* The other states are completely zeroes

---
