---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

MA(2,2) time series Zt is defined by the following equation
$$Zt=2 + 0.8Zt-1 + 0.1Zt-2 + et + 0.4et-1 + 0.1et-2$$ , where et are white noise terms which are IID 
random variables with mean 0 and σ2= 7


#### Exerices 
A.  Simulate 300 observations of the above time series using seed value equal to 100 and plot 
it giving appropriate labels for each axis. Paste the plot into your answer sheet. 

B. Using the data simulated in Part A, compute the mean and standard deviation of 
i) The entire series
ii) The first 150 observations of the series
iii) The last 150 observations of the series 

---

C. Plot the Autocorrelation function (ACF) and Partial Autocorrelation function (PACF) of 
the simulated data in (i), giving appropriate labels for each axis. Paste the plots into your 
answer sheet. 

D.  Comment on the stationarity of data from (ii) and (iii) above. 

E. Comment whether ARMA(2,2) can be inferred from the ACF and PACF plots. 

---

## Solution :

### Part A

```{r}
set.seed(100) 
z=2+arima.sim(n=300,list(ar=c(0.8,0.1),ma=c(0.4,0.1),sd=7^0.5))

plot(z, xlab = "Time", 
  ylab = "Simulated Values from an ARMA(2,2) Time Series", 
  main = "Time Series Plot")

```

---
  
### Part B
#### (i)

```{r}
mean (z)
## [1] 1.520674
sd(z)
## [1] 5.865453

```

---

#### (ii)

```{r}
mean(z[1:150])
## [1] 0.678429
sd(z[1:150])
## [1] 4.997272
```

---

```{r}

mean(z[151:300])
## [1] 2.362919
sd(z[151:300])
## [1] 6.529698
```

---

```{r}

acf(z, xlab = "Lag", 
  ylab = "ACF of Simulated Values from an ARMA(2,2) Time Series", 
main = "ACF") 

```

---

```{r}
pacf(z, xlab = "Lag", 
     ylab = "Partial ACF of Simulated Values from an ARMA(2,2) Time Series", 
     main = "PACF")
```

---
#### Part D.

* The series appears stationary as there are no obvious trends or cycles in the graph of the series and it appears to have constant mean. 
*  However, from Part B, it appears that mean of the two subsets of the data is 
very different.
* Working with a larger subset takes the mean values closer to constant thus 
revealing in stationarity. 

* For a stationary time series, the ACF will drop to zero relatively quickly, 
while the ACF of non-stationary data decreases slowly 
* Hence, we confirm that the series is stationary

---

#### Part E.

* PACF becomes insignificant after two lags and ACF goes gradually to zero in
dicating the strong presence of AR(2)process compared to an ARMA (2,2) models
* In ARMA(2,2) model, we should observe, ACF and PACF to gradually go down to 
zero after a few lags

---

END 
---

---