May 2025 - Question 3

This exercise is a practical demonstration of the **Central Limit Theorem** and its application to the **binomial distribution**.

The process unfolds in three main phases:

1.  **Initial Simulation:** The exercise begins with a small simulation of 100 Bernoulli trials to introduce the concept of a random variable representing a success or failure. This is then scaled up to a larger simulation of 5000 random variables, where each represents the total number of "successes" from 100 trials.

2.  **Theoretical vs. Empirical Analysis:** The simulated results (mean and variance) are compared to their theoretical counterparts, which are derived from the properties of the binomial distribution. The close match between the simulated and theoretical values demonstrates the **law of large numbers**.

3.  **Visual and Conceptual Proof:** The final phase visually confirms the Central Limit Theorem. A histogram of the simulated data is created, and a normal distribution curve (with parameters derived from the binomial properties) is overlaid on it. The strong visual fit between the simulated data's histogram and the theoretical normal curve provides a clear and compelling illustration that for a large number of trials, a binomial distribution can be effectively approximated by a normal distribution.

In summary, the exercise uses simulation to provide a concrete, step-by-step understanding of how the binomial distribution behaves and how its shape can be approximated by the more familiar normal distribution under certain conditions.

**i) Bernoulli Trial Simulation**
A series of 100 Bernoulli trials is simulated, where the probability of success is 0.4. `set.seed(2025)` ensures the results are reproducible.

```{r}
set.seed(2025)
X <- rbinom(n = 100, size = 1, prob = 0.4)
```

**ii) Sum of Successes**
The number of successes (where a "1" is observed) in the 100 trials is calculated.

```{r}
sum(X)
```

The sum is 43.

**iii) Binomial Distribution Simulation**
A simulation of 5000 random variables from a binomial distribution is performed. Each variable represents the number of "successes" in 100 trials, with a success probability of 0.4.

```{r}
set.seed(2025)
applications <- rbinom(n = 5000, size = 100, prob = 0.4)
```

**iv) Sample Mean and Variance**
The mean and variance of the simulated data are calculated.

```{r}
mean(applications)
var(applications)
```

Simulated Mean: 39.9536
Simulated Variance: 23.7778

**v) Theoretical Mean**
The theoretical mean of a binomial distribution is calculated using the formula $\\mu = n \\times p$.
Theoretical Mean = $100 \\times 0.4 = 40$

---
**vi) Comparison of Simulated vs. Theoretical Mean**
The simulated mean (39.9536) is very close to the theoretical mean (40). This is expected because the simulation was run with a large number of trials ($n=5000$), which, according to the law of large numbers, should produce a sample mean that is very close to the population mean.

---

**vii) Histogram of Simulated Data**
A histogram of the simulated data is created to visualize the distribution.

```{r}
hist(applications, main = "Histogram of Low-Income Students in Simulated Applications", 
     xlab = "Number of Low-Income Students", probability = TRUE)
```

---

**viii) Normal Approximation of Binomial Distribution**
A binomial distribution with a large number of trials ($n=100$) can be approximated by a normal distribution.
The parameters for the normal approximation are:

  * Mean: $\\mu = n \\times p = 100 \\times 0.4 = 40$
  * Variance: $\\sigma^2 = n \\times p \\times (1-p) = 100 \\times 0.4 \\times 0.6 = 24$
    So, the distribution can be approximated as $N(\\mu = 40, \\sigma^2 = 24)$.

**ix) Visualizing the Normal Approximation**
A normal density curve is overlaid on the histogram to visually check the quality of the approximation. A density plot of the simulated data is also added for comparison.

```{r}
hist(applications, main = "Histogram of Low-Income Students in Simulated Applications",
     xlab = "Number of Low-Income Students", probability = TRUE)
lines(density(applications), col = "darkgreen")
curve(dnorm(x, mean = 40, sd = sqrt(24)), col = "red", add = TRUE)
```
---

**x) Conclusion on Normal Approximation**
The histogram of the simulated data is bell-shaped and symmetric, centered around the mean of 40. The normal curve fits the shape of the histogram and the simulated density line very well. This visual agreement supports the **Central Limit Theorem**, which states that for a sufficiently large number of trials, the binomial distribution can be accurately approximated by a normal distribution.