---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

1 A researcher records the levels of caffeine present in their bloodstream at various time 
intervals after drinking a cup of coffee. 
T: Time elapsed 
(30-minute 
intervals) 
1 2 3 4 5 6 7 8 9 10 
C: Caffeine in 
blood 
(milligrams) 
52.2 39.3 28.3 19.03 13.96 11.46 8.78 6.55 5.43 5.03 
 
You can enter the values, for T (time) and C (caffeine), into R using: 
```{r}
time = c(1,2,3,4,5,6,7,8,9,10) 
caffeine = c(52.2,39.3,28.3,19.03,13.96,11.46,8.78,6.55,5.43,5.03)
```

##

#### Exercise  1

(i) Plot a scatterplot of the data. 


---

#### Exercise  2

Comment on the relationship between $C$ and $T$ based on the plot in Exercise  1.  

(ii)
The plot shows a non-linear and inverse relationship between C and T (i.e. the level of 
caffeine in the blood does not reduce linearly over time - instead it appears to decay at an 
exponential rate) 
---
#### Exercise  3

Justify why a logarithmic transformation of $C$ is appropriate. 




---
(iii)
Given the shape of the graph in (i), a log transformation should be used on the data (i.e. 
transform C to log(C)) 
This is because the original plot from part (i) has an exponential shape 

---

#### Exercise  4

(iv) Perform the transformation suggested in Exercise  3.  

(iv) > logcaffeine <- log(caffeine) 
> logcaffeine
3.955082 3.671225 3.342862 2.946017 2.636196 2.438863 2.172476 1.879465
1.691939 1.615420 [½]

---

#### Exercise  5

(a) Plot a scatterplot of the transformed data. 
(b) Calculate the Pearson correlation coefficient for the transformed data. 
(c) Comment on the scatterplot and Pearson correlation coefficient. 

---
(v)(a)
> plot(time,logcaffeine,main = "Plot for time and log(caffeine)")



---

Q1
(i)

(ii)
The plot shows a non-linear and inverse relationship between C and T (i.e. the level of caffeine in the blood does not reduce linearly over time - instead it appears to decay at an exponential rate) [3]


---

(iii)
Given the shape of the graph in (i), a log transformation should be used on the data (i.e. transform C to log(C)) 
This is because the original plot from part (i) has an exponential shape 
(iv) logcaffeine <- log(caffeine) 
logcaffeine
3.955082 3.671225 3.342862 2.946017 2.636196 2.438863 2.172476 1.879465
1.691939 1.615420 

---

(v)(a)
plot(time,logcaffeine,main = "Plot for time and log(caffeine)") 
2468101020304050Plot for time and caffeinetimecaffeine
CS1B - Actuarial Statistics - Core Principles - April 2021 - Examiners’ report
CS1B A2021 © Institute and Faculty of Actuaries

(b)
cor(time,logcaffeine,method="pearson") 
# -0.9919573 
(c)
We can see via the plot in part (v)(a) and the calculated Pearson coefficient in part that a strong negative linear relationship exists between the transformed variable and time 
[Total 15]
