---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

3 A study was carried out to estimate the proportion, 𝑝, of workers that commute by 
train to work. A total of n ൌ 200 workers were sampled at random and were asked the 
question: ‘Do you take the train to work?’ 

* The workers’ answers were recorded as a 
binary outcome, yi, for worker i, with 1 for yes and 0 for no. The data are available in 
the file BinaryTrain.RData.

---

Two commuters, Alice and Norman, were interested in the study and proposed 
different prior distributions for the proportion p. 
<p>

Alice assumed a discrete prior distribution g(p) given in the following table: 

p 0.1 0.2 0.3 0.4 0.5 
g(p) 0.5 0.2 0.2 0.05 0.05 
Norman chose to use a beta prior distribution for $p$ , with parameters 3 and 12. 

---

### Exercise  1

(i) (a) Calculate the mean and the standard deviation for Alice’s prior
distribution. 
(b) Generate 10,000 random values from Norman’s prior distribution.
(c) Calculate the mean and standard deviation of the values generated in
part (i)(b).  
(d) Comment on whether or not Alice and Norman have similar prior 
beliefs for p. 

---


Norman’s beta prior distribution for p is adopted for the remainder of the question. 


* The likelihood of the model in the study is given by: 
Lሺpሻ ∝ p\sum yi ሺ1 – pሻn – \sum yi . 
* The posterior density of $p$ is given by: 
f(p|y) ∝ p2 + \sum yiሺ1 – pሻ11 + n – \sumyi ,
where \sum yi is the total sum of all the binary data. 

---

Q3
(i)(a)
```{r}
p = c(0.1, 0.2, 0.3, 0.4, 0.5) 
gp=c(0.5, 0.2, 0.2, 0.05, 0.05) 
mean_A=sum(p*gp)
mean_A
```

```{r}
sd_A = sqrt(sum(gp*p^2)-mean_A^2)
sd_A
```

(b)

```{r}
samples_beta = rbeta(10000, 3, 12) 
```

---

(c)

```{r}
mean_N = mean(samples_beta)
mean_N
```


```{r}
sd_N = sd(samples_beta)
sd_N
```

(d)
The first and second moments of the two priors are very similar
The two prior beliefs are similar despite one being based on a continuous distribution and the 
other on a discrete distribution 
---

### Exercise  2
(ii) Plot the shape of the posterior density of $p$  without identifying it.  

---

(ii)
Read the data in: 
```{r}
load("BinaryTrain.RData")

 
p = seq(0, 1, by = 0.01)
dens = p^(2+sum(y))*(1-p)^(11+length(y)-sum(y))
plot(p, dens, type = "l", ylab ="",xlab="p") 
```


---

```{r}
curve(x^(2+sum(y))*(1-x)^(11+length(y)-sum(y))) 
```
---

### Exercise  3

Plot the density of Norman’s prior distribution by setting `ylim = c(0,14)`. 

The posterior distribution of $p$ is beta with parameters 3 + \sum yi and 12 + n – \sum yi
. 
---

(iii)
Prior distribution
```{r}
plot(p, dbeta(p, 3, 12), type = "l", ylim =c(0, 14), ylab ="")

```

---

Or,

```{r}
curve(dbeta(x, 3, 12), type="l",
ylim=c(0, 14), ylab ="", xlab="p")
```

---

### Exercise  4

(iv) (a) Plot the posterior density of $p$ by adding it to the plot in part (iii). 
(b) Compare the two densities using your answer in part (iv)(a). 
(c) Comment on the extent to which the posterior distribution is affected 
by the prior distribution. 

---

### Exercise  5

(v) Determine a 90% interval estimate for p based on its posterior distribution. 


---

---

### Exercise 6

(vi) Determine the exact posterior probability that p exceeds 0.25. 
(vii) (a) Generate 10,000 samples from the posterior distribution of p. 
(b) Calculate the proportion of sampled values of $p$ that exceed 0.25. 
(c) Compare your answer in part (vii)(b) with your answer in part (vi). 

---

Q3
(i)(a)
p = c(0.1, 0.2, 0.3, 0.4, 0.5) 
gp=c(0.5, 0.2, 0.2, 0.05, 0.05) 
mean_A=sum(p*gp)
mean_A
# 0.195 
>sd_A = sqrt(sum(gp*p^2)-mean_A^2)
>sd_A
# 0.1160819 
(b)
samples_beta = rbeta(10000, 3, 12) 
(c)
mean_N = mean(samples_beta)
mean_N
# 0.2012128 
sd_N = sd(samples_beta)
sd_N
# 0.09913621 
(d)
The first and second moments of the two priors are very similar 
The two prior beliefs are similar despite one being based on a continuous distribution and the other on a discrete distribution 

---

(ii)
Read the data in:
load("BinaryTrain.RData")
p = seq(0, 1, by = 0.01)
dens = p^(2+sum(y))*(1-p)^(11+length(y)-sum(y))
plot(p, dens, type = "l", ylab ="",xlab="p") [3]
Or,
curve(x^(2+sum(y))*(1-x)^(11+length(y)-sum(y)))
Parts (i)-(iii) were generally well answered. Plots in part (i) were varied, with a number of candidates not using reasonable axis labels (or no axis at all), while comments were often missing in (i)(b). A common error in part (ii) was to omit the PAIRED = TRUE parameter, as the question requested. Answers in (iv), (v) were mixed with a range of comments.
CS1B - Actuarial Statistics - Core Principles - April 2021 - Examiners’ report
CS1B A2021 © Institute and Faculty of Actuaries

---

(iii)
Prior distribution
plot(p, dbeta(p, 3, 12), type = "l", ylim =c(0, 14), ylab ="") 
Or,
curve(dbeta(x, 3, 12), type="l",ylim=c(0, 14), ylab ="", xlab="p")

0.00.20.40.60.81.00e+001e-512e-513e-51p
CS1B - Actuarial Statistics - Core Principles - April 2021 - Examiners’ report
CS1B A2021 © Institute and Faculty of Actuaries
(iv)(a)
x = p
curve(dbeta(x, 3+sum(y), 12+length(y)-sum(y)), type = "l", add=TRUE, col="red")

Or,
>lines(p, dbeta(p,3+sum(y),12+length(y)-sum(y)),col="red")

---

(b)
Clearly, the posterior is much narrower than the prior. 
(c)
The posterior distribution is more affected by the data than by the prior. 
(v)
90% interval for p
qbeta(c(0.05, 0.95), 3+sum(y), 12+length(y)-sum(y)) 
# 0.1910197 0.2861843
90% interval for p is (0.191, 0.286). 
(vi)
1 - pbeta(0.25, 3+sum(y), 12+length(y)-sum(y)) 
# 0.3216195

---

(vii)(a)
z = rbeta(10000, 3+sum(y), 12+length(y)-sum(y))  (b)
sum(z>0.25)/10000 
# 0.3271 

---

Or,
length(z[z>0.25])/length(z)
(c) As expected the proportion is very similar to the answer in (vi). 

---

Q4
(i)
load("CS1passenger.RData")
Linear predictor for modelling:
(a) 𝛼𝛼𝑖𝑖+\beta×𝑡𝑡𝑡𝑡𝑡𝑡𝑡𝑡: where the intercept 𝛼𝛼𝑖𝑖,𝑖𝑖=1,2 depends on the semester 
(b) 𝛼𝛼𝑖𝑖+\beta𝑖𝑖×𝑡𝑡𝑡𝑡𝑡𝑡𝑡𝑡: where 𝛼𝛼𝑖𝑖 as above, \beta𝑖𝑖,𝑖𝑖=1,2 also depends on the semester 
(c) 𝛼𝛼𝑖𝑖+\beta𝑖𝑖×𝑡𝑡𝑡𝑡𝑡𝑡𝑡𝑡+𝛾𝛾𝑗𝑗 with 𝛼𝛼𝑖𝑖, \beta𝑖𝑖 as above, 𝛾𝛾𝑗𝑗, j = 1,2 depends on the route. 
Alternative answer:
(a) 𝑦𝑦=𝑎𝑎+𝑏𝑏1x_1+𝑏𝑏2x_2 , where x_1 is temperature and x_2=0 (nonSemester), x_2=1 (Semester)
(b) 𝑦𝑦=𝑎𝑎+𝑏𝑏1x_1+𝑏𝑏2x_2+𝛾𝛾x_1x_2 , where x_1and x_2 as above
(c) 𝑦𝑦=𝑎𝑎+𝑏𝑏1x_1+𝑏𝑏2x_2+𝛾𝛾x_1x_2+𝑏𝑏3x_3 , where x_3=0 (8am), x_3=1 (9am)]


---

(ii)(a)
>Model1<- glm(Passengers~temp*semester + route, family=poisson(link="log"))
>summary(Model1) 
Call:
glm(formula = Passengers ~ temp * semester + route, family = poisson(link = "log"))
Deviance Residuals:
Min 1Q Median 3Q Max
-1.8128 -0.6263 -0.1566 0.5162 1.3991
Coefficients:
Estimate Std. Error z value Pr(>|z|)
(Intercept) 0.40210 0.31155 1.291 0.1968
temp -0.07878 0.03576 -2.203 0.0276 *
semestersemester 0.53514 0.46691 1.146 0.2517 
route9am 0.17370 0.44520 0.390 0.6964
temp:semestersemester 0.10779 0.05741 1.878 0.0604 .
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
A number of candidates did not attempt this question, or only attempted part (i).
For candidates that attempted it, the overall performance was mixed. In part (iii) many candidates plotted the graphs using the simulated values from part (i)(b) or using inappropriate plots, which led to difficulty later in the question.

---

Null deviance: 30.406 on 19 degrees of freedom
Residual deviance: 13.833 on 15 degrees of freedom
AIC: 62.187
(b)
Temperature (temp) is significant 
Semester is not significant 
Route is not significant 
The interaction between temperature (temp) and semester is not significant at 5% significance level 
but it is close to being significant 
(iii)(a)
>Model2<- update(Model1,~.-route) 
Or,
Model2 <- glm(Passengers~temp*semester,family="poisson" (link = "log"))
>summary(Model2)
Call:
glm(formula = Passengers ~ temp + semester + temp:semester, family = poisson(link = "log"))
Deviance Residuals:
Min 1Q Median 3Q Max
-1.84542 -0.66323 -0.06209 0.43732 1.34790
Coefficients:
Estimate Std. Error z value Pr(>|z|)
(Intercept) 0.44284 0.29121 1.521 0.1283
temp -0.07452 0.03387 -2.200 0.0278 *
semestersemester 0.54602 0.46390 1.177 0.2392
temp:semestersemester 0.10012 0.05316 1.883 0.0597 .
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
Null deviance: 30.406 on 19 degrees of freedom
Residual deviance: 13.982 on 16 degrees of freedom
AIC: 60.336

(b)
The AIC has fallen from 62.187 to 60.336 - so new model has improved the initial model 
(iv)(a)
>Model3<- glm(Passengers~temp+temp:semester,family=poisson(link="log"))
>Modela<- glm(Passengers~temp+ semester,family=poisson(link="log")) 
>Modelb<- glm(Passengers~temp*semester,family=poisson(link="log")) 
>Model3$aic
59.65976 
>Modela$aic
62.03591 

>Modelb$aic
60.33588 
Model3 has the lowest AIC compared with the other models. We conclude that Model3 outperforms the other models considered here 
(b)
Model3 doesn’t include both of the main effects. Despite this, the model still suits the data well 
(v)(a)
plot(Model3,1) 

(b)
The residuals plot shows no patterns - exhibiting a fairly random scatter around zero with constant variance and no outliers 
The plot suggests that the model is appropriate 
(vi)
>predict(Model3, data.frame(temp=0,semester="semester",route="8am"),type = "response") [3]
Predicted number is: 1.866568 
[Total 35]
CS1B - Actuarial Statistics - Core Principles - April 2021 - Examiners’ report
CS1B A2021 © Institute and Faculty of Actuaries
