---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

An analysis was carried out to investigate the fairness of two exam markers. They 
both marked the same 150 exam papers, with 10 questions and total possible marks of 
100 for each exam paper. 

The data were collected and arranged into 10 equally spaced 
groups, with marks rounded to the nearest whole number. 

Below are the frequencies of the marks given by each of the exam markers: 

| Exam Marks | 0-10 | 11-20 | 21-30 | 31-40 | 41-50 | 51-60 | 61-70 | 71-80 | 81-90 | 91-100 |
|------------|------|-------|-------|-------|-------|-------|-------|-------|-------|--------|
| Marker 1   | 1    | 8     | 14    | 22    | 33    | 34    | 21    | 9     | 6     | 2      |
| Marker 2   | 0    | 4     | 16    | 25    | 27    | 42    | 23    | 4     | 9     | 0      |

 
---
 
You can enter the exam marks for Marker 1 and Marker 2 into R using: 

```{r}
marks_1 = c(1,8,14,22,33,34,21,9,6,2) 
marks_2 = c(0,4,16,25,27,42,23,4,9,0)  
```


### Exercise 1

Plot two graphs, one for each marker, for the distribution of the exam 
marks given by the two markers. 

[Hint: You may find the R command <tt>barplot()</tt> useful.] 

Compare this plots 
---


### Exercise 1

**Marker 1**
```{r eval=FALSE}
axis = c("0-10","11-20","21-30","31-40","41-50",
         "51-60","61-70","71-80","81-90","91-100")

barplot(marks_1,xlab = "Exam mark groups", 
        ylab = "Number of exams", 
        main = "Exam marks for marker 1",
        col = "Red",
        names = axis, 
        ylim = c(0,40)) 
```

---

### Exercise 1

**Marker 1**
```{r echo=FALSE}
axis = c("0-10","11-20","21-30","31-40","41-50","51-60","61-70","71-80","81-90","91-100")
barplot(marks_1,xlab = "Exam mark groups", ylab = "Number of exams", main = "Exam marks for marker 1",col = "Red",names = axis, ylim = c(0,40)) 
```

---

### Exercise 1

**Marker 2**
```{r eval=FALSE}
barplot(marks_2,xlab = "Exam mark groups", 
        ylab = "Number of exams", 
        main = "Exam marks for marker 2",
        col = "Purple",
        names = axis, 
        ylim = c(0,50)) 
```


---

### Exercise 1

**Marker 2**
```{r echo=FALSE}
barplot(marks_2,xlab = "Exam mark groups", ylab = "Number of exams", main = "Exam marks for marker 2",col = "Purple",names = axis, ylim = c(0,50)) 
```



---

### Exercise 2

One of the marked exam papers is selected at random and the scores given by each of 
the markers are analysed further by question. 

Below are the scores given by each of the exam markers for this selected exam paper for each of the 10 questions: 

| Question | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 |
|----------|---|---|---|---|---|---|---|---|---|----|
| Marker 1 | 4 | 1 | 5 | 1 | 4 | 6 | 4 | 5 | 3 | 6  |
| Marker 2 | 3 | 2 | 4 | 0 | 3 | 4 | 2 | 3 | 3 | 6  |

---


### Exercise 2

Perform a suitable test to determine whether the difference in the mean scores 
of the two markers is zero or not, at the 5% confidence level, taking into 
account that the two markers have marked the same exam paper. 


You can enter the scores for Marker 1 and Marker 2 into R using: 

```{r}
marker_1 = c(4,1,5,1,4,6,4,5,3,6) 
marker_2 = c(3,2,4,0,3,4,2,3,3,6) 
```


#### Hypotheses

* $H_{0}$ ∶ difference in means is zero 

* $H_{1}$ ∶ difference in means is not zero

N.B. This is a paired data design

---

```{r}
t.test(marker_1,marker_2,paired=TRUE) 
```

---

### Exercise 3


* P-value is equal to 0.01872 which is less than the significance level (5%)

* Therefore we shall reject the null hypothesis

* There appears to be difference in the mean scores between the two markers

---

### Exercise 3

Perform the test specified in Exercise 2, ignoring that the two markers have 
marked the same exam paper. 



#### Hypotheses

* $H_{0}$ ∶ difference in means is zero 

* $H_{1}$ ∶ difference in means is not zero


N.B. This is an independent sample design


---

### Exercise 3

```{r}
t.test(marker_1,marker_2) 
```


---

### Exercise 3

* P-value is equal to 0.247 which is more than the significance level (5%)
* Therefore do not reject the null hypothesis
* There appears to be no difference between the mean scores of the two markers

---

### Exercise 4

**Comment on your conclusions from Exercises 2 and 3.**


* Parts 2 and 3 lead to different conclusions

* We would expect the data to not be independent of one another since both markers have 
marked the same exam papers

* Therefore the paired test in part 2 gives smaller variation for the test statistic and leads to 
rejecting the hypothesis of equal means, at the 5% level 
---

### Exercise 5

**Comment on the issues arising when analysing paired data as independent 
samples, and independent samples as though they were paired.** 




* If a paired problem is analysed as though it involved independent samples, then the results would be invalid because the assumption of independence is violated 

* Alternatively, if independent samples are analysed as though they were paired, then the results would be valid although they would be making inefficient use of the data due to the discarding of possible information about the means and variances of the two separate populations 

---

---

Q2

BLANK

---
# Generally very well answered. 
# A common error in part (ii) was failing to refer to the strength of the relationship between caffeine and time. 
# Answers in part (v)(c) were varied, with candidates often giving partial comments.

---

(b)
The distributions of marks look similar, especially for middle scores 
However, there appears to be some differences in marking for low and high scoring exams
The plot for marker 1 resembles a Normal shape (but it is not as clear for maker 2, where there appears to be some skewness) 
Overall, the plots suggest that the two markers are generally consistent 

---

(ii)
$H_0$∶difference in means is zero 𝑣𝑣𝑣𝑣 𝐻𝐻𝐴𝐴∶difference in means is not zero. 

```{r}
t.test(marker_1,marker_2,paired=TRUE) 
```


Paired t-test
data: marker_1 and marker_2
t = 2.862, df = 9, p-value = 0.01872
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
0.1886284 1.6113716
sample estimates:
mean of the differences
0.9 
P-value is equal to 0.01872 
which is less than the significance level (5%) 
Therefore reject the null hypothesis 
There appears to be difference in the mean scores between the two markers 

---

(iii)
$H_0$∶difference in means is zero 𝑣𝑣𝑣𝑣 𝐻𝐻𝐴𝐴∶difference in means is not zero 

```{r}
t.test(marker_1,marker_2) 
```

Welch Two Sample t-test
data: marker_1 and marker_2
t = 1.1968, df = 17.675, p-value = 0.2472
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
-0.6820491 2.4820491
sample estimates:
mean of x mean of y
3.9 3.0 
P-value is equal to 0.2472 
which is more than the significance level (5%) 
Therefore do not reject the null hypothesis 
There appears to be no difference between the mean scores of the two markers 

---

(iv)
Parts (ii) and (iii) lead to different conclusions 
We would expect the data to not be independent of one another since both markers have marked the same exam papers 
Therefore the paired test in (ii) gives smaller variation for the test statistic and leads to rejecting the hypothesis of equal means, at the 5% level 




---