---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

An analysis was carried out to investigate the fairness of two exam markers. They 
both marked the same 150 exam papers, with 10 questions and total possible marks of 
100 for each exam paper. The data were collected and arranged into 10 equally spaced 
groups, with marks rounded to the nearest whole number. 
Below are the frequencies of the marks given by each of the exam markers: 

Exam marks 0–10 11–20 21–30 31–40 41–50 51–60 61–70 71–80 81–90 91–100
Marker 1 1 8 14 22 33 34 21 9 6 2 
Marker 2 0 4 16 25 27 42 23 4 9 0 
 
---
 
You can enter the exam marks for Marker 1 and Marker 2 into R using: 

```{r}
marks_1 = c(1,8,14,22,33,34,21,9,6,2) 
marks_2 = c(0,4,16,25,27,42,23,4,9,0)  
```

---

Part 1

(i) (a) Plot two graphs, one for each marker, for the distribution of the exam 
marks given by the two markers. 
[Hint: You may find the R command barplot() useful.] 
(b) Comment on the graphs produced in part (i)(a). 

---

One of the marked exam papers is selected at random and the scores given by each of 
the markers are analysed further by question. Below are the scores given by each of 
the exam markers for this selected exam paper for each of the 10 questions: 
Question 1 2 3 4 5 6 7 8 9 10 
Marker 1 4 1 5 1 4 6 4 5 3 6 
Marker 2 3 2 4 0 3 4 2 3 3 6 

You can enter the scores for Marker 1 and Marker 2 into R using: 

```{r}
marker_1 = c(4,1,5,1,4,6,4,5,3,6) 
marker_2 = c(3,2,4,0,3,4,2,3,3,6) 
```

---


(ii) Perform a suitable test to determine whether the difference in the mean scores 
of the two markers is zero or not, at the 5% confidence level, taking into 
account that the two markers have marked the same exam paper. 

---

(iii) Perform the test specified in part (ii), ignoring that the two markers have 
marked the same exam paper. 

---

(iv) Comment on your conclusions from parts (ii) and (iii). [2] 
(v) Comment on the issues arising when analysing paired data as independent 
samples, and independent samples as though they were paired. [4] 
