---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

1 A researcher records the levels of caffeine present in their bloodstream at various time 
intervals after drinking a cup of coffee. 
T: Time elapsed 
(30-minute 
intervals) 
1 2 3 4 5 6 7 8 9 10 
C: Caffeine in 
blood 
(milligrams) 
52.2 39.3 28.3 19.03 13.96 11.46 8.78 6.55 5.43 5.03 
 
You can enter the values, for T (time) and C (caffeine), into R using: 

```{r}
time = c(1,2,3,4,5,6,7,8,9,10) 
caffeine = c(52.2,39.3,28.3,19.03,13.96,11.46,8.78,6.55,5.43,5.03)
```


#### Exercise  1

(i) Plot a scatterplot of the data. 


---

#### Exercise  2

Comment on the relationship between $C$ and $T$ based on the plot in Exercise  1.  

(ii)
The plot shows a non-linear and inverse relationship between C and T (i.e. the level of 
caffeine in the blood does not reduce linearly over time - instead it appears to decay at an 
exponential rate) 
---
#### Exercise  3

Justify why a logarithmic transformation of $C$ is appropriate. 




---
(iii)
Given the shape of the graph in (i), a log transformation should be used on the data (i.e. 
transform C to log(C)) 
This is because the original plot from part (i) has an exponential shape 

---

#### Exercise  4

(iv) Perform the transformation suggested in Exercise  3.  

(iv) 

```{r}
logcaffeine <- log(caffeine) 
logcaffeine
```


---

#### Exercise  5

(a) Plot a scatterplot of the transformed data. 
(b) Calculate the Pearson correlation coefficient for the transformed data. 
(c) Comment on the scatterplot and Pearson correlation coefficient. 

---
(v)(a)
> 

```{r}
plot(time,logcaffeine,main = "Plot for time and log(caffeine)")
```


---

**Justifying a Logarithmic Transformation of C**

**Understanding the Problem:**
We're given a dataset that shows a non-linear, inverse relationship between time (T) and caffeine levels (C). This suggests that as time increases, the caffeine level decreases exponentially.

---

**Why Logarithmic Transformation?**

1. **Linearizing the Relationship:**
   * **Exponential Decay:** Exponential decay functions, like the one likely governing caffeine metabolism, often appear curved when plotted on a linear scale.
   * **Logarithmic Transformation:** Taking the logarithm of both sides of an exponential equation can linearize the relationship. This is because logarithms are the inverse of exponential functions.

2. **Normalizing the Data:**
   * **Skewed Data:** Exponential decay often leads to skewed data, where a few large values dominate the distribution.
   * **Log Transformation:** Logarithmic transformations can help normalize skewed data, making it more suitable for linear regression and other statistical analyses.

3. **Improving Model Fit:**
   * **Linear Regression:** Linear regression models assume a linear relationship between variables.
   * **Log Transformation:** By transforming the dependent variable (C) logarithmically, we can often achieve a more linear relationship with the independent variable (T), improving the fit of the linear regression model.

---

**In essence, a logarithmic transformation of C can help:**

* **Simplify the Model:** By transforming the non-linear relationship into a linear one.
* **Improve Model Fit:** By making the data more suitable for linear regression.
* **Enhance Interpretability:** By providing a more straightforward interpretation of the relationship between time and caffeine levels.

---

**Visualizing the Transformation:**

To visualize this, consider plotting the logarithm of C against T. If the resulting plot is more linear than the original plot of C against T, then the logarithmic transformation is justified. 

By understanding these benefits, we can see why a logarithmic transformation of C is an appropriate approach to analyze this specific dataset.



Q1
(i)

(ii)
The plot shows a non-linear and inverse relationship between C and T (i.e. the level of caffeine in the blood does not reduce linearly over time - instead it appears to decay at an exponential rate) 


---

### Part C
(iii)
Given the shape of the graph in (i), a log transformation should be used on the data (i.e. transform C to log(C)) 
This is because the original plot from part (i) has an exponential shape 

---

### Part D

(iv) 

```{r}
logcaffeine <- log(caffeine) 
logcaffeine
```

---

(v)(a)

```{r}
plot(time,logcaffeine,main = "Plot for time and log(caffeine)") 
```

2468101020304050Plot for time and caffeinetimecaffeine

---

(b)
cor(time,logcaffeine,method="pearson") 
# -0.9919573 
(c)
We can see via the plot in part (v)(a) and the calculated Pearson coefficient in part that a strong negative linear relationship exists between the transformed variable and time 

---


# Input data
time <- c(1,2,3,4,5,6,7,8,9,10)
caffeine <- c(52.2,39.3,28.3,19.03,13.96,11.46,8.78,6.55,5.43,5.03)

# (i) Plot a scatterplot
plot(time, caffeine, main = "Caffeine vs. Time",
     xlab = "Time (30-min intervals)", ylab = "Caffeine (mg)")

# (ii) Comment on the relationship
# The relationship appears to be nonlinear, with caffeine levels decreasing rapidly initially and then more slowly over time.

# (iii) Justify logarithmic transformation
# A logarithmic transformation is appropriate because the data appears to be exponentially decreasing. This transformation can help linearize the relationship.

# (iv) Perform the transformation
log_caffeine <- log(caffeine)

# (v)(a) Plot the transformed data
plot(time, log_caffeine, main = "Log(Caffeine) vs. Time",
     xlab = "Time (30-min intervals)", ylab = "Log(Caffeine)")

# (v)(b) Calculate Pearson correlation coefficient
cor(time, log_caffeine)

# (v)(c) Comment on the plot and correlation
# The transformed data appears more linear, suggesting a stronger linear relationship between time and the logarithm of caffeine levels. The Pearson correlation coefficient, which measures the strength of a linear relationship, is closer to -1, indicating a strong negative linear correlation.

**Understanding the Problem:**

We're tasked with analyzing the relationship between time elapsed and caffeine levels in the bloodstream. The goal is to visualize the relationship, transform the data to improve linearity, and quantify the correlation between the transformed variables.

**R Code and Analysis:**

```R
# Input data
time <- c(1,2,3,4,5,6,7,8,9,10)
caffeine <- c(52.2,39.3,28.3,19.03,13.96,11.46,8.78,6.55,5.43,5.03)

# (i) Plot a scatterplot
plot(time, caffeine, main = "Caffeine vs. Time",
     xlab = "Time (30-min intervals)", ylab = "Caffeine (mg)")

# (ii) Comment on the relationship
# The relationship appears to be nonlinear, with caffeine levels decreasing rapidly initially and then more slowly over time.

# (iii) Justify logarithmic transformation
# A logarithmic transformation is appropriate because the data appears to be exponentially decreasing. This transformation can help linearize the relationship.

# (iv) Perform the transformation
log_caffeine <- log(caffeine)

# (v)(a) Plot the transformed data
plot(time, log_caffeine, main = "Log(Caffeine) vs. Time",
     xlab = "Time (30-min intervals)", ylab = "Log(Caffeine)")

# (v)(b) Calculate Pearson correlation coefficient
cor(time, log_caffeine)

# (v)(c) Comment on the plot and correlation
# The transformed data appears more linear, suggesting a stronger linear relationship between time and the logarithm of caffeine levels. The Pearson correlation coefficient, which measures the strength of a linear relationship, is closer to -1, indicating a strong negative linear correlation.
```

**Interpretation:**

1. **Original Plot:** The scatterplot shows a nonlinear, decreasing relationship between time and caffeine levels.
2. **Logarithmic Transformation:** The logarithmic transformation helps linearize the relationship, making it more suitable for linear correlation analysis.
3. **Transformed Plot and Correlation:** The transformed scatterplot shows a clear negative linear relationship between time and the logarithm of caffeine levels. The Pearson correlation coefficient, close to -1, confirms this strong negative linear correlation.

**Conclusion:**

The logarithmic transformation effectively linearizes the relationship between time and caffeine levels. This transformation allows for a more accurate quantification of the correlation using the Pearson correlation coefficient.
