---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

1 A researcher records the levels of caffeine present in their bloodstream at various time 
intervals after drinking a cup of coffee. 
T: Time elapsed 
(30-minute 
intervals) 
1 2 3 4 5 6 7 8 9 10 
C: Caffeine in 
blood 
(milligrams) 
52.2 39.3 28.3 19.03 13.96 11.46 8.78 6.55 5.43 5.03 
 
You can enter the values, for T (time) and C (caffeine), into R using: 
```{r}
time = c(1,2,3,4,5,6,7,8,9,10) 
caffeine = c(52.2,39.3,28.3,19.03,13.96,11.46,8.78,6.55,5.43,5.03)
```

##

#### Exercise  1

(i) Plot a scatterplot of the data. 


---

#### Exercise  2

Comment on the relationship between C and T based on the plot in Exercise  1.  


---
#### Exercise  3

Justify why a logarithmic transformation of C is appropriate. 


---

#### Exercise  4

(iv) Perform the transformation suggested in Exercise  3.  


---

#### Exercise  5

(a) Plot a scatterplot of the transformed data. 
(b) Calculate the Pearson correlation coefficient for the transformed data. 
(c) Comment on the scatterplot and Pearson correlation coefficient. 

---