---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

A statistician wants to model the number of passengers boarding a bus from a bus stop close to a student residential area. They can think of three explanatory variables: 

* Route which route it is (at 8 am or 9 am),
* Semester - if it is during the semester or not, and the 
* Temperature temperature (temp) in degrees Celsius. 

---

* The statistician has data for 20 days, given in 
the file named CS1passenger.RData, and believes that the response variable 
(Passengers) follows a Poisson distribution. After loading the data into R, the data 
frame data_passenger with all variables (Passengers, route, 
semester, temp) will be available. 


---

```{r}
PassengerData <- read.csv("PassengerData.csv")
head(PassengerData)
```


---

###  Exercise 1

State the linear predictor corresponding to models specified with the following R code, explaining all relevant terms: 

a. temp+semester 
b. temp*semester 
c. temp*semester + route 


---

### Part 2

Fit a Poisson Generalised Linear Model (GLM) to the data set for the 
model. Label this model as Model1. Your answer should 
include a summary of the fitted model. Comment on the significance of the parameters of the model. 

---

### Part 3

 Fit an improved model for the model in part (ii)(a), using your answer 
in part (ii)(b). Label this model as Model2. 
(b) Justify why Model2 improves Model1 by referring to the R output. 
You are given a new model (Model3), specified by the following R code: 

---


```{r}
Model3 <- glm(Passengers~temp+temp:semester,data=PassengerData,family=poisson(link="log"))
```


---

### Part 4

Demonstrate that Model3 outperforms the models defined in 
parts (i)(a) and (i)(b). 
(b) Comment on your answer in part (iv)(a). 

---

### Part 5
Draw a suitable plot, for the residuals of Model3, for checking the 
model’s validity. Comment on the plot. 


---

### Part 6
Calculate the predicted number of passengers for an 8 am route during the 
semester at a temperature of 0 degree Celsius, using Model3. [4] 

---

---