---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 1 - The Gamma Distribution

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Use the ‘MotorClaims’ data set (MotorClaims.csv) to answer the following questions:

```{r}
claims <- read.csv("MotorClaims.csv")

head(claims)
```


---


```{r}
summary(claims)
```


```{r}
dim(claims)
```

---

## Exercises


1. Fit Gamma distribution on the dataset provided by determining its scale ( $\theta$ ) and shape ( $\alpha$ )
parameters. State clearly the distribution with the parameters.

2. Simulate 1000 values from the distribution obtained in part 1 and print the first
six values. (Set seed to 100) 

3. Calculate the mean and variance of simulated values. 

4. Obtain a QQ plot for the simulations of 1000 values and a normal distribution. 

5. Add a line to the above plot to show the true position of normal distribution. 

6. Comment on the shape of the distribution and how close it is to a normal distribution.



---

## Solution 1:


```{r}


xbar = mean(claims$Claims)
xbar
```

```{r}
stddev = sd(claims$Claims)
stddev

variance = var(claims$Claims)
variance
```


---



### Method of Moments	


#### Scale Parameter
$${\displaystyle \theta  ={\frac {\operatorname{Var}(X)}{E(X)}}\quad \quad }$$

#### Shape Parameter
$${\displaystyle \alpha ={\frac {E(X)^{2}}{\operatorname{Var}(X)}}\quad \quad }$$ 

---


#### Method of Moments

#### Scale

```{r}
theta  <- variance/xbar
theta 
```

#### Shape

```{r}
alpha <-  (xbar^2)/variance  
alpha
```


---

```{r}
summary(claims)
```


---

### Compare to analytical values

```{r}

qgamma(0.25,shape=alpha,scale=theta)
qgamma(0.50,shape=alpha,scale=theta)
qgamma(0.75,shape=alpha,scale=theta)

```


---

#### Gamma Distributions

$$X \sim Gamma (\alpha = 2.16,\theta = 8639.533)$$


Remark: The Gamma Distribution can also be specified in terms of the shape parameter $\alpha$ and rate parameter $\lambda$ where $\lambda = 1/\theta$

The rate parameter is the "default" argument for R. To use the scale parameter instead, we must specify "***scale=***" in the code.

---


### Using the Rate Parameter

```{r}

lambda = 1/theta
lambda
```

```{r}

qgamma(0.25,shape=alpha, rate=(1/theta))
qgamma(0.50,shape=alpha, rate=(1/theta))
qgamma(0.75,shape=alpha, rate=(1/theta))

```





---

### Exercise 2


```{r}
set.seed(100)
samples <- rgamma(1000,alpha, scale = theta )
head(samples,6)

```


```{r}
summary(samples)
```

---


### Exercise 3


#### mean

```{r}
mean(samples)
```

#### Variance
```{r}
variance <- sd(samples) ^ 2
variance
```

---

### Exercise 4

Normal Probability Plot (a.k.a QQ plots)

<pre><code>
qqnorm(samples)

</code></pre>


---

#### Normal Probability Plot

```{r}
qqnorm(samples)

```


---

### Exercise 5

<pre><code>
qqnorm(samples,
       pch = 16,
       col = "blue")

</code></pre>

---

```{r echo = FALSE}
qqnorm(samples,
       pch = 16,
       col = "blue")

```


---

### Exercise 5

<pre><code>
qqnorm(samples,
       pch = 16,
       col = "blue")

qqline(samples,col="red")
</code></pre>

---

### Exercise 5

```{r echo = FALSE}
qqnorm(samples,
       pch = 16,
       col = "blue")

qqline(samples,col="red")

```


---


### Exercise 6

* Close to normal in the middle values.
* ‘Banana-shaped’ indicates positively skewed.


---

```{r}
hist(samples)
```

---