---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```
#### Exercise 

Data is provided for ‘BMIClaims’ of 150 policyholders and corresponding claim count.
(BMIClaims.csv)


```{r}
BMI <- read.csv("BMIClaims.csv")

head(BMI)
```

---


## Exercise 1

Obtain 95% confidence interval for the standard deviation of BMI (using ``qchisq()``). 

```{r}
var(BMI$BMI)
```
```{r}

sd(BMI$BMI)

```


---

#### Exercise 2

ii) Further test the standard deviation of BMI to be equal to 4 by obtaining p value. State
your conclusion of the test. 


### Inspecting the Dataframe

```{r}
nrow(BMI)
```

---



#### Variance Testing

```{r}

var.test(BMI$BMI,BMI$BMI)

```

---


If obese is defined to be BMI above 30, use binom.test to calculate 99% confidence
interval for proportion of obese people and comment on the likelihood if more than
20 pc are obese.

---

#### Exercise 3

Claim frequency can be calculated as claim count divided by number of policyholders.
Test whether claim frequency is different between obese and others.

---

#### Exercise 4



If obese is defined to be BMI above 30, use <tt>binom.test()</tt> to calculate 99% confidence
interval for proportion of obese people and comment on the likelihood if more than
20 pc are obese.

---

Claim frequency can be calculated as claim count divided by number of policyholders.
Test whether claim frequency is different between obese and others.

Solution 3:
i)

```{r}
BMI <- read.csv("BMIClaims.csv")

n <- length(BMI$BMI)

alpha <- 0.05
```



---



```{r}
sqrt(c((n-1)*var(BMI$BMI)/qchisq(1-alpha/2,df=n-1),(n-1)*var(BMI$BMI)/qchisq(alpha/2,df=n-1)))
```

---

```{r}
sigma <- 4

statistic <- (n-1)*var(BMI$BMI)/sigma^2

statistic

```{r}
> qchisq(alpha/2,n-1)
```

```{r}
qchisq(alpha/2,n-1,lower=FALSE)
```
> 2*(pchisq((n-1)*var(BMI$BMI)/sigma^2,df=n-1,lower.tail=FALSE))
[1] 3.564503e-25
[1]
---
Since p-value is less than 5%, there is sufficient evidence to reject the hypothesis,
i.e. the standard deviation of BMI is not equal to 4.

---

### Exercise 3

```{r}
x <- nrow(BMI[BMI$BMI>30,])
binom.test(x,n,conf.level = 0.99)
```

---

Exact binomial test
data: x and n
number of successes = 10, number of trials = 150, p-value < 2.2e-16
alternative hypothesis: true probability of success is not equal to 0.5
99 percent confidence interval:
0.02522882 0.13728337
sample estimates:
probability of success
0.06666667

---

Since 99% CI for p doesn’t contain $p=0.2$
it is unlikely that the proportion of obese policyholders is more than 20%.. ..

---

#### Exercise 4

```{r}
table(BMI$BMI>30,BMI$ClaimCount)

```
0
FALSE 133
TRUE
7
1
7
3

---


```{r}
y <- c(3,7)
m <- c(10,140)
poisson.test(y,m)
```

---

[1]
Comparison of Poisson rates
data: y time base: m
count1 = 3, expected count1 = 0.66667, p-value = 0.02493
alternative hypothesis: true rate ratio is not equal to 1
95 percent confidence interval:
1.001171 26.282304
sample estimates:
rate ratio

---

Since p-value is less than 5% i.e. 2.5%, there is sufficient evidence to reject the hy
pothesis, i.e. Claim frequency is different between obese and others.
(Alternatively, can use prop.test)

---