---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Q. 4)

The CDC and EISS detect influenza activity through clinical data including Influenza-like
Illness (ILI) physician visits on weekly basis. The objective of this question is to estimate
influenza-like illness (ILI) activity using Google web search logs.

The csv file FluTrain.csv aggregates this data from January 1, 2004 until December 31, 2011
as follows:

* "Week" - The range of dates represented by this observation, in year/month/day format.
* "ILI" - This column lists the percentage of ILI-related physician visits for the corresponding
week.
* "Queries" - This column lists the fraction of queries that are ILI-related for the corresponding
week, adjusted to be between 0 and 1 (higher values correspond to more ILI-related search
queries).

---

#### Exercises

A. Plot a histogram of the dependent variable, ILI. Comment on the shape of the
distribution. 
B. Plot the natural logarithm of ILI versus Queries. What does the plot suggest? 
C. Fit a linear regression model for dependent variable log(ILI). Summarize it. 
D. State the formula of the model fitted in part (iii), explaining all the terms used. 
E. Calculate R-squared and the correlation between the independent and dependent
variable. What is the relationship between the two values? 

---

F. Looking at the time period 2004-2011, which week corresponds to the highest
percentage of ILI-related physician visits? 

G. Based on the linear regression model fitted in question (iii), what is the estimate for
the percentage of ILI-related physician visits for the week computed in question (vi)?

H. What is the relative error between the estimate (prediction calculated in question (vii))
and the actual observed value for the week computed in question (vi)? (4)
vi)
vii)
viii)
---

CS1B_BMIClaims
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_BMIClaims.csv
CS1B_FluTrain
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_FluTrain.csv
CS1B_MotorClaims
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_MotorClaims.csv

---

### Exercise 1


```{r}
FluTrain <- read.csv("FluTrain.csv")
```

```{r}
str(FluTrain)
```

---


```{r}

hist(FluTrain$ILI)
```
The data is positively skewed. Most of the ILI values are small, with a relatively small number of much larger values.


---

### Exercise 2

```{r}
plot(FluTrain$Queries,log(FluTrain$ILI))
```
There is a positive linear relationship between log(ILI) and Queries.
i.e. more the number of the Google search queries, higher the number of ILI-related physician visits.


---

### Part C

```{r}
FluTrend1 = lm(log(ILI) ~ Queries, data = FluTrain)
summary (FluTrend1)
```


---

### Part D

```{r}
coef(FluTrend1)
```

$$ln y = -0.49934 +2.96129x $$

where x is the google search queries and y is the percentage of ILI related physician visits. 


---

### Part E

From the R output, R-squared value is 0.709. 

```{r}
correlation <- cor(FluTrain$Queries,log(FluTrain$ILI))
correlation

```

```{r}
correlation ^ 2
```
Hence, R-squared = Correlation ^ 2


---


#### Exercise 6


```{r}
which.max(FluTrain$ILI)
```

```{r}
FluTrain$Week[303]
```

417 Levels: 2004-01-04 - 2004-01-10 2004-01-11 - 2004-01-17 ... 2011-12-25 - 2011-12-3
1
Week of 18 th October 2009 to 24 th October 2009 corresponds to the highest percentage of ILI-related physician visits.


---


#### Exercise 7

```{r}
PredTest1 = exp(predict(FluTrend1,newdata = FluTrain))
PredTest1[303]
```

---

#### Exercise 8

viii)

```{r}
FluTrain$ILI[303]
```

```{r}
(7.618892-11.72765)/7.618892
```

---


---
It looks like you have a well-structured plan for estimating influenza-like illness (ILI) activity using Google web search logs and clinical data. Here's a step-by-step guide to address the exercises you've outlined:

### Exercise 1: Plot a Histogram of ILI

```r
FluTrain <- read.csv("FluTrain.csv")
hist(FluTrain$ILI, main="Histogram of ILI", xlab="ILI", col="blue", breaks=20)
```
The histogram shows a positively skewed distribution. Most of the ILI values are relatively small, with a few large values.

---

### Exercise 2: Plot the Natural Logarithm of ILI vs. Queries

```r
plot(FluTrain$Queries, log(FluTrain$ILI), main="Log(ILI) vs. Queries", xlab="Queries", ylab="Log(ILI)", col="red", pch=16)
```
The plot suggests a positive linear relationship between log(ILI) and Queries. This means that as the number of ILI-related search queries increases, the log of ILI-related physician visits also increases.

---

### Exercise 3: Fit a Linear Regression Model

```r
FluTrend1 <- lm(log(ILI) ~ Queries, data = FluTrain)
summary(FluTrend1)
```
The summary output will provide the coefficients, R-squared value, and other statistics for the fitted model.

---

### Exercise 4: State the Formula of the Model

Based on the linear regression model, the formula is:

$$ \ln(\text{ILI}) = -0.49934 + 2.96129 \times \text{Queries} $$

Where:
- $\ln(\text{ILI})$ is the natural logarithm of the percentage of ILI-related physician visits.
- $\text{Queries}$ is the fraction of ILI-related search queries.

---

### Exercise 5: Calculate R-squared and Correlation

```r
correlation <- cor(FluTrain$Queries, log(FluTrain$ILI))
R_squared <- correlation^2
R_squared
```
The R-squared value indicates how well the independent variable (Queries) explains the variation in the dependent variable (log(ILI)). In this case, R-squared = 0.709, which means that approximately 70.9% of the variation in log(ILI) is explained by the model.

---

### Exercise 6: Identify the Week with the Highest ILI

```r
max_week_index <- which.max(FluTrain$ILI)
FluTrain$Week[max_week_index]
```
The week corresponding to the highest percentage of ILI-related physician visits is the week of October 18th to October 24th, 2009.

---

### Exercise 7: Estimate ILI for the Week with the Highest ILI

```r
PredTest1 <- exp(predict(FluTrend1, newdata = FluTrain))
PredTest1[max_week_index]
```
This will give you the estimate for the percentage of ILI-related physician visits for that week.

---

### Exercise 8: Calculate the Relative Error

```r
observed_ILI <- FluTrain$ILI[max_week_index]
predicted_ILI <- PredTest1[max_week_index]
relative_error <- abs(predicted_ILI - observed_ILI) / observed_ILI
relative_error
```
The relative error is the absolute difference between the observed and predicted values, divided by the observed value.

I hope this helps with your analysis! If you have any further questions or need additional assistance, feel free to ask. 😊

Would you like me to walk you through the coding process or clarify any other steps?