---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Q. 4)

The CDC and EISS detect influenza activity through clinical data including Influenza-like
Illness (ILI) physician visits on weekly basis. The objective of this question is to estimate
influenza-like illness (ILI) activity using Google web search logs.
The csv file FluTrain.csv aggregates this data from January 1, 2004 until December 31, 2011
as follows:

* "Week" - The range of dates represented by this observation, in year/month/day format.
* "ILI" - This column lists the percentage of ILI-related physician visits for the corresponding
week.
* "Queries" - This column lists the fraction of queries that are ILI-related for the corresponding
week, adjusted to be between 0 and 1 (higher values correspond to more ILI-related search
queries).

---

#### Exercises

i)
Plot a histogram of the dependent variable, ILI. Comment on the shape of the
distribution. (3)
ii) Plot the natural logarithm of ILI versus Queries. What does the plot suggest? 
iii) Fit a linear regression model for dependent variable log(ILI). Summarize it. 
iv) State the formula of the model fitted in part (iii), explaining all the terms used. 
v) Calculate R-squared and the correlation between the independent and dependent
variable. What is the relationship between the two values? 

---
Looking at the time period 2004-2011, which week corresponds to the highest
percentage of ILI-related physician visits? (4)
Based on the linear regression model fitted in question (iii), what is the estimate for
the percentage of ILI-related physician visits for the week computed in question (vi)?
What is the relative error between the estimate (prediction calculated in question (vii))
and the actual observed value for the week computed in question (vi)? (4)
vi)
vii)
viii)
---

CS1B_BMIClaims
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_BMIClaims.csv
CS1B_FluTrain
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_FluTrain.csv
CS1B_MotorClaims
http://www.actuariesindia.org/downloads/Year_2020_November_Exam/CS1B_MotorClaims.csv

---

### Exercise 1


```{r}
FluTrain <- read.csv("FluTrain.csv")
```

```{r}
str(FluTrain)
```
---


```{r}

hist(FluTrain$ILI)
```
The data is positively skewed. Most of the ILI values are small, with a relatively small number of much larger values.


---

### Exercise 2

```{r}
plot(FluTrain$Queries,log(FluTrain$ILI))
```
There is a positive linear relationship between log(ILI) and Queries.
i.e. more the number of the Google search queries, higher the number of ILI-related physician visits.


---

### Exercise 3

```{r}
FluTrend1 = lm(log(ILI) ~ Queries, data = FluTrain)
summary (FluTrend1)
```


---

### Exercise 4

```{r}
coef(FluTrend1)
```

$$ln y = -0.49934 +2.96129x $$

where x is the google search queries and y is the percentage of ILI related physician visits. 


---

### Exercise 5

From the R output, R-squared value is 0.709. 

```{r}
correlation <- cor(FluTrain$Queries,log(FluTrain$ILI))
correlation

```

```{r}
correlation ^ 2
```
Hence, R-squared = Correlation ^ 2


---


#### Exercise 6


```{r}
which.max(FluTrain$ILI)
```

```{r}
FluTrain$Week[303]
```

417 Levels: 2004-01-04 - 2004-01-10 2004-01-11 - 2004-01-17 ... 2011-12-25 - 2011-12-3
1
Week of 18 th October 2009 to 24 th October 2009 corresponds to the highest percentage of ILI-related physician visits.


---


#### Exercise 7

```{r}
PredTest1 = exp(predict(FluTrend1,newdata = FluTrain))
PredTest1[303]
```

---

#### Exercise 8

viii)

```{r}
FluTrain$ILI[303]
```

```{r}
(7.618892-11.72765)/7.618892
```

---


---
