Use the ‘MotorClaims’
(MotorClaims.csv)
i)
to
answer
the
following
questions:

provided

Exercises
===================================================

1. Fit Gamma distribution on the dataset provided by determining its scale and shape
parameters. State clearly the distribution with the parameters.

2. Simulate 1000 values from the distribution obtained in question (i) and print the first
six values. (Set seed to 100) (

3. Calculate the mean and variance of simulated values. 
4. Obtain a QQ plot for the simulations of 1000 values and a normal distribution. 
5. Add a line to the above plot to show the true position of normal distribution. 
6. Comment on the shape of the distribution and how close it is to a normal distribution.




Solution 1:
==========================================

```{r}
claims <- read.csv("MotorClaims.csv")

xbar = mean(claims$Claims)
xbar
```

```{r}
stddev = sd(claims$Claims)
variance = stddev ^ 2
variance
```

```{r}
lambda <- xbar/variance
lambda
```

```{r}
alpha <-xbar * lambda
alpha
```

$$X ~ Gamma (2.16, 0.0001)$$

Exercise 2
=========================================================================
ii)

```{r}
set.seed(100)
samples <- rgamma(1000,alpha,lambda)
head(samples,6)

```
[1] 9305.461 2125.292 25926.442 15685.099 18120.436
8605.442
[2]
[1]
[2]
[5]
iii)
mean(samples)
[1] 18423.47
variance <- sd(samples) ^ 2
variance
[1] 153958637



Exercise 4
=========================================================================
iv)
```{r}
qqnorm(samples)

```

Exercise 5
=========================================================================
v)
qqline(samples,col="red")
[2]
vi)
Close to normal... (1 mark) in the middle values... (1 mark) .
‘Banana-shaped’ indicates positively skewed... (1 mark) .


