---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis
* Probability Distributions

* Question 2 - Principal Component Analysis
* Exam on basis of ***Base R*** 

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

## mtcars

The dataset ‘mtcars’ (built into R) consists of data on various models of car, taken from an American motoring magazine (1974 Motor Trend magazine). 

For each car, there are certain features expressed in varying units.

---

### mtcars

```{r}
head(mtcars)
```

---


### mtcars

<pre><code>
help(mtcars)
</code></pre>

#### Format
A data frame with 32 observations on 11 (numeric) variables.

1. ***mpg*** - 	Miles/(US) gallon
2. ***cyl***	- Number of cylinders
3. ***disp*** - 	Displacement (cu.in.)
4. ***hp*** - 	Gross horsepower
5. ***drat*** - 	Rear axle ratio
6. ***wt*** - 	Weight (1000 lbs)
7. ***qsec*** - 	1/4 mile time
8. ***vs*** - 	Engine (0 = V-shaped, 1 = straight)  [*CATEGORICAL*]
9. ***am*** - 	Transmission (0 = automatic, 1 = manual) [*CATEGORICAL*]
10. ***gear*** - 	Number of forward gears
11. ***carb*** - 	Number of carburetors

---


## Exercises

Load the ‘***mtcars***’ dataset which is built into R. How many observations and variables
are there in this dataset? Your answer should include the R output.

2. Identify the categorical variables from the dataset ‘***mtcars***’ and create a dataset
excluding the categorical variables. 

3. How many observations and variables are there in the new dataset? Your answer
should include the R output. 

4. Carry out a principal component analysis on the new dataset of mtcars by passing two
arguments, ‘center’ and ‘scale’ to be TRUE. Your answer should include a summary
of the principal component analysis. 

5. How many components of the reduced data should be retained using the output
derived in part 4? Also state the reason for the same.

---

#### Exercise 1
```{r}
data("mtcars")
str(mtcars)
```

There are 32 observations (car models) and 11 variables (car features) in the dataset.


---

```{r}
summary(mtcars)
```

---

The two variables ‘vs’ and ‘am’ are categorical variables. (This can be identified using ***str*** or ***summary*** function)

There are the 8th and 9th variables
```{r}
mtcars1 <- mtcars[,c(1:7,10,11)]
```

---

### Exercise 3


```{r}
str(mtcars1)
```

There are 32 observations (car models) and 9 variables (car features) in the dataset.


---

### Exercise 4



```{r}
mtcars1.pca <- prcomp(mtcars1,center = TRUE,scale=TRUE)
summary(mtcars1.pca)
```

---

### Exercise 5

* The R analysis shows that the proportion of variance explained by first three principal components is 91.5% and by
first four variables is 94.5%.

* Thus, it will be appropriate to retain the first three (or four) principal components.

---


---