---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


Use the ‘MotorClaims’ data set (MotorClaims.csv) to answer the following questions:

```{r}
claims <- read.csv("MotorClaims.csv")

head(claims)
```


---


```{r}
summary(claims)
```


```{r}
dim(claims)
```

---

## Exercises


1. Fit Gamma distribution on the dataset provided by determining its scale and shape
parameters. State clearly the distribution with the parameters.

2. Simulate 1000 values from the distribution obtained in question (i) and print the first
six values. (Set seed to 100) 

3. Calculate the mean and variance of simulated values. 

4. Obtain a QQ plot for the simulations of 1000 values and a normal distribution. 

5. Add a line to the above plot to show the true position of normal distribution. 

6. Comment on the shape of the distribution and how close it is to a normal distribution.



---

## Solution 1:


```{r}


xbar = mean(claims$Claims)
xbar
```

```{r}
stddev = sd(claims$Claims)
variance = stddev ^ 2
variance
```


---


```{r}
lambda <- xbar/variance
lambda
```

```{r}
alpha <-xbar * lambda
alpha
```


--- 

$$X ~ Gamma (2.16, 0.0001)$$

---

### Exercise 2


```{r}
set.seed(100)
samples <- rgamma(1000,alpha,lambda)
head(samples,6)

```



---


### Exercise 3

```{r}
mean(samples)
```

```{r}
variance <- sd(samples) ^ 2
variance
```

---

### Exercise 4

Normal Probability Plot (a.k.a QQ plots)

```{r}
qqnorm(samples)

```


---

```{r}
qqnorm(samples,
       pch = 16,
       col = "blue")

```


---

### Exercise 5

```{r}
qqnorm(samples,
       pch = 16,
       col = "blue")

qqline(samples,col="red")

```

---

vi)
Close to normal... (1 mark) in the middle values... (1 mark) .
‘Banana-shaped’ indicates positively skewed... (1 mark) .


