---
title: "Statistics with R"
subtitle: "Modelling Data with R"
author: "Dragonfly Stats"
date: "2016/12/12"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

## mtcars

The dataset ‘mtcars’ (built into R) consists of data on various models of car, taken from an American motoring magazine (1974 Motor Trend magazine). 
For each car, there are certain features expressed in varying units.

```{r}
library(xaringan)
```

---


## Exercises



1. Load the ‘mtcars’ dataset which is built into R. How many observations and variables
are there in this dataset? Your answer should include the R output. (4)
2. Identify the categorical variables from the dataset ‘mtcars’ and create a dataset
excluding the categorical variables. (5)
3. How many observations and variables are there in the new dataset? Your answer
should include the R output. (2)

4. Carry out a principal component analysis on the new dataset of mtcars by passing two
arguments, ‘center’ and ‘scale’ to be TRUE. Your answer should include a summary
of the principal component analysis. (5)
5. How many components of the reduced data should be retained using the output
derived in question (iv)? Also state the reason for the same.

---

Solution 2:
i)
```{r}
data("mtcars")
str(mtcars)
```

There are 32 observations (car models) and 11 variables (car features) in the dataset.


---

```{r}
summary(mtcars)
```

---

The two variables ‘vs’ and ‘am’ are categorical variables. (This can be identified using str or summary function)

```{r}
mtcars1 <- mtcars[,c(1:7,10,11)]
```

---



Exercise 3
==================================================
iii)

```{r}
str(mtcars1)
```

There are 32 observations (car models) and 9 variables (car features) in the dataset.


---


Exercise 4
==================================================
iv)


```{r}
mtcars1.pca <- prcomp(mtcars1,center = TRUE,scale=TRUE)
summary(mtcars1.pca)
```
Importance of components:
PC1 PC2
Standard deviation
PC3
PC4
PC5
PC6
PC7 PC8
[2]
[2]
[1]
PC9
2.3782 1.4429 0.71008 0.51481 0.42797 0.35184 0.32413 0.2419 0.14896
Proportion of Variance 0.6284 0.2313 0.05602 0.02945 0.02035 0.01375 0.01167 0.0065 0.00247
Cumulative Proportion 0.6284 0.8598 0.91581 0.94525 0.96560 0.97936 0.99103 0.9975 1.00000

---

v)
The R analysis shows that the proportion of variance explained by first three principal components is 91.5% and by
first four variables is 94.5%.
Thus, it will be appropriate to retain the first three (or four) principal components.

---
