[24]
The dataset ‘mtcars’ (built into R) consists of data on various models of car, taken from an American motoring magazine (1974 Motor Trend magazine). 
For each car, there are certain features expressed in varying units.

Exercises
===============================================
Q. 2)
dataset
Load the ‘mtcars’ dataset which is built into R. How many observations and variables
are there in this dataset? Your answer should include the R output. (4)
Identify the categorical variables from the dataset ‘mtcars’ and create a dataset
excluding the categorical variables. (5)
How many observations and variables are there in the new dataset? Your answer
should include the R output. (2)
Carry out a principal component analysis on the new dataset of mtcars by passing two
arguments, ‘center’ and ‘scale’ to be TRUE. Your answer should include a summary
of the principal component analysis. (5)
How many components of the reduced data should be retained using the output
derived in question (iv)? Also state the reason for the same.



Solution 2:
i)
data("mtcars")
> str(mtcars)
'data.frame': 32 obs. of 11 variables:
$ mpg : num 21 21 22.8 21.4 18.7 18.1 14.3 24.4 22.8 19.2 ...
$ cyl : num 6 6 4 6 8 6 8 4 4 6 ...
$ disp: num 160 160 108 258 360 ...
$ hp : num 110 110 93 110 175 105 245 62 95 123 ...
$ drat: num 3.9 3.9 3.85 3.08 3.15 2.76 3.21 3.69 3.92 3.92 ...
$ wt : num 2.62 2.88 2.32 3.21 3.44 ...
$ qsec: num 16.5 17 18.6 19.4 17 ...
$ vs : num 0 0 1 1 0 1 0 1 1 1 ...
$ am : num 1 1 1 0 0 0 0 0 0 0 ...
$ gear: num 4 4 4 3 3 3 3 4 4 4 ...
$ carb: num 4 4 1 1 2 1 4 2 2 4 ...
There are 32 observations (car models) and 11 variables (car features) in the dataset.
[4]
ii)
summary(mtcars)
mpg
Min.
:10.40
1st Qu.:15.43
Median :19.20
Mean
:20.09
3rd Qu.:22.80
Max.
:33.90
wt
Min.
:1.513
1st Qu.:2.581
Median :3.325
Mean
:3.217
3rd Qu.:3.610
cyl
Min.
:4.000
1st Qu.:4.000
Median :6.000
Mean
:6.188
3rd Qu.:8.000
Max.
:8.000
qsec
Min.
:14.50
1st Qu.:16.89
Median :17.71
Mean
:17.85
3rd Qu.:18.90
disp
Min.
: 71.1
1st Qu.:120.8
Median :196.3
Mean
:230.7
3rd Qu.:326.0
Max.
:472.0
vs
Min.
:0.0000
1st Qu.:0.0000
Median :0.0000
Mean
:0.4375
3rd Qu.:1.0000
hp
Min.
: 52.0
1st Qu.: 96.5
Median :123.0
Mean
:146.7
3rd Qu.:180.0
Max.
:335.0
am
Min.
:0.0000
1st Qu.:0.0000
Median :0.0000
Mean
:0.4062
3rd Qu.:1.0000
drat
Min.
:2.760
1st Qu.:3.080
Median :3.695
Mean
:3.597
3rd Qu.:3.920
Max.
:4.930
gear
Min.
:3.000
1st Qu.:3.000
Median :4.000
Mean
:3.688
3rd Qu.:4.000
Page 3 of 8IAI
Max.
:5.424
carb
Min.
:1.000
1st Qu.:2.000
Median :2.000
Mean
:2.812
3rd Qu.:4.000
Max.
:8.000
Max.
:22.90
Max.
:1.0000
Max.
:1.0000
Max.
CS1B-1120
:5.000
The two variables ‘vs’ and ‘am’ are categorical variables. (This can be identified using str or summary function)
mtcars1 <- mtcars[,c(1:7,10,11)]






Exercise 3
==================================================
iii)

```{r}
str(mtcars1)
```

There are 32 observations (car models) and 9 variables (car features) in the dataset.

Exercise 4
==================================================
iv)


```{r}
mtcars1.pca <- prcomp(mtcars1,center = TRUE,scale=TRUE)
summary(mtcars1.pca)
```
Importance of components:
PC1 PC2
Standard deviation
PC3
PC4
PC5
PC6
PC7 PC8
[2]
[2]
[1]
PC9
2.3782 1.4429 0.71008 0.51481 0.42797 0.35184 0.32413 0.2419 0.14896
Proportion of Variance 0.6284 0.2313 0.05602 0.02945 0.02035 0.01375 0.01167 0.0065 0.00247
Cumulative Proportion 0.6284 0.8598 0.91581 0.94525 0.96560 0.97936 0.99103 0.9975 1.00000
[2]
[5]
v)
The R analysis shows that the proportion of variance explained by first three principal components is 91.5% and by
first four variables is 94.5%.
Thus, it will be appropriate to retain the first three (or four) principal components.
[3]
[19 Marks]
Solution 3:
i)
> BMI <- read.csv("BMIClaims.csv")
> n <- length(BMI$BMI)
> alpha <- 0.05
...
[2]
> sqrt(c((n-1)*var(BMI$BMI)/qchisq(1-alpha/2,df=n-1),(n-1)*var(BMI$BMI)/qchisq(alpha/2
,df=n-1)))
[2]
[1] 5.920028 7.434763
[2]
[6]
Page 4 of 8IAI
CS1B-1120
ii)
> sigma <- 4
> statistic <- (n-1)*var(BMI$BMI)/sigma^2
> statistic
[1] 404.5421
> qchisq(alpha/2,n-1)
[1] 117.098
> qchisq(alpha/2,n-1,lower=FALSE)
[1] 184.687
> 2*(pchisq((n-1)*var(BMI$BMI)/sigma^2,df=n-1,lower.tail=FALSE))
[1] 3.564503e-25
[1]
[2]
[1]
Since p-value is less than 5%, there is sufficient evidence to reject the hypothesis,
i.e. the standard deviation of BMI is not equal to 4.
[2]
[6]
iii)
> x <- nrow(BMI[BMI$BMI>30,])
> binom.test(x,n,conf.level = 0.99)
[1]
[2]
Exact binomial test
data: x and n
number of successes = 10, number of trials = 150, p-value < 2.2e-16
alternative hypothesis: true probability of success is not equal to 0.5
99 percent confidence interval:
0.02522882 0.13728337
sample estimates:
probability of success
0.06666667
Since 99% CI for p doesn’t contain p=0.2
it is unlikely that the proportion of obese policyholders is more than 20%.. ..
[1]
[1]
[1]
[6]
iv)
> table(BMI$BMI>30,BMI$ClaimCount)
0
FALSE 133
TRUE
7
1
7
3
> y <- c(3,7)
> m <- c(10,140)
> poisson.test(y,m)
[2]
[1]
Comparison of Poisson rates
data: y time base: m
count1 = 3, expected count1 = 0.66667, p-value = 0.02493
alternative hypothesis: true rate ratio is not equal to 1
95 percent confidence interval:
1.001171 26.282304
sample estimates:
rate ratio
6
Since p-value is less than 5% i.e. 2.5%, there is sufficient evidence to reject the hy
pothesis, i.e. Claim frequency is different between obese and others.
(Alternatively, can use prop.test)
[2]
[6]
