---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3

---

### Exercise

Consider a sample of 10,000 car insurance policies sold to drivers who have had a 
driving licence for no more than 5 years. 
<br>
The data are given in the file 
<tt>CarInsurance.Rdata</tt>, which contains the following four variables: 

* ***age***: a factor taking the values 0 for young and 1 for old; this refers to the age of 
the policyholder. 
* ***LY***: the number of years the policyholder has had a driving licence before the start 
of the current year. 
* ***NCD***: the number of years the policyholder has had a no claims discount before the 
start of the current year. 
* ***claims***: the number of claims the policyholder submitted during the current year. 

<br>
It has been suggested that drivers belonging to the old age category have more years 
of no claims discount than younger drivers. 

---

### Load the Data

```{r include=FALSE}
rm(list=ls())
```

```{r}
load("CarInsurance.Rdata")
ls()
```

---

## Exercises

**A.** Plot two histograms for the number of years with no claims discount, one for 
young policyholders and another for old policyholders. 

**B.** Comment on the suggestion that drivers belonging to the old age category 
have more years of no claims discount than young drivers, using the 
histograms from part (A).  

**C.** Calculate the proportion of policyholders with more than 2 years of no claims discount among 
young policyholders. 

**D.** Calculate the proportion of policyholders with more than 2 years of no claims discount among old 
policyholders. 

**E.** Test the hypothesis that the proportion of policyholders with more than 2 
years of no claims discount is equal in both groups (young and old 
policyholders). 


---

### Exercise A.

Plot histograms for no claims discount years:

```{r eval=FALSE}
NCD_old = NCD[age==1] 
NCD_young = NCD[age==0] 

hist(NCD_old, prob=TRUE, breaks = c(0,1,2,3,4,5)) 
```

---

```{r echo=FALSE}
NCD_old = NCD[age==1] 
NCD_young = NCD[age==0] 

hist(NCD_old, prob=TRUE, breaks = c(0,1,2,3,4,5)) 
```

---

Alternative R code (this produces more accurate plots):

```{r eval = FALSE}
hist(NCD_young, breaks=seq(-0.5,5.5,by=1))

hist(NCD_old, breaks=seq(-0.5,5.5,by=1))
```

---

Alternative R code (this produces more accurate plots):

```{r echo = FALSE}
hist(NCD_young, breaks=seq(-0.5,5.5,by=1))

hist(NCD_old, breaks=seq(-0.5,5.5,by=1))
```

---

### Exercise B.

* The distribution for old policyholders is more concentrated on the right, so the number of years of NCD seems to be higher for old policyholders. 

---

### Exercise C.

Calculate proportions:

```{r}
prop_young = sum(NCD_young > 2)/length(NCD_young) 
prop_young
```

```{r}
prop_young = mean(NCD_young > 2)
prop_young
```


Proportion is 0.5736150 

---

### Exercise D.

Alternative R code:

```{r}
length(NCD_old[NCD_old>2])/length(NCD_old)
```

```{r}
prop_old = sum(NCD_old > 2)/length(NCD_old) 
prop_old

```
Proportion is 0.9704187 

---

### Exercise D.

Alternative R code:

```{r}

length(NCD_old[NCD_old>2])/length(NCD_old)
```


---

### Exercise E.

Test the hypothesis that prop of $NCD >2$ is equal for young and old, against alternative that 
proportions are not equal: 

```{r}
NCDgreaterTwo = c(sum(NCD_young > 2), sum(NCD_old > 2)) 

head(NCDgreaterTwo)
```

```{r}
YoungOld = c(length(NCD_young), length(NCD_old)) 

head(YoungOld)
```

---

### Exercise E.

Two-sample test for equality of proportions:

```{r}
prop.test(NCDgreaterTwo, YoungOld, correct=F) 
```


The $p$-value is very small, therefore we reject the null hypothesis of equal proportions. 



---

### Exercises

An analyst assumes that the number of claims per year has a Poisson distribution. The 
analyst wishes to fit a generalised linear model (GLM) to the observed claim counts 
using age, LY and NCD as explanatory variables. 


**F.** Fit a GLM to the claim counts with age, LY and NCD as the main effect explanatory variables.


**G.** To simplify the integration of the model into business processes, an insurer wishes to 
use only models with at most two explanatory variables. 

Determine which two explanatory variables (out of age, LY and NCD) should 
be used in a GLM for the claim counts if only two explanatory variables are 
allowed. 

**H.** Comment on the comparison of the fit of the various candidate models.  

---


### Exercise F.


**Fit a GLM to the claim counts with age, LY and NCD as the main effect 
explanatory variables.**

```{r eval=FALSE}
glm(claims ~ factor(age) + LY + NCD, family = poisson) 

summary(glm(claims ~ age + LY + NCD, family = poisson)) 


```

---

```{r echo=FALSE}
glm(claims ~ factor(age) + LY + NCD, family = poisson) 
summary(glm(claims ~ age + LY + NCD, family = poisson)) 


```

---

### Exercise G.

We need to fit the three possible models with two variables each and compare the AIC 
values:

```{r}
glm(claims ~ LY + NCD, family = poisson)$aic
# 23102.41 
```

---

### Exercise G

```{r}
glm(claims ~ age+LY, family = poisson)$aic
# 23125.08 

```

---

### Exercise G

```{r}
glm(claims ~ age+NCD, family = poisson)$aic
# 23131.97 
```

Model LY + NCD has the smallest AIC value and should be preferred. 

---

### Exercise H.

* The AIC for model LY + NCD (23102) is slightly lower than the AIC for
age + LY + NCD (23104), but the AIC values are very close. 

* We would prefer model LY + NCD as it also involves fewer variables (it is a smaller 
model). 
---

---


---