---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

### Exercise 

An engineer is considering the maximum number of people who can occupy a lift. 

In particular, the engineer wants to assess the probability of exceeding a maximum 
weight when eight people are allowed to use the lift at the same time, compared to 
nine people. 

Assume that: 
* the total weight of eight people chosen at random follows a normal distribution 
with a mean of 560 kg and standard deviation of 57 kg 

* the total weight of nine people chosen at random follows a normal distribution 
with a mean of 630 kg and standard deviation of 61 kg.

---

### Exercises

A. Calculate the probability that the total weight of eight people exceeds 650 kg.
  
B. Calculate the probability that the total weight of nine people exceeds 650 kg.
  
C. Comment on your answers to parts (A) and (B).  

D.  Calculate the interval for the central region containing 80% of the distribution 
of the total weight of eight people.

---

### Exercises

Assume now that the total weight of eight people chosen at random follows a gamma 
distribution with parameters $\alpha = 96.5220$ and $\lambda = 0.1724$. 

E. Calculate the interval for the central region containing 80% of the 
distribution of the total weight of eight people in this case. 

F. Comment on your answers to parts (D) and (E). 


---

### Part A

Let $X$ be the total weight of 8 people. 

By the assumption: $X \sim N(560 , 57^2)$. We are 
interested in the probability $P(X > 650)$:

```{r}
# Upper Tail / Right hand side of Bell Curve
pnorm(x=650,mean=560,sd=57,lower=FALSE) 
```

```{r}
# Complement of LHS of Bell Curve
1 - pnorm(x=650,mean=560,sd=57) 
```

The probability that the total weight of 8 people exceeds 650kg is equal to 0.05717406. 

---

### Part B

Let $Y$ be the total weight of 9 people. By the assumption: $Y \sim N(630 , 61^2)$. 

We are interested in the probability $P(Y > 650)$:


```{r}
pnorm(650,mean=630,sd=61, lower=FALSE) 
```

```{r}
1 - pnorm(650,mean=630,sd=61) 
```


* We get that the probability that the total weight of 9 people exceeds 650kg is much higher 
and is equal to 0.3715054. 

---

#### Exercise C

* Parts (A) & (B) show that as more people enter the lift, the probability of exceeding 
650kg increases. 
* While the probability of exceeding 650kg is small with 8 people, 
exceeding the maximum weight is considerably more likely with 9 people. 
---

#### Exercise D

* Again, $X \sim N(560, 57^2)$, where $X$ is the total weight of 8 people. 

* In order to find the central region that contains 80% of the distribution we need to identify the 
10%-percentile and the 90%-percentile of X.

* We use the function “<tt>qnorm()</tt>" in the code:

```{r}
qnorm(p=0.1,mean=560,sd=57) 
```

```{r}
qnorm(0.9,560,57) 
```


The requested interval is **[486.9516, 633.0484]**. 
---

#### Gamma Distribution

* Shape parameter $\alpha$ - R argument is <tt>shape=</tt>.

* Rate parameter $\lambda$ - R argument is <tt>rate=</tt>.

* Scale parameter is the inverse of the rate parameter - R argument is <tt>scale=</tt>.

---

#### Exercise E



We now have $Y \sim G(96.5220, 0.1724)$, where $Y$ is the total weight of 8 people. 

The calculation this time produces:

```{r}
qgamma(p = 0.1, shape= 96.5220, rate = 0.1724) 

```


```{r}
qgamma(p = 0.9, shape = 96.5220, rate = 0.1724) 
```

and the interval is **[488.195, 634.0333]**. 

---

Using Scale parameter

```{r}
qgamma(p = 0.1, shape= 96.5220, scale = (1/0.1724)) 

```


```{r}
qgamma(p = 0.9, shape = 96.5220, scale = (1/0.1724)) 
```

---

#### Exercise F

* The intervals are very similar. 
* The mean and standard deviation of the two distributions are (approximately) equal. 
* As the first (shape) parameter of the gamma distribution is large, the distribution is close 
to a normal distribution. 

---

END


---

END

---

Candidates overall answered well this question. 
In parts (i) and (ii) a common error was calculating the probability at the wrong tail, e.g. 
using pnorm() instead of 1-pnorm().
In part (iv) some candidates provided a wrong answer giving a single value with 
qnorm(0.8, …). 
In part (v) (b), well prepared candidates referenced the gamma shape parameter.
