O1-Question-1.Rmd

Q. 1) The Gym instructor participated in global survey of weights used for biceps curl machine. 
You are required to help him analyse the data to share the results.

#### Exercises

A. Load the data “weights.csv” into R and print the summary of the data.
B. Obtain a 90% confidence interval for the standard deviation of weights lifted by Males
(M).
C. Assuming that the weights follow normal distribution with mean and standard deviation 
of data provided for Males.
Test the Hypothesis that µ=12 kgs in R and obtain the p-value using pnorm or 
otherwise. 
D. Further, assuming that the weights follow normal distribution with mean and standard 
deviation of data provided for Females (F).
Test the Hypothesis that $\mu=7$ kgs in R and obtain the p-value using <tt>pnorm()</ttor otherwise.
E. Additionally, generate datasets for Males and Females of size 10. Use set.seed(2022)
to set seed value for generating sample data. 
Perform paired t test to test the hypothesis that average weights used by Females are 5 
kgs less than weights used by Males.
Comment whether average weights used by Females is less than 5kgs or not. 


#################################################################################
Solution 1:
i) data<-read.csv("weights.csv")
summary(data)
Gender Weights Day 
Length:216 Min. : 5.00 Min. : 1.00 
Class :character 1st Qu.: 7.00 1st Qu.: 9.00 
Mode :character Median :10.00 Median :17.00 
 Mean :10.22 Mean :17.16 
 3rd Qu.:13.00 3rd Qu.:25.25 
 Max. :16.00 Max. :34.00
(1)
(1)

---

ii) 

```{r}
weight_M <- subset(data$Weights, data$Gender == "M" , select = c(data$Weights), drop = 
FALSE)
```

Alternate:
M_subset<-data[data$Gender=="M",]
weight_M <- M_subset$Weights
Marks given for other valid alternate solutions. 
nm <- length(weight_M) -1 /* n-1
sm <- sd(weight_M)
nm
[1] 107
sm
[1] 2.306458
sm*sqrt(nm/qchisq(c(0.95,0.05),nm))
[1] 2.075453 2.601171
(1)
(0.5)
(0.5)
(1)
---

iii) mum<-12

```{r}
xbarm <-mean(weight_M)
nm <- length(weight_M)
Statisticm <- ((xbarm-mum)/(sm/sqrt(nm)))
Statisticm
[1] 3.295867
1- pnorm(Statisticm)
[1] 0.0004905918
p- value <5%, there is no significant evidence to accept the null hypothesis
```

---

Alternate:
One Sample t-test
data: weight_M
t = 3.2959, df = 107, p-value = 0.001332
alternative hypothesis: true mean is not equal to 12
95 percent confidence interval:
12.29151 13.17145
sample estimates:
mean of x 
12.73148
p- value = .001332 <5%, there is no significant evidence to accept the null hypothesis

---
[Max 3]
iv) weight_F <- subset(data$Weights, data$Gender == "F" , select = c(data$Weights), drop = 
FALSE)
(1)
IAI CS1B-1222
 Page 3 of 13
nf <- length(weight_F)
xbarf
[1] 7.703704
sf <- sd(weight_F)
nf
[1] 108
sf
[1] 2.079087
Statisticf <- ((xbarf-muf)/(sf/sqrt(nf)))
Statisticf
[1] 3.517459
pnorm(Statisticf)
[1] 0.0002178499

---
p- value <5%, there is no significance evidence to accept the null hypothesis
Alternate:
F_subset<-data[data$Gender=="F",]
weight_F <- F_subset$Weights
t.test(weight_F,mu=7,alternative = "two.sided")
One Sample t-test
data: weight_F
t = 3.5175, df = 107, p-value = 0.0006407
alternative hypothesis: true mean is not equal to 7
95 percent confidence interval:
7.307108 8.100300
sample estimates:
mean of x 
7.703704
The required p value is 0.0006407.
p- value <5%, there is no significance evidence to accept the null hypothesis

----

### Part (E)

v) set.seed(2022)
male<-c(rnorm(10,mum,sm))
female<-c(rnorm(10,muf,sf))
mean(male)
[1] 10.69983
mean(female)
[1] 7.015142
t.test(male,female,paired=TRUE,alternative="less",mu=5)
 Paired t-test
data: male and female
t = -1.7111, df = 9, p-value = 0.06061
alternative hypothesis: true mean difference is less than 5
95 percent confidence interval:
 -Inf 5.093811
sample estimates:
mean difference 
 3.684693

---

Since p-value >5%,there is no strong evidence to reject null hypothesis
The average weights used by females is 3.7kg lesser than weights used by male

---
