---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 2

---

Q. 2) An investor wants to invest in a portfolio comprising of two stocks. He is evaluating four 
stocks for the purpose of his investment viz ABC Oil, XYZ Airways, PQR Realty and LMN 
Bank. 
Annual returns (%) from these stocks over the past 10 years have been tabulated below –
Year ABC
Oil XYZ Airways PQR Realty LMN Bank
1 -6 8 -20 14
2 -10 0 19 4
3 3 -4 3 5
4 18 10 13 15
5 -10 20 20 19
6 1 8 7 9
7 3 0 -3 6
8 -13 10 13 10
9 -14 5 20 7
10 13 -19 1 16

---

i) Calculate the mean and standard deviation of the annual returns from each stock using 
appropriate functions in R. (5)
ii) Calculate the sample correlation coefficient between annual returns from various pairs 
of stocks viz (ABC, XYZ), (ABC, PQR), (ABC, LMN), (XYZ, PQR), (XYZ, LMN), 
(PQR, LMN). Use Pearson’s method. (3)
IAI CS1B-1222

---

iii) The investor feels that in between the following pairs of stocks, there should exist a 
perfect negative correlation –
a) Changes in oil prices affect the prices of aviation fuel thus affecting aviation 
profits. Thus, the investor feels that returns from ABC Oil and XYZ Airways 
should have perfect negative correlation.
b) He also feels that changes in interest rates lead to equal and opposite effects on 
the profits of the real estate sector and financial services sector. Thus, he is also 
contemplating a perfect negative correlation between returns from PQR Realty 
and LMN Bank. 
---

Calculate a 95% two-sided confidence interval for the population correlation coefficient for 
the above two pairs using Pearson’s Method and hence test the validity of the investor’s 
premonitions. [Hint: Use cor.test function in R] (4)

--- 
The investor is considering two alternative investment strategies for setting up his portfolio 
as follows –
Strategy A: Investing 50% of his funds in ABC Oil and balance 50% in XYZ Airways.
Strategy B: Investing 25% of his funds in LMN Bank and balance 75% in PQR Realty.
He has asked to develop the following metrics to assist with the final choice between 
Strategy A or Strategy B:
Metric 1: Strategy with highest Mean Returns would be selected.
Metric 2: Strategy with lowest Risk (Lowest Standard Deviation) would be selected.
Metric 3: Strategy with highest “Risk-Adjusted Returns” would be selected.
Risk-Adjusted Return = (Mean Return)/(Standard Deviation of Return)

---

iv) Calculate the mean return for the above two strategies. Which strategy will be chosen 
based on Metric 1 i.e. highest Mean Returns? (2)
v) Investor feels that both the strategies are equally risky. Using var.test function in R 
(two-sided), check whether variance of returns from Strategy A is equal to the variance 
of returns from Strategy B at 10% level of significance. Your answer should clearly 
mention the null and alternate hypothesis, the p-value and the outcome of the test. 
[Hint: Create two new vectors for strategies A and B using the proportions given 
earlier.] (4)

---

Investor has come up with an alternative formula to calculate the standard deviation of both 
the strategies which is dependent on the correlation coefficients calculated in part (iii). 
Using this formula, Standard deviation for a portfolio of two stocks 1 and 2 is to be 
calculated using the following method:
SD(1,2) = (P1
2
* SD1
2 + 2 Corr(1,2) * P1 * P2 * SD1 * SD2 + P2
2
* SD2
2
) ½ 
Where – P1 and P2 denote the proportion of invested funds in stocks 1 and 2 respectively. 
IAI CS1B-1222
 Page 4 of 7
vi) Using the confidence interval for population correlation coefficient calculated in part 
(iii), calculate interval for the standard deviation of returns for Strategy A and Strategy 
B using the above formula. Which strategy will you choose based on Metric 2 i.e. 
Lowest Standard Deviation of Returns? (4)
vii) Which strategy outperforms the other in terms of Metric 3 i.e. “Higher Risk-Adjusted 
Returns”? Use means calculated in part (iv) and interval of standard deviations 
calculated in part (vi) to calculate interval for risk-adjusted returns. 
Also, compare your choices from part (iv), (vi) and (vii) and comment on the same. (2)


##########################Solution 2
:
i
)
> abc <
- c(
-6,
-10,3,18,
-10,1,3,
-13,
-14,13)
> xyz <
- c(8,0,
-4,10,20,8,0,10,5,
-19)
> pqr <
- c(
-20,19,3,13,20,7,
-3,13,20,1)
> lmn <
- c(14,4,5,15,19,9,6,10,7,16)
> 
> abc_mean = mean(abc)
> xyz_mean = mean(xyz)
> pqr_mean = mean(pqr)
> lmn_mean = mean(lmn)
> 
> print(abc_mean)
[1]
-1.5

--- 

> print(pqr_mean)
[1] 7.3
> print(xyz_mean)
[1] 3.8
> print(lmn_mean)
[1] 10.5

```{r} 
abc_sd = sd(abc)
xyz_sd = sd(xyz)
pqr_sd = sd(pqr)
lmn_sd = sd(lmn)
```
> print(abc_sd)
[1] 11.00757
> print(pqr_sd)
[1] 12.62317
> print(xyz_sd)
[1] 10.46476
> print(lmn_sd)
[1] 5.190804 


---

ii
)
> r_abc_xyz = cor(abc,xyz)
> r_abc_pqr = cor(abc,pqr)
> r_abc_lmn = cor(abc,lmn)
> r_xyz_pqr = cor(xyz,pqr)
> r_xyz_lmn = cor(xyz,lmn)
> r_pqr_lmn = cor(pqr,lmn)
> 
> print(r_abc_xyz)
[1]
-0.4456345
> print(r_abc_pqr)
[1]
-0.2842739
> print(r_abc_lmn)
[1] 0.2634939
> print(r_xyz_pqr)
[1] 0.2923746
> print(r_xyz_lmn)
[1] 0.2577296
> print(r_pqr_lmn)
[1]
-0.08393819
[3]

---

iii)
> cor.test(abc,xyz,method="pearson",alternative="two.sided",conf.level = 0.95)
Pearson's product-moment correlation
data: abc and xyz
t = -1.408, df = 8, p-value = 0.1968
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
-0.8396649 0.2557513
sample estimates:
 cor
-0.4456345
> cor.test(pqr,lmn,method="pearson",alternative="two.sided",conf.level = 0.95)
Pearson's product-moment correlation
data: pqr and lmn
t = -0.23825, df = 8, p-value = 0.8177
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
-0.6777458 0.5761368
sample estimates:
 cor
-0.08393819

---

Confidence interval for correlation coefficient between returns of ABC Oil and XYZ Airways is 
(-0.84,0.26). Since it does not contain -1, we can say that there is no possibility of a perfect 
negative correlation at 5% level of significance. 
Confidence interval for correlation coefficient between returns of PQR Realty and LMN Bank 
is (-0.68,0.58). Since it does not contain -1, we can say that there is no possibility of a perfect 
negative correlation at 5% level of significance. 

---

iv)
> pABC = 0.50
> pXYZ = 0.50
> pPQR = 0.75
> pLMN = 0.25 
> 
> strat_A = pABC * abc + pXYZ * xyz
> print(mean(strat_A))
[1] 1.15
> 
> strat_B = pPQR * pqr + pLMN * lmn
> print(mean(strat_B))
[1] 8.1
Since, strategy B gives higher mean returns, it would be wise to go with strategy B, if mean 
returns is the metric to be targeted. 
(1.5)
(0.5)

---

v) 

H0: Variance of Returns from Strategy A = Variance of Returns from Strategy B
H1: Variance of Returns from Strategy A ≠ Variance of Returns from Strategy B

---

```{r}
var.test(x=strat_A,y=strat_B,conf.level = 0.90)
```

F test to compare two variances
data: strat_A and strat_B
F = 0.35856, num df = 9, denom df = 9, p-value = 0.1426
alternative hypothesis: true ratio of variances is not equal to 1
90 percent confidence interval:
0.112795 1.139835
sample estimates:
ratio of variances
 0.3585634
 
---

#### conclusion

Since the p-value 0.1426 > 10%, we have sufficient evidence to accept the null hypothesis at 
the 10% level of significance. Hence, the investor’s assumption of the two strategies being 
equally risky seems reasonable. 
(2)
(1)
[4]
vi)
> sd_A1 = (pABC^2 * abc_sd^2 + 2 * pABC * pXYZ * abc_sd * xyz_sd * -0.83 + pXYZ^2 * xyz_s
d^2)^(1/2)
> sd_A2 = (pABC^2 * abc_sd^2 + 2 * pABC * pXYZ * abc_sd * xyz_sd * 0.26 + pXYZ^2 * xyz_sd
^2)^(1/2)
> 
> print(sd_A1)
[1] 3.140851
> print(sd_A2)
[1] 8.523165
> sd_B1 = (pPQR^2 * pqr_sd^2 + 2 * pPQR * pLMN * pqr_sd * lmn_sd * -0.68 + pLMN^2 * lm
n_sd^2)^(1/2)
> sd_B2 = (pPQR^2 * pqr_sd^2 + 2 * pPQR * pLMN * pqr_sd * lmn_sd * 0.58 + pLMN^2 * lmn
_sd^2)^(1/2)
> 
> print(sd_B1)
[1] 8.637509
> print(sd_B2)
[1] 10.27457
The limits for standard deviation of Strategy A are (3.14, 8,52) and the limits for standard 
deviation of Strategy B are (8.64, 10.27). 
Based on Metric 2, since Strategy A has lower range of standard deviation, Strategy A would 
be selected.
[Max 4]
vii) We develop limits for the risk-adjusted returns under Strategies A and B using the following 
code:
> RAR_A1 = mean(strat_A)/sd_A1
> RAR_A2 = mean(strat_A)/sd_A2
> RAR_B1 = mean(strat_B)/sd_B1
> RAR_B2 = mean(strat_B)/sd_B2
> 
> print(RAR_A1)
[1] 0.3661428
IAI CS1B-1222
 Page 7 of 13
> print(RAR_A2)
[1] 0.1349264
> print(RAR_B1)
[1] 0.9377704
> print(RAR_B2)
[1] 0.788354
Based on the above the limits for Risk-Adjusted Return for Strategy A are (0.13,0.37) and limits 
for Strategy B are (0.79,0.94). Since Strategy B gives higher risk-adjusted returns, we should go 
for Strategy B using Metric 3. 
Strategy B gives higher returns and has higher risk, but it eventually ends up giving higher risk 
adjusted returns (i.e. more returns per unit of risk) as compared to Strategy A.
(1)
(1)
[2]
