---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```




* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 4

---

Q. 4) Government of “Actuaria” is operating a popular jackpot lottery called “***Power Ball***”. 

The 
game consists of two draws as given below 

* Draw 1 = Drawing 5 White Balls from a set of 69 White Balls.
* Draw 2 = Drawing 1 Red Ball (Power Ball) from a set of 26 Red Balls (Power Balls).

After drawing 5 white balls and the power ball, a ticket is issued for INR 0.50. 

A specimen ticket is given below –

<pre><code>
GOVERNMENT OF ACTUARIA
“POWER BALL LOTTERY”
Draw 1 – White Balls
05 18 29 37 61
Draw 2 – Power Ball
21
Ticket Price - INR 0.50
</code></pre>


---

The Government of Actuaria publishes its draw of 5 White Balls and the Power Ball. If all 
the 5 White Balls and the Power Ball drawn by the ticketholder exactly match with the balls 
drawn by the Government, then the ticketholder is entitled to a jackpot. The amount of 
jackpot is INR 2 crore. 
* Kindly note that for white balls, the order is not considered relevant 
while matching.



There are other prizes available depending on the white/power balls matched 
(notwithstanding the order for white balls). The prize money for various matches can be 
found in the following table as per the official website of Government of Actuaria: -

---

Case White Balls Matched Power Ball Matched Prize Money in INR
A 5 1 Jackpot
B 5 0 10,00,000
C 4 1 50,000
D 4 0 100
E 3 1 100
F 3 0 7
G 2 1 7
H 1 1 4
I 0 1 4

---

* Let X denote number of matching draws for White Balls and Y denote number of matching 
draws for Power Ball. 
* Under the given circumstances, hyper-geometric distribution is considered suitable to 
calculate probabilities.
* In R, probability of matching white balls / power ball using hyper-geometric distribution is 
determined using the dhyper function.
* For instance, under Case E, probability of matching 3 white balls is determined using the 
following code:

```{r}
dhyper(3,m=5,n=64,k=5)
```

Here the arguments of dhyper function can be explained as below –

* x denotes the value of white balls to be matched.
* m denotes the maximum number of successes.
* n denotes the maximum number of failures.
* k denotes the total number of draws.

---

### Exercises

i)
a) Create two separate vectors w_draw and r_draw for the white balls and power ball 
to be matched under cases A to I.
b) State the probability mass function of hyper-geometric distribution in terms of the 
arguments x, m, n and k as defined earlier.
c) Using the vectors created in part (i)(a) and the dhyper function as illustrated earlier, 
create two separate result vectors prob_w_draw and prob_r_draw showing the 
probabilities of matching white balls and power ball under Cases A to I. 
Probability of matching white balls is independent of the probability of matching the power 
ball.

---

d) Calculate the joint probabilities of matching white balls and power balls using result 
vectors prob_w_draw and prob_r_draw created in part (i)(c).
e) Create a vector prize showing the prize money under cases A to I. 

---

### Exercises

f) Determine the total expected prize money pay-out using the joint probabilities 
determined in part (i)(d) and the prize money vector created in part (i)(e).

g) Assuming that there are no other costs to the Government in operating the lottery 
expect the prize money pay-out, calculate the profits, if any, implicit in the ticket 
price. 

ii) If instead of using the hyper-geometric distribution, we decided to use the binomial 
distribution for modelling the probabilities of matching white balls and power ball.
a) State the probability mass function of binomial distribution in terms of arguments x, 
m, n and k as defined earlier. 

---

### Exercises

b) Re-determine the joint probabilities using the dbinom function and arrive at the total 
expected prize money pay-out. (4)
iii) Point out the difference, if any, in the outcomes under part (i) and (ii).
iv) Determine the mean and variance of both the binomial distribution and hyper-geometric distribution for random variable X (matching white balls) and comment on 
the differences in the two.
v) Which distribution is more suitable to model these matching probabilities – hyper-geometric or binomial? Substantiate with reasons. 

---

### Part A

Solution 4:
i) a) We define two data sets <tt>w_draw<ttand <tt>r_draw</ttcorresponding to the white balls and red balls 
to be matched for cases A to I. 
R Code and Output:

```{r}
w_draw <- c(5,5,4,4,3,3,2,1,0) # white balls to be matched corresponding to cases A to I
r_draw <- c(1,0,1,0,1,0,1,1,1) # red balls to be matched corresponding to cases A to I
```
---

### Part B

b) Formula for the probability mass function of Hyper-Geometric Distribution in terms of the 
arguments specified in the question is given below:
$$P(X=x) = mCx * nC(k-x) / (m+n)Ck$$
Where –

x = 0, 1, 2, 3, ………………
0 < p < 1 
(p = m/(m+n)) 

---

### Part C

c) Further we define variables prob_w_draw and prob_r_draw to determine the probabilities 
under hypergeometric distribution where w_draw and r_draw will be used as input, m will be 
the successes (5 and 1), n will be the failures (64 and 25) and k will be the number of balls 
drawn i.e. sample size (5 and 1)

---

R Code and Output:

```{r}
prob_w_draw <- dhyper(w_draw,m=5,n=64,k=5)
print(prob_w_draw)
```
8.897974e-08 8.897974e-08 2.847352e-05 2.847352e-05 1.793832e-03
[6] 1.793832e-03 3.707252e-02 2.826780e-01 6.784271e-01

```{r}
prob_r_draw <- dhyper(r_draw,m=1,n=25,k=1)
print(prob_r_draw)
```


---

### Part D

d) Finally, we define prob_draw to be the multiplication of prob_w_draw and prob_r_draw 
assuming the draws are independent of each other.
R Code and Output:

```{r}
prob_draw = prob_w_draw * prob_r_draw # multiplication of probabilities
print(prob_draw)
```


---

e) We define another data set prize which defines the prize amounts for Cases A to I.
R Code and Output:

```{r}
prize <- c(20000000,1000000,50000,100,100,7,7,4,4)
print(prize)
```

 
---

### Part F

We then define amt by multiplying prob_draw with prize to arrive at the expected amount one 
can win from the lottery / jackpot.

<tt>R Code and Output:</tt>
```{r}
amt = prize * prob_draw
print(amt)
```

 0.068445956 0.085557445 0.054756765 0.002737838 0.006899352 0.012073867
[7] 0.009981063 0.043488918 0.104373403
print(sum(amt))
 0.3883146 
 
---

g) Since, the expected prize amount is INR 0.39 and the prize of the lottery ticket is INR 0.50, 
there is a profit of INR 0.11 implicit in the ticket price. 

------

### Part G

ii)a) Formula for the probability mass function of Binomial Distribution in terms of the arguments 
specified in the question is given below:

$$P(X=x) = kCx * (p)x * (1-p)(k-x)$$
Where –
* x = 0, 1, 2, 3, ………………
* 0 < p < 1 
* (p = m/(m+n)) 


---

### Part H

 b) The joint probabilities and the expected prize money pay-out is re-determined using binomial 
distribution for determining matching probabilities:
R Code and Output:
```{r}
prob_w_draw_bin <- dbinom(w_draw,5,5/69,log=FALSE)
print(prob_w_draw_bin)

```
 1.998042e-06 1.998042e-06 1.278747e-04 1.278747e-04 3.273592e-03
[6] 3.273592e-03 4.190197e-02 2.681726e-01 6.865219e-01

```{r}
prob_r_draw_bin <- dbinom(r_draw,1,1/26,log=FALSE)

print(prob_r_draw_bin)
```


---


```{r}
prob_draw_bin = prob_w_draw_bin * prob_r_draw_bin
print(prob_draw_bin)
```

 7.684776e-08 1.921194e-06 4.918257e-06 1.229564e-04 1.259074e-04
[6] 3.147684e-03 1.611614e-03 1.031433e-02 2.640469e-02
amt_bin = prize * prob_draw_bin

```{r}
print(amt_bin)
```
 1.53695522 1.92119403 0.24591284 0.01229564 0.01259074 0.02203379
[7] 0.01128130 0.04125733 0.10561876
print(sum(amt_bin))
 3.90914

---

### Part I

If we compare binomial probabilities with hyper-geometric probabilities:

1. For the power ball, the probabilities for all cases are same under both binomial and 
hyper-geometric distribution.

2. For white balls, the probabilities under binomial distribution tend to be higher than 
those under hyper-geometric distribution. 

3. Due to this the expected pay-out determined using binomial distribution tends to be 
on the higher side as compared to one determined using hyper-geometric distribution.


---

iv) We use the formulae from the tables to determine the mean and variance for X using both 
binomial and hyper-geometric distributions.
R Code and Output:

```{r}
mean_x_hyper = 5 * 5/69
mean_x_binom = 5 * 5/69
print(mean_x_hyper)
```

 0.3623188
print(mean_x_binom)

 0.3623188

```{r}
var_x_hyper = 5*5*(69-5)*(69-5)/((69-1)*69^2)
print(var_x_hyper)
```

```{r}
var_x_binom = 5 * 5/69 * (1-5/69)
print(var_x_binom)
```

 0.3360639

* Mean is the same in both cases.
* However, variance under hyper-geometric distribution is lower as compared to the variance 
under the binomial distribution. 
* This is because hyper-geometric distribution is a without 
replacement alternative of the binomial distribution. 
* Since, after a trial, that observation is not 
replaced, the variability in the results is reduced.


---

v) 
* In reality, for lotteries, draws are done without replacement and hence hyper-geometric 
distribution would be more suitable in modelling matching probabilities.

* However, as the size of the population (m+n) goes on increasing, binomial distribution 
provides a good approximation for hyper-geometric probabilities.


---

BLANK 


---