---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```



* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3

---

### Health Cover

A company wants to launch a new product PA to provide special health cover. 
* Under this health cover, the company will pay customer INR 5000 for each claim. 
* They have asked an Actuary to derive price for the product for different segment of customers.
* The Actuary has collected a sample data, named “***PA_Data***”, for 20 customers showing 
number of claims (Claim) made by each customer along with the Gender, Age and Health
condition (Diabetic or not) of the customer.
* Actuary assumes that the response variable – number of claims (Claim) follows Poisson 
distribution. 

---

### Exercises

A.  Load the “<tt>PA_Data</tt>” datset and fit the Poisson Generalised Linear Model (GLM) to 
the dataset, specified by the following parameter of R code for GLM: 
<tt>Claim ~ Gender*Health + Age</tt>
 Include summary of the fitted model and label the model as Model1. 

B.  Write down the linear predictor for the model given in part (i) (i.e. Model1). Explain 
all the terms clearly.

C.  Comment on the significance of the parameters of the fitted model in part (i) (i.e. 
Model1). Also state the scaled deviance and compute Log Likelihood of the model. 

---

### Part A 
i)
ii)
#i.
```{r}
PA <- read.csv("PA_Data.csv")
model1 <- glm(Claim~Gender*Health+Age,family=poisson(lin="log"),data=PA)
```

---


```{r}
summary(model1)
```
--- 

* linear predictor for Model 1 is
*  a+b1X1+b2X2+b3X3+b4X1X2 where 

$$ Alt: 0.17452 + 0.06235\;X1 -1.31680\;X2 + 0.02248\;X3 -0.10401\;X1X2$$
* X1 = 0 for Female Gender and 1 for Male Gender
* X2 = 0 for Diabetic and 1 for Non Diabetic
* X3 is Age
* X1X2 indicates interaction term between Gender and Health Condition

---

### Exercise 3



* Gender is not significant
* Health Condition is significant
* Age is not significant
* Interaction term between Gender and Health condition is not significant

* Scaled deviance = 13.179
* AIC = - 2LogL(Model) + 2*Parameters

$$LogL(Model) = Parameters - AIC/2 $$


---

```{r}
L<- 4- model1$aic/2
L

```
Log Likelihood of Model1 is -27.07302

---

### Exercise 3

```{r}
model2<-glm(Claim~Health+Age,family=poisson(lin="log"),data=PA)

model2$aic < model1$aic

# Model 2 AIC is lower than Model 1 showing Model2 outperforms Model1
```

---

### Exercise 5

```{r}
summary(model2)
```



* Age is not significant and thus can be dropped to improve the model
* Give full marks in case reached to same conclusion using alternate methods.

---

### Exercise 4

Fit another model (labelled as <tt>Model2</tt>) specified by following R code:

<tt> Claim~Health+Age</tt>

Demonstrate that <tt>Model2</tt> outperforms the model defined in **exercise 1**.
 
---

### Exercise 5

* Examine the significance of <tt>Model2</tt> and fit an improved model. 

* Label this model as <tt>Model3</tt>. Also justify why <tt>Model3</tt> is an improvement. 

---


### Exercise 6

* Compute the price for the product for below customers using Model2.
* Note that Price is equal to expected aggregate claim amount per customer. 
* <i>(In other 
words, **expected claim amount x expected number of claims**) </i>

|Customer |Age| Health |Gender|
|:----------:|:--------:|:-----:|:--------:|
|Student1 |30 |Diabetic |Male|
|Actuary2| 50 |Non-Diabetic |Female|

* Which customer price will be higher under the models.
* Provide the reason for the same. 

---


### Exercises

Company is planning modification to the product to make it affordable and attractive. 

Under 
this modification, INR 6000 will be paid for 1st claim and then INR 4000 for 2nd claim 
onwards. 

<i>
Hint: Under this product, essentially INR 4000 amount is paid for each claim plus additional 
INR 2000 amount if there are claim(s). 
</i>



--- 

#### Part 6

vi)
```{r}
model3<-glm(Claim~Health,family=poisson(lin="log"),data=PA)

summary(model3)
```

---

Call:
glm(formula = Claim ~ Health, family = poisson(lin = "log"), 
 data = PA)
Deviance Residuals: 
 Min 1Q Median 3Q Max 
-1.22474 -0.81754 -0.07119 0.27453 1.44149 
Coefficients:
 Estimate Std. Error z value Pr(>|z|) 
(Intercept) 1.1394 0.2000 5.697 1.22e-08 ***
HealthNonDiabetic -1.4271 0.3887 -3.671 0.000241 ***
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
(Dispersion parameter for poisson family taken to be 1)
 Null deviance: 30.147 on 19 degrees of freedom
Residual deviance: 14.436 on 18 degrees of freedom
AIC: 57.403
Number of Fisher Scoring iterations: 5

---


```{r}
model3$aic < model2$aic
```


* Under Model 3. AIC reduced from 58.17 to 57.40


---

#vi.

```{r}
Student1<- data.frame(Gender ="M",Health ="Diabetic",Age=30)
Actuary2<- data.frame(Gender ="F",Health ="NonDiabetic",Age=50)
```

```{r}
# Price = 5000 * expected number of claims
# Price of student1 is 12240.06 for Model 2 and 15625 for Model 3 
5000*predict.glm(model2,newdata = Student1,type= "response")

```

 1 
12240.06 
5000*predict.glm(model3,newdata = Student1,type= "response")
 1 
15625 


*  Price of Actuary2 is 4797.457 for Model 2 and 3750 for Model 3 

```{r}
5000*predict.glm(model2,newdata = Actuary2,type= "response")
```

```{r}
5000*predict.glm(model3,newdata = Actuary2,type= "response")
```



---


* Under both models, Student1 price is coming higher.
* Since student is diabetic. Under both models, health condition is significant and for non-diabetic, parameter value is negative implying lower claims for non-diabetic

---



### Exercise 7
vii)
a) Compute the price under this modification for Student1 and Actuary2 (as per details 
given in **exercise 6** for Model3.
b) Compare and explain why price for Student1 is lower for modified product whereas 
for Actuary2 is higher.


---

vii) a)
 b)
#vii. a.

```{r}
Student1_mean=predict.glm(model3,newdata = Student1,type= "response")
Actuary2_mean=predict.glm(model3,newdata = Actuary2,type= "response")

#price in case of modified product is
# 4000* expected number of claims + 2000 * probability of 1 or more claim.

```


```{r}
#price of Student1:

#compute probability of 0 
sp0<-dpois(0,Student1_mean)

4000*Student1_mean + 2000*(1-sp0)
```
14412.13 

```{r}
#price for Actuary2
ap0<-dpois(0,Actuary2_mean)
4000*Actuary2_mean + 2000*(1-ap0)
```

4055.267 

---

#vii. b.
Student1_mean
 1 
3.125 

* For student1, expected claims are 3.125. 
* Since, more claims are expected for Student1, reduction of payment lead to more saving than extra payment for 1st claim and thus, less price


---

Actuary2_mean
 1 
0.75 

* Whereas for Actuary2, expected claims are 0.75 close to 1.
* thus, more payment expected for Actuary2 resulting in higher price.

---

BLANK
---