---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

2 A life insurance company has issued a 1-year group insurance policy covering 10,000
employees categorized into four age bands of 2,500 employees each.
The following amounts are paid upon each death under the policy:
* Death from pandemic diseases mentioned in the policy document: £80,000.
* Death from cause other than above: £40,000.
The premium charged for each group was estimated as the mean of a compound
binomial distribution.
The ‘CS2B_A23_Qu_2_Data.csv’ file contains the following data for each group:
group: name of the group
age_band: age range in the group
Mort: expected probability of death
p_dth: proportion of deaths due to pandemic diseases
Before answering this question, the ‘CS2B_A23_Qu_2_Data.csv’ file should be
loaded into R and assigned to a data frame called ‘Prem’.
(i) Calculate the premium that company would have charged under each of the
groups, ignoring expenses and profit margin, assigning your output to a vector
called ‘Prem_charged’ and displaying ‘Prem_charged’ in your answer script.

The head of the department wishes to project the actual claims over the next year
using random numbers.
(ii) Generate a sample of size 2,500 under each of the above groups from a
Bernoulli distribution, simulating whether death occurs for each of the
employees using the expected probability of death from the table ‘Prem’. You
should output your results to a data frame called ‘Tab_R’ consisting of four
column vectors ‘G1_R’to ‘G4_R’. The random number generator seed to be
used is 123.
Use the summary() function to display your results. 
(iii) Generate a data frame ‘Tab_U’ of observations from the Uniform(0,1)
distribution, consisting of four column vectors, ‘G1_U’ to ‘G4_U’, with 2,500
components each. The random number generator seed to be used is 300. Use
the R function head(), to display the first six rows of ‘Tab_U’. 
---
(iv) Generate a data frame ‘Tab_V’, consisting of four column vectors ‘G1_V’ to
‘G4_V’, with 2,500 components each, representing claim amounts under each
group.
If death has occurred, the relevant entry of ‘Tab_V’ should be set to 80,000 if
the random variable generated in part (iii) is less than or equal to the
proportion of pandemic deaths specified in the table ‘Prem’ for the relevant
group, or 40,000 otherwise.
Use the summary() function to display your results. [8]
(v) Calculate the aggregate claim amount under each group, displaying your
output as a vector. 
(vi) Calculate the projected profit or loss under each group and the policy as a
whole. 
One of the employees suggests that the premium rate charged should be based on the
actual claims obtained from part (v) above.
(vii) Comment on the validity of the employee’s suggestion, assuming that the
distribution used, the parameters used and the methodology adopted are
reasonable. 
[Total 33]
CS2B A2023–6


Q2
(i)
Prem<-read.table("CS2B_A23_Qu_2_Data.csv", TRUE, ",") 
Prem_charged= 
2500* 
(Prem$Mort*Prem$p_dth*80000 
+Prem$Mort*(1-Prem$p_dth)*40000) 
Prem_charged 
 187500 260000 640000 080000 
(ii)
set.seed(123) 
Tab_R<- data.frame(cbind(G1_R=rbinom(2500,1,Prem$Mort),G2_ R=rbinom(2500,1,Prem$Mort[2]),G3_R=rbinom(2500,1,Prem$ Mort[3]),G4_R=rbinom(2500,1,Prem$Mort[4]))) [4½]
summary(Tab_R) 
Or
set.seed(123) 
G1_R=rbinom(2500,1,Prem$Mort) 
G2_R=rbinom(2500,1,Prem$Mort[2]) 
G3_R=rbinom(2500,1,Prem$Mort[3]) 
G4_R=rbinom(2500,1,Prem$Mort[4]) 
Tab_R<- data.frame(cbind(G1_R,G2_R,G3_R,G4_R)) 
summary(Tab_R) 
Or
set.seed(123) 
B = matrix(nrow = 2500, ncol = 4) 
for (i in 1:4) { 
q = Prem$Mort[i] 
B[ ,i] = rbinom(2500,1, q ) 
}
Tab_R=data.frame (G1_R=B[,1] , G2_R=B[,2], G3_R =B[,3], G4_R=B[,4] ) 
summary(Tab_R) 
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2023 - Examiners’ report
CS2B A2023 © Institute and Faculty of Actuaries
G1_R G2_R G3_R G4_R
Min. :0.0000 Min. :0.000 Min. :0.0000 Min. :0.0000
1st Qu.:0.0000 1st Qu.:0.000 1st Qu.:0.0000 1st Qu.:0.0000
Median :0.0000 Median :0.000 Median :0.0000 Median :0.0000
Mean :0.0016 Mean :0.002 Mean :0.0028 Mean :0.0068
3rd Qu.:0.0000 3rd Qu.:0.000 3rd Qu.:0.0000 3rd Qu.:0.0000
Max. :1.0000 Max. :1.000 Max. :1.0000 Max. :1.0000 
(The function “rbern” is not a part of the base R installation. However, if the candidate installs “Rlab” package and uses “rbern” function, full marks are to be awarded assuming no other errors)

```{r}
install.packages("Rlab")
library("Rlab")
set.seed(123) 
Tab_R<- data.frame(cbind(G1_R=rbern(2500,Prem$Mort),G2_R=rbern(2500,Prem $Mort[2]),G3_R=rbern(2500,Prem$Mort[3]),G4_R=rbern(2500,Prem$Mort[4])) [4½]
summary(Tab_R) 
```
G1_ R G2_ R G3_ R G4_R
Min. :0.0000 Min. :0.000 Min. :0.0000 Min. :0.0000
1st Qu.:0.0000 1st Qu.:0.000 1st Qu.:0.0000 1st Qu.:0.0000
Median :0.0000 Median :0.000 Median :0.0000 Median :0.0000
Mean :0.0016 Mean :0.002 Mean :0.0028 Mean :0.0068
3rd Qu.:0.0000 3rd Qu.:0.000 3rd Qu.:0.0000 3rd Qu.:0.0000
Max. :1.0000 Max. :1.000 Max. :1.0000 Max. :1.0000 


--- 
(ii)
set.seed(300) 
U = matrix(runif(10000), ncol = 4) [2]
Tab_U=data.frame (G1_U=U[,1] , G2_U=U[,2], G3_U =U[,3], G4_U=U[,4] ) 
head(Tab_U) 
Or
set.seed(300) 
U = matrix(nrow = 2500, ncol = 4) 
for (j in 1:4) { 
U[ ,j] = runif(2500) 
}
Tab_U=data.frame (G1_U=U[,1] , G2_U=U[,2], G3_U =U[,3], G4_U=U[,4] ) 
head(Tab_U) 
G1_U G2_U G3_U G4_U
1 0.9152467 0.9633111 0.4436952 0.4876556
2 0.7633293 0.5037355 0.3548937 0.9567698
3 0.8056856 0.8878896 0.5736878 0.3526416
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2023 - Examiners’ report
CS2B A2023 © Institute and Faculty of Actuaries
4 0.7337780 0.3721702 0.5864507 0.8759781
5 0.6820679 0.6179086 0.8117664 0.5643849
6 0.0120303 0.3733971 0.4308092 0.2999071 
(iv)
M = matrix(nrow = 2500, ncol = 4)
dth1 =80000 
dth2 =40000 
for (i in 1:4) { 
M[,i] = 
Tab_R[,i] * 
ifelse( 
Tab_U[,i]<= Prem$p_dth[i], 
dth1, 
dth2) 
}
Tab_V=data.frame (G1_V=M[,1] , G2_V=M[,2], G3_V =M[,3], G4_V=M[,4] ) 
summary(Tab_V) 
G1_V G2_V G3_V G4_V
Min. : 0 Min. : 0 Min. : 0 Min. : 0
1st Qu.: 0 1st Qu.: 0 1st Qu.: 0 1st Qu.: 0
Median : 0 Median : 0 Median : 0 Median : 0
Mean : 96 Mean : 96 Mean : 192 Mean : 464
3rd Qu.: 0 3rd Qu.: 0 3rd Qu.: 0 3rd Qu.: 0Max. :80000Max. :80000 Max. :80000 Max. :80000 
(v)
Aggregate<- c(sum(Tab_V$G1_V) , sum(Tab_V$G2_V), sum(Tab_V$G3_V), sum(Tab_V$G4_V) ) 
Aggregate 
1,] 240000 240000 480000 1160000 
(vi)
Prem_charged-Aggregate 
 -52500 20000 160000-80000 
sum(Prem_charged-Aggregate) 
 47500 
(vii)
The employee‘s suggestion is not valid 
as the aggregate claim amount in part (v) under each group represents one sample 
The above process would have to be performed many times, and averages taken, to
obtain reasonable premium rates 
[Total 33]
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2023 - Examiners’ report
CS2B A2023 © Institute and Faculty of Actuaries
This question was well answered, and the average mark was considerably higher than that for the other two questions on the paper.
In part (i) some candidates forgot to multiply by 2500 policies to produce total premiums and instead calculated premiums per policy.
In part (ii) use of rbern rather than rbinom gained full marks if applied correctly. Candidates are reminded to include the code for setting the seed value in these simulations questions. The examiners understand that where a candidate has the seed but needs to run their code more than once to get a correct answer then the numerical values produced will be different to those above. Candidates are not penalised for this.
In part (vii) marks were awarded for a range of sensible arguments in support of the correct conclusion.
