---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

1 Before answering this question, the ‘markovchain’ R package should be loaded
into R using the following code:
install.packages("markovchain")
library(markovchain)
A three-state Markov chain model consisting of healthy (‘H’), sick (‘S’), and dead
(‘D’) states has the following weekly transition probabilities:
pHH = 0.98
pHS = 0.019
pHD = 0.001
pSS = 0.75
pSH = 0.2
pSD = 0.05
(i) Construct a ‘markovchain’ object for the transition matrix for the above
Markov chain model, calling it ‘transitions_markov’. 
(ii) Calculate the probability that, given a life is healthy now, the life will be sick
at time 2 weeks. 
(iii) Calculate the probability that, given a life is healthy now, the life will be sick
at some point in the next 52 weeks. 
(iv) Calculate the probability that, given a life is healthy now, the life will remain
healthy for the entire year. 

---

A new, highly contagious disease has become prevalent in the country and it is
proposed that the HSD model above be amended to model the effects of this disease.
Lives become ill for a period of time and then either recover or die. Once recovered, a
life is deemed immune and cannot become ill again.
It has been decided to use a Markov jump model instead of a chain model. The
Markov model consists of four states: healthy (‘H’), sick (‘S’), recovered (‘R’) and a
dead (‘D’) state. The following daily transition rates have been estimated from a
recent investigation:
μSD = 4%
μSR = 13%

---

The rate from healthy to infected, μHS, is equal to b multiplied by it where b is a
constant and it is the proportion of lives in the sick state at time t. All other transition
rates are zero.
(v) Explain why it may be preferable to set μHS = b it (as above), rather than using
a constant value (as has been done for the other rates). 
(vi) Give an example of a scenario that would be expected to result in a
particularly high value of b. 


The proposed value of b is 0.3, and it is assumed that at time 0, 1% of the population
are in the sick state and the remaining 99% are in the healthy state.
(vii) Using this revised model and a step length of 0.01 days, calculate the
occupancy probabilities in each of the four states from t = 0.01 to t = 100 days
inclusive. You should output your answers for each value of t to successive
rows of a matrix called ‘mat_ans’. 
(viii) Plot a graph showing the probabilities in part (vii) with suitably labelled axes
and making clear which plot components correspond to which state. 
(ix) Calculate the probability that a life healthy at time 0 is sick after:
(a) 5 days.
(b) 30 days.

(x) Calculate the expected present value of a daily rate of £1 payable while sick
using a force of interest of 6% p.a. 

---


Q1 (i) 

```{r}
transitions <- matrix(  c(0.98, 0.019,0.001,
                           0.2, 0.75, 0.05,
						   0,  0, 1),
   byrow=TRUE,nrow=3,
   dimnames=list (c("healthy","sick", "dead"), 
                  c("healthy","sick","dead"))) 
(transitions_markov <- new("markovchain", transition
Matrix=transitions)) 

```

A 3-dimensional discrete Markov Chain defined by the following states:
healthy, sick, dead
The transition matrix (by rows) is defined as follows:
healthy sick dead
healthy 0.98 0.019 0.001
sick 0.20 0.750 0.050
dead 0.00 0.000 1.000

---

(ii) c(1,0,0)*transitions_markov^2 
healthy sick dead
[1,] 0.9642 0.03287 0.00293 
The probability that the life will be sick at t=2 weeks is 3.287% 

---

(iii)
transitions_sick<-matrix(c(0.98, 0.019,0.001,0,1, 0,0,0,1),byrow=TRUE,nrow=3, dimnames=list(c("healthy","sick","dead"),c("healthy",
"sick","dead"))) 
transitions_sickmarkov <- new("markovchain", transitionMatrix=transitions_sick) 
c(1,0,0)*transitions_sickmarkov^52 
healthy sick dead
[1,] 0.3497486 0.6177389 0.03251257 
the probability that the life will be sick at some point in the next 52 weeks is
61.77% 

--- 

(iv)

```{r}
0.98^52
The probability a healthy life will have remained healthy for the entire year is
34.97% 
```

(v)
It is reasonable to assume that the rate at which healthy lives become ill is
proportional to the number of lives which are infectious.

--- 

or other similar reasonable comment about the nature of this transition 
(vi) b reflects measures imposed for transmissivity e.g. masks, social distancing, lockdowns. A high b weak measures i.e. no masks, no social distancing [2]
(vii) calculation of occupancy probabilities
#initial populations
s<-.01
h<-.99
r<-0
dead<-0 
b<-.3 
#par value
#daily rates
k<-.13 
j<-.04 
#just set up 4x4
days<-100 
dt1=0.01 
#in days
no_steps<-days/dt1 
#100 days
mat_ans<-matrix(0,no_steps,4) 
for (qq in 1:no_steps) { 
mu12=b*s
#mu12=b*s, or try constant rate
mu13=0
mu14=0 
mu21=0
mu23=k #recovery rate
mu24=j #death rate 
mu31=0.0
mu32=0.0
mu34<-0 
mu43<-0
mu41=0
mu42=0 
mu11=-mu12-mu13-mu14
mu22=-mu21-mu23-mu24

mu33=-mu32-mu31-mu34
mu44=-mu42-mu41-mu43 
t_rates=c(mu11,mu12,mu13,mu14,mu21,mu22,mu23,mu24,mu31, mu32,mu33,mu34,mu41,mu42,mu43,mu44) 
gen_matrix=matrix(t_rates,4,4,byrow=TRUE) 
Po=c(h,s,r,dead) 
#starting proportions at each loop
P_dt =diag(4) + gen_matrix*dt1 
#gets prob matrix each loop
occ_probs=Po%*%P_dt 
#occupancy probs end of each period given starting states
h<-occ_probs
s<-occ_probs[2]
r<-occ_probs[3]
dead<-occ_probs[4] 
mat_ans[qq,]<-occ_probs 
#posts answers in matrix for plotting

---

(viii)
#plots
x = seq(from = 1/100, to = 100, by = 1/100) 
plot(x,mat_ans[,1],ylim=c(0,max(mat_ans)),col=1,lwd=3, lty=1,xlab="",ylab="",type="l",main="HSRD model") 
lines(x,mat_ans[,2],lwd=1,lty=1) 
#sick
lines(x,mat_ans[,3],lwd=1,lty=2) 
#recovered
lines(x,mat_ans[,4],lwd=1,lty=3) 
# dead
legend(x="topright",legend = c("healthy","sick", "recovered","dead"),lwd = c(3,1,1,1),lty = c(1,1,2,3)) 

---

(ix)(a)
mat_ans[5/dt1,2] 
 0.01860525 
The probability that a life healthy at time 0 is sick after 5 days is 0.01860525 
(b) mat_ans[30/dt1,2] 
 0.1171159 
The probability that a life healthy at time 0 is sick after 30 days is 0.1171159 
(x)
yyy<-mat_ans[,2] 
sum(0.01*yyy*exp(-0.06*seq(1,10000)/36500)) [2]
 4.227194 
The required expected present value is £4.23 

---

BLANK
---

* This question on the application of Markov processes to survival analysis in R was not well answered. 
* Candidates need to be ready to use R to solve problems and simply learning R commands will not be sufficient. 
* Many candidates failed to see that a new matrix was required in part (iii).
Part (vii) was not well answered. 
* Alternative methods for generating the mat_ans matrix are possible here and can obtain full marks. 
* In general marks were be awarded in ½ mark increments for a wide variety of approaches in the following 5 stages to a solution for this part:
step 1 - parameterising the solution using figures given in the question (max 2 marks)
CS2B - Risk Modelling and Survival Analysis - Core Principles - April 2023 - Examiners’ report
CS2B A2023 © Institute and Faculty of Actuaries
step 2 - setting the time sequence and initiating the loop (max 2 marks)
step 3 - calculating the mu values (max 3 marks)
step 4 - converting the mu’s into probabilities (max 3 marks)
step 5 - outputting the matrix (1 mark)
Similarly for part (x) a wide variety of approaches to incorporating an interest rate element were given full credit.
