---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Q. 1) Consider a random variable X which represents the project completion time for a process
automation project for a FMCG company (in years). Three scenarios have been contemplated
for completion of the project which are as follows:

|Scenario |Probability Distribution|
|:--------------:|:----------------------|
|Scenario 1 | X has a beta distribution with parameters α =5, β=1;|
|Scenario 2 | X has a beta distribution with parameters α =1, β=5;|
|Scenario 3 | X has a beta distribution with parameters α =3, β=3.|


### Exercises

#### Part 1

Calculate the following probabilities for X under Scenario 1:

A. Use pbeta function to calculate P(0.2 < X < 0.8);

B. Use <tt>qbeta</ttto find x such that P(X x) = 0.65.


---

ii) Use the following code to calculate the coefficient of skewness of X under all the three
scenarios:
$$Skew = 2*((β-α)/(α+β+2))* sqrt((α+β+1)/(α*β)) $$


---



Histograms for these simulated samples have been given below:

#### Scenario 1
X1 ~ Beta (α =5, β=1)

#### Scenario 2
X2 ~ Beta (α =1, β=5)

#### Scenario 3
X3 ~ Beta (α =3, β=3)

---

Solution 1:
(i)
a) pbeta(0.8,5,1) - pbeta(0.2,5,1) (1)
[1] 0.32736 (1)
b) qbeta(0.65,5,1,lower= FALSE) or Alternate: qbeta(0.35,5,1
) 
 (1)
 [1] 0.8106131 (1)
 
---

(ii) 
a=c(5,1,3)
b=c(1,5,3) 
skew=2*((b-a)/(a+b+2))*sqrt((a+b+1)/(a*b)) (2)
skew
[1] -1.183216 1.183216 0.000000 (1)
 (3) 
Or Alternatively
skew1=2*((1-5)/(5+1+2))*sqrt((5+1+1)/5)
skew1
[1] -1.183216
skew2=2*((5-1)/(1+5+2))*sqrt((5+1+1)/5)
skew2
[1] 1.183216
skew3=2*((3-3)/(3+3+2))*sqrt((3+3+1)/9)
skew3
[1] 0

---

iii) Under all the three scenarios, write a code to simulate a sample of 12,000 values from a
Beta distribution.

* Use the command set.seed(421967) to initialize the random number
generator, before you start the simulation. 
* Simulate this sample by defining vectors “x1”,
“x2” and “x3” for Scenario1, Scenario2 and Scenario3 and write code to draw histogram
for each of the scenarios.

<i>You are NOT required to execute the code or print the result of these vectors or reproduce
the histograms.</i>

---

### Exercise 4

iv) Comment on the shape of the histograms plotted above by taking into consideration
coefficient of skewness calculated in part (ii). 

---



Histogram of 12000 values from Beta (5,1) Distribution
x1
Frequency
0.2 0.4 0.6 0.8 1.0
0 500 1500 2500
Histogram of 12000 values from Beta (1,5) Distribution
x2
Frequency
0.0 0.2 0.4 0.6 0.8
0 500 1500 2500
Histogram of 12000 values from Beta (3,3) Distribution
x3
Frequency
0.0 0.2 0.4 0.6 0.8 1.0
0 200 600 1000

Histograms of sample means for the three scenarios have been given below:
Scenario 1: x1_bar Scenario 2: x2_bar Scenario 3: x3_bar


---

 
(iii) 
```{r}
set.seed(421967) 
x1=rbeta(12000,5,1) 
hist(x1) 
```

---

```{r}
x2=rbeta(12000,1,5) 
hist(x2) 
```

---

```{r}
x3=rbeta(12000,3,3)
hist(x3) 
```
 
---


* As Alpha is greater than 1 and Beta is equal to 1, the histogram is heavily negatively skewed and strictly increasing as evident from the result obt
ained in (ii) above and from the shape of the graph.

* As Alpha is equal to 1 and Beta is greater than 1, the histogram is he
avily positively skewed and strictly decreasing as evident from the result obt
ained in (ii) above and from the shape of the graph.

* As both the parameters alpha and beta are equal, the graph is roughly 
symmetrical as evident from the graph and the value of the skewness obtained
in (ii) above.

---

### Exercise 5

Write a code to perform 1,200 repetitions of the simulations in ***Exercise 3*** for all the three
scenarios. 

* You should compute and store the value of the mean of the sample (<tt>x1_bar</tt>), (<tt>x2_bar</tt>) and (<tt>x3_bar</tt>) for each repetition. 

* Use the same command <tt>set.seed(421967)</ttto
initialize the random number generator, before you start the simulations.

<i>You are NOT required to execute the code or print the output of these simulations or
histogram of these simulations. </i>

---

### Exercise 5

(v) 

```{r}
set.seed(421967)
x1_bar <-replicate (1200, mean(rbeta (12000,5,1))) 
```


Or alternatively
```{r}
set.seed(421967)
x1_bar=rep(0,1200)
for(i in 1:1200){x1<-rbeta(12000,5,1);x1_bar[i]<-mean(x1)} 
```

---

set.seed(421967)
x2_bar <-replicate (1200, mean (rbeta (12000,1,5)))
Or alternatively
set.seed(421967)
x2_bar=rep(0,1200)
for(i in 1:1200){x2<-rbeta(12000,1,5);x2_bar[i]<-mean(x2)}

---

```{r}
set.seed(421967)

x3_bar <-replicate(n = 1200, 
                   mean(rbeta (12000,3,3))
                   )

```

---

Or alternatively

```{r}
set.seed(421967)

x3_bar=rep(0,1200)
for(i in 1:1200){
  x3<-rbeta(12000,1,5);
  x3_bar[i]<-mean(x3)
}

```
  
```{r}

#First 20 Observations

head(x3_bar,20)
```

---

### Exercise 6


Comment on the shape of the histograms, by referring to the central limit theorem. Also,
compare and contrast with your observations in ***Exercise 4***.

-------------------------

* The distribution of sample mean is roughly symmetrical in all the three ca
ses irrespective of the values of the shape parameters (alpha and beta).

* These 
shape parameters(alpha and beta) do not significantly affect the sample mean 
of large sample size, which is in line with the central limit theorem.

* Irrespective of the population distribution of the random variable from which the sample is selected, for a large sample size the distribution of the sample means is approximately normal. 

--- 
