---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Q.4)
The following data represents the average claim size (motor third party - accident injury) under motor insurance policies settled in six different states across a country.
State i Year j 2018-2019 2019-2020 2020-2021 2021-2022 2022-2023 Assam
219458
240371
289307
264439
279704 Bihar
216594
231311
261915
286211
291934 Chattisgarh
213871
231461
264519
261279
270058 Delhi
389197
400926
393130
391921
373129 Odisha
224879
243361
276718
292564
300135 Kerala
194814
230113
258101
276876
287973
Enter the data in R in the form of a matrix using the following code:
Claims<-matrix(c(219458, 216594, 213871, 389197,224879,194814,240371,231311,231461,400926,243361,230113, 289307, 261915, 264519, 393130, 276718, 258101, 264439,286211, 261279, 391921, 292564, 276876, 279704,291934, 270058, 373129, 300135, 287973), nrow=6, ncol=5)

Page 6 of 6
Or
(You can copy the R code from provided file ‘Q4 Reference_RCode.docx)
i)
Calculate, using Empirical Bayes Credibility Theory (EBCT) Model 1, the following:
a)
E[m(ɵ)]
b)
E[𝑠2(ɵ)]
c)
Var [m(ɵ)]
(2)
(2)
(4)
ii)
Calculate the credibility factors Zi and the credibility premiums for Delhi and Kerala.
(5)
iii)
Comment on the relationship between n and Zi in case of EBCT Model 1.
(2)
The number of motor third party claims per policy is modelled as a random variable X with a Poisson distribution with unknown parameter ʎ. The log likelihood function for estimating ʎ is given by:
L(ʎ) = log(ʎ) × Σx− ʎn
where n is the number of observations in the sample data.
There is data for a total of 150 observations and the total number of motor third party claims is 280.
The log likelihood function for the values of ʎ = 0, 0.1, 0.2, ………….., 1.8, 1.9, 3.3 has been plotted below:
iv)
Determine an approximate maximum likelihood estimate for ʎ using this plot.
(2)
v)
Determine the exact maximum likelihood estimate for ʎ and compare your answer with the approximate estimate obtained in part (iv).
(3)
[20]
*************
0.00.51.01.52.02.53.0-600-400-200Log-likelihood of lambdalambdaLog_likelihood
4 Consider a random variable, 𝑋, following a modified exponential distribution with
Cumulative Distribution Function (CDF):
𝐹􁈺𝑥􁈻 􀵌 􀵜
0, 𝑥 􀵏 0
1 􀵆 exp􁈺􀵆𝜆𝑥􀬶􁈻, 𝑥􀵒0
(i) Plot the CDF 𝐹􁈺𝑥􁈻 as a function of 𝑥 for 𝑥 􀵌 0, 0.1, 0.2, … ,9.9, 10 when
λ 􀵌 0.2. [4]
(ii) Plot the density function 𝑓􁈺𝑥􁈻 of 𝑋 as a function of 𝑥 for
𝑥 􀵌 0, 0.1, 0.2, … ,9.9, 10 when λ 􀵌 0.2. [5]
A random sample of 100 values of X is provided in randomSample.Rdata.
Loading the sample data into R will generate a vector x with 100 values representing
the sample.
(iii) Calculate the value of the log likelihood function for the parameter λ at the
point λ 􀵌 0.2 based on this random sample. [3]
(iv) Plot the values of the log likelihood function for the parameter λ based on the
sample in randomSample.Rdata. Your plot of the log likelihood function
must be for values of λ 􀵌 0.01, 0.02, … , 0.99, 1. [7]
The maximum likelihood estimator for the parameter λ based on a random sample
𝑋􀬵, … , 𝑋􀯇 is given by:
λ􀷠
􀵌
𝑁
Σ 𝑋􀯜
􀯇 􀬶
􀯜􀭀􀬵
(v) Estimate the value of λ using the maximum likelihood estimator given above
and the sample in randomSample.Rdata. [3]
(vi) Comment on the plot in part (iv) and the estimate in part (v). [2]
[Total 24]
CS1B A2024–6


Solution 4:
(i) 
a) > m <- mean(rowMeans(Claims)) (1)
 > m
 [1] 278542.3 (1)
(2) 
b) > s<-mean(apply(Claims,1,var)) (1)
 > s
 [1] 846425572 (1)
(2) 
c) > n <-ncol(Claims)
 > n
 [1] 5 
 (1)
 > v<-var(rowMeans(Claims))-mean(apply(Claims,1,var))/n (2)
 > v
 [1] 2842778626 (1)
(4) 
(ii) 
> Z <- n/(n+s/v) (1)
> Z
[1] 0.9437976 (1)
> PurePremium <-Z*rowMeans(Claims)+(1-Z)*m (1)
> PurePremium
[1] 259773.5 258770.4 249940.8 383415.5 268150.2 251203.4 (1)
Credibility premium for Delhi is 383415.5 and for Kerala is 251203.40 (1)
(5) 
(iii) 
Z is an increasing function of n. In the formula for credibility factor Z = n
/ (n + s/v), with an increasing value of n, Z will tend to increase. Intuitive
ly also, it is true because as the number of observations for the particular r
isk under consideration are more, more reliable is the specific data from that
particular risk and hence credibility factor Z would be higher indicative of m

Page 11 of 11
ore weightage to the specific data(mean for the specific risk) rather than the
collateral data (overall mean for all risks)
 (2)
(iv)
Based on the graph, the approximate maximum likelihood estimate i.e. the value
at which the log likelihood is maximum is around 1.8 to 1.9.
 (2)
(v) 
Exact Maximum Likelihood Estimate λ is
λ = Σ𝑥𝑖/𝑛
 = 280/150
 = 1.87 (2)
So, the actual maximum likelihood estimate calculated using first principles
is close to the approximate maximum likelihood determined based on the graph.
 (1)
 (3) 
 [20] 
**************