---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Q. 2) 
The following exercises relate to a regression analysis of maximum systolic blood pressure against age. A recent study suggests the relationship is given by the equation:

**Maximum Systolic Blood Pressure** = 100 + **Age** (in years)

---

### **Exercises**

1.  If we decide to fit a regression line with maximum systolic blood pressure as the response variable (Y) and age as the explanatory variable (X), what should be the values of the regression coefficients $\alpha$ and $\beta$ in light of the above equation?

---

Suppose this is to be empirically proven, and 20 people of varying ages are tested for their maximum systolic blood pressure.

---

The following data has been collected:

| Age (X) | Maximum Systolic Blood Pressure (Y) | Age (X) | Maximum Systolic Blood Pressure (Y) |
| :---: | :---: | :---: | :---: |
| 28 | 132 | 60 | 149 |
| 37 | 140 | 55 | 154 |
| 41 | 155 | 29 | 117 |
| 52 | 160 | 43 | 146 |
| 57 | 167 | 36 | 142 |
| 49 | 148 | 50 | 168 |
| 38 | 128 | 34 | 144 |
| 25 | 131 | 40 | 156 |
| 23 | 118 | 26 | 114 |
| 48 | 139 | 33 | 133 |

---

The data of age in years (X) and corresponding maximum systolic blood pressure (Y) can be entered in R using the below code:
```{r}
x=c(28, 37, 41, 52, 57, 49, 38, 25, 23, 48, 60, 55, 29, 43, 36, 50, 34, 40, 26, 33)

y=c(132, 140, 155, 160, 167, 148, 128, 131, 118, 139, 149, 154, 117, 146, 142, 168, 144, 156,
114, 133)
```

Enter the data in R by copying the above code.
---

### Exercise 2

ii) Prepare a scatter plot of the data and briefly comment on the same. (3)

Histogram of 1200 simulations of x1_bar
x1_bar
Frequency
0.830 0.832 0.834 0.836 0.838
0 50 100 150
Histogram of 1200 simulations of x2_bar
x2_bar
Frequency
0.162 0.164 0.166 0.168 0.170
0 50 100 150
Histogram of 1200 simulations of x3_bar
x3_bar
Frequency

0.494 0.496 0.498 0.500 0.502 0.504
0 50 100 150 200 250

---

### Exercise 3

Plot the fitted line for regression of Y on X.

---

### Exercise 4

Using anova function check whether the slope parameter is significant.

---

Solution 2:
(i)
The given equation is 
 Maximum Systolic Blood Pressure = 100 + Age (in years)
 It can be written as
 y = 100 + x
 y = 100 + 1*(x)
 y = α + βx
 𝛼 = 100 𝑎𝑛𝑑 𝛽 = 1
 (2)

---
(ii)
x=c(28, 37, 41, 52, 57, 49, 38, 25, 23, 48, 60, 55, 29, 43, 36, 50, 34, 40, 
26, 33)
y=c(132, 140, 155, 160, 167, 148, 128, 131, 118, 139, 149, 154, 117, 146, 1
42, 168, 144, 156, 114, 133)
plot(x,y) 
(1)
 
 (1)
The age in years(x) and the systolic blood pressure(y) are positively correlat
ed.

---

### Exercise 3

* Fiting a Linear Model using the <tt>lm()</tt> command.

```{r}
lm.result=lm(y~x) 
```

---

```{r}
plot(lm((y~x)))
abline(lm(y~x)) 
```

---

### Exercise 4 

```{r}
anova(lm.result)
``` 

Analysis of Variance Table
Response: y
 Df Sum Sq Mean Sq F value Pr(>F) 
x 1 3082.9 3082.94 33.591 1.717e-05 ***
Residuals 18 1652.0 91.78 


#### Conclusion


From the above, it is clear that the slope parameter is significant.

---

### Exercise 5
Obtain a summary for the linear regression model fitted in **Exercise 4** and clearly state the estimates for the values of the coefficients α and β.


---

### Exercise 5 

```{r}
summary(lm.result) 

```

Call:
lm(formula = y ~ x)
Residuals:
 Min 1Q Median 3Q Max
-15.485 -6.504 1.177 5.979 14.846
Coefficients:
 Estimate Std. Error t value Pr(>|t|) 
(Intercept) 96.4994 8.1460 11.846 6.21e-10 ***
x 1.1331 0.1955 5.796 1.72e-05 ***

---

Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Residual standard error: 9.58 on 18 degrees of freedom
Multiple R-squared: 0.6511, Adjusted R-squared: 0.6317
F-statistic: 33.59 on 1 and 18 DF, p-value: 1.717e-05 
 (1)
The value of the estimates of the co-efficient are
 Alpha(α) = 96.4994 Beta(β) = 1.1331 (1)

---

### Exercise 6
vi)
Test the veracity of the above equation as promulgated in the recent research, by referring to your answers in **Exercise 1** and **Exercise 5**.



---

### Exercise 6

* The values of alpha and beta are expected to be 100 and 1 respectively as per (i). 

* But empirical test results fetch the values as 96.4994 and 1.1331
respectively, which is close to the expected values of 100 and 1 respectively. 

* Hence when empirically tested, we find that the claim made by the research
about the maximum systolic blood pressure is valid.


---

### Exercise 7

Calculate the sample Pearson correlation coefficient, sample Kendall correlation coefficient and sample Spearman correlation coefficient.

----------------------------------------------


#### Pearson Exercise 7


```{r}
cor(x,y,method="pearson")
```

#### Spearman

```{r}
cor(x,y,method="spearman")
```

#### Kendall Tau

```{r}
cor(x,y,method="kendall")
```


---

### Exercise 8 

viii)
Comment on each method of calculating correlation coefficient and also comment on the sample correlation coefficients obtained in part (vii).




---

### Exercise 8

(viii) 
 Pearson’s correlation co-efficient measures the strength of the linear r
elationship between the two variables, whereas Spearman Correlation method mea
sures the strength of monotonic but not necessarily linearity between two vari
ables.

*  Since Spearman considers the rank than the actual values, the value of t
he coefficient is less affected by extreme values/outliers in the data than Pearson’s Correlation Coefficient. Hence it is more robust.

* Kendall’s correlation coefficient is considered to have better statistical
properties when the data set is small and have more tied ranks, though it considers the relative values between the data set and not actual values.


* Generally, the value of Kendall’s coefficient is lower than the Spearman’s
rank coefficient.

* Based on the sample correlation coefficients calculated in part (vii), we
conclude Spearman Rank Coefficient Pearson’s Coefficient Kendall’s Coeffi
cient

---

### Exercise 9

One of your colleagues says that there exists perfect positive correlation between age and maximum systolic blood pressure.

* Perform a hypothesis test to test the above statement at 5% level of significance. 

* You should state the null and alternate hypotheses, report the p-value of the test and arrive at a clear conclusion.

<i>(Hint: Use Pearson Method)</i>



---

### Exercise 9

(ix) 

#### Hypotheses

𝐻0 : ρ = 1
𝐻1 : ρ ≠ 1 

---

```{r}
cor.test(x,y,method="pearson")  
```

#### Output Details
* p-value = 1.717e-05
* alternative hypothesis: true correlation is not equal to 0


* 95 percent confidence interval: 0.5667661 0.9206793
sample estimates:
 cor
0.8069094 (1)
p-value is 1.717e-05 

---

#### Interpretation

Since 95% confidence interval (0.5667661, 0.9206793) does not include
the value 1, there is sufficient evidence to reject the hypothesis that there
is perfect correlation between Age and Systolic Blood Pressure though there is
strong positive correlation between the age and systolic Blood Pressure.

---

BLANK

---