---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

Q. 3)  
i. 
install.packages("markovchain", type = "binary") 
install.packages("dplyr", type = "binary") 
library(markovchain) 
library(dplyr) 
Page 14 of 19 
IAI                                                                                                                                                  
CS2B-1124 
MC <- new("markovchain", states = c("1","2","3"), 
transitionMatrix = matrix(data = c(0.1,0.6,0.3,0.4,0.1,0.5,0.2,0.3,0.5),  
byrow = TRUE , nrow = 3), name = "MC") 
MC 
A  3 - dimensional discrete Markov Chain defined by the following states:  
1, 2, 3  
The transition matrix  (by rows)  is defined as follows:  
1   2   3 
1 0.1 0.6 0.3 
2 0.4 0.1 0.5 
3 0.2 0.3 0.5 
ii. 
markovchain::summary(MC) 
MC  Markov chain that is composed by:  
Closed classes:  
1 2 3  
Recurrent classes:  
{1,2,3} 
Transient classes:  
NONE  
The Markov chain is irreducible  
The absorbing states are: NONE 
Alternate Answer 
markovchain::is.irreducible(MC) 
TRUE 
iii. 
markovchain::meanFirstPassageTime(MC) 
1        2        3 
1 0.000000 2.051282 2.631579 
2 3.333333 0.000000 2.280702 
3 4.000000 2.820513 0.000000 
markovchain::meanRecurrenceTime(MC) 
1        2        3  
4.200000 3.230769 2.210526  
iv. 
InitialState <- c(1,0,0) 
After2steps <- InitialState * (MC^2) 
After4steps <- InitialState * (MC^4) 
After6steps <- InitialState * (MC^6) 
(2) 
(1) 
(4) 
Page 15 of 19 
IAI                                                                                                                                                  
CS2B-1124 
>After2steps 
1    2    3 
[1,] 0.31 0.21 0.48 
> After4steps 
1      2      3 
[1,] 0.2491 0.2931 0.4578 
> After6steps 
1        2        3 
[1,] 0.239851 0.306891 0.453258 
>steadyStates(MC) 
1         2        3 
[1,] 0.2380952 0.3095238 0.452381 
v. 
stat_dist <- data.frame(c(After2steps),c(After4steps),c(After6steps),c(steadyStates(MC))) 
stat_dist 
q = c(2,4,6,15) 
(4) 
plot(q,stat_dist[1,],type = "l",main = "Movement towards stationary state",col = "red", ylim = 
c(0.2,0.5)) 
lines(q,stat_dist[2,], col = "black") 
lines(q,stat_dist[3,], col = "blue") 
legend("top",legend = c("State 1", "State 2", "State 3"),col = c("red", "black", "blue"), 
pch = 7) 
After 6 steps the Markov chain almost becomes stationary.  
Page 16 of 19 
IAI                                                                                                                                                  
CS2B-1124 
(6) 
vi. 
set.seed(42) 
n_steps <- 10000 
states <- numeric(n_steps) 
states[1] <- 1 
for (i in 2:n_steps) { 
states[i] <- sample(1:3, 1, prob = MC[states[i-1], ]) 
} 
head(states) 
[1] 1 1 1 2 1 3 
vii. 
triplet_counts <- c( 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 1 & states[-1] == 1), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 1 & states[-1] == 1), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 1 & states[-1] == 1), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 2 & states[-1] == 1), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 2 & states[-1] == 1), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 2 & states[-1] == 1), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 3 & states[-1] == 1), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 3 & states[-1] == 1), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 3 & states[-1] == 1), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 1 & states[-1] == 2), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 1 & states[-1] == 2), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 1 & states[-1] == 2), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 2 & states[-1] == 2), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 2 & states[-1] == 2), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 2 & states[-1] == 2), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 3 & states[-1] == 2), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 3 & states[-1] == 2), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 3 & states[-1] == 2), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 1 & states[-1] == 3), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 1 & states[-1] == 3), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 1 & states[-1] == 3), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 2 & states[-1] == 3), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 2 & states[-1] == 3), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 2 & states[-1] == 3), 
sum(states[-c(1, 2)] == 1 & states[-length(states)] == 3 & states[-1] == 3), 
sum(states[-c(1, 2)] == 2 & states[-length(states)] == 3 & states[-1] == 3), 
sum(states[-c(1, 2)] == 3 & states[-length(states)] == 3 & states[-1] == 3) 
) 
ijk 
(4) 
<- 
c(111,112,113,121,122,123,131,132,133,211,212,213,221,222,223,231,232,233,311,312,313
 ,321,322,323,331,332,333) 
triplets_test <- data.frame(ijk,triplet_counts) 
triplets_test 
Page 17 of 19 
IAI                                                                                                                                                  CS2B-1124 
Page 18 of 19 
 
ijk triplet_counts 
1  111             30 
2  112            158 
3  113             69 
4  121            111 
5  122            787 
6  123            346 
7  131            115 
8  132            520 
9  133            275 
10 211            582 
11 212            151 
12 213            732 
13 221            117 
14 222             25 
15 223            174 
16 231            545 
17 232            140 
18 233            648 
19 311            132 
20 312            207 
21 313            351 
22 321            297 
23 322            487 
24 323            770 
25 331            482 
26 332            639 
27 333           1109 
(6) 
  
 
 viii. 
double_counts <- c( 
  sum(states[-length(states)] == 1 & states[-1] == 1), 
  sum(states[-length(states)] == 2 & states[-1] == 1), 
  sum(states[-length(states)] == 3 & states[-1] == 1), 
  sum(states[-length(states)] == 1 & states[-1] == 2), 
  sum(states[-length(states)] == 2 & states[-1] == 2), 
  sum(states[-length(states)] == 3 & states[-1] == 2), 
  sum(states[-length(states)] == 1 & states[-1] == 3), 
  sum(states[-length(states)] == 2 & states[-1] == 3), 
  sum(states[-length(states)] == 3 & states[-1] == 3) 
) 
 
ij <- c(11,12,13,21,22,23,31,32,33) 
double_counts <- data.frame(ij,double_counts) 
double_counts$ij <- as.character(double_counts$ij) 
double_counts 
  ij double_counts 
1 11           257 
2 12          1244 
3 13           910 
4 21          1465 
5 22           316 
IAI                                                                                                                                                  CS2B-1124 
Page 19 of 19 
 
6 23          1333 
7 31           690 
8 32          1554 
9 33          2230 
(2) 
 
 ix. 
triplets_test$ij <- substr(triplets_test$ijk,1,2) 
merged_df <- left_join(triplets_test,double_counts, by = "ij") 
 
tp <- c(0.1,0.6,0.3,0.4,0.1,0.5,0.2,0.3,0.5) 
jk <- c(11,12,13,21,22,23,31,32,33) 
 
t_prob <- data.frame(jk,tp) 
t_prob$jk <- as.character(t_prob$jk) 
 
 
merged_df$jk <- substr(triplets_test$ijk,2,3) 
 
merged_df1 <- left_join(merged_df,t_prob, by = "jk") 
 
merged_df1$ef <- merged_df1$double_counts*merged_df1$tp 
 
merged_df1$ts <- (((merged_df1$triplet_counts - merged_df1$ef)^2)/merged_df1$ef) 
 
chi_square_test <- chisq.test(merged_df1$triplet_counts,p = 
merged_df1$ef/sum(merged_df1$ef)) 
chi_square_test 
 
#Chi-squared test for given probabilities 
 
data:  merged_df1$triplet_counts 
X-squared = 7971.8, df = 26, p-value < 2.2e-16 
 
 
if (chi_square_test$p.value < 0.05) { 
  cat("Reject the null hypothesis: The observed frequencies do not match the expected 
frequencies.\n") 
} else { 
  cat("Fail to reject the null hypothesis: The observed frequencies match the expected 
frequencies.\n") 
} 
 
Reject the null hypothesis: The observed frequencies do not match the expected frequencies. 
(10) 
[39 Marks] 
 
 