---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


Q. 2)  
i. 
install.packages("survival", type = "binary") 
install.packages("survminer", type = "binary") 
install.packages("ggplot2", type = "binary") 
library(survival) 
library(survminer)  
library(ggplot2) 
Cruise<- read.csv("<path>/Cruise.csv") 
summary(Cruise$time) 
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.  
2.0   114.2   300.0   203.0   300.0   300.0  
table(Cruise$status) 
0   1  
105  95  
ii. 
km_fit <- survfit(Surv(time,status) ~ 1, data = Cruise) 
summary(km_fit) 
(3) 
[33 Marks] 
(2) 
Page 7 of 19 
IAI                                                                                                                                                  CS2B-1124 
Page 8 of 19 
 
Call: survfit(formula = Surv(time, status) ~ 1, data = Cruise) 
 
 time n.risk n.event survival std.err lower 95% CI upper 95% CI 
    2    200       1    0.995 0.00499        0.985        1.000 
    3    199       1    0.990 0.00704        0.976        1.000 
    4    198       1    0.985 0.00860        0.968        1.000 
    5    197       2    0.975 0.01104        0.954        0.997 
    8    195       1    0.970 0.01206        0.947        0.994 
   12    194       3    0.955 0.01466        0.927        0.984 
   20    191       2    0.945 0.01612        0.914        0.977 
   26    189       2    0.935 0.01743        0.901        0.970 
   28    187       1    0.930 0.01804        0.895        0.966 
   29    186       2    0.920 0.01918        0.883        0.958 
   34    184       1    0.915 0.01972        0.877        0.954 
   39    183       1    0.910 0.02024        0.871        0.951 
   41    182       2    0.900 0.02121        0.859        0.943 
   45    180       1    0.895 0.02168        0.854        0.939 
   46    179       2    0.885 0.02256        0.842        0.930 
   47    177       1    0.880 0.02298        0.836        0.926 
   48    176       1    0.875 0.02339        0.830        0.922 
   49    175       2    0.865 0.02416        0.819        0.914 
   50    173       1    0.860 0.02454        0.813        0.909 
   55    172       1    0.855 0.02490        0.808        0.905 
   57    171       1    0.850 0.02525        0.802        0.901 
   58    170       1    0.845 0.02559        0.796        0.897 
   60    169       2    0.835 0.02625        0.785        0.888 
   66    167       2    0.825 0.02687        0.774        0.879 
   71    165       1    0.820 0.02717        0.768        0.875 
   76    164       1    0.815 0.02746        0.763        0.871 
   80    163       1    0.810 0.02774        0.757        0.866 
   83    162       1    0.805 0.02802        0.752        0.862 
   90    161       1    0.800 0.02828        0.746        0.857 
   91    160       1    0.795 0.02855        0.741        0.853 
   92    159       1    0.790 0.02880        0.736        0.849 
   94    158       2    0.780 0.02929        0.725        0.840 
   98    156       1    0.775 0.02953        0.719        0.835 
   99    155       1    0.770 0.02976        0.714        0.831 
  103    154       1    0.765 0.02998        0.708        0.826 
  107    153       1    0.760 0.03020        0.703        0.822 
  111    152       1    0.755 0.03041        0.698        0.817 
  112    151       1    0.750 0.03062        0.692        0.812 
  115    150       1    0.745 0.03082        0.687        0.808 
  116    149       1    0.740 0.03102        0.682        0.803 
  118    148       1    0.735 0.03121        0.676        0.799 
  120    147       2    0.725 0.03157        0.666        0.790 
  122    145       1    0.720 0.03175        0.660        0.785 
  123    144       1    0.715 0.03192        0.655        0.780 
  125    143       1    0.710 0.03209        0.650        0.776 
  126    142       1    0.705 0.03225        0.645        0.771 
  127    141       2    0.695 0.03256        0.634        0.762 
  131    139       1    0.690 0.03270        0.629        0.757 
  132    138       1    0.685 0.03285        0.624        0.752 
  135    137       2    0.675 0.03312        0.613        0.743 
  137    135       2    0.665 0.03337        0.603        0.734 
  138    133       1    0.660 0.03350        0.598        0.729 
  139    132       1    0.655 0.03361        0.592        0.724 
  140    131       2    0.645 0.03384        0.582        0.715 
  143    129       1    0.640 0.03394        0.577        0.710 
  144    128       1    0.635 0.03404        0.572        0.705 
  145    127       1    0.630 0.03414        0.567        0.701 
  147    126       1    0.625 0.03423        0.561        0.696 
  148    125       2    0.615 0.03441        0.551        0.686 
  149    123       1    0.610 0.03449        0.546        0.681 
  150    122       1    0.605 0.03457        0.541        0.677 
  151    121       1    0.600 0.03464        0.536        0.672 
  155    120       1    0.595 0.03471        0.531        0.667 
  156    119       1    0.590 0.03478        0.526        0.662 
  157    118       2    0.580 0.03490        0.515        0.653 
  158    116       1    0.575 0.03496        0.510        0.648 
  161    115       1    0.570 0.03501        0.505        0.643 
  165    114       1    0.565 0.03506        0.500        0.638 
  167    113       1    0.560 0.03510        0.495        0.633 
  168    112       2    0.550 0.03518        0.485        0.623 
  169    110       2    0.540 0.03524        0.475        0.614 
  170    108       1    0.535 0.03527        0.470        0.609 
IAI                                                                                                                                                  CS2B-1124 
Page 9 of 19 
 
  174    107       1    0.530 0.03529        0.465        0.604 
  177    106       1    0.525 0.03531        0.460        0.599  
ggsurvplot(km_fit,data = Cruise,title = "Kaplan-Meier", combine = TRUE, conf.int = TRUE, 
risk.table = TRUE) 
 
 
 
(4) 
 
 iii. 
km_fit_treatment <- survfit(Surv(time, status) ~ Pclass, data = Cruise) 
summary(km_fit_treatment) 
Call: survfit(formula = Surv(time, status) ~ Pclass, data = Cruise) 
 
                Pclass=1  
 time n.risk n.event survival std.err lower 95% CI upper 95% CI 
    2     65       1    0.985  0.0153        0.955        1.000 
    8     64       1    0.969  0.0214        0.928        1.000 
   34     63       1    0.954  0.0260        0.904        1.000 
   39     62       1    0.938  0.0298        0.882        0.999 
   58     61       1    0.923  0.0331        0.861        0.990 
   60     60       1    0.908  0.0359        0.840        0.981 
   71     59       1    0.892  0.0384        0.820        0.971 
   94     58       1    0.877  0.0407        0.801        0.961 
   99     57       1    0.862  0.0428        0.782        0.950 
  107     56       1    0.846  0.0448        0.763        0.939 
  111     55       1    0.831  0.0465        0.744        0.927 
  135     54       1    0.815  0.0481        0.726        0.915 
  140     53       1    0.800  0.0496        0.708        0.903 
  155     52       1    0.785  0.0510        0.691        0.891 
  161     51       1    0.769  0.0523        0.673        0.879 
  165     50       1    0.754  0.0534        0.656        0.866 
  169     49       1    0.738  0.0545        0.639        0.853 
  177     48       1    0.723  0.0555        0.622        0.840 
IAI                                                                                                                                                  CS2B-1124 
Page 10 of 19 
 
 
                Pclass=2  
 time n.risk n.event survival std.err lower 95% CI upper 95% CI 
    3     67       1    0.985  0.0148        0.956        1.000 
    4     66       1    0.970  0.0208        0.930        1.000 
    5     65       2    0.940  0.0289        0.885        0.999 
   12     63       1    0.925  0.0321        0.865        0.990 
   26     62       2    0.896  0.0374        0.825        0.972 
   28     60       1    0.881  0.0396        0.806        0.962 
   29     59       1    0.866  0.0417        0.788        0.951 
   41     58       2    0.836  0.0453        0.752        0.929 
   46     56       2    0.806  0.0483        0.717        0.906 
   47     54       1    0.791  0.0497        0.699        0.895 
   49     53       2    0.761  0.0521        0.666        0.870 
   66     51       2    0.731  0.0542        0.633        0.846 
   76     49       1    0.716  0.0551        0.616        0.833 
  103     48       1    0.701  0.0559        0.600        0.820 
  112     47       1    0.687  0.0567        0.584        0.807 
  122     46       1    0.672  0.0574        0.568        0.794 
  123     45       1    0.657  0.0580        0.552        0.781 
  125     44       1    0.642  0.0586        0.537        0.768 
  126     43       1    0.627  0.0591        0.521        0.754 
  137     42       1    0.612  0.0595        0.506        0.740 
  143     41       1    0.597  0.0599        0.490        0.727 
  157     40       2    0.567  0.0605        0.460        0.699 
  158     38       1    0.552  0.0608        0.445        0.685 
  168     37       1    0.537  0.0609        0.430        0.671 
  169     36       1    0.522  0.0610        0.415        0.657 
  174     35       1    0.507  0.0611        0.401        0.642 
 
                Pclass=3  
 time n.risk n.event survival std.err lower 95% CI upper 95% CI 
   12     68       2    0.971  0.0205        0.931        1.000 
   20     66       2    0.941  0.0285        0.887        0.999 
   29     64       1    0.926  0.0317        0.866        0.991 
   45     63       1    0.912  0.0344        0.847        0.982 
   48     62       1    0.897  0.0369        0.828        0.972 
   50     61       1    0.882  0.0391        0.809        0.962 
   55     60       1    0.868  0.0411        0.791        0.952 
   57     59       1    0.853  0.0429        0.773        0.941 
   60     58       1    0.838  0.0447        0.755        0.930 
   80     57       1    0.824  0.0462        0.738        0.919 
   83     56       1    0.809  0.0477        0.721        0.908 
   90     55       1    0.794  0.0490        0.704        0.896 
   91     54       1    0.779  0.0503        0.687        0.884 
   92     53       1    0.765  0.0514        0.670        0.872 
   94     52       1    0.750  0.0525        0.654        0.860 
   98     51       1    0.735  0.0535        0.638        0.848 
  115     50       1    0.721  0.0544        0.621        0.836 
  116     49       1    0.706  0.0553        0.605        0.823 
  118     48       1    0.691  0.0560        0.590        0.810 
  120     47       2    0.662  0.0574        0.558        0.784 
  127     45       2    0.632  0.0585        0.528        0.758 
  131     43       1    0.618  0.0589        0.512        0.745 
IAI                                                                                                                                                  CS2B-1124 
Page 11 of 19 
 
  132     42       1    0.603  0.0593        0.497        0.731 
  135     41       1    0.588  0.0597        0.482        0.718 
  137     40       1    0.574  0.0600        0.467        0.704 
  138     39       1    0.559  0.0602        0.452        0.690 
  139     38       1    0.544  0.0604        0.438        0.676 
  140     37       1    0.529  0.0605        0.423        0.662 
  144     36       1    0.515  0.0606        0.409        0.648 
  145     35       1    0.500  0.0606        0.394        0.634 
  147     34       1    0.485  0.0606        0.380        0.620 
  148     33       2    0.456  0.0604        0.352        0.591 
  149     31       1    0.441  0.0602        0.338        0.576 
  150     30       1    0.426  0.0600        0.324        0.562 
  151     29       1    0.412  0.0597        0.310        0.547 
  156     28       1    0.397  0.0593        0.296        0.532 
  167     27       1    0.382  0.0589        0.283        0.517 
  168     26       1    0.368  0.0585        0.269        0.502 
  170     25       1    0.353  0.0580        0.256        0.487 
 
 
ggsurvplot(km_fit_treatment, data = Cruise, conf.int = TRUE, xlab = "Ti
 me (minutes)", ylab = "Survival Probability", title = "Kaplan-Meier Sur
 vival Curve by Pclass",pval = TRUE, risk.table = TRUE) 
 
  
 
Passenger travelling in Class 1 had the highest survival probability followed by Passenger travelling in 
Class 2 & 3. 
(4) 
 
 iv. 
cox_fit <- coxph(Surv(time, status) ~ Age + Sex + Pclass + Mstatus + Parch + SibSp + 
Embarked, data = Cruise) 
summary(cox_fit) 
Call: 
coxph(formula = Surv(time, status) ~ Age + Sex + Pclass + Mstatus +  
    Parch + SibSp + Embarked, data = Cruise) 
 
  n= 200, number of events= 95  
 
IAI                                                                                                                                                  
CS2B-1124 
coef exp(coef)  se(coef)
 Age
      z
 Pr(>|z|)     
       0.002217  1.002219  0.005214  0.425   0.6707     
Sex       0.477555  1.612128  0.217117  2.200   0.0278 *   
Pclass    0.564965  1.759386  0.135786  4.161 3.17e-05 *** 
Mstatus  -0.010950  0.989110  0.105248 -0.104   0.9171     
Parch    -0.212472  0.808583  0.106457 -1.996   0.0460 *   
SibSp     
0.181520  1.199038  0.132582  1.369   0.1710     
Embarked -0.200554  0.818277  0.128254 -1.564   0.1179     --- 
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
exp(coef) exp(-coef) lower .95 upper .95 
Age
         1.0022     0.9978    
0.9920    1.0125 
Sex         1.6121     0.6203    1.0534    2.4672 
Pclass      1.7594     0.5684    1.3483    2.2958 
Mstatus    
 0.9891     1.0110    0.8047    1.2157 
Parch       0.8086     1.2367    0.6563    0.9962 
SibSp       1.1990     0.8340    0.9247    1.5548 
Embarked    0.8183     1.2221    0.6364    1.0521 
Concordance= 0.646  (se = 0.028 ) 
Likelihood ratio test= 28.77  on 7 df,   p=2e-04 
Wald test      
      =
 27.36  on 7 df,   p=3e-04 
Score (logrank) test = 28.67  on 7 df,   p=2e-04 
(3) 
v. 
(a)  The coef column represents the log hazard ratios (log(HR)). These coefficients tell us how much 
each unit increase in the predictor variable affects the hazard rate (risk of the event). A positive 
coefficient indicates that an increase in the predictor leads to an increased hazard (higher risk), while 
a negative coefficient suggests a decreased hazard (lower risk). 
Age: The coefficient is 0.002217, suggesting a very slight increase in hazard with age. 
Sex: The coefficient is 0.477555, meaning being male (since male is coded as 1 and female as 0) 
increases the hazard compared to being female. 
Pclass: The coefficient is 0.564965, indicating a higher passenger class (i.e. Passenger class 2 & 3) is 
associated with a higher hazard as compared to Passenger class 1. 
Mstatus: The coefficient is -0.010950, indicating as the factor increases and moves towards 3 i.e. 
Married the risk reduces marginally. 
Parch:  The coefficient is -0.212472, indicating as the number of parents/children aboard the ship 
increases the risk reduces sharply. 
SibSp: The coefficient is   0.181520, indicating as the spouse/higher number of siblings being aboard 
the ship increases the risk reduces increases. 
Embarked: The coefficient is   -0.200554, indicating the risk decreases for passengers boarding last. 
(b)  Small p-values (usually < 0.05) suggest that the predictor is statistically significant in the model. 
Thus, Sex, Parch & Pclass are the most statistically significant as their p values are less than 0.05. 
Page 12 of 19 
IAI                                                                                                                                                  
CS2B-1124 
(c)  The lower .95 and upper .95 columns show the 95% confidence interval for the hazard ratio. If 
this interval excludes 1, it indicates that the predictor likely has a significant effect. 
Sex, Pclass and Parch excludes the confidence interval 1 and thus are the significant effect in line with 
our conclusion in previous sub part. 
(7) 
vi. 
exp(coef(cox_fit)) 
Age               
Sex     
Pclass    
Mstatus      
Parch      
1.0022192  
SibSp   
1.6121277 
Embarked  
1.7593859 0.9891096 0.8085829 
1.1990382 
0.8182769 
The exp(coef) column shows the hazard ratios (HR), which are easier to interpret than the raw 
coefficients. The hazard ratio tells us the relative risk of the event per unit increase in the predictor. 
exp(coef)  
Age         
1.0022 
Sex         
Pclass      
1.6121  
1.7594 
Mstatus     0.9891 
Parch      
SibSp       
 0.8086    
1.1990    
Embarked    0.8183 
Age (HR = 1.0022): For each additional year of age, the hazard (risk of event) increases by 0.22%. 
Sex (HR = 1.6121): Males have a 61.2% higher risk of the death compared to females. 
Pclass (HR = 1.7594): For each unit increase in the Pclass factor i.e. Pclass changing from 1 to 2, the 
hazard increases by 75.94%. 
Mstatus (HR = 0.9891): For each unit increase in the Mstatus factor i.e. Mstatus changing from 0 to 
1, the hazard decreases by 1.09%. 
Parch (HR = 0.8086): For each unit increase in the number of parents/children of the passenger 
being aboard, the hazard decreases by 19.14%. 
SibSp (HR = 1.1990): For each unit increase in the number of siblings or spouse travelling with the 
passenger being aboard, the hazard increases by 19.90%. 
Embarked (HR = 0.8183): For each unit increase in the Embarking station factor, the hazard decreases 
by 18.17%. Since the embarking station has been numbered in a manner that last embarking station 
has the highest factor, we can say that as with each subsequent embarking station the risk decreases 
by 18.17%. 
vii. 
ggsurvplot(list(KM = km_fit, Cox = survfit(cox_fit)),data = Cruise,title = "Kaplan-Meier vs Cox 
Model",combine = TRUE, conf.int = TRUE, pval = TRUE, risk.table = TRUE) 
(4) 
Page 13 of 19 
IAI                                                                                                                                                  
CS2B-1124 
viii. 
Use Kaplan-Meier When: 
• You want to visualize and describe survival times between two or more groups. 
• You have a simple comparison between groups based on one categorical variable. 
• You’re working with small datasets or want a simple descriptive analysis. 
• The focus is on estimating survival probabilities over time without adjusting for other 
factors. 
Use CoxPH When: 
• You need to adjust for multiple covariates (e.g., age, treatment, gender) to assess their 
independent effect on survival. 
(2) 
• You want to estimate hazard ratios to quantify the relative risk associated with covariates, 
as calculated in (iv).  
• You’re building a predictive model to estimate survival times based on individual 
characteristics. 
• You’re interested in testing interactions between covariates. 
• The dataset contains complex factors that need to be considered simultaneously. 
(2) 
[28 Marks] 
