---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---

                                                                                                                                       
CS2B-1124 
Q. 1)  
install.packages("dplyr", type = "binary") 
install.packages("randomForest", type = "binary") 
install.packages("caret", type = "binary") 
install.packages("rpart", type = "binary") 
install.packages("rpart.plot", type = "binary") 
library(dplyr) 
library(randomForest)  
library(caret) 
library(rpart) 
library(rpart.plot) 
i) 
heart_disease <- read.csv("<path>/heart_disease.csv") 
heart_disease$sex <- as.factor(heart_disease$sex) 
heart_disease$cp <- as.factor(heart_disease$cp) 
heart_disease$fbs <- as.factor(heart_disease$fbs) 
heart_disease$restecg <- as.factor(heart_disease$restecg) 
heart_disease$exang <- as.factor(heart_disease$exang) 
heart_disease$slope <- as.factor(heart_disease$slope) 
heart_disease$ca <- as.factor(heart_disease$ca) 
heart_disease$thal <- as.factor(heart_disease$thal) 
heart_disease$target <- as.factor(heart_disease$target) 
Alternate Solution 
ii) 
heart_disease <- heart_disease %>% mutate(across(c(sex, cp, fbs, restecg, exang, slope, ca, 
thal, target), factor)) 
(3) 
train_index <- c(1:(nrow(heart_disease)*0.7)) 
train_data <- heart_disease[train_index, ] 
test_data <- heart_disease[-train_index, ] 
iii)  
set.seed(123) 
(2) 
tree_model <- rpart(target ~ age + sex + cp + trestbps + chol + fbs + restecg + thalach + exang 
+ oldpeak + slope + ca + thal,data = train_data, method = "class") 
rpart.plot(tree_model) 
Page 2 of 19 
IAI                                                                                                                                                  
CS2B-1124 
Inferences 
• Patients have been split basis the value of thalassemia. If the value is less than 3, then 
they go to left node else to the right node. Thus, 45% of the data has moved to right 
node and remaining to the left node. 
iv)  
• On the left node, a further split has been done basis the age of the patient. If the age  is 
less than 59, then it has branched leftward. This represents 35% of the total data. Thus, 
the probability that a heart disease is diagnosed when thalassemia is 3 and age of the 
patient is less than 59 is 0.08. However, with thalassemia being 3, age being greater 
than 59 and chest pain with 4, the probability of diagnosis the heart diseases is 0.73 
• The right node has been split basis the value of the chest pain. If the value is less than 
or equal to 3, it has been further split basis the number of major vessels. 
• Thus, if the thalassemia is not normal but, chest pain being less than equal to 3 and 
number of colored vessels being 0, the probability of getting heart diseases diagnosed 
is 0.24. The probability increases to 0.75 if number of colored vessels is greater than 0. 
• However, if thalassemia is not normal and chest pain is 4, the probability of getting 
heart diseases diagnosed increases to 0.89. 31% of total population falls here. 
(3) 
pred_tree <- predict(tree_model, test_data, type = "class") 
confusionMatrix(pred_tree, test_data$target) 
Confusion Matrix and Statistics 
Reference 
Prediction  0  1 
0 39 14 
1  7 30 
Accuracy : 0.7667           
95% CI : (0.6657, 0.8494) 
No Information Rate : 0.5111           
Page 3 of 19 
IAI                                                                                                                                                  CS2B-1124 
Page 4 of 19 
 
    P-Value [Acc > NIR] : 5.619e-07        
                                           
                  Kappa : 0.5315           
                                           
 Mcnemar's Test P-Value : 0.1904           
                                           
            Sensitivity : 0.8478           
            Specificity : 0.6818           
         Pos Pred Value : 0.7358           
         Neg Pred Value : 0.8108           
             Prevalence : 0.5111           
         Detection Rate : 0.4333           
   Detection Prevalence : 0.5889           
      Balanced Accuracy : 0.7648           
                                           
       'Positive' Class : 0            
 
confusion_matrix_tree <- confusionMatrix(pred_tree, test_data$target) 
 
confusion_matrix_tree$overall['Accuracy'] 
 Accuracy  
0.7666667   
 
confusion_matrix_tree$byClass['Precision'] 
Precision  
0.7358491 
 
confusion_matrix_tree$byClass['Recall'] 
   Recall  
0.8478261  
(4) 
  
 v) 
set.seed(123) 
rf_model_1 <- randomForest(target ~ age + sex + cp + trestbps + chol + fbs + restecg + thalach + exang 
+ oldpeak + slope + ca + thal,data = train_data, ntree = 100, mtry = 13) 
 
pred_rf_1 <- predict(rf_model_1, test_data) 
confusionMatrix(pred_rf_1, test_data$target) 
 
Confusion Matrix and Statistics 
 
          Reference 
Prediction  0  1 
         0 42 14 
         1  4 30 
                                           
               Accuracy : 0.8              
                 95% CI : (0.7025, 0.8769) 
    No Information Rate : 0.5111           
    P-Value [Acc > NIR] : 1.329e-08        
                                           
                  Kappa : 0.5978           
                                           
 Mcnemar's Test P-Value : 0.03389          
IAI                                                                                                                                                  CS2B-1124 
Page 5 of 19 
 
                                           
            Sensitivity : 0.9130           
            Specificity : 0.6818           
         Pos Pred Value : 0.7500           
         Neg Pred Value : 0.8824           
             Prevalence : 0.5111           
         Detection Rate : 0.4667           
   Detection Prevalence : 0.6222           
      Balanced Accuracy : 0.7974           
                                           
       'Positive' Class : 0  
(4) 
 
 vi)  
The default value of mtry is d/3 or √𝑑   ~ 4 (rounded off) 
 
set.seed(123) 
rf_model_2 <- randomForest(target ~ age + sex + cp + trestbps + chol + fbs + restecg + thalach 
+ exang + oldpeak + slope + ca + thal,data = train_data, ntree = 100) 
 
pred_rf_2 <- predict(rf_model_2, test_data) 
confusionMatrix(pred_rf_2, test_data$target) 
 
Confusion Matrix and Statistics 
 
          Reference 
Prediction  0  1 
         0 41 12 
         1  5 32 
                                           
               Accuracy : 0.8111           
                 95% CI : (0.7149, 0.8859) 
    No Information Rate : 0.5111           
    P-Value [Acc > NIR] : 3.351e-09        
                                           
                  Kappa : 0.6207           
                                           
 Mcnemar's Test P-Value : 0.1456           
                                           
            Sensitivity : 0.8913           
            Specificity : 0.7273           
         Pos Pred Value : 0.7736           
         Neg Pred Value : 0.8649           
             Prevalence : 0.5111           
         Detection Rate : 0.4556           
   Detection Prevalence : 0.5889           
      Balanced Accuracy : 0.8093           
                                           
       'Positive' Class : 0  
 
Alternate answer 
 
Confusion Matrix and Statistics 
 
IAI                                                                                                                                                  
CS2B-1124 
Reference 
Prediction  0  1 
0 42 13 
1  4 31 
Accuracy : 0.8111           
95% CI : (0.7149, 0.8859) 
No Information Rate : 0.5111           
P-Value [Acc > NIR] : 3.351e-09        
Kappa : 0.6203           
Mcnemar's Test P-Value : 0.05235          
Sensitivity : 0.9130           
Specificity : 0.7045           
Pos Pred Value : 0.7636           
Neg Pred Value : 0.8857           
Prevalence : 0.5111           
Detection Rate : 0.4667           
Detection Prevalence : 0.6111           
Balanced Accuracy : 0.8088           
'Positive' Class : 0                                   
vii) 
tree_accuracy <- confusion_matrix_tree$overall['Accuracy'] 
cat("Decision Tree Accuracy:", tree_accuracy, "\n") 
rf_accuracy_1 <- confusionMatrix(pred_rf_1, test_data$target)$overall['Accuracy'] 
rf_accuracy_2 <- confusionMatrix(pred_rf_2, test_data$target)$overall['Accuracy'] 
cat("Random Forest Accuracy mtry 13:", rf_accuracy_1, "\n") 
cat("Random Forest Accuracy default mtry:", rf_accuracy_2, "\n") 
Decision Tree Accuracy: 0.7666667 
Random Forest Accuracy mtry 13: 0.8 
Random Forest Accuracy default mtry: 0.8111111  
(4) 
The accuracy for the decision tree model is the least, however Random Forest models have 
better accuracy in this particular scenario. 
Among the 2 random forest models, the model with default value of mtry has better accuracy 
than the model which uses all the parameters. 
Using all predictor variables while fitting Random Forest is bagging. While using 1 predictor 
variable will overfit the model. Here, Default value (13/3 ~ rounded off to 4) chooses the 4 
predictor variables at random and fits the model to give the output. Thus, this increases the 
accuracy and at the same time does not over fits. 
(3) 
Page 6 of 19 
IAI                                                                                                                                                  
CS2B-1124 
viii) 
Decision Trees:  
a) Prone to high variance,  
b) Sensitive to data,  
c) Can easily overfit due to their flexibility and  
d) Ability to capture detailed patterns (including noise) in the training data. 
Random Forest:  
a) Combines multiple randomized trees and combines their predictions,  
b) This reduces variance and thus overfitting by diluting the impact of noise and over
specific patterns learned by individual trees.  
c) The randomness in feature selection and bootstrapped data sampling ensures diversity 
in the trees 
(3) 
ix) 
ROC Curve 3 has the highest AUC of 0.99. This is a measure of model performance.  
Also, ROC Curve 3 has strictly higher True Positives rates for false positives rates at 5% 
indicating high predictive power. 
Thus, ROC Curve 3 indicates best model performance among the given three, due to the above 
stated reasons.  
AUC for random classifier is 0.5 
AUC for perfect classifier is 1. 
