---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)

options(scipen = 2, digits = 3) #set to two decimal 
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

## Exercise 1

Monthly totals of airline passengers are given in time series data ‘***AirPassengers***’.

Plot the data as time series with frequency =12.

---

### AirPassengers 

Monthly Airline Passenger Numbers 1949-1960

#### Description
The classic Box & Jenkins airline data. Monthly totals of international airline passengers, 1949 to 1960.

#### Usage
```{r eval=FALSE}
AirPassengers
```

#### Format
A monthly time series, in thousands.

#### Source

Box, G. E. P., Jenkins, G. M. and Reinsel, G. C. (1976) Time Series Analysis, Forecasting and Control. Third Edition. Holden-Day. Series G.

---

#### AirPassengers

```{r}
AirPassengers
```


---

```{r eval =FALSE}
apts <- ts(AirPassengers, frequency=12)

plot(apts)
```


---

```{r echo=FALSE}
apts <- ts(AirPassengers, frequency=12)

plot(apts)
```


---

## Exercise 2

Plot the acf function.



```{r eval=FALSE}

acf(apts)
```


---


```{r echo=FALSE}

acf(apts)
```


---


## Exercise 3

Decompose the trend, seasonality and randomness for the data.

(Hint: use  ***decompose()*** function.)



```{r eval=FALSE}
f <- decompose(apts)
plot(f)

```

---

```{r echo=FALSE}
f <- decompose(apts)
plot(f)

```

---

## Exercise 4

Fit Arima model.
(Hint: order=c(1,0,0), list(order=c(2,1,0), period=12 as parameters.)



```{r}
fit <- arima(AirPassengers, 
             order=c(1,0,0), 
             list(order=c(2,1,0), period=12))
fit
```

---


```{r}
class(fit)
```




```{r}
summary(fit)
```

---

## Exercise 5

Predict the data for next 24 months.



```{r}
# "fore" for forecast

fore <- predict(fit, n.ahead=24)

fore
```

---


## Exercise 6

Find upper and lower bound of fitted data and plot the same along with predicted data with legends.

```{r}
U <- fore$pred + 2*fore$se 
L <- fore$pred - 2*fore$se 

c(U,L)
```

```{r eval=FALSE}
ts.plot(AirPassengers, fore$pred, U, L, 
        col=c(1,2,4,4), 
        lty = c(1,1,2,2))

legend("topleft", 
       c("Actual", "Forecast", "Error Bounds (95% Confidence)"),
       col=c(1,2,4), 
       lty=c(1,1,2))
```

---


```{r echo=FALSE}
ts.plot(AirPassengers, fore$pred, U, L, 
        col=c(1,2,4,4), 
        lty = c(1,1,2,2))

legend("topleft", 
       c("Actual", "Forecast", "Error Bounds (95% Confidence)"),
       col=c(1,2,4), 
       lty=c(1,1,2))
```

---