---
title: "Risk Modelling and Survival Analysis with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Risk Modelling and Survival Analysis with R


#### CS2B Curriculum

* It develops knowledge of and the ability to apply statistical methods for risk modelling, time series analysis methods, stochastic processes (especially Markov chains and Markov jump processes), survival analysis (including regression methods applied to duration data), and graduation methods  It also includes a high-level introduction to machine learning. 

* Question 2 Part (a)
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

```{r echo=FALSE, include=FALSE}
library(magrittr)


options(digits=4)
```


---
Q. 4)

An life insurance company is selling Endowment, Pure Term and Unit Linked products
wherein 80% of the business comes from Endowment, 5% from Term Insurance and 15%
from Unit Linked products. The sum assured ranges between 30000 to 200000 with multiple
of 5000 . The sum assured are normally distributed with below parameters - Endowment
(mean 50000, standard deviation 10000), Pure Term (mean 100000, standard deviation
25000), Unit Linked (mean 60000, standard deviation 5000). One policy is randomly picked
from the system and the sum assured of that policy is 75000.
---

1. 
Create a sequence ranging from 30000 to 200000 with an increase of 5000 and plot the
density functions of the above three distributions in a single graph.
2.  Find the probability that the policy (75000 sum assured) comes from the Endowment
product portfolio and the probability that the policy comes from Unit Linked Product
portfolio.


---



```


```{r}

Solution 4:

[7 Marks]
1. 
y<-seq(30000,200000,5000)
y
FE<-dnorm(y,50000,10000)
FPT<-dnorm(y,100000,25000)
FUL<-dnorm(y,60000,5000)
plot(y,FE,typ="l",col="darkgreen",lwd=2)
lines(y,FPT,col="darkred",lwd=2)
lines(y,FUL,col="darkblue",lwd=2)

```


```{r}

2. 
PE<-0.8
PPT<-0.05
PUL<-0.15
PxE<-dnorm(75000,50000,10000)
PxPT<-dnorm(75000,100000,25000)


```

---

```{r}

PxUL<-dnorm(75000,60000,5000)
P1<-PE*PxE / (PE*PxE+PPT*PxPT+PUL*PxUL)
P2<-PUL*PxUL / (PE*PxE+PPT*PxPT+PUL*PxUL)
P1
P2


```
