---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


1 Before answering this question, the following R code should be run:
```{r}
set.seed(912)
y=arima.sim(list(order=c(0,1,0)),n=400)
fit=arima(y,order=c(1,0,0))
fit
```

(i) Comment briefly, in your own words, on each line of R code above. 
(ii) (a) State the standard error of the ar1 parameter estimate in the fit object
created by the R code above.
(b) Determine the corresponding 95% confidence interval.

(iii) Comment on your answer to part (ii). 
(iv) Calculate the predicted values using the model fit, the future values of y for
ten steps ahead. 
(v) Generate, and display in your answer script, a matrix, A, of dimension
10 × 2, which contains the predicted values in part (iv) together with the
corresponding standard errors. 
(vi) Construct R code to generate a plot that contains the time series data
y, together with the ‘ten steps ahead’ predictions from part (iv) and their 95%
prediction intervals. 
(vii) Construct R code to display, next to each other, the sample AutoCorrelation
Function (sample ACF) and sample Partial AutoCorrelation Function (sample
PACF) for the data set y. 
(viii) Construct R code to display, next to each other, the sample ACF and sample
PACF for the residuals of the model fit. 
(ix) Comment on the graphical output of parts (vii) and (viii). 
(x) Perform the Ljung and Box portmanteau test for the residuals of the model
fit with four, six and twelve lags. 
(xi) Comment, based on your answers to parts (ix) and (x), on whether there is
enough evidence to conclude that the model fit is appropriate. 

