---
title: "Statistics with R"
subtitle: "Modelling Data with R"
author: "Dragonfly Stats"
date: "2016/12/12"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---




Q. 2) The file “Data.csv” contains records of premium received and claim paid between certain 
time period. The ‘Day’ information represents the date on which the transaction occurred, 
calculated from the first transaction date. 
i) Separate the data into two tables – premium and claim, based on the transaction type 
and calculate 
a) mean premium; 
b) mean claim amount; 
c) standard deviation of premium; and 
d) standard deviation of claim amount. 

---

ii) It is assumed that the waiting times between consecutive events follow a Poisson 
process. Expand the ‘claim’ table to include a column for observed waiting time (in 
days) for all the claims and calculate a Poisson parameter value. (7)


---
 Solution 2: 
i)
data<-read.table("Data.csv",sep=",",header=T) 
premium <- data[data$Type=="Premium",] 
premium IAI CS2B-0321 
Page 5 of 9
> premium 
 Day Amount Type 
1 1 934 Premium 
5 3 1563 Premium 
7 8 7568 Premium 
9 11 1569 Premium 
11 14 5107 Premium 
…………………… 
…………………… 
188 349 5559 Premium 
190 350 2366 Premium 
191 352 5517 Premium 
192 353 5371 Premium 
 [2] 
 
 ---
 
mean.premium=mean(premium[,2]) 
mean.premium 
> mean.premium 
[1] 5189.613 
 [1] 
sd.premium=sqrt(var(premium[,2])) 
sd.premium 
> sd.premium 
[1] 2878.675 
 [1]
claim <- data[data$Type=="Claims",] 

---

```{r}
head(claim )
```
> claim 
 Day Amount Type 
2 1 160854 Claims 
3 1 35968 Claims 
4 2 376784 Claims 
6 5 259155 Claims 
8 8 60865 Claims 

--- 

```{r}
mean.claim=mean(claim[,2]) 
mean.claim 
```

> mean.claim 
[1] 259644.1 
 [1] 

```{r}
sd.claim=sqrt(var(claim[,2])) 
sd.claim 
```
 
> sd.claim 
[1] 155027 
 [1] 

---

```{r}
temp <- cbind(claim,c(NA,diff(claim[,1]))) IAI CS2B-0321 

colnames(temp)[4] <- "Diff" 
claim=temp 
claim 
```

---

```{r}
mean.wait.time=mean(claim[,4][2:length(claim[,4])]) 
Poiss.param.claims=1/mean.wait.time 
Poiss.param.claims 
> Poiss.param.claims 
[1] 0.3314763 
```