---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


Q. 5) A primary school has two wings A and B. The students can change their Wing freely. It was 
observed that the rate of movement is dependent on the age ‘x’ of the student (ranging from 
5 to 15). The rate of movement from A to B = [0.08*x – 0.2] and the rate of movement from 
B to A = [0.05 * x + 0.1]. 

---

i) Write a code to calculate the generator matrix G(x). Calculate the rate at which a 12 
year old student will move from wing B to A and from A to B. 


--

Solution 5: 
i)
```{r}
sigmaAB=function(x){ 
 rate=0.08*x-0.2
 rate} 
```

```{r}
sigmaBA=function(x){ 
 rate=0.05*x+0.1 
 rate} 
``` 

---

```{r}
gen.matrix=function(x){ 
 muAB=sigmaAB(x) 
 muBA=sigmaBA(x) 
 muAA=-muAB 
 muBB=-muBA 
 movement.rates = c(muAA,muBA,muAB,muBB) 
 X =(matrix(movement.rates, 2, 2)) 
 X} 
 
``` 

---

```{r}
gen.matrix(12) 
``` 

---
#### Exercise 2

Over a very small time period h, the transition probability matrix P(t,t+h) can be 
approximated as P(t,t+h) = I + h*G(t); I is the identity matrix. 

Using this, calculate the 
probability that a 8 year old will move to Wing B in next 5 years in case h=1/365. 

---

```{r}
calc.prob.matrix=function(s,t,h){ 
 Ph=diag(2)+gen.matrix(s)*h 
 temp.matrix=Ph 
 for (j in 1:((t-s)/h-1)){ 
 temp.matrix=temp.matrix%*%Ph 
 Ph=diag(2)+gen.matrix(s+h*j)*h 
 } 
 temp.matrix 
} 
```

---

```{r}
calc.prob.matrix(8,5,1/365) 
```

The rate of moving to Wing B = 0.3682517. 

---