---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

* This presentation features Time Series Analysis, and is split into two parts

---


### Part A

The data file “***Expenses.csv***” contains records of last 5 years monthly expenses paid by a 
company starting from January 2015. 

### Exercises

1. Import the data and classify it as a time series object, specifying the relevant dates. 

2. Assuming that the data shows yearly seasonal variation, remove the seasonality from 
the data using seasonal differencing. 

3. Plot a graph showing the actual data, the trend line, the trend + seasonal line and any 
remaining un-trended number. Plot them in the same graph and use different colour 
lines for each of the series. 


---

### Exercise 1

```{r}
expenses <- read.csv("Expenses.csv") 

class( expenses)

dim(expenses)
```

```{r}
expenses <- ts(expenses ,frequency=12,start=c(2015,1)) 

class( expenses)
```

---

```{r}
expenses 
```
---

```{r}
plot(expenses) 
```

---

```{r}
decomposed <-(decompose(expenses,type="additive")) 

plot(decomposed)

```

---

```{r}
trend <- decomposed$trend 

seasonal <- decomposed$seasonal 

random <- decomposed$random 
```

---


```{r}

random 
```

---


```{r}

plot(random )
```

---

### Exercise 3

```{r eval=FALSE}
ts.plot(expenses,ylab="",main="Components of time series ", 
        col="dark grey") 

lines(trend,col="red") 
lines(seasonal+trend,col="blue") 
lines(random,col="yellow") 
```


---

```{r echo =FALSE}
ts.plot(expenses,ylab="",main="Components of time series ", 
        col="dark grey") 

lines(trend,col="red") 
lines(seasonal+trend,col="blue") 
lines(random,col="yellow") 
```
---


### Exercise 3

```{r eval=FALSE}
ts.plot(expenses,ylab="",main="Components of time series ", 
        col="dark grey") 
lines(seasonal+trend,col="green")
```


---

### Trend and Seasonal Component

```{r eval =FALSE}
ts.plot(seasonal,ylab="",
        main="Components of time series ", 
        col="grey",
        ylim=c(0,1300)) 
lines(trend,col="blue")  
```


---

```{r echo =FALSE}
ts.plot(seasonal,ylab="",
        main="Components of time series ", 
        col="grey",
        ylim=c(0,1300)) 
lines(trend,col="blue")  
```

---

```{r eval=FALSE}
ts.plot(random,ylab="",main="Components of time series ", 
        col="dark green") 

```

---

```{r echo =FALSE}
ts.plot(random,ylab="",main="Components of time series ", 
        col="dark green") 

```

---

### Part B
Import the time series data file “ARIMA.csv” and answer the below questions. 

#### Exercises

4. Plot a graph for the above time series data “***ARIMA.csv***”. 

5. Plot two graphs to show the ACF and PACF for the time series provided. 

6. Comment on the stationarity of the data based on the above graphs. 

---
  
### Exercise 1

```{r}
ARIMAdata<-read.csv("ARIMA.csv") 
```

```{r}
head(ARIMAdata)
```

---
  
### Exercise 1

```{r}
ARIMAdata<-read.csv("ARIMA.csv",header=FALSE) 
```

```{r}
head(ARIMAdata)
```

---

```{r}
class(ARIMAdata)
```


```{r}
dim(ARIMAdata)
```
---


```{r}
ts.plot(ARIMAdata,main="Time series data",ylab="Data") 
```

---
 
### Exercise 2

```{r eval = FALSE}
 acf(ARIMAdata, main="Time series data ",ylab="sample ACF") 
```

---

```{r echo=FALSE}
 acf(ARIMAdata, main="Time series data ",ylab="sample ACF") 
```

---

```{r eval = FALSE} 
 pacf(ARIMAdata, 
      main=" Time series data",
	  
	  ylab="sample PACF") 
```
---

```{r, echo = FALSE} 
 pacf(ARIMAdata, 
      main=" Time series data",
	  
	  ylab="sample PACF") 
```
---
### Exercise 3

The ACF drops to zero relatively quickly, denoting the series is stationary. 

---

---