---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


The file “***InsuranceData.csv***” contains records of premium received and claim paid between certain 
time period. 
* The ‘Day’ information represents the date on which the transaction occurred, 
calculated from the first transaction date. 

#### Exercise 1

Separate the data into two tables – premium and claim, based on the transaction type 
and calculate 

1. mean premium; 
2. mean claim amount; 
3. standard deviation of premium; and 
4. standard deviation of claim amount. 

---

#### Exercise 2

It is assumed that the waiting times between consecutive events follow a Poisson 
process. 

Expand the ‘claim’ table to include a column for observed waiting time (in 
days) for all the claims and estimate a Poisson rate parameter value. 



---


```{r}

InsuranceData <- read.csv("InsuranceData.csv")

head(InsuranceData)
```

---


```{r warning=FALSE}

library(magrittr)
library(janitor)

InsuranceData %>% tabyl(Type)

```

---

```{r}
premium <- InsuranceData[InsuranceData$Type=="Premium",] 
```

```{r}
summary(premium)
```

---

#### Premiums Claims

```{r} 
mean.premium=mean(premium[,2]) 
mean.premium 
```

```{r}
sd.premium=sd(premium[,2]) 
sd.premium 
```
 
---

####  Claims Data
 
```{r} 
claims <- InsuranceData[InsuranceData$Type=="Claims",] 

```


```{r}
head(claims )
```
 

---

#### Mean and Variance of Claims

```{r}
mean.claim=mean(claims[,2]) 
mean.claim 
```

```{r}
sd.claim= sd(claims[,2]) 
sd.claim 
```
 
---


```{r}
# using "tidyverse"

# library(dplyr)

InsuranceData %>% group_by(Type) %>%
  summarize(mean.value = mean(Amount),
            std.dev = sd(Amount)
            )


```

---

### Computing the Waiting Times

* Time Differences

```{r}
waiting_time <- diff(claims[,1]) 

waiting_time

```


---

Check that dimensions fit

```{r}
length(waiting_time)
```


```{r}
nrow(claims)
```

---


```{r}
head(claims)
```



```{r}
waiting_time <- c(NA,waiting_time)

length(waiting_time)

```
 

---

#### Update Claims Data

```{r}
claims <- cbind(claims,waiting_time)

head(claims) 
```


```{r}
tail(claims)
```

---

#### Mean Time Difference


```{r}
mean.wait.time <- mean(claims$waiting_time,na.rm=TRUE)
mean.wait.time

```

#### Estimate Poisson Rate Parameter

```{r}
Poiss.param.claims <- 1/mean.wait.time 

Poiss.param.claims 


```

---


---