---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---


The force of mortality ‘mu(x)’ for a particular group of life can be calculated using the 
formula as `A*B^x`; where 

* `A=0.00002`
* `B=exp(c)` 
* `c=0.09`
* x denotes the age of the life.

---

#### Exercises

1. Write down a formula for mu(x) in R and calculate the value of mu(50). 

2. If survival probability ***px*** can be approximated as `[exp(-mu(x+0.5))]`, calculate the 
probability that a life aged 50, will remain alive after 10 years. 

3. Create a table of mortality rate ***qx*** for ages between 40 to 50. 

---

```{r}
A <- 0.00002 
C <- 0.09 
B <- exp(C) 

mu<-function(x) { 
 A * B^x 
} 

mu(50) 
```

---

```{r}
px<-function(x) { 
 exp(-mu(x+0.5)) 
} 

 
```

Ten year range starting from 50

```{r}
x <- 50 

t <- 1:10

x+t-1
```

---

```{r}

x <- 50 
npx <- 1


for(t in 1:10){
  alive=px(x+t-1);
  
  npx=npx*alive

  } 
npx 


```
---

```{r}

for(t in 1:10){
  alive=px(x+t-1);
  
  npx=npx*alive
  cat("\nAge:",x+t-1,"\tFor year:", alive,"\tCumulative:",npx)
  } 



```
---

```{r}


qx<-function(x) { 
 1-px(x) 
} 

```

```{r}

qx(40:50)
```

---




```{r}
cbind(40:50,qx(40:50))
```


---

```{r}
xlist <- seq(40,50) 

qlist <- sapply(xlist,qx) 
```


```{r}
cbind(xlist,qlist) 
```


