---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

#### Exercise

Q. 5) A primary school has two wings A and B. The students can change their Wing freely. It was 
observed that the rate of movement is dependent on the age ‘x’ of the student (ranging from 
5 to 15). 

* The rate of movement from A to B = [0.08*x – 0.2] 
* The rate of movement from B to A = [0.05 * x + 0.1]. 

---

#### Exercise 1 

Write a code to calculate the ***generator matrix*** G(x). 
Calculate the rate at which a 12 year old student will move from wing B to A and from A to B. 


* write a function called <tt>gen.matrix()</tt>

---


**Understanding the Problem and Generator Matrix**

In this problem, we have a two-state system (wings A and B) where the transition rates between the states depend on the age of the student. 
A generator matrix, G(x), is a matrix that captures these transition rates for a given age x.

**Constructing the Generator Matrix**

Let's denote the state of being in wing A as state 1 and the state of being in wing B as state 2. Then, the generator matrix G(x) can be defined as:

```
G(x) = [[-q11(x)  q12(x)]
        [ q21(x) -q22(x)]]
```

Where:

* `q11(x)`: Rate of leaving state 1 (wing A) = Rate of moving from A to B = 0.08x - 0.2
* `q12(x)`: Rate of moving from state 1 (wing A) to state 2 (wing B) = 0.08x - 0.2
* `q21(x)`: Rate of moving from state 2 (wing B) to state 1 (wing A) = 0.05x + 0.1
* `q22(x)`: Rate of leaving state 2 (wing B) = Rate of moving from B to A = 0.05x + 0.1

**Calculating Rates for a 12-year-old Student**

For a 12-year-old student (x = 12):

* Rate of movement from A to B: 0.08 * 12 - 0.2 = 0.76
* Rate of movement from B to A: 0.05 * 12 + 0.1 = 0.7

---

**R Code Implementation**

```R
generator_matrix <- function(x) {
  q11 <- 0.08 * x - 0.2
  q12 <- 0.08 * x - 0.2
  q21 <- 0.05 * x + 0.1
  q22 <- 0.05 * x + 0.1

  matrix(c(-q11, q12, q21, -q22), nrow = 2, byrow = TRUE)
}

# For a 12-year-old student:
G_12 <- generator_matrix(12)
print(G_12)
```

This R code will output the generator matrix for a 12-year-old student, which can be further used to analyze the dynamics of student movement between the two wings.


--

Solution 5: 
i)
```{r}
sigmaAB=function(x){ 
 rate=0.08*x-0.2
 rate} 
```

```{r}
sigmaBA=function(x){ 
 rate=0.05*x+0.1 
 rate} 
``` 

---

```{r}
gen.matrix=function(x){ 
 muAB = sigmaAB(x) 
 muBA = sigmaBA(x) 
 muAA = -muAB 
 muBB = -muBA 
 movement.rates = c(muAA,muBA,muAB,muBB) 
 X =(matrix(movement.rates, 2, 2)) 
 X
 } 
 
``` 

---

```{r}
gen.matrix(12) 
``` 

---
#### Exercise 2

* Over a very small time period $h$, the transition probability matrix P(t,t+h) can be 
approximated as P(t,t+h) = I + h*G(t); I is the identity matrix. 

* Using this, calculate the 
probability that a 8 year old will move to Wing B in next 5 years in case $h=1/365$. 

---

#### Exercise 2
```{r}
calc.prob.matrix=function(s,t,h){ 
 
  Ph=diag(2)+gen.matrix(s)*h 
  temp.matrix=Ph 
 
  for (j in 1:((t-s)/h-1)){ 
    temp.matrix = temp.matrix %*% Ph 
    Ph = diag(2) + gen.matrix(s+h*j)*h 
    } 
  
  temp.matrix 
} 
```

---

```{r}
calc.prob.matrix(8,5,1/365) 
```

The rate of moving to Wing B = 0.3682517. 

---

sigmaBA=function(x){  
  rate=0.05*x+0.1 
  rate}  
 
gen.matrix=function(x){  
  muAB=sigmaAB(x)  
  muBA=sigmaBA(x)  
  muAA=-muAB 
  muBB=-muBA 
  movement.rates = c(muAA,muBA,muAB,muBB) 
  X =(matrix(movement.rates, 2, 2)) 
  X} 
gen.matrix(12) 
 
> gen.matrix(12) 
      [,1]  [,2] 
[1,] -0.76  0.76 
[2,]  0.70 -0.70 
 
Thus, the rates are 0.70 (from B to A) and 0.76 (from A to B). 
             [8] 
 
ii)  
calc.prob.matrix=function(s,t,h){  
  Ph=diag(2)+gen.matrix(s)*h  
  temp.matrix=Ph  
  for (j in 1:((t-s)/h-1)){  
    temp.matrix=temp.matrix%*%Ph  
    Ph=diag(2)+gen.matrix(s+h*j)*h  
  }  
  temp.matrix  
} 
calc.prob.matrix(8,5,1/365) 
 
> calc.prob.matrix(8,5,1/365) 
          [,1]      [,2] 
[1,] 0.6317483 0.3682517 
[2,] 0.5260447 0.4739553 
 
The rate of moving to Wing B = 0.3682517. 
---

To calculate the generator matrix \( G(x) \) and the transition probabilities for movement between the wings for a student of age \( x \), let's break it down into smaller steps and provide the necessary R code.

### Step 1: Calculating the Generator Matrix \( G(x) \)
The generator matrix \( G(x) \) is defined as:
\[ G(x) = \begin{bmatrix} 
-q_{11}(x) & q_{12}(x) \\
q_{21}(x) & -q_{22}(x) 
\end{bmatrix} \]

Where:
- \( q_{11}(x) = q_{12}(x) = 0.08x - 0.2 \)
- \( q_{21}(x) = q_{22}(x) = 0.05x + 0.1 \)

#### R Code to Calculate Generator Matrix \( G(x) \)
```R
sigmaAB <- function(x) { 
  rate <- 0.08 * x - 0.2
  return(rate)
} 

sigmaBA <- function(x) { 
  rate <- 0.05 * x + 0.1 
  return(rate)
} 

gen.matrix <- function(x) { 
  muAB <- sigmaAB(x)
  muBA <- sigmaBA(x)
  muAA <- -muAB
  muBB <- -muBA
  movement.rates <- c(muAA, muBA, muAB, muBB)
  X <- matrix(movement.rates, nrow = 2, byrow = TRUE)
  return(X)
} 

# Calculate the generator matrix for a 12-year-old student
gen.matrix(12)
```

### Step 2: Calculating Transition Probabilities Over Time
Using the generator matrix \( G(t) \), we can approximate the transition probability matrix \( P(t,t+h) \) over a very small time period \( h \) as:
\[ P(t,t+h) \approx I + h \cdot G(t) \]

#### R Code to Calculate the Transition Probability Matrix
```R
calc.prob.matrix <- function(s, t, h) { 
  Ph <- diag(2) + gen.matrix(s) * h 
  temp.matrix <- Ph 

  for (j in 1:((t - s) / h - 1)) { 
    temp.matrix <- temp.matrix %*% Ph 
    Ph <- diag(2) + gen.matrix(s + h * j) * h 
  } 

  return(temp.matrix)
} 

# Calculate the transition probability matrix for an 8-year-old over 5 years
calc.prob.matrix(8, 5, 1/365)
```

### Explanation of Results
1. **Generator Matrix for a 12-year-old**:
   This matrix captures the transition rates between wings A and B for a 12-year-old student. The rates are:
   - From A to B: \( 0.76 \)
   - From B to A: \( 0.7 \)

2. **Transition Probability Matrix for an 8-year-old over 5 years**:
   This matrix provides the probabilities of moving between wings A and B over a 5-year period for an 8-year-old student, given a small time interval \( h = 1/365 \).

I hope this helps! If you have any further questions or need more assistance, feel free to ask.