---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

The Claim size $X$ for a class of insurance follows a Gamma (10,0.2) distribution. 

#### Exercises

1. Calculate $f{30}$. 
2. Calculate the median of the function. 

---

### Gamma Distribution

#### Parameters	

* shape: $\alpha > 0$  (code as "a")
* rate: $\beta > 0$   (code as "b")

PDF
$${\displaystyle f(x)={\frac {\beta ^{\alpha }}{\Gamma (\alpha )}}x^{\alpha -1}e^{-\beta x}}$$
CDF	
$${\displaystyle F(x)={\frac {1}{\Gamma (\alpha )}}\gamma (\alpha ,\beta x)}$$

---


#### Gamma Distribution

Mean		
$${\displaystyle {\frac {\alpha }{\beta }}}$$


Variance	
$${\displaystyle {\frac {\alpha }{\beta ^{2}}}}$$




---

```{r}
a <- 10 
b <- 0.2 

x <- 30 

```

```{r}
dgamma(x,shape = a, rate= b) 
```

---



```{r}

### rate and shape are the default arguments
dgamma(x,a, b) 
```
```{r}

### can specify rate and scale
dgamma(x,shape = a, scale = 1/b) 
```



---

### Exercise 2

We want to find M such that $P(X <= M) = 0.5$ 

```{r}

qgamma(0.5,a,b) 
```

---

### Exercise 3

A quota share reinsurance is in place with 75% retained proportion. 

Estimate the mean and variance of the amount paid by the re-insurer in respect of a single claim. 

#### Remark 

- Re-insurer pays 25% of claim 

- we will denote the amount paid by the insurer as $Y$
 

 
---
 
```{r}
r <- 0.75 

ExpX = a/b 
ExpX

VarX = a/b^2 
VarX
```


---

#### Expected Value

```{r}
  
ExpY <- (1-r)*ExpX 
```

```{r}
ExpY 
```

---

#### Variance 


Recall : $\operatorname{Var}(aX)= a^2\operatorname{Var}(X)$
```{r}
VarY <- (1-r)^2*VarX 
```



```{r}
VarY 
```


---
[use seed = 250 for below 3 questions]. 


### Exercise 4

The aggregate claim follows a compound distribution with below parameters: 

* Number of claims follow Poisson distribution with parameter 500 
* Individual claim size follows a Gamma distribution with parameters (600,0.3) 


Generate 1000 observations and calculate the 500th simulated value. 


---

```{r}
set.seed(250) 

n <- rpois(1000,500) 
s <- numeric(1000) 

for(i in 1:1000) {
  x <- rgamma(n[i],shape=600,rate=0.3) 
  s[i] <- sum(x)
}

s[500] 


```
 
---
 
### Exercise 5
The insurer wants to limit its claim liability and hence has modified the terms of the 
policy. It would now only pay for any amount in excess of an amount 500 on each 
claim. 

Calculate the 500th simulated value for the insurer under the revised condition. 


---

### Exercise 5

```{r}
set.seed(250) 
M <-500 
n <- rpois(1000,500) 
s <- numeric(1000) 
```

```{r}
for(i in 1:1000) {
  x <- rgamma(n[i],shape=600,rate=0.3) 
  z <- pmax(0,x-M) 
  s[i] <- sum(z)
}

s[500] 
```
 
---



### Exercise 6

The insurer wants to analyze another claim sharing option. Under this, insurer will pay 
75% of the claim amount and the rest 25% will be borne by the policyholder. 

Calculate the 500th simulated value for the insurer under this option. 


---

```{r}
set.seed(250) 
M <- 0.25 
n <- rpois(1000,500) 
s <- numeric(1000) 

for(i in 1:1000) {
  x <- rgamma(n[i],shape=600,rate=0.3) 
  z <- (1-M)* x 
  s[i] <- sum(z)
  } 

s[500] 

```

---



---