---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

#### Exercise

Q. 5) A primary school has two wings A and B. The students can change their Wing freely. It was 
observed that the rate of movement is dependent on the age ‘x’ of the student (ranging from 
5 to 15). 

* The rate of movement from A to B = [0.08*x – 0.2] 
* The rate of movement from B to A = [0.05 * x + 0.1]. 

---

#### Exercise 1 

Write a code to calculate the ***generator matrix*** G(x). 
Calculate the rate at which a 12 year old student will move from wing B to A and from A to B. 


* write a function called <tt>gen.matrix()</tt>

---


**Understanding the Problem and Generator Matrix**

In this problem, we have a two-state system (wings A and B) where the transition rates between the states depend on the age of the student. A generator matrix, G(x), is a matrix that captures these transition rates for a given age x.

**Constructing the Generator Matrix**

Let's denote the state of being in wing A as state 1 and the state of being in wing B as state 2. Then, the generator matrix G(x) can be defined as:

```
G(x) = [[-q11(x)  q12(x)]
        [ q21(x) -q22(x)]]
```

Where:

* `q11(x)`: Rate of leaving state 1 (wing A) = Rate of moving from A to B = 0.08x - 0.2
* `q12(x)`: Rate of moving from state 1 (wing A) to state 2 (wing B) = 0.08x - 0.2
* `q21(x)`: Rate of moving from state 2 (wing B) to state 1 (wing A) = 0.05x + 0.1
* `q22(x)`: Rate of leaving state 2 (wing B) = Rate of moving from B to A = 0.05x + 0.1

**Calculating Rates for a 12-year-old Student**

For a 12-year-old student (x = 12):

* Rate of movement from A to B: 0.08 * 12 - 0.2 = 0.76
* Rate of movement from B to A: 0.05 * 12 + 0.1 = 0.7

---

**R Code Implementation**

```R
generator_matrix <- function(x) {
  q11 <- 0.08 * x - 0.2
  q12 <- 0.08 * x - 0.2
  q21 <- 0.05 * x + 0.1
  q22 <- 0.05 * x + 0.1

  matrix(c(-q11, q12, q21, -q22), nrow = 2, byrow = TRUE)
}

# For a 12-year-old student:
G_12 <- generator_matrix(12)
print(G_12)
```

This R code will output the generator matrix for a 12-year-old student, which can be further used to analyze the dynamics of student movement between the two wings.


--

Solution 5: 
i)
```{r}
sigmaAB=function(x){ 
 rate=0.08*x-0.2
 rate} 
```

```{r}
sigmaBA=function(x){ 
 rate=0.05*x+0.1 
 rate} 
``` 

---

```{r}
gen.matrix=function(x){ 
 muAB = sigmaAB(x) 
 muBA = sigmaBA(x) 
 muAA = -muAB 
 muBB = -muBA 
 movement.rates = c(muAA,muBA,muAB,muBB) 
 X =(matrix(movement.rates, 2, 2)) 
 X
 } 
 
``` 

---

```{r}
gen.matrix(12) 
``` 

---
#### Exercise 2

* Over a very small time period $h$, the transition probability matrix P(t,t+h) can be 
approximated as P(t,t+h) = I + h*G(t); I is the identity matrix. 

* Using this, calculate the 
probability that a 8 year old will move to Wing B in next 5 years in case $h=1/365$. 

---

#### Exercise 2
```{r}
calc.prob.matrix=function(s,t,h){ 
 
  Ph=diag(2)+gen.matrix(s)*h 
  temp.matrix=Ph 
 
  for (j in 1:((t-s)/h-1)){ 
    temp.matrix = temp.matrix %*% Ph 
    Ph = diag(2) + gen.matrix(s+h*j)*h 
    } 
  
  temp.matrix 
} 
```

---

```{r}
calc.prob.matrix(8,5,1/365) 
```

The rate of moving to Wing B = 0.3682517. 

---


---