
[Total 10]4
Data were collected on average alcohol and cigarette consumption per adult per year
for a number of countries. The data are given in the file smoking_data.RData
and contain the following information:
country: the country concerned;
alcohol: alcohol consumption per adult per year (litres/year);
cigarettes: number of cigarettes consumed per adult per year.
(i)
(a)
(b)
(ii)
Construct a scatterplot of the data with alcohol consumption on the
x axis.
Comment on the relationship between the data on alcohol and cigarette
consumption based on your plot in part (i)(a).

Calculate Pearson’s correlation coefficient between the data on alcohol and
cigarette consumption.

An analyst suggests using the following R code to modify the data:
```{r}
alcohol.2 = alcohol[-c(6,16)]
cigarettes.2 = cigarettes[-c(6,16)]
```
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(i) (a)

```{r}
load("smoking_data.RData")
plot(alcohol, cigarettes, pch = 19, main="Alcohol versus
Cigarettes",
xlab = "Alcohol per adult (litres/year)",
ylab = "Cigarettes per adult per year")
```


(b) There is a moderate positive association between average alcohol consumption
and average cigarette consumption.

The overall shape of the scatterplot is roughly linear

with the exception of two points that indicate higher average cigarette consumption
relative to alcohol consumption than the other countries in this sample.

(ii) r1 = cor(alcohol, cigarettes, method="pearson")
Pearson’s sample correlation coefficient is equal to 0.485.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%





(iii)
Explain what the above code does and give a justification for its use.

For the remainder of the question, use the modified data
(alcohol.2, cigarettes.2), as produced by applying the R
code above.
(iv)
(a)
Construct a scatterplot with alcohol consumption on the x axis.
(b)
Calculate Pearson’s correlation coefficient between the new data on
alcohol and cigarette consumption.
(c)
Comment on your answers in parts (ii) and (iv)(b).
[6]
(v)
Perform a hypothesis test for the null hypothesis that Pearson’s population
correlation coefficient is equal to zero, against the alternative that it is positive.
You should report the p-value of the test and a clear conclusion.

A media bulletin has reported that ‘ ‘' h‘igher alcohol consumption causes higher
cigarette consumption ’ .
(vi)
Comment on whether this report is justified based on your analysis in parts (iv)
and (v).

CS1B S2020–4(vii)
(viii)
(ix)
(x)


Perform a simple linear regression analysis on the new data using a
model of the form y = α + βx + ε (cigarette consumption, y, on alcohol
consumption, x), where the error terms ε independently follow a
N(0, s 2 ) distribution.
		Your answer should show the fitted line plotted on the data scatterplot
and report the estimate of parameter s. 
State the proportion of the total variability of the responses explained
by the model, based on your output in part (vii). 
Plot a graph of the residuals of the model fitted in part (vii)
against the explanatory variable. 
Comment on the validity of the model, based on your output in
part (ix) .



Q4


(iii)
alcohol.2 = alcohol[-c(6,16)]
cigarettes.2 = cigarettes[-c(6,16)]
The code removes the data in the 6 th and 16 th element of the data vectors.

These correspond to the two countries that produce the largest deviations from the overall
pattern of the plot (outliers): Greece (10.4, 8.3) and USA (4.7, 4.9).

(iv) (a)
plot(alcohol.2, cigarettes.2, pch = 19, main="Alcohol
versus Cigarettes
(GR,
USA omitted)", xlab = "Alcohol per adult
(litres/year)", ylab =
------------------------------------

"Cigarettes per adult per year")

(b) r2 = cor(alcohol.2, cigarettes.2, method="pearson")
Pearson’s sample correlation coefficient is now equal to 0.862.


(c) The correlation coefficient increases when we remove the two countries producing
large deviations

because the remaining observations follow a fairly strong linear pattern.

(v) The R code is
cor.test(alcohol.2, cigarettes.2, method = "pearson",
alternative = "greater")

The p-value is equal to 3.738 × 10 −5 , showing very strong evidence against the null
hypothesis. We reject that Pearson’s correlation coefficient is equal to zero, in favour of
the alternative that it is positive.

(vi) Although the analysis suggests statistical association, we have not shown causation here.
So the report is not accurate.

(vii)
(a) The R code is
m1 = lm(cigarettes.2 ~ alcohol.2)
summary(m1)
------------------------------------


R Output:
Call:
lm(formula = cigarettes.2 ~ alcohol.2)
Residuals:
Min
1Q
Median
-1.04695 -0.41333 -0.01515
3Q
0.48439
Max
0.91801
Coefficients:
Estimate Std. Error t value Pr(>|t|)
(Intercept) -0.29845
0.75715 -0.394
0.7
alcohol.2
0.42207
0.07178
5.880 7.48e-05 ***
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Residual standard error: 0.6288 on 12 degrees of freedom
Multiple R-squared: 0.7424,
Adjusted R-squared: 0.7209
F-statistic: 34.58 on 1 and 12 DF, p-value: 7.476e-05
Draw the fitted line:
abline(m1)


From the R output the estimate of parameter σσ is 0.6288.

(viii) From the R output the proportion of the total variability of the responses explained by
the model is 0.7424.

(ix) R code:
------------------------------------

plot(alcohol.2,residuals(m1), main="Residuals plot (GR,
USA omitted)", xlab = "Alcohol per adult (litres/year)",
ylab = "Residuals")

(x) The mean of the errors seems to be close to zero,
but their variance increases with x.
This suggests that the assumed model may not be appropriate.



[Total 34]
Part (i) was well answered. A number of candidates did not provide appropriate or adequate
comments on the linearity of the plot. Parts (ii), (iii) and (iv) were generally well answered. In
part (v) many candidates did not mention the strength of the evidence against the null
hypothesis, while a few performed a 2-sided test. In part (vi) a number of candidates
erroneously indicated that correlation implies causation. Part (vii) was reasonably well
answered, while in part (viii) some candidates quoted the adjusted R-squared. Parts (ix) and
(x) were poorly answered with several incorrect plots given and a number of candidates not
attempting these parts.
------------------------------------
