---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3 - Hypothesis Testing with ***t.test()***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```



```{r}
load("smoking_data.rdata")
```

4 Data were collected on average alcohol and cigarette consumption per adult per year
for a number of countries. The data are given in the file smoking_data.RData
and contain the following information:
country: the country concerned;
alcohol: alcohol consumption per adult per year (litres/year);
cigarettes: number of cigarettes consumed per adult per year.

---
### Exercise 1
(i) (a) Construct a scatterplot of the data with alcohol consumption on the
x axis.
(b) Comment on the relationship between the data on alcohol and cigarette
consumption based on your plot in part (i)(a).
---
### Exercise 2
(ii) Calculate Pearson’s correlation coefficient between the data on alcohol and
cigarette consumption. 
---
### Exercise 3
An analyst suggests using the following R code to modify the data:
(iii) Explain what the above code does and give a justification for its use. [3]
For the remainder of the question, use the modified data
(alcohol.2, cigarettes.2), as produced by applying the R
code above.
---
### Exercise 4

(iv) (a) Construct a scatterplot with alcohol consumption on the x axis.
(b) Calculate Pearson’s correlation coefficient between the new data on
alcohol and cigarette consumption.
(c) Comment on your answers in parts (ii) and (iv)(b).

---
### Exercise 5
Perform a hypothesis test for the null hypothesis that Pearson’s population
correlation coefficient is equal to zero, against the alternative that it is positive.
You should report the p-value of the test and a clear conclusion. 

---
### Exercise 6
A media bulletin has reported that ‘'‘h‘igher alcohol consumption causes higher
cigarette consumption .
(vi) Comment on whether this report is justified based on your analysis in parts (iv)
and (v). 

---

```{r}
alcohol.2 = alcohol[-c(6,16)]
cigarettes.2 = cigarettes[-c(6,16)]
```

---
### Exercise 7

Perform a simple linear regression analysis on the new data using a
model of the form $$y = α + βx + ε$$ (cigarette consumption, y, on alcohol
consumption, x), where the error terms ε independently follow a
N(0, s2) distribution.
Your answer should show the fitted line plotted on the data scatterplot
and report the estimate of parameter s. 


---
### Exercise 8 

State the proportion of the total variability of the responses explained
by the model, based on your output in part (vii). [1]
(ix) Plot a graph of the residuals of the model fitted in part (vii)
against the explanatory variable. 


---
### Exercise 10

Comment on the validity of the model, based on your output in
part (ix) . 
---


Great, you have a detailed set of exercises here aimed at understanding the relationship between alcohol and cigarette consumption using statistical methods in R. The objectives of these exercises are to:

1. **Visualize Data**:
   - Create scatterplots to visually assess relationships between variables (alcohol and cigarette consumption).

2. **Calculate Correlations**:
   - Compute Pearson’s correlation coefficient to quantify the strength and direction of the relationship between alcohol and cigarette consumption.

3. **Data Modification and Justification**:
   - Modify datasets by excluding outliers or specific data points and justify the reasons for these modifications.

4. **Hypothesis Testing**:
   - Perform hypothesis tests (e.g., using the `t.test()` function) to evaluate whether the observed correlations are statistically significant.

5. **Linear Regression Analysis**:
   - Fit a linear regression model to the data, interpret the results, and assess the goodness-of-fit.

6. **Critically Analyze Media Statements**:
   - Evaluate and comment on claims made in media reports based on the statistical analysis conducted.

7. **Residual Analysis**:
   - Analyze residuals from the regression model to assess the validity and assumptions of the model.

#### Here's a brief overview of what each exercise entails:

### Exercise 1
- **(i) (a)**: Construct a scatterplot of alcohol consumption (x-axis) vs. cigarette consumption (y-axis).
- **(i) (b)**: Comment on the observed relationship based on the scatterplot.

### Exercise 2
- **(ii)**: Calculate Pearson’s correlation coefficient between alcohol and cigarette consumption.

### Exercise 3
- **(iii)**: Explain and justify the modification of the data using the provided R code.

### Exercise 4
- **(iv) (a)**: Create a scatterplot with the modified data.
- **(iv) (b)**: Calculate Pearson’s correlation coefficient for the modified data.
- **(iv) (c)**: Compare and comment on the correlations from the original and modified data.

### Exercise 5
- **(v)**: Perform a hypothesis test to determine if the population correlation coefficient is zero.

### Exercise 6
- **(vi)**: Comment on the media report’s claim that higher alcohol consumption causes higher cigarette consumption based on your analysis.

### Exercise 7
- **(vii)**: Conduct a simple linear regression analysis on the modified data and plot the fitted line on the scatterplot.

### Exercise 8
- **(viii)**: Calculate the proportion of variability explained by the regression model.
- **(ix)**: Plot the residuals against the explanatory variable.

### Exercise 10
- **(x)**: Comment on the validity of the regression model based on the residual plot.

The goal of these exercises is to equip you with the skills to perform and interpret statistical analyses using R, understand the relationships between variables, and critically evaluate statistical claims.

Is there anything specific you’d like to dive deeper into or any part of the exercise you need further assistance with?
