---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 2 - Chi Square Test

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

5 A waiting time random variable X follows an Exponential distribution with rate l
parameterised as f (x) = le–lx (x > 0, l > 0).
The rate l has a Gamma prior distribution with parameters $a$ and $s$. 

* A Bayesian credibility model provides that the posterior mean of 1/l can be expressed as
Z3x + (1– Z)3
s
α – 1
, where Z = n
α + n – 1
with n being the number of past waiting times observed.

* Assume that the parameters of the prior Gamma distribution are $a = 5$ and $s = 1$.


---

(i) Determine an estimate of the posterior mean of $1/l$ assuming $n = 10$ by
implementing M = 3,000 Monte Carlo repetitions. 

---

#### Exercise 1

```{r}
M = 3000
n = 10
alpha = 5
s = 1
pml = numeric(M)
```
---

```{r}
Z = n/(alpha + n - 1) 

for (i in 1:M){ 
 lam = rgamma(1, alpha, s) 
 x = rexp(n, lam) 
 pml[i] = Z*sum(x)/n + (1-Z)*s/(alpha-1) 
 }
```

---

```{r}
head(pml)
```
 
```{r}
mean(pml)
```
A point estimate is given as the mean of the produced estimates, i.e.
#mean(pml) 
which gives 0.248. 
---
### Exercise 2

Determine an estimate of the posterior mean of $1/l$ and the value of x when
n = 1,000, by implementing M = 3,000 Monte Carlo repetitions. 

This time we use the same code as in (i) but we are recording the mean of x for the case n =
1000.

```{r}
M = 3000
n = 1000
alpha = 5
s = 1
pml = numeric(M)
mx = numeric(M)
Z = n/(alpha + n - 1)
```

---

```{r}
for (i in 1:M){
 lam = rgamma(1, alpha, s)
 x = rexp(n, lam)
 mx[i] = mean(x)
 pml[i] = Z*sum(x)/n + (1-Z)*s/(alpha-1)
}

```

---

```{r}
mean(pml)
mean(mx)
```

With this new run, both estimates are equal to 0.245.

---

### Exercise 3

Plot the histograms of the samples of the posterior mean of $1/l$ and of x
obtained in Part 2.


```{r eval=FALSE}
hist(pml, main = "Histogram of posterior means", xlab = "Posterior means")
hist(mx, main ="Histogram of the distribution of the means of x", 
xlab = "mean of x")

```
---

```{r echo=FALSE}
hist(pml, main = "Histogram of posterior means", 
 xlab = "Posterior means",
 breaks=50)
```
---

```{r echo=FALSE}
hist(mx, main ="Histogram of the distribution of the means of x", 
xlab = "mean of x",
breaks=50)

```
---

### Exercise 4

Compare, by visual inspection of the graphs in part (iii), the distributions
of the posterior mean of 1/l and the distribution of x when n = 1,000.

#### Conclusion

The distributions look identical.
---

### Exercise 5

Comment on the behaviour of the credibility model as n increases, relating
your answer to your findings in Exercise 4. 

#### Conclusion

We should expect that the distributions are the same, 
since the credibility factor $Z$ tends to 1 as $n$ increases and therefore the posterior mean of 1⁄𝜆𝜆
becomes $\bar{x}$.

---

END

---