---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

An employee at a café has been trained to set the coffee machine so that an espresso
coffee portion results in 2.0 grams of coffee being placed into a cup. 

Knowing that
variations are expected, the employee pours eight portions and measures the amounts
to be 
$$\{1.95, 1.80, 2.10, 1.82, 1.75, 2.01, 1.83, 1.90\}$$.

#### Exercises

1. Calculate an 95% confidence interval for the mean size of espresso coffee
portions. 
2. Comment on whether the mean portion of coffee is equal to 2.0 grams. 
3. Repeat Exercises 1 and 2 for an 80% confidence interval

---

```{r}
amounts <- c(1.95, 1.80, 2.10, 1.82, 1.75, 2.01, 1.83, 1.90) 


### First 6 items
head(amounts)


### Last 6 items
tail(amounts)

```

---

#### Inspecting the Data

```{r}
length(amounts)

class(amounts)

mode(amounts)
```

---

### Use ***t.test()***:
```{r}
t.test(amounts,conf.level=0.95) 

```

---

### Conclusion
From output:

  95 percent confidence interval: (1.796, 1.994) 




The employee sees that value 2.0 is not in the interval and therefore a hypothesis of a 2 
gram portion would be rejected at the 5% level.


---

```{r}
t.test(amounts,conf.level=0.80) 

```

* From output:
80 percent confidence interval: (1.836, 1.954) 

#### Conclusion
The employee sees that value 2.0 is not in the interval and therefore a hypothesis of a 2 
gram portion would be rejected at the 20% level.

---

```{r}
t.test(amounts,conf.level=0.99) 

```
---