---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3 - Hypothesis Testing with ***t.test()***

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A machine in a sweet factory fills bags of sweets to weigh 500 grams. The actual
weight of the sweet bags is known to follow a Normal distribution. 

The sweet manufacturer believes that the machine is under-filling the sweet bags. A sample of
10 sweet bags is taken and weighed, as summarised below.

|  |  |  |  |  ||  |  |  |  |  |
|---|---|---|---|---|---|---|---|---|---|---|
|Bag | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 |
|Weight (grams) | 474.11 | 512.01 | 493.64 | 495.03 | 518.13 | 486.03 | 494.48 | 501.76 | 498.83 | 503.02 |

---


The data can be entered into R using the following code:


```{r}
weight<-c(474.11,512.01,493.64,495.03,518.13,486.03,
494.48,501.76,498.83,503.02)
```

#### Exercises


1. Perform a suitable t-test to determine whether the sweet bags are being
consistently under-filled, stating the hypotheses and the level of significance
used in the test. 

2. Propose an interpretation of your conclusion in part 1. 


---


The test has the null hypothesis that the mean sweet bag weight is equal to 500 grams, and 
the alternative hypothesis that the mean sweet bag weight is less than 500 grams. 
```{r}
t.test(weight,mu=500,alternative="less")
```
Remark: we run a one-sided t-test (as we are concerned with the one 
sided scenario of ‘under-filling’)

---


#### Interpreting the p-value

The test has been performed at the 5% significance level. Since the p-value of 0.287 is not 
less than the significance level of 0.05, we cannot reject the null hypothesis. 


#### Conclusion

Since we cannot reject the null hypothesis that the mean sweet bag weight is 500 grams, in 
favour of the alternative that the mean sweet bag weight is less than 500 grams, 
there is insufficient evidence that the sweet bags are being under-filled.


---