---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 2 - Chi Square Test

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```
A researcher has collected the following data on a group of students, regarding
whether they passed or failed an exam and whether or not they attended tutorials:


| Number of students  | Exam Passed | Exam Failed |  
|---|---|---|---|---|
| Attended tutorials  | 132 | 27 |  
| Did not attend tutorials  | 120 | 51 |  
|  |  |  |  |  |




---

The data can be entered into R in matrix form using the following code:

```{r}
exam.success = matrix(c(132,120,27,51),ncol=2,nrow=2)
```

```{r}

exam.success
```


```{r}

addmargins(exam.success)
```
---

#### Exercises
The researcher wants to establish whether tutorial attendance is independent of exam
success, using a chi-square test.

1. State the hypotheses of this test. 

2. Calculate the expected frequencies for the data under the null hypotheses
in part 1.

3. Perform the test stating clearly your conclusion. 

---

```{r}
chisq.test(exam.success)
```

```{r}
chisq.test(exam.success)$expected
```
---

```{r}
names( chisq.test(exam.success) )
```

### Residuals 

Residuals are the difference between observed and expected values

```{r}
chisq.test(exam.success)$residuals
```

---

### Chi-square test, 

Removing the Yates' continuity correction

```{r}
chisq.test(exam.success, correct=FALSE) 
```

The p-value is significant (e.g. at the 1% level), since 0.006069 < 0.01 – therefore there is 
evidence to reject the null hypothesis and we conclude that tutorial attendance and exam
success are not independent.

---

---