---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1 - Data Analysis
* Exam on basis of ***Base R*** ( I will mention ***tidyverse***)

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

A sample dataset $X$ is comprised of the following 10 numbers:

|       |        |        |        |        |        |
|-------|--------|--------|--------|--------|--------|
| 0.5820 | 0.04981 | 0.1552 | 0.1555 | 0.9036 |
| 0.8501 | 0.9288  | 0.4408 | 0.9688 | 0.6300 |


### Exercises

1. Find the mean value of $X$ 

2. Calculate the standard deviation of $X$ 

3. Find the median of $X$ 

---



```{r}
X  <-  c(0.5820, 0.04981, 0.1552, 0.1555, 0.9036, 0.8501, 0.9288, 0.4408, 0.9688, 0.6300)
```

---

#### Summary Statistics

```{r}
mean(X)
```

```{r}
sd(X)
```

```{r} 
median(X)
```

---

#### Sort

```{r}
sort(X)
```

#### Reverse the order

```{r}
rev(X)
```



---

### Summary

```{r}
summary(X)
```

### Five Number Summary

Tukey's five number summary (minimum, lower-hinge, median, upper-hinge, maximum)

```{r}
fivenum(X)
```

---

#### <tt>which.max()</tt> and <tt>which.min()</tt>

```[r]
X
```

```{r}
which.min(X)
```

```{r}
which.max(X)
```

---

#### Rank and Order



```{r}
data.frame(X,rank = rank(X))
```

<tt>order()</tt> returns the indices of the vector in sorted order

```{r}
order(X) 
```

---


```{r}
X
```


```{r}
X[order(X)]
```


---