---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

A general insurance company has sold 5000 Accident Benefit Policies to independent lives.

* The number of claims follows a Poisson distribution with parameter 500. 

* The amount of claim from individual policy is following a Gamma distribution with parameters $\alpha = 250$ and $\lambda = 0.5$.

* Calculate the aggregate claim amount paid by the insurer by simulating 5000 values if the insurer retains only 500 (assume the aggregate claim amount follows a compound Poisson distribution). 

*Use seed number 4321.*


---

***Poisson Distribution***

* Simulating the number of claims
* Poisson Random Variable with mean $\lambda = 500$

```{r}
set.seed(1234)

rpois(10,lambda=500)

```


```{r}
set.seed(4321)

num_claims <- rpois(1,lambda=500)

num_claims
```

---


***Individual claim amounts***

Gamma Distribution

* Shape Parameter $\alpha = 250$
* Rate Parameter $\lambda = 0.5$
* Scale Parameter $\beta = \frac{1}{\lambda} = 2$

```{r}

rgamma(1,shape = 250, scale = 2)

```


```{r}

claims <- rgamma(num_claims,shape = 250, scale= 2)

head(claims)
```


---


***Parallel Maxima and Minima***

```{r}

claims <- c( 530, 480, 490, 505, 515, 525, 475)

pmax(claims,500)
```


```{r}
pmin(claims,500)

```


---


***Simulation of 5000 observations***


````{r}
set.seed(1234)
n <- rpois(5000,lambda = 500)

Agg_Claim <- rep(0,5000)

for(j in 1:5000){
    y<-rgamma(n[j], shape = 250,  scale = 0.5)
    
    z<-pmin(y,500)
    
    Agg_Claim[j] <-sum(z)
    }

```

---


```{r}

head(Agg_Claim,20)

```


---

```{r}

summary(Agg_Claim)

```

---


```{r}

hist(Agg_Claim,breaks=100)

```

---
