---
title: "Statistics with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---



### Data
Before answering this question, the data file must be generated in R using the following code:


```{r}
set.seed(12345); 
y = arima.sim( list(order = c(1,1,0), ar = 0.7), n = 300)

set.seed(12345); 
y = arima.sim( list(order = c(1,1,0), ar = 0.7), n = 300)
```

---

Plot the time series giving appropriate labels for each axis and paste the chart into your answer.


<pre><code>
ts.plot( y, 
    xlab = "Time", 
    ylab = "Simulated Values from an ARIMA(1,1,0) Time Series")
</code></pre>

---

```{r}
ts.plot( y, 
    xlab = "Time", 
    ylab = "Simulated Values from an ARIMA(1,1,0) Time Series")
```

---

```{r}
plot( y, 
  
     xlab = "Time", 
  
     ylab = "Simulated Values from an ARIMA(1,1,0) Time Series")

x <- 1:301

```

---

```{r}
plot(x, y, 
    
    type="l", 
    
    xlab = "Time", 
  
    ylab = "Simulated Values from an ARIMA(1,1,0) Time Series")
```


---

(b) Comment on the general features of your chart.

---

Exercise 2
(a) Determine the best least squares linear fit, adding it to your chart.

(b) Explain whether this least squares linear trend can be removed such that a zero mean stationary distribution is appropriate for the residuals.


---

```{r}
x = 1:301
leastsquaresfit = lm(y~x)
```

---
```{r}
leastsquaresfit$coefficients
```

(ii) (a)
​
Intercept: 16.4405647
Slope: -0.2436929 

---
abline(leastsquaresfit)
OR:
lines(leastsquaresfit$fitted.values) [11⁄2]
Error in parse(text = x, srcfile = src): <text>:4:22: unexpected input
3: Intercept: 16.4405647
4: Slope: -0.2436929 [11<e2>
  ^
Traceback:


​
​
(b)
plot(leastsquaresfit$res, xlab = "Time", ylab="Residuals")
​
 
It is clear that the residuals are not stationary as they are negative in the first third
followed by positive residuals in the middle part and then negative in the last part.
​
OR:
acf(leastsquaresfit$res)
The residuals are not stationary as the ACF of the residuals seems to generate a unit root as
the ACF values are very slowly decaying.
​
Error in parse(text = x, srcfile = src): <text>:7:4: unexpected symbol

---

Exercise 3
========================================
Plot the sample Autocorrelation function (ACF) and sample Partial Autocorrelation function (PACF) of the original data, giving appropriate labels for each axis and paste the charts into your answer.

```{r}
acf(y, xlab = "Lag", ylab = "ACF of Simulated Values
from an ARIMA(1,1,0) Time Series", main = "ACF")
pacf(y, xlab = "Lag", ylab = "Partial ACF of Simulated
Values from an ARIMA(1,1,0) Time Series", main = "PACF")   
```


(iv)
(v)
(a)
(b) Comment, by visually inspecting these plots, on the possible modelling
strategy which could be adopted.
​
(a) Perform an appropriate transformation to the data such that a stationary
model is possible, pasting any relevant charts into your answer.
​
(b) Comment on your answer to part (iv)(a).
​
[5]
(b)


There is clearly not a constant mean / any mean-reverting feature in the data ... 
...so stationarity does not hold.
​
There is perhaps an initial upwards trend...
​
...but overall there seems to be a downwards trend.
​
[7, Max. 4]
​
​
 

---
 
(iii)(b)
Despite the PACF showing no significance past lag 2 which could indicate
stationarity...
​
... clearly the ACF is not behaving as a stationary ARMA process should.
​
There is a slow decay in the ACF values suggesting a unit root process
​
So we need to perform differencing.

---

Exercise 4
=============================



```{r}
library(forecast)
tsdisplay(diff(y))  
```



```{r}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
ts.plot(diff(y), main = "diff(y)")
points(diff(y),cex=0.4)
acf(diff(y))
pacf(diff(y))
```



​
​
​
These plots indicate that the differenced data could be stationary and both ACF and
PACF seem to decay fast towards zero.
​
​
 
The plots do not indicate any seasonality.

%55
(v)(a)
The chosen model for the transformed data is ARIMA(1,0,0) since the differenced
data looks stationary, and PACF is close to zero from lag 2.
​


```{r}
fit10=arima(diff(y),order = c(1,0,0));fit10
```


Call:
arima(x = diff(y), order = c(1,0,0))
Coefficients:
ar1 intercept
0.7140
-0.2324
s.e. 0.0402
0.1951
sigma^2 estimated as 0.9493:
aic = 842.46
log likelihood = -418.23,
​
The proposed model parameters for the transformed data are:
α = 0.714, intercept = -0.2324



(a) Propose an appropriate model for the transformed data.
​
(b) Justify the choice of model in part (v)(a) by performing an appropriate
diagnostics procedure and comparisons with alternative models.


5

(v)(b)
By inspecting the ACF and PACF plots of differenced data, alternative models can be
considered by changing the values of p and q. In the following we fit 6 models:
​


```{r}
fit11=arima(diff(y),order= c(1,0,1));fit11$aic
# 844.452
fit10=arima(diff(y),order= c(1,0,0));fit10$aic
# 842.4563
fit01=arima(diff(y),order= c(0,0,1));fit01$aic
# 911.0564
```

```{r}
fit21=arima(diff(y),order= c(2,0,1));fit21$aic
# 845.8664
fit12=arima(diff(y),order= c(1,0,2));fit12$aic
# 846.3837
fit22=arima(diff(y),order= c(2,0,2));fit22$aic
# 847.6294
```



*     This confirms that the ARIMA(1,0,0) model is a good fit compared with
these alternatives.

    To check we can use a diagnostic testing procedure of:
tsdiag(fit10)

​
 
...The ACF of residuals together with the corresponding Ljung-Box test output
...(i.e. high p-values observed suggesting good fit - residuals close to white noise)
...suggest that this is a correct model.
