---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---


Solution 1:
i) i
.

```{r}
set.seed(052023)
u<-runif(150)
round(mean(u),2)
```
[1] 0.52
This shows sample mean ~0.52.

---

ii) 

```{r}
chi <- 1 -qchisq(u,2) 
```

iii) i
i
i
.

```{r}
gam<-qgamma(u,1,1/2)
sum(chi-gam)
```

[1] 0
(1)

---

We know the property that if $X ~ Gamma(α,ʎ)$ then 2ʎX has χ2 distribution with 2α degrees of 
freedom.
* We have X ~ χ2 with 2 degrees of freedom.
* Above can be written as (2ʎX/2ʎ) ~ χ2 with 2α degrees of freedom where α=1,ʎ =1/2
Thus,
X ~ Gamma(α,ʎ) with α=1,ʎ =1/2
X ~ Gamma(1,1/2)
This is why both samples are same.


---

iv) .
a) hist(chi, main ="Histogram of chi square distribution sample") a (1)
(1)
#Histogram shows positive skewed distribution (1)
[3]
b) b

---

> 

```{r}
summary(chi)
```

 Min. 1st Qu. Median Mean 3rd Qu. Max. 
0.006184 0.603187 1.534968 2.217946 2.983161 12.350562
(1)
Alternate:
mean(chi)
median(chi)
#mean = 2.218
#median = 1.535
#Mean is greater than median since it is positively skewed distribution
(0.5)
(0.5)
(1)
[2]

---

v) 

```{r}
set.seed(052023) 
y = rep(0,1000)
for(i in 1:1000){
 y[i] = sum(rchisq(150,2))
}
```


---

vi) 

```{r}
hist(y, main ="Histogram of Samples Sum") 
```


* The distribution of sample sums is roughly symmetrical. 
* This displays Central Limit Theorem property. 
* As the sample size gets large , distribution move towards normality. 



---

END

---


---
