---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1

---

Solution 3:
i) i
.
#m1x0 = E[X] and 
#m2x0=E[x^2] and var = E[x^2] - E[x]^2 >> m2x0 = var + E[x]^2 
prior_mean=60
prior_sd=5


m1x0= prior_mean
m2x0 = prior_sd^2 + prior_mean^2
(0.5)


---

(1)
> m1x0
[1] 60
> m2x0
[1] 3625
(0.5)
(0.5)

---

ii) .
a) a
.
# theta follows N(prior mean,prior variance)
# Random Variable X follows N(theta,variance)
# posterior distribution of theta follows Normal with
# post mean = (n*sample mean/variance +prior mean/prior variance)/
# (n/variance + 1/prior variance)

---

#post variance = 1/(n/variance + 1/prior variance) (1)
Max 2
n<-5
sample_mean<- 340/5
sdev<-20
 
post_mean = (n*sample_mean/sdev^2 + prior_mean/prior_sd^2)/(n/sdev^2 + 1/prior_sd^2)
post_var= 1/(n/sdev^2 + 1/prior_sd^2)
(2)
(1)
> post_mean
[1] 61.90476
> post_var
[1] 19.04762
> sqrt(post_var)
[1] 4.364358

---

b) b
.
sample2_mean<-3400/50
n2<-50
post2_mean = (n2*sample2_mean/sdev^2 + prior_mean/prior_sd^2)/(n2/sdev^2 + 1/prior_sd^2)
post2_var= 1/(n2/sdev^2 + 1/prior_sd^2)

> post2_mean
[1] 66.06061
> post2_var
[1] 6.060606
> sqrt(post2_var)
[1] 2.46183

---

iii) .
a) a
.

```{r}
x<-60+seq(-3,3,by=0.2)*5
y<-dnorm(x,mean=60,sd=5)
plot(x,y,ylim=c(0,.2)) 

```

b) b
.

```{r}
px1<-post_mean+seq(-3,3,by=0.2)*sqrt(post_var)
py1<-dnorm(px1,mean=post_mean,sd=sqrt(post_var))
lines(px1,py1)
```

---

c) c
.

```{r}
px2<-post2_mean+seq(-3,3,by=0.2)*sqrt(post2_var)
py2<-dnorm(px2,mean=post2_mean,sd=sqrt(post2_var))
lines(px2,py2)
```

---

d) d
.
(1)
The posterior distribution with sample size =5 is close to prior distribution. There is slight shift to 
mean towards sample mean and similar dispersion. 
(1)

* When the sample size increased, the posterior distribution moves towards sample mean and 
dispersion. 
* More weight is given to sample where sample is big. Further, the variation reduced with larger 
sample size. 

---