---
title: "Risk Modelling and Survival Analysis with R"
subtitle: "R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Risk Modelling and Survival Analysis with R


#### CS2B Curriculum

* It develops knowledge of and the ability to apply statistical methods for risk modelling, time series analysis methods, stochastic processes (especially Markov chains and Markov jump processes), survival analysis (including regression methods applied to duration data), and graduation methods  It also includes a high-level introduction to machine learning. 

* Question 2 Part (a)
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

```{r echo=FALSE, include=FALSE}
library(magrittr)


options(digits=4)
```

## Question 2 Part (a)

* An exponential distribution has a parameter of $\lambda = 0.4$.

* Use the in-built functions in R to perform the following tasks.


---


### Exercise 1


Simulate 1,000 values from this distribution, assigning this to a variable called ***Exp_Vector*** and calculate the mean and variance of
the simulated values. 

Paste the results of your calculation into your answer.

```{r}
rexp(5,rate = 0.4)
```

---


### Exercise 1


```{r}
options(digits=4)

rexp(5,rate = 0.4)
```


---

### Exercise 1


```{r}
rexp(5,0.4)
```


```{r}
rexp(5,0.4)
```


---


### Using ***set.seed()***

```{r}
set.seed(1234)
rexp(5,0.4)
```

```{r}
set.seed(1234)
rexp(5,0.4)
```

```{r}
set.seed(1234)
rexp(5,0.4)

```


---

#### Mean and Variance

```{r}
set.seed(1234)
Exp_Vector <- rexp(1000,0.4)
```

```{r}
mean(Exp_Vector)
```

```{r}
var(Exp_Vector)
```


#### Numerical Summary

```{r}
summary(Exp_Vector)
```

---

#### Analytical Results

The mean and variance will vary due to the random number generation. 

If the sample size was large enough, the mean and variance should be close the underlying distribution (exponential with parameter 0.4) as follows:

Mean = 2.5

$$ E(X) = \frac{1}{\lambda} = 2.5$$

Variance = 6.25

$$ \operatorname{Var}(X) = \frac{1}{\lambda^2} = 6.25$$

---

### Part 2



Plot a histogram of ***Exp_Vector*** showing the frequencies and paste the plot into your answer.

---

### Part 2

```{r}
hist(Exp_Vector,breaks=50)

```

---

```{r}
hist(Exp_Vector, breaks=50,col=c("lightblue","pink"))

```


---

#### Part 3

Plot the probability density function for this distribution as:

1. a scatter plot
2. a line graph.


#### Sequence

```{r}
x <- seq(0,20,by=0.5)
x
```

---

### Probability Density Function


```{r}
PDF <- dexp(x, 0.4)

head(PDF)

```

---

### Probability Density Function - Scatterplot


```{r}
plot(x, PDF)
```


---

### Probability Density Function - Lineplot


```{r}
plot(x, PDF, type="l")
```

---


### Probability Density Function - Lineplot


```{r}
plot(x, PDF, type="l", col="red")
```

---

### Probability Density Function - Lineplot


```{r}
plot(x, PDF, type="l", col="red", lwd=2.4)
```

---


---



---


Exercise  2
============================================================

A lognormal distribution has parameters $\mu = 0$ and $\sigma^2 = 1$.

%Use the in-built functions in R to:
% subsubsection*{Task a}

Simulate 1,000 values from this distribution, assigning this to a vector called ``\texttt{LNorm\_Vector}`` and calculate the mean and variance of the simulated values. 




```{r}
set.seed(1234)
LNorm_Vector <- rlnorm(10000, meanlog = 0, sdlog = 1)
```
```{r}
mean(LNorm_Vector) 

var(LNorm_Vector) 
```


\includegraphics[scale=1.2]{00-A2/images/lognormal_1_estimates.JPG}


* The mean and variance will vary due to the random number generation. 

* If the sample size was large enough, the mean and variance should be close the underlying
distribution (lognormal with parameters $ \mu = 0$, $\sigma^2 = 1$) as follows:\\
Mean = 1.649\\
Variance = 4.6708\\


---

Plot a histogram of ``\texttt{LNorm\_Vector}`` showing the frequencies.



\includegraphics[scale=0.4]{00-A2/images/output_23_0.jpeg}


---

Task (c)
======================================

Plot a second histogram in a new graph of ``\texttt{LNorm\_Vector}`` showing the probability densities, setting the y-axis range from 0 to 0.7 for this graph.


```{r}
hist(LNorm_Vector, breaks=25, freq = FALSE, xlim = c(0,25), ylim = c(0,1))

grid = seq(0,25,0.1)
lines(grid, plnorm(grid,0,1),type="l",xlab="x",ylab="f(x)", col="black")
lines(ecdf(LNorm_Vector),col="red")
legend("bottomright",c("True cumulative distribution","Estimate"),lty=1,col=c("black", "red"))
```




```{r}
hist(LNorm_Vector)
```


\includegraphics[scale=0.4]{00-A2/images/output_24_0.jpeg}




```{r}
hist(LNorm_Vector, freq = FALSE, xlim = c(0,25), ylim = c(0,0.7))
```


\includegraphics[scale=0.4]{00-A2/images/output_25_0.png}


---

#### Exercise  (d)
Add the cumulative density function of \texttt{LNorm\_Vector} to the chart
in Exercise  (ii)(c) and paste the plot into your answer.

#### Exercise  (e)

Add the theoretical lognormal (0,1) distribution to the chart in Exercise  (ii)(d) to highlight the difference to the sample, including
appropriate labels and legend and paste the plot into your answer.




```{r}

hist(LNorm_Vector, freq = FALSE, xlim = c(0,25), ylim = c(0,0.7))

lines(grid, dlnorm(grid,0,1),type="l",xlab="x",ylab="f(x)", col="black")

lines(density(LNorm_Vector), col="red")

legend("topright",c("True Density", "Estimate"),lty=1,col=c("black", "red"))

```


---




The output is:
(iii) (a)

```{r}
rpareto <- function(n,alpha,lambda) {
rp <- lambda*( (1-runif(n))^(-1/alpha) -1 )
rp
}
```

---

```{r}

LNorm_Vector = rpareto(1000, 3, 1)
mean(LNorm_Vector)
var(LNorm_Vector)
```

---

The mean and variance will vary due to the random number generation. If the sample size
was large enough, the mean and variance should be close the underlying distribution
(Pareto α = 3, λ = 1) as follows:
Mean = 0.5
Variance = 0.75


---

Note: The correct R code receives full marks.
Candidates are not required to paste theirs simulated sample.
Note: Alternative solutions to (iii) are possible. For example,

```{r}
rpareto <- function(alpha, lambda) {
rp <- lambda*( (1-runif(1))^(-1/alpha) -1 )
rp
}

LNorm_Vector = replicate( 1000, rpareto(3,1))
mean(LNorm_Vector)
var(LNorm_Vector)

```






