---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

2 Before answering this question, the R package for copulas should be loaded into R 
using the following code: 
<pre><code>
install.packages("copula") 
library(copula) 
</code></pre>
A copula model is to be constructed of the dependence between annual average 
temperature $X$, in degrees Celsius, and annual rainfall Y, in millimetres, in a region at 
risk of desertification. The marginal distributions of X and Y are: 
X ~ N൫20, 32
൯, Y ~ N൫200, 502
൯. 
---

Two copulas are under consideration to model the dependence of X and Y: 

* Copula 1: A Gaussian copula with ρ = –0.5. 
* Copula 2: A Student’s t copula with 3 degrees of freedom and ρ = –0.5. 

The following R code generates n simulations of jointly distributed random variables 
U and V, where the marginal distributions of U and V are the uniform distribution on 
[0, 1] and the dependency between U and V is given by a Gaussian copula with 
correlation ρ (rho): 
```{r}
rCopula(n, normalCopula(param = rho)) 
```
The corresponding code for the Student’s t-copula with nu degrees of freedom is: 
```{r}
rCopula(n, tCopula(param = rho, df = nu)) 
```

The output from the function rCopula is in the format of an n × 2 matrix whose 
rows are the observations of (U, V). 
Before answering part (i), generate the matrix, GaussXY, in R using the following 
code: 
```{r}
set.seed(3) 
GaussUV = rCopula(200000, normalCopula(param = -0.5)) 
GaussXY = matrix(nrow = 200000, ncol = 2) 
GaussXY[,1] = qnorm(GaussUV[,1], mean = 20, sd = 3) 
GaussXY[,2] = qnorm(GaussUV[,2], mean = 200, sd = 50) 
```
The 200,000 × 2 matrix, GaussXY, represents 200,000 observations of (X, Y) under 
Copula 1. 

---
(i) Generate 200,000 observations of (X, Y) using Copula 2 and the marginal 
distributions of X and Y, assigning the observations to a 200,000 × 2 matrix 
named t3XY. You should set a random number generator seed of 3 before 
generating the observations and use the R function, <tt>head()</tt>, to display the 
first six rows of t3XY in your answer script.  

---
(ii) Plot, on a single graph, line graphs of the estimated conditional mean of Y
given that X ൒ x against x for values of x at intervals of 1 from 10 to 30 
inclusive for each of the two generated copulas. You should use separate 
colours to identify each of the two line graphs and set the y-axis range from 50 
to 250.  
(iii) Comment on the key features of your plot in part (ii).  
(iv) Comment on the implications of your plot in part (ii) if the distribution of 
temperatures, X, is expected to become more weighted towards higher values 
as a result of climate change. 

---
 
Q2
install.packages("copula")
library(copula)
set.seed(3)
GaussUV = rCopula(200000, normalCopula(param = -0.5))
GaussXY = matrix(nrow = 200000, ncol = 2)
GaussXY[,1] = qnorm(GaussUV[,1], mean = 20, sd = 3)
GaussXY[,2] = qnorm(GaussUV[,2], mean = 200, sd = 50)

---
(i)
set.seed(3) 
t3UV = rCopula(200000, tCopula(param = -0.5, df = 3)) 
t3XY = matrix(nrow = 200000, ncol = 2) 
t3XY[,1] = qnorm(t3UV[,1], mean = 20, sd = 3) 
t3XY[,2] = qnorm(t3UV[,2], mean = 200, sd = 50) 
head(t3XY) 

---
(ii)
x = seq(from = 10, to = 30, by = 1) 
y1 = vector(length = 21) 
y2 = vector(length = 21) 
for (i in 1:21) { 
y1[i]= mean(GaussXY[,2][GaussXY[,1] >= x[i]]) [3]
y2[i]= mean(t3XY[,2][t3XY[,1] >= x[i]]) 
}
---
plot( 
x, 
y1, 
ylim = c(50, 250), 
type = "l", 
ylab = "Conditional Mean of Y | X >= x", 
main = "Conditional Mean of Y | X >= x for Two Copulas with rho = -
0.5") 
lines( 
x, 
y2, 
col = "red") 
legend("topright", 
legend = c("Gaussian Copula", "Student's t Copula with 3 df"), 
col = c("black", "red"), 
pch=7)
---

(iii)
For the lower values of x, the conditional means for both copulas are similar 
and close to the unconditional mean of 200 
The conditional means for both copulas decrease with increasing x 
because of the negative value of rho 
Since the t3 copula exhibits positive tail dependence and the Gaussian copula has zero tail 
dependence 
the graph for the t3 copula slopes downwards more steeply than for the Gaussian copula 
The order of the Gaussian and t3 copulas is not consistent for all values of x (i.e. the
red line is slightly above the black line from around x = 18 to x = 23) 

---
#### Exercise (iv)
In order to draw any conclusions here, it is necessary to assume that the conditional 
distribution of Y given that X >= x remains unchanged for each value of x 
The graph in part (ii) indicates that the mean level of rainfall can be expected to decrease 
with the extent of the decrease being greater for the Student’s t3 copula than for the
Gaussian copula 
---