---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

(iv) Comment on how realistic your answer to part (iii) is in terms of what the 
bank may expect in practice, suggesting how the parameter a could be 
changed to make it more realistic.  
The R code for fitting and plotting a decision tree object called ***tree***, that predicts 
whether a customer will default or not, is as follows: 
```{r}
tree = rpart(default ~ ., data = specimen, method = "class")
rpart.plot(tree, digits = 3, type = 5, extra = 106) 
```

---
(v) Plot the decision tree object, called tree, using the R code above.  
(vi) State, using the decision tree plot, whether a customer with features f1 = 0 and 
f2 = –0.425 is expected to default or not.  
Before launching the product, the Actuary gathers actual customer data from another 
bank, that is available in the public domain, to assess the predictive power of the 
decision tree model. The ‘***CS2B_Sept_21_Qu_3_Data.csv***’ file contains the actual 
data for 100 customers in the required format for the decision tree model. 
Before answering the remainder of this question, the ‘***CS2B_Sept_21_Qu_3_Data.csv***’ 
file should be loaded into R and assigned to a data frame called actual. 



---


(iv)
The answer to part (iii) is not likely to be realistic 
as the mean default probability of their potential customers would be expected to be 
significantly less than 0.5 
A negative value of a would be better 
so as to reduce the expected probability of default 

---


### Part E


```{r}
tree = rpart(default ~ ., data = specimen, 
             method = "class")

rpart.plot(tree, digits = 3, type = 5, extra = 106)

```

---

### Part (vi)

According to the decision tree, this customer is NOT expected to default 
