---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* ### Question  1
---

1 For an investigation into the buying behaviour of consumers shopping at an online
store, data about the age, the time per month spent on the online shop’s website and
the number of purchases from the shop per year are recorded.
This dataset is saved in the file ‘onlineShop.Rdata’. After loading the data into R,
using the command load("onlineShop.RData"), the data frame
onlineShop will be available. The data frame onlineShop contains the
following three variables:
 age – the age (in years) of the customer
 <tt>time_spent_on_website</tt> – the time per month spent on the shop’s website
(in hours)
 purchases_per_year – the number of purchases from the shop per year.
(i) Fit a Generalised Linear Model (GLM) to the data assuming a Poisson
distribution with a log link function, using purchases_per_year as the
response variable and age and <tt>time_spent_on_website</tt> as explanatory
variables without an interaction term. Your answer should include the
estimated coefficients, the deviance and the Akaike’s Information Criterion
(AIC) of the fitted model. 
(ii) State an expression for the expected number of purchases per year in terms of
the variables age and <tt>time_spent_on_website</tt> using the fitted model
from part (i). 


---

It is suggested that a log link function may not be appropriate, and a square root link
function should be used instead, that is 𝑔􁈺𝜇􁈻 􀵌 √𝜇. 

In R, a square root link function
can be specified using <tt>family = poisson(link = "sqrt")</tt>.

#### More Exercises

(iii) Fit a GLM to the data assuming a Poisson distribution with a square root link
function using purchases_per_year as the response variable and age
and <tt>time_spent_on_website</tt> as explanatory variables without an
interaction term. Your answer should include the estimated coefficients, the
deviance and the AIC of the fitted model.
(iv) State an expression for the expected number of purchases per year in terms of
the variables age and <tt>time_spent_on_website</tt> using the fitted model
from part (iii). 
(v) Compare the models in parts (i) and (iii) and decide which model should be
preferred using the deviance. 
It is suggested that the deviance is not an appropriate measure to compare the two
models in parts (i) and (iii), and that the AIC should be used instead.
(vi) Comment on this suggestion in the context of comparing the two models in
parts (i) and (iii). In particular, identify any advantages that the AIC may
have compared to the deviance in this context. 

---

**1. GLM for Online Shopping Behavior**

**1(i) Poisson GLM with Log Link**

```{r}
load("onlineShop.Rdata")

# Fit the Poisson GLM with log link
model_log <- glm(purchases_per_year ~ age + <tt>time_spent_on_website</tt>, 
                 data = onlineShop, family = poisson)

# Summary of the model
summary(model_log)
```

**1(ii) Expected Number of Purchases with Log Link**

The expected number of purchases per year, E(Y), can be expressed as:

```
E(Y) = exp(β₀ + β₁*age + β₂*<tt>time_spent_on_website</tt>)
```

where β₀, β₁, and β₂ are the estimated coefficients from the model.

**1(iii) Poisson GLM with Square Root Link**

```{r}
# Fit the Poisson GLM with square root link
model_sqrt <- glm(purchases_per_year ~ age + <tt>time_spent_on_website</tt>, 
                  data = onlineShop, family = poisson(link = "sqrt"))

# Summary of the model
summary(model_sqrt)
```

**1(iv) Expected Number of Purchases with Square Root Link**

The expected number of purchases per year, E(Y), can be expressed as:

```
E(Y) = (β₀ + β₁*age + β₂*<tt>time_spent_on_website</tt>)²
```

**1(v) Model Comparison using Deviance**

We can compare the two models using the deviance. A lower deviance indicates a better fit. 
However, it's important to note that deviance can be misleading when comparing models with different link functions.

**1(vi) Comparing Models using AIC**

The AIC is a more reliable measure for comparing models with different link functions. 
It penalizes for model complexity, making it suitable for comparing models with different numbers of parameters. 
In this case, both models have the same number of parameters, but the AIC can still be used to compare their overall fit.

The AIC takes into account both the goodness of fit and the complexity of the model. A lower AIC indicates a better model. 

By comparing the AIC values of the two models, we can determine which model provides a better balance of fit and parsimony.
