---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---



4 Two members of an interview panel (A and B) have scored a large number of
applicants (on a scale from 1 to 40), with the purpose of ranking the applicants
according to their suitability for a job position. The scores for a sample of twelve
applicants are given below, with SA and SB denoting the scores of interview panel
member A and B, respectively:
 SA = c(35,14,28,33,29,22,19,36,21,30,15,18)
 SB = c(38,18,25,30,22,17,23,29,32,31,15,19).
(i) Plot a suitable graph for assessing the agreement in the two panel members’
scores. [2]
(ii) Comment on the agreement in the two panel members’ scores, based on the
plot in part (i). [2]
(iii) Calculate Pearson’s correlation coefficient r for these data. [1]
(iv) Calculate Spearman’s rank correlation coefficient rs for these data. [1]
(v) Calculate Kendall’s rank correlation coefficient 𝜏 for these data. [1]
(vi) Comment on the suitability of the three correlation coefficients used in parts
(iii)–(v), with respect to the purpose of the scoring. [2]
(vii) Perform three statistical tests for a suitable hypothesis of no association
between the scores of the two panel members, using one of the correlation
coefficients in parts (iii)–(v) for each test.
For each test, your answer should include the p-value and a conclusion. [11]
(viii) Comment on the validity of the tests in part (vii) for testing the hypothesis of
no association between the scores of the two panel members. [2]
[Total 22]
