---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


2 An investment analyst is interested in the relationship between the share price of a
company and its Earnings-Per-Share (EPS).
A sample of data for eight companies has been collected. This is stored in the file
‘CompanySample.RData’. After loading this data into R, the data frame
CompanySample will be available that contains two columns:
 price – the company’s share price (in $)
 EPS – the company’s earnings-per-share (in $).
(i) Fit a linear model to the sample, with price as the dependent variable and
EPS as the independent variable, and state the estimated coefficients of the
model. [3]
(ii) State the value of the coefficient of determination, R2, for the model fitted
in part (i). [1]
(iii) Perform an analysis of variance, using the anova(…) command to test if
there is a linear relationship between price and EPS. Your answer should
include the F-statistic and p-value of the test. [4]
(iv) Comment on how well the linear model in part (i) fits the data, using your
answers to parts (ii) and (iii). [2]
The analyst believes that the observation where price = 99 is an anomaly because of
the large size of the value of price and should be removed from the sample.
(v) Fit a linear model to the sample without the observation price = 99 and state
the value of R2. [3]
[Hint: The code CompanySample[-n,] returns the data frame with the
nth observation removed.]
(vi) Compare the fit of the models from parts (i) and (v). [3]
(vii) Discuss if removing the observation where price = 99 is appropriate. [3]


---

