---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---



4 Two members of an interview panel (A and B) have scored a large number of
applicants (on a scale from 1 to 40), with the purpose of ranking the applicants
according to their suitability for a job position. The scores for a sample of twelve
applicants are given below, with SA and SB denoting the scores of interview panel
member A and B, respectively:
 SA = c(35,14,28,33,29,22,19,36,21,30,15,18)
 SB = c(38,18,25,30,22,17,23,29,32,31,15,19).
(i) Plot a suitable graph for assessing the agreement in the two panel members’
scores. [2]
(ii) Comment on the agreement in the two panel members’ scores, based on the
plot in part (i). [2]
(iii) Calculate Pearson’s correlation coefficient r for these data. [1]
(iv) Calculate Spearman’s rank correlation coefficient rs for these data. [1]
(v) Calculate Kendall’s rank correlation coefficient 𝜏 for these data. [1]
(vi) Comment on the suitability of the three correlation coefficients used in parts
(iii)–(v), with respect to the purpose of the scoring. [2]
(vii) Perform three statistical tests for a suitable hypothesis of no association
between the scores of the two panel members, using one of the correlation
coefficients in parts (iii)–(v) for each test.
For each test, your answer should include the p-value and a conclusion. [11]
(viii) Comment on the validity of the tests in part (vii) for testing the hypothesis of
no association between the scores of the two panel members. [2]
[Total 22]


**Understanding the Problem:**

We're tasked with:
1. **Visualizing Agreement:** Plotting a graph to compare the two sets of scores.
2. **Calculating Correlation Coefficients:** Computing Pearson's, Spearman's, and Kendall's correlation coefficients.
3. **Hypothesis Testing:** Performing statistical tests to determine the significance of the correlation.
4. **Interpreting Results:** Drawing conclusions about the agreement between the two panel members' scores.

**R Code Implementation:**

```R
# Data
SA <- c(35, 14, 28, 33, 29, 22, 19, 36, 21, 30, 15, 18)
SB <- c(38, 18, 25, 30, 22, 17, 23, 29, 32, 31, 15, 19)

# (i) Plot a scatter plot
plot(SA, SB, xlab = "Panel Member A's Score", ylab = "Panel Member B's Score")

# (ii) Comment on agreement
# ... (Visually inspect the scatter plot for agreement)

# (iii) Pearson's correlation
cor(SA, SB, method = "pearson")

# (iv) Spearman's rank correlation
cor(SA, SB, method = "spearman")

# (v) Kendall's rank correlation
cor(SA, SB, method = "kendall")

# (vi) Comment on suitability of correlation coefficients
# ... (Discuss the appropriateness of each coefficient for ordinal and interval data)

# (vii) Hypothesis testing
# Pearson's correlation test
cor.test(SA, SB, method = "pearson")

# Spearman's rank correlation test
cor.test(SA, SB, method = "spearman")

# Kendall's rank correlation test
cor.test(SA, SB, method = "kendall")

# (viii) Comment on the validity of the tests
# ... (Discuss the assumptions of each test and their potential violations)
```

**Interpreting the Results:**

**Parts (i) and (ii):**
- A scatter plot can visually assess the agreement between the two sets of scores.
- A strong positive correlation would indicate high agreement, while a weak or negative correlation would suggest disagreement.

**Parts (iii) to (v):**
- Pearson's correlation measures the linear relationship between two continuous variables.
- Spearman's and Kendall's correlation coefficients measure the monotonic relationship between two ranked variables.

**Part (vi):**
- Pearson's correlation is suitable for interval data, while Spearman's and Kendall's are suitable for ordinal data.
- In this case, while the scores are technically on an interval scale, they might be better treated as ordinal due to potential non-linear relationships and subjective judgments.

**Part (vii):**
- The hypothesis tests for each correlation coefficient assess the null hypothesis of no correlation.
- The p-value indicates the significance of the correlation.
- A low p-value suggests a significant correlation.

**Part (viii):**
- The validity of the tests depends on the assumptions of each test.
- For example, Pearson's correlation assumes a linear relationship and normality.
- Spearman's and Kendall's correlation tests are more robust to deviations from normality.
- It's important to consider the nature of the data and the specific research question when choosing the appropriate test.

By following these steps and interpreting the results, you can gain insights into the agreement between the two panel members' scores and make informed decisions about the selection process.
