---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


5 A sample of 100 daily sales amounts at a store is collected, with the aim to investigate
if daily sales amounts are affected by whether or not the store has substantial
discount offers available on a particular day. The data are available in the file
‘sales_data.RData’, which contains the following variables for 100 days:
 sales.amount – the daily sales amount
 discount – an indicator, showing whether or not the store has substantial
discount offers available on each day (discount = 1 if discounts are available;
discount = 0 otherwise).

---
(i) Plot appropriate boxplots for comparing the sales amounts on discount and
non-discount days. [3]
[Hint: You may find the boxplot(formula, …) R command useful.]
(ii) Comment on the effect of discount offers on sales amounts using the boxplots
produced in part (i). [2]
A sales analyst considers two distributional assumptions: a normal and a gamma
model for the daily sales amounts.
(iii) Fit two separate Generalised Linear Models (GLMs) to the data, one for each
of the two distributional assumptions, to investigate the dependence of the
daily sales amounts on availability of substantial discount offers on the day.
For each model you should use the canonical link function. 

---

(iv) Comment on the impact of the availability of substantial discount offers on
daily sales amounts based on the output of the two models fitted in part (iii).
Your answer should include a relevant p-value and interpretation of the
estimate of the coefficient of the discount variable for each fitted model. 

---

(v) Determine which of the two models fitted in part (iii) should be preferred for
investigating the dependence of daily sales amounts on the availability of
substantial discount offers on the day. [3]
(vi) Determine the expected sales amount on a discount day under each of the
GLMs fitted in part (iii), explicitly using the estimated coefficients from
each model. 
---
(vii) Comment on the comparison of the answers in part (vi) taking into account
your answer in part (v). [1]
[Total 23]
END OF PAPER