---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---




3 A random variable X with values in the interval [0,1] has the following Cumulative
Distribution Function (CDF):
ğ¹ôˆºğ‘¥ôˆ» ô€µŒ 1 ô€µ† ôˆº1 ô€µ† ğ‘¥ô€¯”ôˆ»ô€¯•, ğ‘¥ âˆˆ ôˆ¾0,1ôˆ¿
and inverse CDF:
ğ¹ô€¬¿ô€¬µôˆºğ‘¢ôˆ» ô€µŒ ô€µ¬1 ô€µ† ôˆº1 ô€µ† ğ‘¢ôˆ»
ô€¬µ
ô€¯•ô€µ°
ô€¬µô€¯”
where ğ‘ ô€µ 0 and ğ‘ ô€µ 0 are parameters. The Probability Density Function (PDF) is
then given by:
ğ‘“ôˆºğ‘¥ôˆ» ô€µŒ ğ‘ğ‘ğ‘¥ô€¯”ô€¬¿ô€¬µôˆº1 ô€µ† ğ‘¥ô€¯”ôˆ»ô€¯•ô€¬¿ô€¬µ, ğ‘¥ âˆˆ ôˆ¾0,1ôˆ¿
You do not need to check the form of the inverse CDF or PDF.

Exercies

(i) Plot the inverse CDF ğ¹ô€¬¿ô€¬µôˆºğ‘¢ôˆ» for an appropriate range of values of u when
a = 0.7 and b = 0.5. [5]
ôˆºiiôˆ» Simulate 1,000 values of X for a = 0.7 and b = 0.5 using the inverse transform
method and store the 1,000 simulated values in a vector in R for later use.

You must use the command set.seed(123) to initialise the random
number generator before you start the simulation. [3]
(iii) Plot a histogram of the 1,000 simulated values obtained in part (ii) using
relative frequencies. [2]
(iv) Plot the PDF f(x) of X for a = 0.7 and b = 0.5 for values of ğ‘¥ âˆˆ ôˆ¾0,1ôˆ¿. The
graph of the PDF should be superimposed on the histogram produced in
part (iii). 

(v) Comment on the plot in part (iv) comparing the simulated values of X with the
PDF, also taking into account the size of the simulated sample. [2]
(vi) Plot the PDF f(x) of X for a = 0.7 and b = 0.5 for values of ğ‘¥ âˆˆ ôˆ¾0,1ôˆ¿ in a new
plot, and add the PDF f(x) for a = 3 and b = 2 to the same plot. [4]
A statistician wants to use one of the PDFs in part (vi) as a prior density for a
Bayesian analysis.
(vii) Comment on the prior beliefs about the underlying parameter that the two
PDFs represent. 

---