---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


5 A sample of 100 daily sales amounts at a store is collected, with the aim to investigate
if daily sales amounts are affected by whether or not the store has substantial
discount offers available on a particular day. The data are available in the file
‘sales_data.RData’, which contains the following variables for 100 days:
 sales.amount – the daily sales amount
 discount – an indicator, showing whether or not the store has substantial
discount offers available on each day (discount = 1 if discounts are available;
discount = 0 otherwise).

---
(i) Plot appropriate boxplots for comparing the sales amounts on discount and
non-discount days. [3]
[Hint: You may find the boxplot(formula, …) R command useful.]
(ii) Comment on the effect of discount offers on sales amounts using the boxplots
produced in part (i). [2]
A sales analyst considers two distributional assumptions: a normal and a gamma
model for the daily sales amounts.
(iii) Fit two separate Generalised Linear Models (GLMs) to the data, one for each
of the two distributional assumptions, to investigate the dependence of the
daily sales amounts on availability of substantial discount offers on the day.
For each model you should use the canonical link function. 

---

(iv) Comment on the impact of the availability of substantial discount offers on
daily sales amounts based on the output of the two models fitted in part (iii).
Your answer should include a relevant p-value and interpretation of the
estimate of the coefficient of the discount variable for each fitted model. 

---

(v) Determine which of the two models fitted in part (iii) should be preferred for
investigating the dependence of daily sales amounts on the availability of
substantial discount offers on the day. [3]
(vi) Determine the expected sales amount on a discount day under each of the
GLMs fitted in part (iii), explicitly using the estimated coefficients from
each model. 
---
(vii) Comment on the comparison of the answers in part (vi) taking into account
your answer in part (v). [1]
[Total 23]

**Understanding the Problem:**

We're tasked with:
1. **Visualizing Data:** Creating boxplots to compare sales on discount and non-discount days.
2. **Fitting GLMs:** Building two GLMs (Normal and Gamma) to model the effect of discount offers on sales.
3. **Model Comparison:** Choosing the better-fitting model.
4. **Predicting Sales:** Calculating expected sales for a discount day using both models.

**R Code Implementation:**

```R
# Load the data
load("sales_data.RData")

# (i) Plot boxplots
boxplot(sales.amount ~ discount, data = sales_data, 
        xlab = "Discount Offer", ylab = "Sales Amount")

# (ii) Comment on the effect of discount offers
# ... (Visually inspect the boxplots)

# (iii) Fit GLMs
# Normal GLM
model1 <- glm(sales.amount ~ discount, data = sales_data, family = gaussian)

# Gamma GLM
model2 <- glm(sales.amount ~ discount, data = sales_data, family = Gamma(link = "inverse"))

# (iv) Comment on the impact of discount offers
summary(model1)
summary(model2)
# ... (Interpret the coefficients and p-values)

# (v) Determine the better model
# ... (Compare AIC, BIC, or use hypothesis testing)

# (vi) Expected sales on a discount day
# Normal GLM
expected_sales1 <- predict(model1, newdata = data.frame(discount = 1))

# Gamma GLM
expected_sales2 <- predict(model2, newdata = data.frame(discount = 1), type = "response")

# (vii) Comment on the comparison
# ... (Compare the expected sales from both models and discuss the implications)
```

**Interpreting the Results:**

**Parts (i) and (ii):**
- The boxplots visually compare the distribution of sales amounts for discount and non-discount days.
- If the boxplot for discount days is significantly higher, it suggests that discounts have a positive impact on sales.

---

**Parts (iii) and (iv):**
- The GLMs model the relationship between sales and the discount indicator.
- The coefficient of the `discount` variable represents the expected change in sales when the discount is present.
- The p-value associated with the coefficient indicates the statistical significance of the effect.

---

**Part (v):**
- Model comparison can be done using information criteria (AIC, BIC) or hypothesis testing.
- The model with the lower AIC or BIC is generally preferred.
- Hypothesis testing can be used to compare the models' fit to the data.

---

**Parts (vi) and (vii):**
- The expected sales on a discount day can be calculated using the predicted values from the GLMs.
- The comparison of the expected sales from the two models can provide insights into the potential impact of the choice of distribution on the estimated effect of discounts.

---


**Additional Considerations:**
- **Model Assumptions:** Ensure that the assumptions of the GLMs (e.g., linearity, independence of errors, normality of residuals for the Normal GLM) are met.
- **Outliers and Influential Points:** Identify and handle any outliers or influential points that might affect the model's fit.
- **Model Diagnostics:** Use diagnostic plots (e.g., residual plots, Q-Q plots) to assess the model's fit and identify potential issues.

By following these steps and carefully interpreting the results, you can gain valuable insights into the impact of discount offers on daily sales.

END OF PAPER