---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 1
---


2 An investment analyst is interested in the relationship between the share price of a
company and its Earnings-Per-Share (EPS).
A sample of data for eight companies has been collected. This is stored in the file
‘CompanySample.RData’. After loading this data into R, the data frame
CompanySample will be available that contains two columns:
 price – the company’s share price (in $)
 EPS – the company’s earnings-per-share (in $).
(i) Fit a linear model to the sample, with price as the dependent variable and
EPS as the independent variable, and state the estimated coefficients of the
model. [3]
(ii) State the value of the coefficient of determination, R2, for the model fitted
in part (i). [1]
(iii) Perform an analysis of variance, using the anova(…) command to test if
there is a linear relationship between price and EPS. Your answer should
include the F-statistic and p-value of the test. [4]
(iv) Comment on how well the linear model in part (i) fits the data, using your
answers to parts (ii) and (iii). [2]
The analyst believes that the observation where price = 99 is an anomaly because of
the large size of the value of price and should be removed from the sample.
(v) Fit a linear model to the sample without the observation price = 99 and state
the value of R2. [3]
[Hint: The code CompanySample[-n,] returns the data frame with the
nth observation removed.]
(vi) Compare the fit of the models from parts (i) and (v). [3]
(vii) Discuss if removing the observation where price = 99 is appropriate. [3]

---

**Understanding the Problem**

We're tasked with:
1. **Fitting a Linear Model:** Building a linear regression model to predict share price based on EPS.
2. **Model Evaluation:** Assessing the model's fit using R-squared and ANOVA.
3. **Outlier Analysis:** Identifying and handling a potential outlier.
4. **Model Comparison:** Comparing the original model with the one after outlier removal.

**R Code Implementation**

```R
# Load the data
load("CompanySample.RData")

# (i) Fit the linear model
model1 <- lm(price ~ EPS, data = CompanySample)
summary(model1)

# (ii) R-squared
summary(model1)$r.squared

# (iii) ANOVA
anova(model1)

# (iv) Comment on the fit
# ... (Interpret R-squared and p-value from ANOVA)

# (v) Remove the outlier and fit the model
CompanySample_new <- CompanySample[-which(CompanySample$price == 99), ]
model2 <- lm(price ~ EPS, data = CompanySample_new)
summary(model2)$r.squared

# (vi) Compare the models
# ... (Compare R-squared values and discuss the impact of the outlier)

# (vii) Discuss outlier removal
# ... (Discuss the justification for removing the outlier and potential consequences)
```

**Interpreting the Results**

**Parts (i) and (ii):**
- The estimated linear regression model is of the form:
  `price = intercept + slope * EPS`
- The estimated coefficients (intercept and slope) can be obtained from the `summary(model1)` output.
- R-squared measures the proportion of variance in the dependent variable (price) explained by the independent variable (EPS).

**Part (iii):**
- The ANOVA test assesses the significance of the regression model.
- The F-statistic and p-value from the `anova(model1)` output can be used to determine if the linear relationship is statistically significant.

**Part (iv):**
- A high R-squared value indicates a good fit, meaning the model explains a large proportion of the variation in price.
- A low p-value from the ANOVA test suggests that the linear relationship is statistically significant.

**Parts (v) and (vi):**
- By removing the outlier, we can assess its impact on the model's fit.
- Compare the R-squared values of the two models to see if the outlier significantly affected the model's performance.
- If the R-squared value increases significantly after removing the outlier, it suggests that the outlier was indeed influential.

**Part (vii):**
- Discuss the potential reasons for the outlier (e.g., measurement error, unusual event).
- Consider the impact of removing the outlier on the model's generalizability.
- If the outlier is truly an anomaly, removing it may improve the model's predictive accuracy. However, if it's a legitimate data point, removing it could lead to bias.
- It's often recommended to use robust regression techniques or other methods to handle outliers, rather than simply removing them.

**Note:**
- The specific numerical results and interpretations will depend on the actual data in the `CompanySample.RData` file.
- Always visualize the data (e.g., scatter plot) to gain insights into the relationship between the variables and identify potential outliers.
- Consider other diagnostic tools like residual plots to assess the model's assumptions and identify other potential issues.


---

