---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
Q. 2) You have been asked to analyse the number of deaths per month in the country $X$ from lung 
disease over the period from Jan 2015 to Dec 2020. This information is contained in a time 
series called <tt>Lung_Deaths.csv<tt>.

1.  Convert the data into a time series format representing the above period after loading it into R. 
2.  Plot this time series giving appropriate labels for each axis.
3.   Plot, on two separate graphs, the sample autocorrelation function (sample ACF) and sample partial autocorrelation function (sample PACF) of the series with appropriate 
labelling of the axes. 
4.  Justify the presence of seasonality in the data based on (2.  and (3.   above.
5.  Apply seasonal differencing to the time series to remove the element of seasonality. 
6.   Plot the ACF and PACF of this time series generated in (5.  above up to lag k = 5 years, 
giving appropriate labels for each axis. 
7.  Comment on your plots in part (6.  , making reference to the stationarity.

---


### Part A
```{r}
Lung_Deaths <- read.csv("D:\\IAI ### Question  Paper\\July 2022\\CS2B_Final\\Lung_
Deaths.csv")
```



```{r}

Lung_Deaths<-ts(Lung_Deaths, start =c(2015,1), frequency = 12)
Lung_Deaths
```
## Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec
## 2015 3035 2552 2704 2554 2014 1655 1721 1524 1596 2074 2199 2512
## 2016 2933 2889 2938 2497 1870 1726 1607 1545 1396 1787 2076 2837
## 2017 2787 3891 3179 2011 1636 1580 1489 1300 1356 1653 2013 2823
## 2018 3102 2294 2385 2444 1748 1554 1498 1361 1346 1564 1640 2293
## 2019 2815 3137 2679 1969 1870 1633 1529 1366 1357 1570 1535 2491
## 2020 3084 2605 2573 2143 1693 1504 1461 1354 1333 1492 1781 1915

---

### Part B

```{r}
plot(Lung_Deaths, ylab = "Lung Deaths", main = "Lung deaths between 2015 and 
2020")
```

---
# (iiPart A

```{r}
acf(Lung_Deaths, lag.max = 30,main = "ACF of Lung Deaths")
pacf(Lung_Deaths, lag.max = 30,main = "PACF of Lung Deaths")
```

---

* There is seasonality in monthly data for which high values tend always to 
occur in some particular months and low values tend always to occur in other 
particular months. In this case, S = 12 (months per year) is the span of the 
periodic seasonal behavior
* If significant correlation persists over many lags in the ACF functions, It 
could either be a truly random autocorrelation in the series or there is a 
fixed effect or trend that hasn't been removed (i.e. the series is not 
stationary). Hence non stationary data

* Time Series of Lung Deaths is establishing seasonality which is visible 
from highs and lows at uniform intervals of time
* ACF is exhibiting a high positive serial correlation at Lag 1 indicating 
the presence of seasonality in the data

---
### Part E

```{r}
diff12=diff(Lung_Deaths, lag = 12, differences=1)
plot(diff12)

````

---
### Part F

```{r}
acf(diff12,lag.max = 30)
```


```{r}
pacf(diff12,lag.max = 30)
````

---

### Part G

* One summary statistic of a stationary time series is the auto-correlation function, or the ACF

* If significant correlation persists over many lags in the ACF functions, It could either be a truly random autocorrelation in the series or there is a fixed effect or trend that hasn't been removed (i.e. the series is not stationary). Here that is not the case and hence stationary data


* Even after the seasonal difference, we observe significant correlation with 
lag 12 both in the ACF and the PACF, indicating seasonal AR first order could 
be an appropriate model for the data

* The blue dotted lines on the ACF and PACF indicate cut-offs for significance.

#For a stationary time-series the ACF should decay to zero quickly and display no signs of oscillation.

#The ACF looks to cut out at lag 2 and does not contain any periodic oscillation so this would indicate stationarity.
#The PACF shows no significance past lag 2. This again, indicates stationarit
y.

---


**1. Loading and Converting Data to Time Series Format**

```{r}
# Load the data
lung_deaths <- read.csv("Lung_Deaths.csv")

# Convert to a time series object
lung_deaths_ts <- ts(lung_deaths, start = c(2015, 1), end = c(2020, 12), frequency = 12)
```

**2. Plotting the Time Series**

```{r}
plot(lung_deaths_ts, xlab = "Year", ylab = "Number of Deaths", main = "Monthly Lung Deaths")
```

**3. Plotting ACF and PACF**

```{r}
par(mfrow=c(1,2))
acf(lung_deaths_ts, lag.max = 24, main = "ACF of Lung Deaths")
pacf(lung_deaths_ts, lag.max = 24, main = "PACF of Lung Deaths")
```

---

**4. Justifying Seasonality**

The ACF plot shows significant spikes at lags 12, 24, and potentially higher multiples of 12. 
This indicates a strong seasonal pattern with a period of 12 months. The PACF plot also supports this, showing significant spikes at lags 1 and 12.

---

**5. Applying Seasonal Differencing**

```{r}
lung_deaths_diff <- diff(lung_deaths_ts, lag = 12)
```

---

**6. Plotting ACF and PACF of the Differenced Series**

```{r}
par(mfrow=c(1,2))
acf(lung_deaths_diff, lag.max = 60, main = "ACF of Differenced Lung Deaths")
pacf(lung_deaths_diff, lag.max = 60, main = "PACF of Differenced Lung Deaths")
```

---


**7. Commenting on the Plots**

The ACF and PACF plots of the differenced series show a more rapid decay, suggesting that the seasonal component has been effectively removed. 
However, there might still be some residual autocorrelation, particularly at lower lags. This could indicate the need for further differencing or other transformations to achieve stationarity. 

It's important to note that while seasonal differencing can help to remove seasonal patterns, it can also introduce other issues, such as increased variability in the data. Therefore, it's crucial to carefully consider the impact of differencing on the time series and to explore other potential modeling approaches, such as SARIMA models, to capture both seasonal and non-seasonal patterns.

---