---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
[25]
Q. 3) The number of customers arriving to a grocery store can be modelled by a Poisson process 
with rate of 30 customers per hour.

#### Exercises 
A.  Find the probability that there are 2 customers between 10:00 AM and 10:15 AM. 
2.  Find the probability that there are 7 customers between 11:00 AM and 11:20 AM
and 15 customers between 11:20 AM and 12 noon. (
3.   Prepare a probability distribution table for different number of customers (0, 1, 2…20) at 
any given 10 minute time interval. (5)

---

Solution 3
### Part A 
dpois(2,7.5)
## [1] 0.0155555
 [2]
 
---

# Part B

* Assume Independent Events
* **A**:  7 customers between 11:00 AM and 11:20 AM
* **B**:  15 customers between 11:20 AM and 12 noon

dpois(7,10)*dpois(15,20)
## [1] 0.004652489
 [4]
 
---
 
# Part C

```{r}
probs<-c()
for (i in 0:20) {
   p <- dpois(i,5)
   probs<-c(probs,p)
}
```

output <- matrix(NA, nrow = 21, ncol = 2)
output<-as.data.frame(output)
names(output)<-c("Customers","Probability")

output$Customers<-0:20
output$Probability<-probs
output

---
## Customers Probability
## 1 0 6.737947e-03
## 2 1 3.368973e-02
## 3 2 8.422434e-02
## 4 3 1.403739e-01
## 5 4 1.754674e-01
## 6 5 1.754674e-01
## 7 6 1.462228e-01
## 8 7 1.044449e-01
## 9 8 6.527804e-02
## 10 9 3.626558e-02
## 11 10 1.813279e-02
## 12 11 8.242177e-03
## 13 12 3.434240e-03
## 14 13 1.320862e-03
## 15 14 4.717363e-04
## 16 15 1.572454e-04
## 17 16 4.913920e-05
## 18 17 1.445271e-05
## 19 18 4.014640e-06
## 20 19 1.056484e-06
## 21 20 2.641211e-07

---


```{r}
ppois(0:20,5)
```

## A. Probability of 2 customers in 15 minutes

**Given:**
* Average rate (λ) = 30 customers/hour
* Time interval (t) = 15 minutes = 0.25 hours

**Calculating λt:**
* λt = 30 * 0.25 = 7.5

**Using Poisson probability formula:**
P(X = 2) = (e^(-λt) * (λt)^x) / x!
= (e^(-7.5) * 7.5^2) / 2!
≈ 0.0842

## B. Probability of 7 customers in 20 minutes and 15 customers in 20 minutes

**Given:**
* λ = 30 customers/hour
* Time interval (t) = 20 minutes = 1/3 hour

**Calculating λt:**
* λt = 30 * (1/3) = 10

**Using Poisson probability formula:**
P(X = 7 in 20 minutes) = (e^(-10) * 10^7) / 7! ≈ 0.0901
P(X = 15 in 20 minutes) = (e^(-10) * 10^15) / 15! ≈ 0.0347

Since these events are independent, the probability of both occurring is the product of their individual probabilities:
P(7 in 20 minutes and 15 in 20 minutes) = 0.0901 * 0.0347 ≈ 0.0031


---

## C. Probability Distribution Table for 10-Minute Intervals

**Given:**
* λ = 30 customers/hour
* Time interval (t) = 10 minutes = 1/6 hour

**Calculating λt:**
* λt = 30 * (1/6) = 5

---

**Creating the probability distribution table:**

| Number of Customers (X) | Probability P(X) |
|---|---|
| 0 | 0.0067 |
| 1 | 0.0337 |
| 2 | 0.0842 |
| 3 | 0.1404 |
| 4 | 0.1755 |
| 5 | 0.1755 |
| 6 | 0.1463 |
| 7 | 0.1042 |
| 8 | 0.0652 |
| 9 | 0.0363 |
| 10 | 0.0181 |
| 11 | 0.0082 |
| 12 | 0.0034 |
| 13 | 0.0013 |
| 14 | 0.0005 |
| 15+ | 0.0003 |

**Note:** To calculate the probabilities for X ≥ 15, we can use the cumulative distribution function (CDF) of the Poisson distribution or approximate it using the normal distribution.

---


---