---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

[11]
Q .4) Refer to “data_1.csv”.
1.  Write down formulae for the estimated integrated hazard and its estimated variance, using 
the Nelson‐Aalen model.
2.  Load the table into R. Add two new columns to the table and populate them with estimated 
integrated hazard, and the estimated variance for each tj of the given data.
3.   Produce a scatterplot showing the values of the estimated integrated hazard across time tj, 
for j = 0, 1,2, 3,…,20, together with the corresponding 90% confidence interval values.
You need to do a proper labelling of the axes. 

---

Solution 4
# (Part A
# Estimated Integrated Hazard:
# Estimated variance of the estimator of the integrated hazard is calculated 
as:
 [3]
 
---

### Part A

```{r}
file <- read.csv("data_1.csv")

head(file)
```

---

```{r}
dj <- file$dj
nj <- file$nj
tj <- file$tj

```

```{r}
names(file)[5]<-"Censored"
Lambda=cumsum(dj/nj)
Var = cumsum(dj*(nj-dj)/nj^3)
file$Lambda = Lambda
file$Variance = Var
file
```

---


<pre><code>
## j tj nj dj Censored Lambda Variance
## 1 1 3 200 2 NA 0.01000000 0.0000495000
## 2 2 6 198 5 NA 0.03525253 0.0001738173
## 3 3 8 193 3 NA 0.05079657 0.0002531045
## 4 4 11 190 2 NA 0.06132288 0.0003079230
## 5 5 14 188 4 5 0.08259948 0.0004186884
## 6 6 18 179 10 NA 0.13846540 0.0007133528
## 7 7 21 169 3 NA 0.15621688 0.0008165266
## 8 8 23 166 1 NA 0.16224098 0.0008525977
## 9 9 24 165 1 NA 0.16830158 0.0008891060
## 10 10 26 164 2 15 0.18049670 0.0009625597
## 11 11 30 147 3 NA 0.20090487 0.0010985574
## 12 12 34 144 4 NA 0.22868264 0.0012861003
## 13 13 38 140 1 NA 0.23582550 0.0013367563
## 14 14 39 139 2 NA 0.25021399 0.0014387812
## 15 15 40 137 4 2 0.27941107 0.0016456761
## 16 16 43 131 8 NA 0.34047977 0.0020833808
## 17 17 45 123 4 4 0.37300010 0.0023391756
## 18 18 50 115 5 NA 0.41647836 0.0027008095
## 19 19 53 110 2 NA 0.43466018 0.0028630935
## 20 20 55 108 1 NA 0.44391944 0.0029480336
</code></pre>

---

### Part C

```{r eval=FALSE}
sdLambda=sqrt(Var)

plot(c(0,tj),c(0,Lambda),xlim=c(0,55),ylim=c(0,0.5),type="s",
 main="Nelson-Aalen estimator of the integrated hazard",
 xlab="Time t",ylab="Integrated hazard")

lines(tj,Lambda-1.64485*sdLambda, type="s",lty=2)
lines(tj,Lambda+1.64485*sdLambda, type="s",lty=2)
```


---

### Part C

```{r echo=FALSE}
sdLambda=sqrt(Var)

plot(c(0,tj),c(0,Lambda),
 xlim=c(0,55),ylim=c(0,0.5),type="s",
 main="Nelson-Aalen estimator of the integrated hazard",
 xlab="Time t",ylab="Integrated hazard")

lines(tj,Lambda-1.64485*sdLambda, type="s",lty=2)
lines(tj,Lambda+1.64485*sdLambda, type="s",lty=2)
```


---


---


---