---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Q. 6) Refer to the data set “Cricket.csv”. The data contains the following columns
A. Player – Name of the Player
B. Team – The team he represented in IPL 2022
C. Init_Group – His initial group (Batsman/ Bowler/ Alrounder)
D. Runs – Number of runs scored in IPL 2022
E. Ave_Bat – Batting Average
F. Wickets – Number of wickets taken as a bowler
G. Economy – Economy as a bowler

---

Load the data into R and answer the following questions. Name the data frame as “Cricket”.
i) What is the average of “Runs” and “Wickets” for three groups of players based on the 
initial grouping.
ii) Create a copy of “Cricket” and rename it as “Cricket1”. In “Cricket1” keep only the four 
numerical columns and remove the rest of them.

---
Perform a feature scaling on Cricket1 by executing the following code 
```{r}
Cricket1 = as.data.frame(scale (Cricket1))
```
After scaling, set a seed value of 100 using the following code <tt>set.seed(100)</tt>.

--=
iii) Execute k-means clustering algorithm on “Cricket1” and assign it to a variable 
“clust_Cricket”. Print the cluster means of the generated clusters.
iv) Add the cluster memberships of each of the players to the original data frame (Cricket) by 
creating a new column called “***Clust_Membership***”.
v) Rename the cluster memberships as “Batsman”, “***Bowler***” and “Alrounder” by following 
the rules given below:
* The cluster with the highest average runs is renamed as “Batsman”
* The cluster with the highest average number of wickets is renamed as “***Bowler***”
* The remaining cluster is renamed as “Alrounder”
vi) Compute the error rate in clustering by comparing the initial grouping with the clusters 
created.
vii)Comment on the possible reasons for misclassification in the clustering.
[20]
********************