---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Q. 6) Refer to the data set “Cricket.csv”. 

The data contains the following columns

A. Player – Name of the Player
B. Team – The team he represented in IPL 2022
C. Init_Group – His initial group (Batsman/ Bowler/ Alrounder)
D. Runs – Number of runs scored in IPL 2022
E. Ave_Bat – Batting Average
F. Wickets – Number of wickets taken as a bowler
G. Economy – Economy as a bowler

---

### Exercises

Load the data into R and answer the following Questions. Name the data frame as “Cricket”.

1.  What is the average of “Runs” and “Wickets” for three groups of players based on the 
initial grouping.

2.  Create a copy of “Cricket” and rename it as “Cricket1”. In “Cricket1” keep only the four 
numerical columns and remove the rest of them.

---
Perform a feature scaling on Cricket1 by executing the following code 
```{r}
Cricket1 = as.data.frame(scale (Cricket1))
```
After scaling, set a seed value of 100 using the following code <ttset.seed(100)</tt.

---

3.   Execute k-means clustering algorithm on “Cricket1” and assign it to a variable 
“clust_Cricket”. Print the cluster means of the generated clusters.
4.  Add the cluster memberships of each of the players to the original data frame (Cricket) by 
creating a new column called “***Clust_Membership***”.

5.  Rename the cluster memberships as “Batsman”, “***Bowler***” and “Alrounder” by following 
the rules given below:
* The cluster with the highest average runs is renamed as “Batsman”
* The cluster with the highest average number of wickets is renamed as “***Bowler***”
* The remaining cluster is renamed as “Alrounder”

6.   Compute the error rate in clustering by comparing the initial grouping with the clusters 
created.

7.  Comment on the possible reasons for misclassification in the clustering.

---


### Part A

```{r}
Cricket <- read.csv("D:\\IAI ### Question  Paper\\July 2022\\CS2B_Final\\Cricket.c
sv")

```

```{r}
### Part A
runs_avg <- aggregate(Runs~Init_group, data = Cricket, FUN = "mean")
runs_avg
## Init_group Runs
## 1 Alrounder 143.66667
## 2 Batsman 213.91892
## 3 Bowler 16.44444
```
---

```{r}
wickets_avg <- aggregate(Wickets~Init_group, data = Cricket, FUN = "mean")
wickets_avg
## Init_group Wickets
## 1 Alrounder 4.3095238
## 2 Batsman 0.2432432
## 3 Bowler 9.0138889
```

----

### Part B

```{r}
Cricket1 <- Cricket
Cricket1 <- Cricket1[,4:7]
```

---

### Part C

```{r}
Cricket1 = as.data.frame(scale (Cricket1))
set.seed(100)
clust_Cricket<-kmeans(Cricket1,3)
clust_Cricket

```

```{r}
## K-means clustering with 3 clusters of sizes 39, 62, 87
## 
## Cluster means:
## Runs Ave_Bat Wickets Econonmy
## 1 -0.4626058 -0.3628813 1.7012202 -0.8011925
## 2 1.1193502 1.0687167 -0.6221781 0.8735751
## 3 -0.5903229 -0.5989433 -0.3192247 -0.2633925

```

---

```{r}


## Clustering vector:
## [1] 3 2 3 3 3 2 1 3 2 3 2 3 2 3 3 2 2 1 3 3 2 1 3 1 2 3 3 3 3 3 2 3 2 2 

3 1 2
## [38] 3 3 2 1 2 2 3 3 2 3 3 2 3 1 2 3 2 3 3 1 2 2 1 1 3 3 1 3 3 2 1 1 3 3 
2 3 3
## [75] 3 2 1 3 1 2 1 3 3 3 1 2 1 3 2 3 3 3 3 1 1 1 1 3 2 2 2 1 3 2 2 2 3 3 
3 2 1
## [112] 3 2 3 3 3 2 3 2 3 1 1 2 1 3 3 2 3 3 3 2 3 3 2 3 1 3 1 1 2 2 2 2 3 2 
3 2 2
## [149] 2 3 3 3 2 3 3 2 3 3 2 3 2 1 2 3 1 2 3 2 2 1 1 1 2 3 3 3 1 1 2 3 3 3 
3 3 2
## [186] 1 2 1
## 
## Within cluster sum of squares by cluster:
## [1] 50.32514 141.94279 108.10393
## (between_SS / total_SS = 59.8 %)
## 
## Available components:
## 
## [1] "cluster" "centers" "totss" "withinss" "tot.withi
nss"
## [6] "betweenss" "size" "iter" "ifault"

---

### Part D

```{r}
Cricket$Clust_Membership <-clust_Cricket$cluster
```


---

### Part E

```{r}
Cricket$Clust_Membership <-factor(Cricket$Clust_Membership, 
	labels = c("Bowler","Batsman","Alrounder"))
```

---
### Part F

```{r}
table(Cricket$Clust_Membership,Cricket$Init_group)

```
## 
## Alrounder Batsman Bowler
## Bowler 7 0 32
## Batsman 11 51 0
## Alrounder 24 23 40
error<- (32+11+24+23)/nrow(Cricket)
error
## [1] 0.4787234

---

# Part G
# Possible reasons
# 1. Many batsman in realty might have played very few matches and scored ver
y few runs. They might not have got classified as batsmen by the cluster
# 2. Similarly many bowlers also

# 3. When there were only 42 players who were alrounders in the real data, cl
uster considered 87 as alrounders because the players who do not fall into ba
tsmen and bowers were treated as alrounders
# 4. No batsman gor classified as a bowler and vice versa indicating the misclassification is with respect to alrounders
# Rather than considering all players, if we would have considered players wh
o played a minimum number of matches, the clustering accuracy would have impr
oved significantly

---

BLANK

---

Here's the R code to answer the questions and perform the requested analysis:

```{r}
# Load the data (assuming the data is in a CSV file named "Cricket.csv")
Cricket <- read.csv("Cricket.csv")

# 1. Average runs and wickets by group
group_averages <- aggregate(Runs + Wickets ~ Initial.Grouping, Cricket, FUN = mean)
print(group_averages)

# 2. Copy and select numerical columns
Cricket1 <- Cricket[, c(2:5)]

---

# 3. Feature scaling and setting seed
Cricket1 <- as.data.frame(scale(Cricket1))
set.seed(100)

# 4. K-means clustering
clust_Cricket <- kmeans(Cricket1, centers = 3, nstart = 20)
print(clust_Cricket$centers)  # Print cluster means


---

# 5. Add cluster membership and rename
Cricket$Clust_Membership <- clust_Cricket$cluster
Cricket$Clust_Membership <- factor(Cricket$Clust_Membership, 
                                    levels = 1:3, 
                                    labels = c("Batsman", "Bowler", "Alrounder")[order(mean(Cricket[, 2:3] ), decreasing = c(TRUE, FALSE))])

---

# 6. Error rate calculation (assuming "Initial.Grouping" is the true grouping)
misclassified <- sum(Cricket$Clust_Membership != Cricket$Initial.Grouping)
error_rate <- misclassified / nrow(Cricket) * 100
print(paste("Error rate:", error_rate, "%"))

---

# 7. Reasons for misclassification
# Possible reasons for misclassification include:
* **Data limitations:** The data might not perfectly capture the true abilities of players. Factors like playing conditions, team composition, and individual performances on specific days can influence runs and wickets.
* **Number of clusters:** Using only three clusters might not perfectly represent the diverse skillsets of players. More clusters could lead to more accurate groupings.
* **Feature selection:** Only runs and wickets were used for clustering.  Additional features like batting average, bowling average, and strike rate could provide more information for differentiating player types.
* **Choice of distance metric:** K-means uses Euclidean distance. Alternative distance metrics might provide better separation between player types.
```

---


This code performs the following steps:

1. **Loads the data.**
2. **Calculates the average runs and wickets for each group.**
3. **Creates a copy of the data with only numerical columns and performs feature scaling.**
4. **Applies K-means clustering with 3 clusters and prints the cluster means.**
5. **Adds a new column named "Clust_Membership" to the original data frame and assigns cluster labels based on average runs and wickets.**
6. **Calculates the error rate by comparing the initial grouping with the generated clusters.**
7. **Discusses potential reasons for misclassification in the clustering process.**

This approach provides insights into player skillsets, explores potential misclassifications, and highlights the importance of data quality and appropriate feature selection for effective clustering.
