---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---



### Part A

1.  Write a general function to compute the truncated moments for a lognormal distribution 
using k, σ, µ, L (the lower bound) and U (the upper bound)
 Given
σ= 0.7
µ = 1.2
L = 10
U = ∞
2.  Use the function generated in (1.  to compute the first and second order truncated moments.
3.   Using the function in (1.  or otherwise, compute the first two moments for a non-truncated 
normal distribution with the same µ and σ.
4.  Compare the results generated in (2.  and (3.   with appropriate reasoning.

---

Solution 5
# (i)

```{r}
trunc_logN <- function(k, mu, sig, L, U){
 
 Lk <- (log(L) - mu)/sig - k*sig
 Uk <- (log(U) - mu)/sig - k* sig
 cumulate <- exp(k*mu+.5*k^2*sig^2) * (pnorm(Uk) - pnorm(Lk))
 print(paste("First term:", exp(k*mu+.5*k^2*sig^2)))
 print(paste("Phi(Uk): ", pnorm(Uk)))
 print(paste("Phi(Lk): ", pnorm(Lk)))
 
 print(paste("Final Value:", cumulate))
 
}
```

---


# (ii)
```{r}
trunc_logN(1, 1.2, 0.7, 10,Inf)
## [1] "First term: 4.24185214282043"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0.809246116780569"
## [1] "Final Value: 0.809149768285661"
```

```{r}
trunc_logN(2, 1.2, 0.7, 10,Inf)
## [1] "First term: 29.3707711132894"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0.569507941751428"
## [1] "Final Value: 12.6438837089077"
```
---


### Part C
# (iii)
*  First moment of a normal distribution = mu = 1.2
* Second moment of a normal distribution = 0.7^2+1.2^2 = 1.93
* The question should have been "Log normal distribution" instead of normal distribution
* Assuming Lognormal distribution
* Moments of Lognormal distribution are 

```{r}
exp(1.2+0.5*0.7^2)
## [1] 4.241852
exp(2*1.2+0.7^2)*(exp(0.7^2)-1)+(exp(1.2+0.5*0.7^2))^2
## [1] 29.37077
```

---


trunc_logN(1, 1.2, 0.7, 0,Inf)
## [1] "First term: 4.24185214282043"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0"
## [1] "Final Value: 4.24185214282043"
trunc_logN(2, 1.2, 0.7, 0,Inf)
## [1] "First term: 29.3707711132894"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0"
## [1] "Final Value: 29.3707711132894"
----

#### Part D

* The results are not comparable as (ii) is associated with Lognormal distribution while (iii) is associated with normal distribution
* Assuming (iii) is log normal distribution
* The mean and variance of a truncated distribution are very much lower than the non-truncated because
# (i) Beyond 10, the probabilities are very less and the actual range of valu
es in excess of 10 is very limited
# (ii) The consistency improves because of a narrower range and hence lower v
ariance and second order movement

---
