---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
#library(dplyr)
library(janitor)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---



### Part A

1.  Write a general function to compute the truncated moments for a lognormal distribution 
using k, σ, µ, L (the lower bound) and U (the upper bound)
 Given
σ= 0.7
µ = 1.2
L = 10
U = ∞

2.  Use the function generated in (1.  to compute the first and second order truncated moments.

3.   Using the function in (1.  or otherwise, compute the first two moments for a non-truncated 
normal distribution with the same µ and σ.

4.  Compare the results generated in (2.  and (3.   with appropriate reasoning.

---


### Exercise 1

```{r}
trunc_logN <- function(k, mu, sig, L, U){
 
 Lk <- (log(L) - mu)/sig - k*sig
 Uk <- (log(U) - mu)/sig - k* sig
 cumulate <- exp(k*mu+.5*k^2*sig^2) * (pnorm(Uk) - pnorm(Lk))
 print(paste("First term:", exp(k*mu+.5*k^2*sig^2)))
 print(paste("Phi(Uk): ", pnorm(Uk)))
 print(paste("Phi(Lk): ", pnorm(Lk)))
 
 print(paste("Final Value:", cumulate))
 
}
```

---


### Exercise 2

```{r}
trunc_logN(1, 1.2, 0.7, 10,Inf)
## [1] "First term: 4.24185214282043"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0.809246116780569"
## [1] "Final Value: 0.809149768285661"
```

```{r}
trunc_logN(2, 1.2, 0.7, 10,Inf)
## [1] "First term: 29.3707711132894"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0.569507941751428"
## [1] "Final Value: 12.6438837089077"
```
---

### Exercise 3

*  First moment of a normal distribution = mu = 1.2
* Second moment of a normal distribution = 0.7^2+1.2^2 = 1.93
* The question should have been "Log normal distribution" instead of normal distribution
* Assuming Lognormal distribution
* Moments of Lognormal distribution are 

```{r}
exp(1.2+0.5*0.7^2)
```


```{r}
exp(2*1.2+0.7^2)*(exp(0.7^2)-1)+(exp(1.2+0.5*0.7^2))^2
## [1] 29.37077
```

---

```{r}
trunc_logN(1, 1.2, 0.7, 0,Inf)
## [1] "First term: 4.24185214282043"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0"
## [1] "Final Value: 4.24185214282043"
trunc_logN(2, 1.2, 0.7, 0,Inf)
## [1] "First term: 29.3707711132894"
## [1] "Phi(Uk): 1"
## [1] "Phi(Lk): 0"
## [1] "Final Value: 29.3707711132894"

```

----

### Exercise 4

* The results are not comparable as (Part B) is associated with Lognormal distribution while (Part C) is associated with normal distribution
* Assuming (Part C) is log normal distribution
* The mean and variance of a truncated distribution are very much lower than the non-truncated because
* (Part A) Beyond 10, the probabilities are very less and the actual range of valu
es in excess of 10 is very limited
* (Part BThe consistency improves because of a narrower range and hence lower variance and second order movement

---

BLANK

---

---