---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---
[25]
Q. 3) The number of customers arriving to a grocery store can be modelled by a Poisson process 
with rate of 30 customers per hour.

#### Exercises 
A.  Find the probability that there are 2 customers between 10:00 AM and 10:15 AM. 
2.  Find the probability that there are 7 customers between 11:00 AM and 11:20 AM
and 15 customers between 11:20 AM and 12 noon. (
3.   Prepare a probability distribution table for different number of customers (0, 1, 2…20) at 
any given 10 minute time interval. (5)

---

Solution 3
### Part A 
dpois(2,7.5)
## [1] 0.0155555
 [2]
 
---

# Part B

* Assume Independent Events
* **A**:  7 customers between 11:00 AM and 11:20 AM
* **B**:  15 customers between 11:20 AM and 12 noon

dpois(7,10)*dpois(15,20)
## [1] 0.004652489
 [4]
 
---
 
# Part C

```{r}
probs<-c()
for (i in 0:20) {
   p <- dpois(i,5)
   probs<-c(probs,p)
}
```

output <- matrix(NA, nrow = 21, ncol = 2)
output<-as.data.frame(output)
names(output)<-c("Customers","Probability")

output$Customers<-0:20
output$Probability<-probs
output

---
## Customers Probability
## 1 0 6.737947e-03
## 2 1 3.368973e-02
## 3 2 8.422434e-02
## 4 3 1.403739e-01
## 5 4 1.754674e-01
## 6 5 1.754674e-01
## 7 6 1.462228e-01
## 8 7 1.044449e-01
## 9 8 6.527804e-02
## 10 9 3.626558e-02
## 11 10 1.813279e-02
## 12 11 8.242177e-03
## 13 12 3.434240e-03
## 14 13 1.320862e-03
## 15 14 4.717363e-04
## 16 15 1.572454e-04
## 17 16 4.913920e-05
## 18 17 1.445271e-05
## 19 18 4.014640e-06
## 20 19 1.056484e-06
## 21 20 2.641211e-07

---


```{r}
ppois(0:20,5)
```