---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

***CS2B – Risk Modelling and Survival Analysis ***


* The emphasis is placed on being able to apply statistical methods to actuarial problems using real data sets and the open-source software environment R. 

* Time Series. Probability Distributions, Survival Analysis

---

Solution 1:
Corn_Prices<-read.csv("D:\\Monthly_corn.csv")
(i) Close<-ts(Corn_Prices$Close,start = c(2017,1),frequency=12)


```{r}
plot(Close, ylab = "Corn Monthly prices", 
	main = "Monthly plot of corn prices")

```


---

### Part C

The series is not stationary as it appears to be trending upwards with 
time. The mean at different periods appears not to be constant For periods be
tween 2017 to 2020, it appeared to be stationary.

---

### Part D

(iv) acf(Close, main = "ACF", xlab = "Lags", ylab = "Serial Correlation", la
g.max=60)


```{r}
pacf(Close, main = "PACF", xlab = "Lags", ylab = "Serial Correlation", lag.ma
x=60)
```


---

(v) If a series is stationary, the ACF should decay to zero quickly and sho
uld not display any oscillation.The series is not stationary because auto cor
relation function is not decaying to zero quickly. Also, if the number of lag
s is increased, oscillation is also observed, hence cannot be stationary

---

(vi) Create another time series "monthly_returns" by using the formula log(P
t/Pt-1), where Pt and Pt-1 correspond to the closing prices of month t and t1 respectively.
returns<-diff(log(Close))
 
(vii) acf(returns, main = "ACF", xlab = "Lags", ylab = "Serial Correlation", 
lag.max=60)
[1 for the plot, 0.5 for proper labeling of title and axes]
pacf(returns, main = "PACF", xlab = "Lags", ylab = "Serial Correlation", lag.
max=60)

(viii) The monthly returns appear stationary. If a series is stationary, the A
CF should decay to zero quickly and should not display any oscillation.
 
(ix) The most appropriate ARMA will be ARMA (0,0) as the ACF and PACF do not 
show significance at any of the lags [1]
(x) library(forecast)
## Warning: package 'forecast' was built under R version 4.0.5
## Registered S3 method overwritten by 'quantmod':
## method from
## as.zoo.data.frame zoo


m1<-arima(returns,order = c(0,0,0))
m2<-arima(returns,order = c(1,0,0))
m3<-arima(returns,order = c(0,0,1))
m4<-arima(returns,order = c(1,0,1))
m1$aic
## [1] -164.6838
m2$aic
## [1] -163.5281
m3$aic
## [1] -163.3079
m4$aic
## [1] -165.2896
auto.arima(returns)
## Series: returns 
## ARIMA(0,0,0)(1,0,0)[12] with zero mean 
## 
## Coefficients:
## sar1
## 0.2548
## s.e. 0.1120
## 
## sigma^2 estimated as 0.005687: log likelihood=86.37
## AIC=-168.74 AICc=-168.58 BIC=-164.14
# ARMA (1,1) has the lowest AIC.
[0.5 for fitting each of the four models, 2 for identifying the best model]

---

(xi) The deviation is observed because seasonality is not considered in the 
model. ACF is actually demonstrating a small significant serial correlation a
t lag 12, which might give rise to seasonal ARMA. 
# If we are forcefully fitting non seasonal models to a seasonal data, the di
screpancy is possible.
 [1 mark for each valid observation, Max 2]
[24 Marks]