---
title: "Statistics with R"
subtitle: "Statistical Modelling with R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* In a study done at the National Institute of Science and Technology, asbestos fibers on 
filters were counted as part of a project to develop measurement standards for asbestos 
concentration.

* An operator counted the number of fibers in each of 23 grid squares, yielding the 
following counts:


|    |    |    |    |    |    |
|----|----|----|----|----|----|
| 31 | 29 | 19 | 18 | 31 | 28 |
| 34 | 27 | 34 | 30 | 16 | 18 |
| 26 | 27 | 27 | 18 | 24 | 22 |
| 28 | 24 | 21 | 17 | 24 |    |


* Assume that the Poisson distribution with unknown parameter lambda describes the 
variability from each of the grid squares. 

---

#### Exercises

1.  Calculate Q1, Q3 and Inter-quartile range.

2.  Plot histogram of sample data and label it appropriately.

3.   Use the method of maximum likelihood to estimate the parameter lambda.

4.  Test the hypothesis whether the mean fiber count is equal to 25. Comment on the 
results.

5.  Calculate the standard error of parameter lambda. 

6.   Calculate the 90% confidence interval for standard error.

7.  Calculate the probability of fiber count exceeding 30, with the help of Central Limit 
Theorem.

---

### Set Up Data

```{r}
fiber_count <- c(31,29,19,18,31,28, 
           34,27,34,30,16,18, 
           26,27,27,18,24,22, 
           28,24,21,17,24)
```

---


### Exercise 1

- **Q1, Q3, and IQR:**

  - `quantile()` function calculates the quartiles.

  - IQR is calculated as Q3 - Q1.

  - `IQR()` calculates the Inter-quartile range directly
   
```{r}   
Q1 <- quantile(fiber_count,0.25)
Q1
```

```{r}
Q3 <- quantile(fiber_count,0.75)
Q3

```

---


### Exercise 1

- **Q1, Q3, and IQR:**


```{r}
IQR(fiber_count)
```


---

### Exercise 2

**Histogram:**

 - `hist()` function creates a histogram of the fiber counts.

```{r eval=FALSE}
hist(fiber_count)
```

---

### Plot Histogram

```{r eval=FALSE}
hist(fiber_count, 
     main = "Histogram of Fiber Counts", 
     xlab = "Fiber Count", 
     ylab = "Frequency",
     breaks=8)
```

---


### Plot Histogram

```{r echo=FALSE}
hist(fiber_count, 
      main = "Histogram of Fiber Counts", 
     xlab = "Fiber Count", 
     ylab = "Frequency",breaks=8)
```

---

### Exercise 3

**Maximum Likelihood Estimation:**

 - For Poisson distribution, the MLE of lambda is the sample mean.
   

```{r}
lambda.hat=mean(fiber_count)

print(lambda.hat)
```

---

### Exercise 4

**Hypothesis Testing:**
   - `t.test` performs a one-sample t-test to compare the sample mean with the hypothesized mean of 25.

$H_0$: The mean fiber count is 25

$H_1$: Mean fiber count is not equal to 25

```{r}
t.test(fiber_count,mu=25)

```

---

* Based on the p-value the null hypothesis Ho that “the mean fiber count is 25” cannot be rejected. 
* Also 25 lies within the 95% confidence interval.


---

### Exercise 5

**Standard Error of Lambda:**
   - The standard error of the sample mean for a Poisson distribution is the square root of the sample mean divided by the sample size.

```{r}
lambda.hat.sterror = sqrt(lambda.hat/length(fiber_count))

print(lambda.hat.sterror)

```

Remark: For a Poisson Distribution, the standard deviation is always equal to the square root of the mean.i.e σ = μ

Above expression is coded in form: $$\sqrt{\frac{\sigma^2}{n}}.$$ 


---

### Exercise 6

**Confidence Interval for Standard Mean:**
   - We assume a normal approximation for the distribution of the standard error.
   - The confidence interval is calculated using the z-score for the desired confidence level.


```{r}
#quantile
Zs <- c(-1,1)*qnorm(.95)
Zs
```

```{r}
lambda.CI.Limits = lambda.hat + Zs*lambda.hat.sterror

print(lambda.CI.Limits)

```

---

***..Just out of curiosity***

```{r}
t.test(fiber_count,conf.level = 0.90)
```


---

### Exercise 7

```{r}
probability = pnorm(30,
                    mean = lambda.hat,
                    sd = sqrt(lambda.hat),
                    lower.tail = FALSE)

cat("Probability of fiber count exceeding 30:", probability)
```

**Interpretation:**

* The calculated probability represents the likelihood of observing a sample mean greater than 30, assuming the true population mean is equal to the sample mean. 


---

**Note:**

- The CLT works best when the sample size is large enough. In this case, a sample size of 23 might not be considered very large, but it's still reasonable to apply the CLT as an approximation.

- The Poisson distribution is often used to model count data like fiber counts. 

- However, the CLT allows us to approximate the distribution of the sample mean using a normal distribution, even if the underlying distribution is not normal.

---

**Interpretation of Results:**

- The histogram will visualize the distribution of fiber counts.
- The MLE of lambda provides an estimate of the average fiber count per grid square.
- The hypothesis test will determine if there's evidence to reject the null hypothesis that the mean fiber count is 25.
- The standard error quantifies the variability of the MLE.
- The confidence interval for the standard error provides a range of plausible values for the true standard error.

---

---


---
