---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
lib_dir: libs
nature:
  highlightStyle: github
  countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

* Introduction to R for Actuarial Students

* CS1B Curriculum

* Introduction to R programming
* Fundamentals of Statistical Analysis

* Question 3
---

### Annual Average Rainfall Comparison

An analysis was carried out to investigate the annual average rainfall of two countries. 

Here is the rainfall data for Iran and Belgium from 1974 to 1991. 

Some information for Iran is missing.



| Year | Iran | Belgium | Year | Iran | Belgium |
|------|------|---------|------|------|---------|
| 1974 | 128  | 160     | 1983 |      | 150     |
| 1975 | 125  | 128     | 1984 |      | 182     |
| 1976 | 133  | 169     | 1985 |      | 158     |
| 1977 | 104  | 105     | 1986 |      | 156     |
| 1978 | 146  | 151     | 1987 |      | 123     |
| 1979 | 132  | 164     | 1988 |      | 141     |
| 1980 | 125  | 162     | 1989 | 118  | 176     |
| 1981 |      | 177     | 1990 | 129  | 162     |
| 1982 |      | 185     | 1991 | 124  | 172     |

---

The analysis will use the recorded values.

```{r}
Iran <- c(128,125,133,104,146,132,125,118,129,124)

Belgium <- c(160,128,169,105,151,164,162,177,185,150,
             182,158,156,123,141,176,162,172)
```

---


#### Exercises

**A**.  Perform a suitable test to determine whether the rainfall in both countries has equal 
variance or not, at the 5% significance level. 

**B**.  Test whether the mean rainfall in both countries is equal or not, at 5% significance
level. 

**C**.   Calculate the 95% confidence interval for difference in means. 

**D**.  Comment on the results in exercises C and D.


---


### Exercise A. 

**Equal Variance Test**

Performing an F-test to determine if the variances of rainfall between Iran and Belgium are equal.


#### Hypotheses

$H_0$: Variances of annual average rainfall of Belgium and Iran are same

$H_1$: Variances of annual average rainfall of Belgium and Iran are different


```{r eval=FALSE}
var.test(Iran, Belgium)
```

---

```{r echo=FALSE}
var.test(Iran, Belgium)
```


### Conclusion

With the p-value < 0.05, we reject the null hypothesis, concluding that the variances are not equal.



---

### Exercise B.

**Mean Rainfall Hypothesis Test**

Using a t-test with unequal variances to test if the mean rainfall is the same:

```{r}
t.test(Iran, Belgium, var.equal = FALSE)
```
Since the p-value < 0.05, we reject the null hypothesis and conclude that the mean rainfall in Iran and Belgium is different with 95% confidence.

---


```{r echo=FALSE}
t.test(Iran, Belgium, var.equal = FALSE)
```



### Conclusion

Since, P-value<0.05 we reject the null hypothesis and can conclude that both cities have different amount 
of rainfall with 95% confidence.

---

### Exercise C

* The 95% confidence interval for the difference in means is provided in the output of the t-test:

```
95 percent confidence interval:

-42.79403 to -17.85041
```

---

### Exercise D

**Commentary**

- The confidence interval (-42.8, -17.8) does not contain 0, indicating that the mean rainfall is significantly different between the two countries.

- This result aligns with the conclusion from the t-test in Exercise B, affirming that Iran and Belgium have different average annual rainfall.


---

---